### 2024-04-02 Notes/Notes.tex
Initial commit after a bit of work

\documentclass[12pt]{article}
\usepackage{}
\title{A480 Notes}
\author{Brayden Leicester}
\begin{document}
\maketitle

\section{Things to Check}
The named asteroids in \cite{Harris1994} (most notably 4179 Toutatis) should be checked for TESS observations. As these should tumble.

288 Glauke \cite{Harris2015}

\section{Background writing}
%TODO 1I observation paragraph/s

The first interstellar object discovered in late 2017 has come to be called 1I/'Omuamua. \dots

%TODO TESS intro
The Transiting Exoplanet Survey Satellite (TESS) \cite{Ricker2014} is a large area, high cadence, imaging \dots

\section{Method Notes}

\bibliographystyle{apalike}
\bibliography{bibfile.bib}
\end{document}

### 2024-04-02 Test Code/expectedPositions.py

@@ -1,0 +1,112 @@
+"""expectedPostions.py
+
+Queries MPC and gets the positions of asteriods in a TESS field in a defined number of timesteps. Plots them in RA and Dec space
+
+B Leicester, 2/4/2024
+
+Last Edited 2/4/24
+
+"""
+#TODO speed up MPC query.
+#! problem is keeping them sequential
+
+#TODO colour by time better
+#! problem is colour is chosen from cmap uniformly from number of coord pair for that asteriod, not from times themselves
+
+import numpy as np
+import matplotlib.pyplot as plt
+import itertools
+from astropy.coordinates import SkyCoord
+import astropy.units as u
+from astropy.time import Time
+from find_asteroid import MPCquery
+import pandas as pd
+
+
+def plotExpectedPos(posArr, times, lims, names):
+    """Plots each asteroid as a sepeate (hopfully) marker
+    Input:
+        posArr: A 3d ndArray, with each object having its own RA and DEC collumns that are the lenght of *times*, to get each asteriod as a "line" of points
+
+        times: A 1d arrayLike, meant as a sequence for colouring, actual time is not needed, as that is defined by TESS sector. Though it can be used as the same list should be used to query MPC
+
+        lims: 4 float list, for the x and y min and max limits for the plot
+
+        names: 1d arrayLike, should be the same lenght as the 1st dimension of *posArr*    
+    
+        
+    Outputs:
+        plt.figure of the tracks of the asteroids
+    """
+    #Sets up plot
+    fig, ax = plt.subplots(1,figsize =(12,12))
+    ax.set_xlabel("RA")
+    ax.set_ylabel("Dec")
+    ax.set_xlim([lims[0],lims[1]])
+    ax.set_ylim([lims[2],lims[3]])
+
+    markers = itertools.cycle((".","x","^","*","D","+","v","o", "<", ">","H","1","2","3","4","8","X","p","d","s","P","h")) #marker cycle
+    
+    #plots each asteroid
+    for i in range(posArr.shape[0]):
+        ax.scatter(posArr[i,0],posArr[i,1], marker=next(markers), c=times, cmap="plasma", label = names[i])
+    
+    plt.scatter(ra_i, dec_i, marker="+", s=100, c="g") #Center of seach area
+
+    # ax.legend() #! Not really working, stops after a few symbols
+
[…]

### 2024-04-02 Test Code/find_asteroid.py

@@ -1,0 +1,198 @@
+from astropy.wcs import WCS
+from astropy.coordinates import SkyCoord
+import astropy.units as u
+import numpy as np
+from astroquery.jplhorizons import Horizons
+from astropy.utils.data import download_file
+import pandas as pd
+from astropy.time import Time
+from tqdm import tqdm
+import requests
+
+def locate_asteroid(id,wcs,epoch,obscode='474',MPC=True,maglim=30):
+
+    if type(id) == str:
+        number = False
+    else:
+        number = True
+    radius = _get_max_angle(wcs)
+    ra, dec = _get_center_coord(wcs)
+    if MPC:
+        results = MPCquery(ra,dec,epoch,radius*60,obscode=obscode,limit=str(maglim))
+    else:
+        results = Skybotquery(ra,dec,epoch,radius*2,obscode)
+
+    if type(results) == type(None):
+        print('No asteroids found')
+        return None
+    if MPC:
+        ind = np.zeros(len(results)) > 1
+        for i in range(len(results)):
+            if str(id).lower() in results.iloc[i]['name'].lower():
+                ind[i] = True
+    else:
+        if number:
+            ind = results['Num'].values == ' {} '.format(id)
+        else:
+            ind = results['Name'].values == ' {} '.format(id)
+    if sum(ind) == 0:
+        print('no targets by that id. Possible targets are shown below')
+        print(results)
+    elif sum(ind) > 1:
+        print('multiple targets by that id. Possible targets are shown below, please specify')
+        print(results[ind])
+    asteroid = results.iloc[ind]
+    c = SkyCoord(asteroid['RA'].values,asteroid['Dec'].values,unit=(u.hourangle, u.deg))
+    position = wcs.all_world2pix(c.ra.deg,c.dec.deg,0)
+    return np.array(position).flatten()
+
+
+
+def _get_max_angle(wcs):
+    foot = wcs.calc_footprint()
+    ra = foot[:,0]
+    dra = ra[:,np.newaxis] - ra[np.newaxis,:]
+    rang = np.nanmax(dra)
+
+    dec = foot[:,1]
+    ddec = dec[:,np.newaxis] - dec[np.newaxis,:]
+    dang = np.nanmax(ddec)
+    radius = np.nanmax([rang,dang])
+    return radius
+
+def _get_center_coord(wcs):
+    y,x = wcs.array_shape
+    ra, dec = wcs.all_pix2world(x/2,y/2,0)
[…]

### 2024-04-04 Test Code/expectedPositions.py
skybot and MPC tracks

@@ -15,6 +15,7 @@ Last Edited 2/4/24
 
 import numpy as np
 import matplotlib.pyplot as plt
+import matplotlib.colors as mplc
 import itertools
 from astropy.coordinates import SkyCoord
 import astropy.units as u
@@ -46,16 +47,17 @@ def plotExpectedPos(posArr, times, lims, names):
     ax.set_ylim([lims[2],lims[3]])
 
     markers = itertools.cycle((".","x","^","*","D","+","v","o", "<", ">","H","1","2","3","4","8","X","p","d","s","P","h")) #marker cycle
-    
+    cnorm = mplc.Normalize(np.min(times), np.max(times))
     #plots each asteroid
     for i in range(posArr.shape[0]):
-        ax.scatter(posArr[i,0],posArr[i,1], marker=next(markers), c=times, cmap="plasma", label = names[i])
+        ax.plot(posArr[i,0],posArr[i,1], c="grey", alpha=0.4)
+        ax.scatter(posArr[i,0],posArr[i,1], c=times, marker=next(markers), cmap="plasma", norm=cnorm, label = names[i])
     
     plt.scatter(ra_i, dec_i, marker="+", s=100, c="g") #Center of seach area
 
     # ax.legend() #! Not really working, stops after a few symbols
 
-    fig.savefig(f"./posTracks_ra{ra_i}_dec{dec_i}_t{t_i}.png")
+    # fig.savefig(f"./posTracks_ra{ra_i}_dec{dec_i}_t{t_i.mjd}.png")
 
 #sets up positions and time
 #TODO get _i vals as inputs?    

### 2024-04-05 Test Code/expectedPositions.py
skybot making sense now, index problems.

@@ -53,11 +53,11 @@ def plotExpectedPos(posArr, times, lims, names):
         ax.plot(posArr[i,0],posArr[i,1], c="grey", alpha=0.4)
         ax.scatter(posArr[i,0],posArr[i,1], c=times, marker=next(markers), cmap="plasma", norm=cnorm, label = names[i])
     
-    plt.scatter(ra_i, dec_i, marker="+", s=100, c="g") #Center of seach area
+    ax.scatter(ra_i, dec_i, marker="+", s=100, c="g") #Center of seach area
 
     # ax.legend() #! Not really working, stops after a few symbols
 
-    # fig.savefig(f"./posTracks_ra{ra_i}_dec{dec_i}_t{t_i.mjd}.png")
+    fig.savefig(f"./posTracks_ra{ra_i}_dec{dec_i}_t{t_i.mjd}_Mv{magLim}.png")
 
 #sets up positions and time
 #TODO get _i vals as inputs?    
@@ -66,11 +66,12 @@ numTimesteps = 27 #! often small as takes a long time
 t_i=Time("2020-04-17T00:00:00.000", format='isot', scale='utc')
 dt = (27/numTimesteps)*u.day
 timeList = t_i + dt*np.arange(0,numTimesteps)
+magLim = 20.0
 
 #MPC query/ies
 resList = []
 for pos, time in enumerate(timeList.jd):  
-    queryRes = MPCquery(ra_i,dec_i,time, 300, limit="18.0",obscode="C57") #! This is the time sink, especially with so many times
+    queryRes = MPCquery(ra_i,dec_i,time, 300, limit=str(magLim),obscode="C57") #! This is the time sink, especially with so many times
     resList.append(queryRes)
 
 #Gets the unique asteroid names, to be used to get tracks of the same asteroid with time

### 2024-04-05 Test Code/find_asteroid.py

@@ -128,7 +128,7 @@ def Skybotquery(ra, dec, times, radius=10/60, location='474',
             else:
                 df = df.append(res)
     if df is not None:
-        df.reset_index(drop=True)
+        df.reset_index(drop=True) #! should have inplace=True...
     return df
 
 

### 2024-06-02 Notes/Notes.tex
some writing

@@
 \documentclass[12pt]{article}
-\usepackage{}
+\usepackage{siunitx, natbib, hyperref,graphicx}
+
+\DeclareSIUnit{\arcsec}{''}
+
 \title{A480 Notes}
@@
 
+%*Pos gathering
+To check for asteroids in the TESS data, the positions of the asteroids with time are required.
+For most asteroids, their orbital elements are well known, so it is a matter of looking them up and cross-matching with transients in the TESS data.
+However, querying APIs for timesteps of \qty{30}{\min} or shorter is
+prohibitively expensive, especially when they rate limit their calls.
+Python was used to make API calls to {Skybot}\footnote{\href{https://vo.imcce.fr/webservices/skybot/}{Skybot}} to get positions of asteroids in a cone search box in RA and Dec space.
+As TESS sectors are 27%?Check
+days long, querying every \qty{12}{\hour} is manageable.
+These positions are still very sparsely spaced in time compared to the TESS data, so an interpolation is needed to bridge the gap.
+With TESS data coming in $\frac12\unit{\hour}$ chunks, 24 interpolated points are needed between each API call.
+Assuming this is fine is justified, as asteroids move at close to a TESS pixel per TESS frame \citep{Pal2018,Pal2020}.
+For the faster TESS data, more interpolated points are needed, but a smaller the change in position between each point.
+
+In another series of API calls, this time querying {JPL Horizons}\footnote{\href{https://ssd.jpl.nasa.gov/horizons/}{JPL Horizons}} by object name, the orbital elements of the asteroids can be found.
+From these elements the type of asteroids (i.e.
+main belt, NEOs, Jupiter Trojan etc.) can be determined by plotting the semi-major axis, $a$, against the inclination, $i$, and the eccentricity, $e$, of the body.
+Osculating elements, such as the heliocentric distance, $r_h$ and the distance to the observer, delta $\delta$%?check
+can also be found from Horizons.
+Plotting inclination against a change in these distances with time can give an idea of the motion of the bodies.
+The JPL queries can also be used to check the validity of the interpolation from Skybot.
+By geting the RA and Dec values of a named asteroid for all the timesteps interpolated, a $\Delta$RA against $|Delta$Dec plot can be made.
+This was heavily rate limited, as the epoch querying did not want to play ball with date formats, but what was produced showed that the interpolation was accurate to within an arcsecond most of the time, the earliest points were the most astray, but this was still within three or four arcseconds, much smaller than the \qty{21}{\arcsec} TESS pixels.
+
+%*Interpolation
+
+%*Cross matching
+Having interpolated these positions, there was a well sampled set of RA, Dec and time values of where these asteroids should be in the TESS data.
+They should show up in a pixel for a small number of frames, of order $\sim1$.
+This is the same as a lot of other transient events, a sharp rise in brightness and then disapering quickly again.
+The number of frames do change, type Ia supernova will brighten is a matter of a few hours and then dim for days, while stellar flares are of similar profile by a smaller max brightness and a corrsepondingly shorter decay time.
+Asteroids are very short, however detection pipelines are robust.
+These pipelines have already found the aforementioned supernova and stellar flares, the job of this work is to catch all the asteroids in the set of all the transients.
+To do this, catelogue matching is in order.
+
+%*Period fit
+
 \bibliographystyle{apalike}

### 2024-06-28 Notes/Notes.tex
same as last commit that didn't grab everything??????????

@@
 \documentclass[12pt]{article}
-\usepackage{siunitx, natbib, hyperref,graphicx}
+\usepackage{natbib, hyperref,graphicx}
+
+\usepackage[separate-uncertainty=true,multi-part-units=single]{siunitx}
 
 \DeclareSIUnit{\arcsec}{''}
+
+\usepackage[utf8]{inputenc}
+\usepackage{newunicodechar}
+%\usepackage{libertine}
+
+\DeclareRobustCommand{\okina}{%
+  \raisebox{\dimexpr\fontcharht\font`A-\height}{%
+    \scalebox{0.8}{`}%
+  }%
+}
+\newunicodechar{ʻ}{\okina}
+
+\newcommand{\omuamua}{\okina Omuamua}
 
@@
 \section{Background writing}
+
+%TODO TESS intro
+The Transiting Exoplanet Survey Satellite (TESS) \cite{Ricker2014} is a large area, high imaging cadence, space telescope. 
+TESS is tasked with staring at the same piece of sky for \qty{27}{\day} at a time (a sector), delivering \qtyproduct{96 x 24}{\degree} full frame images (FFIs) at regular intervals. 
+These FFIs are built from stacked \qty{2}{\second} exposures, leveraging the short readout times of the 16 CCD cameras onboard. 
+With the initial cadence for these full frame images set to \qty{30}{\minute}, the time resolution of TESS is unparalleled. 
+This does come at the cost of spatial resolution, as the pixels are each \qty{21}{\arcsec} square. 
+A Nyquist frequency of \unit{\per\hour} is well sampled enough to characterise most variable stars, as well as find the orbital periods of exoplanets to a high precision. 
+After the initial mission had mapped the entire sky, TESS started to reduce the time of the full frame images, down to \qty{10}{\minute} and then down to only \qty{200}{\second}.
+
+Of interest here is what such a high sampling rate can do for statistic on the asteroid population. 
+For bright asteroids the rotation periods should be able to be easily determined from this vast dataset. 
+The shortest FFIs will be able to accurately determine the rotation periods of all but the fastest rotating asteroids, most of which will be too dim to see in the TESS data. 
+There have been attempts before to find and classify the asteroids in TESS data before by \citet{Pal2018, Pal2020}.
+This work aims to extend there study to more sectors, and a to use a different data reduction method, using the \texttt{TESSreduce} package \citep{Ridden-Harper2021}.
+As part of a full sky transient survey using TESS, \texttt{TESSELLATE} (Ridden-Harper and Roxburgh et al., in prep), asteroids pop out as transient objects that are spikes in brightness of pixels for only a few frames.
+They can confuse pipelines looking for short time stellar transients such as flare stars and supernova. 
+As such, a way of filtering them out is required, as well as this filtering will get the properties of the lightcurves, such as the rotation periods and amplitude variation of these small solar system bodies.
+A full sky, self consistent catalogue of these properties is important for many reasons. 
+The asteroid population statistics are useful in their own right, understanding the orbits of near earth objects is useful for planetary defence, and these bodies \dots
+
 %TODO 1I observation paragraph/s
 
-The first interstellar object discovered in late 2017 has come to be called 1I/'Omuamua. \dots
+The first interstellar object (ISO) was discovered in late 2017, it has come to be called 1I/\omuamua \citep[see][for a review]{Bannister2019}.
+\omuamua was determined to be spectroscopically red \citep{Fitzsimmons2017, Meech2017}, and having a photometric colour in the neutral end of the solar system range \citep{Bannister2017}. 
+It appeared to be tumbling \citep[e.g.][]{Drahus2018,Fraser2018}, the double peaked light curve was found to have a rotation period of \qty{8.67(0.34)}{\hour} \citep{Belton2018}.
 
-%TODO TESS intro
-The Transiting Exoplanet Survey Satellite (TESS) \cite{Ricker2014} is a large area, high cadence, imaging \dots
+Peak to peak LC of 2.5 mag %TODO cite
+
+Combing the tumbling with an elongation ratio of up to \qty{6(1)}{}:1 \citep{McNeill2018}, 1I is said to have a cigar shape \citep{Belton2018}.
+
+1I was classed as asteroid due to lack of a coma, and no noticable activity %TODO cite
 
@@
 The JPL queries can also be used to check the validity of the interpolation from Skybot.
-By geting the RA and Dec values of a named asteroid for all the timesteps interpolated, a $\Delta$RA against $|Delta$Dec plot can be made.
+By geting the RA and Dec values of a named asteroid for all the timesteps interpolated, a $\Delta$RA against $\Delta$Dec plot can be made.
 This was heavily rate limited, as the epoch querying did not want to play ball with date formats, but what was produced showed that the interpolation was accurate to within an arcsecond most of the time, the earliest points were the most astray, but this was still within three or four arcseconds, much smaller than the \qty{21}{\arcsec} TESS pixels.
@@
 \bibliography{bibfile.bib}
+
 \end{document}

### 2024-06-28 Test Code/compareResults.py

@@ -41,6 +41,9 @@ def useKDTree(df1, df2, cols, maxDist:float=0.01,maxTimeSep:float=0.1,k:int=1):
 
     df2 = df2.sort_values("distToMatch").drop_duplicates(subset=["IDMatch"]).sort_index() #drops all the duplicate IDs, as very point should be 1:1 not n:1
 
+    df2.reset_index(drop=False, inplace=True)
+    df2.rename(columns={"index":"DetectIDs"}, inplace=True)
+
     return df2
  
 
@@ -66,7 +69,13 @@ def useKDTree(df1, df2, cols, maxDist:float=0.01,maxTimeSep:float=0.1,k:int=1):
 #     return cat1_id,cat2_id
 
 
-interpRes = pd.read_csv("./InterpolatedQuerryResultTESSdata_Sector2_Cam1_Ccd1_Cut1of16_wcs.csv")
+sector = 29
+cam = 1
+ccd = 3
+cut = 7
+
+
+interpRes = pd.read_csv(f"./InterpolatedQuerryResult_{sector}_{cam}_{ccd}_{cut}.csv")
 
 
 # #TODO get csv from Ryan of problems. 
@@ -88,14 +97,14 @@ interpRes = pd.read_csv("./InterpolatedQuerryResultTESSdata_Sector2_Cam1_Ccd1_Cu
 # randRes = pd.DataFrame({"Namerand": resForRand["Name"],"RArand":randRAs, "Decrand":randDecs, "epochrand":randTimes}) #into a df.
 #detectedSources = randRes
 
-detectedSources = pd.read_csv("./TESSdata_Sector2_Cam1_Ccd1_Cut1of16_detected_sources.csv", usecols=["ra", "dec", "mjd", "flux", "Type"])
+detectedSourcesAll = pd.read_csv(f"../OzData/{sector}_{cam}_{ccd}_{cut}_detected_sources.csv", usecols=["ra", "dec", "mjd", "flux", "Type"])
 
 #     indexs= interpRes.index[interpRes["Name"]==name]
 #     cutItrpDf = interpRes.loc[indexs]
 
-indexsOfInterest = detectedSources.index[detectedSources["Type"] == "0"]
+indexsOfInterest = detectedSourcesAll.index[detectedSourcesAll["Type"] == "0"]
 
-detectedSources = detectedSources.loc[indexsOfInterest]
+detectedSources = detectedSourcesAll.loc[indexsOfInterest]
 
 detectedSources["jd"] = detectedSources["mjd"] +2400000.5
[…]

### 2024-06-28 Test Code/expectedPositionsSkyBot.py

@@ -243,10 +243,18 @@ def plotExpectedPos(posDf: pd.DataFrame, timeList: npt.ArrayLike, targetPos: lis
         norm=cnorm, cmap=cmap), ax=ax, pad=0.1)
     clb.ax.set_title("Time in MJD", fontsize=16, y=1.03)
     clb.ax.ticklabel_format(useOffset=False)
-    centerPix = SkyCoord(ra_i, dec_i, unit=(u.deg, u.deg)).to_pixel(
-        w)  # center of seach area in pixels
+    centerPix = SkyCoord(ra_i, dec_i, unit=(u.deg, u.deg)).to_pixel(w)  # center of seach area in pixels
     ax.scatter(centerPix[0], centerPix[1], marker="+",
                s=100, c="g")  # center marker
+    
+    #*BOX to check 3.2 deg 
+    # corners = np.array([[ra_i+1.6, dec_i+1.6],[ra_i-1.6, dec_i+1.6],[ra_i-1.6, dec_i-1.6],[ra_i+1.6, dec_i-1.6],[ra_i+1.6, dec_i+1.6]])
+
+    # cornersPix =  SkyCoord(corners, unit=(u.deg, u.deg)).to_pixel(w) 
+
+    # ax.plot(cornersPix[0], cornersPix[1])
+
+
     ax.text(-37, -10, s=scaleStr, fontsize=18)
 
     fig.tight_layout()
@@ -348,7 +356,13 @@ def plotHorizons(nameList: list[str], t_i:Time, t_f:Time| None = None,loc:str="5
 
     return fig
 
-fname = "./TESSdata_Sector2_Cam1_Ccd1_Cut1of16_wcs.fits"
+sector = 29
+cam = 1
+ccd = 3
+cut = 7
+
+
+fname = f"../OzData/{sector}_{cam}_{ccd}_{cut}_wcs.fits"
 
 targetWSC = fits.open(fname)[0]
 
@@ -364,8 +378,9 @@ res, times = querySB(myTargetPos, magLim=magLim, numTimesteps=54, qRad=3.2)
 
 # res.to_csv(f"./querryResult_ra{myTargetPos[0]}_dec{myTargetPos[1]}_t{myTargetPos[2].mjd}_Mv{magLim}.csv")
 
-# posFig = plotExpectedPos(res, times, myTargetPos, magLim=magLim, scaleAlpha=True)
+posFig = plotExpectedPos(res, times, myTargetPos, magLim=magLim, scaleAlpha=True)
[…]

### 2024-06-28 Test Code/periodDetection.py

@@ -1,0 +1,48 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas as pd 
+import lightkurve as lk
+
+def detect_period(times, fluxes, fluxErrs):
+    """
+    Calculates the periodogram and returns the period at the max power.
+    """
+    lc =lk.LightCurve(data=None, time=times, flux=fluxes, flux_err=fluxErrs)
+    lc = lc.normalize()
+    pg = lc.to_periodogram()
+    period = pg.period_at_max_power
+
+    return period.value
+
+def load_matches(sector, cam, ccd, cut):
+    matchesDF = pd.read_csv(f"./{sector}_{cam}_{ccd}_{cut}_InterpAndDetect_Matches.csv")
+    matchesDF.drop(columns="Unnamed: 0", inplace=True)
+    return matchesDF
+
+def name_cut(df, name:str, colName:str="NameMatch"):
+    "take a df and gives  only the values where the colName == name"
+    return df.loc[df.index[df[colName]==name]]
+
+
+matchesDF = load_matches(29,1,3,7)
+
+unqNames = pd.unique(matchesDF["NameMatch"])
+
+periodDict = {}
+
+
+for name in unqNames:
+    nameDf = name_cut(matchesDF, name)
+    
+    if nameDf.shape[0]<2:
+        continue
+    time = nameDf["mjd"].values
+    flux = nameDf["flux"].values
+    fluxErr = np.ones_like(flux)*1e-4
+    period = detect_period(time,flux,fluxErr)
+    periodDict[name] = period
+
+
+print(periodDict)
+
+plt.scatter(np.arange(len(periodDict)),periodDict.values())

### 2024-06-28 Test Code/periodDetection.py
the largest periods are concering

@@ -30,11 +30,15 @@ unqNames = pd.unique(matchesDF["NameMatch"])
 
 periodDict = {}
 
+badCount = 0
 
 for name in unqNames:
     nameDf = name_cut(matchesDF, name)
     
-    if nameDf.shape[0]<2:
+    numObser = 2
+    if nameDf.shape[0]<numObser:
+        badCount+=1
+        print(f"{badCount}. not enough points (<{numObser}) for {name}")
         continue
     time = nameDf["mjd"].values
     flux = nameDf["flux"].values
@@ -44,5 +48,15 @@ for name in unqNames:
 
 
 print(periodDict)
+print(len(unqNames))
 
-plt.scatter(np.arange(len(periodDict)),periodDict.values())
+fig, ax = plt.subplots()
+
+ax.scatter(np.arange(len(periodDict)),periodDict.values())
+ax.set_ylabel("Period [days]")
+ax.set_xlabel("Index")
+
+
+maxPName = max(periodDict, key=periodDict.get)
+maxP = periodDict[maxPName]
+print(maxPName, maxP)

### 2024-07-04 Notes/Notes.tex
some more notes

@@
 To do this, catelogue matching is in order.
+Using the \texttt{KDTree} algorithm \citep{Maneewongvatana1999} as implemented in \texttt{SciPy} \citep{2020SciPy-NMeth} %TODO cite
+, the RA and Dec values of the interpolated asteroid positions can be queried for each of the unknown detections in the small cut of a TESS sector of interest. 
+The smallest time gap between any match is \qty{0.1}{\day}, this is to stop any matches that are coincident spatialy but very far apart in time and is intentionally longer than one TESS FFI. 
+Then, if multiple detection match to the same interpolated point, a cut is made by smallest distance, in coordinate space, and any distance must be smaller than \qty{0.01}{\degree} (\qty{36}{\arcsec}), which is within 2 TESS pixels.
 
 %*Period fit
+From these matches, the flux of the detection can be plotted against the time of the event, this is a textbook lightcurve.
+Using the \texttt{Lightkurve} package \citep{Lightkurve2018}, a Lomb-Scargle periodigram can be formed, and the most likely period can be determined.
+\texttt{Lightkurve} was built for TESS data (as well as for the Kepler mission), so it is the best fit for the analysis here.
 

### 2024-07-04 Test Code/compareResults.py

@@ -23,7 +23,7 @@ def useKDTree(df1, df2, cols, maxDist:float=0.01,maxTimeSep:float=0.1,k:int=1):
     dists, indices =  tree.query(df2[cols].values, k=k)
 
     fts = [c for c in df1.columns]
-    df2["distToMatch"] = dists
+    df2["distToMatch"] = dists 
     for c in fts:
         df2[f"{c}Match"]= df1[c].values[indices]
 

### 2024-07-11 Test Code/compareResults.py
another cut with better? wcs

@@ -69,16 +69,16 @@ def useKDTree(df1, df2, cols, maxDist:float=0.01,maxTimeSep:float=0.1,k:int=1):
 #     return cat1_id,cat2_id
 
 
-sector = 29
+sector = 22
 cam = 1
 ccd = 3
-cut = 7
+cut = 8
 
 
 interpRes = pd.read_csv(f"./InterpolatedQuerryResult_{sector}_{cam}_{ccd}_{cut}.csv")
 
 
-# #TODO get csv from Ryan of problems. 
+# #// TODO get csv from Ryan of problems. 
 # # will have to rename vars to be readable
 # #*gets random offset to make matches from
 # obsPos = [301.60, -38.68, Time("2020-04-17T00:00:00.000", format='isot', scale='utc')]
@@ -200,7 +200,7 @@ for name in unqNames:
         plt.scatter(nameCut["Time"], nameCut["flux"], label=name)
         plt.legend()
 
-    if name == " 2001 UR124 " or name ==" Henry ":
+    if name == " Ruff " or name ==" Lincoln ":
         nameCut.to_csv(f"./{name}Matches.csv")
 
 

### 2024-07-11 Test Code/expectedPositionsSkyBot.py

@@ -121,8 +121,17 @@ def querySB(targetPos: list, qRad: float = 10.0, qLoc: str = "C57", numTimesteps
     dt = (27/numTimesteps)*u.day  # a dt in days
     # list of times constructed sequentially
     timeList = t_i + dt*np.arange(0, numTimesteps)
-    result = _Skybotquery(ra_i, dec_i, timeList.jd,
+    result = pd.DataFrame()
+    
+    while len(result)<numTimesteps:
+        
+        try:
+            result = _Skybotquery(ra_i, dec_i, timeList.jd,
                           radius=qRad, location=qLoc, cache=True)
+        except:
+            continue
+    
+
     brightResult = result.loc[result["Mv"] <= magLim].reset_index(drop=True)
 
     coords = SkyCoord(
@@ -356,10 +365,10 @@ def plotHorizons(nameList: list[str], t_i:Time, t_f:Time| None = None,loc:str="5
 
     return fig
 
-sector = 29
+sector = 22
 cam = 1
 ccd = 3
-cut = 7
+cut = 8
 
 
 fname = f"../OzData/{sector}_{cam}_{ccd}_{cut}_wcs.fits"

### 2024-07-11 Test Code/periodDetection.py

@@ -3,6 +3,16 @@ import matplotlib.pyplot as plt
 import pandas as pd 
 import lightkurve as lk
 
+plt.rcParams.update({
+    "font.size": 18,
+    "font.family": "serif",
+    "figure.autolayout": True,
+    "axes.grid": False,
+    # "xtick.minor.visible": True,
+    # "ytick.minor.visible": True,
+})
+
+
 def detect_period(times, fluxes, fluxErrs):
     """
     Calculates the periodogram and returns the period at the max power.
@@ -24,7 +34,7 @@ def name_cut(df, name:str, colName:str="NameMatch"):
     return df.loc[df.index[df[colName]==name]]
 
 
-matchesDF = load_matches(29,1,3,7)
+matchesDF = load_matches(22,1,3,8)
 
 unqNames = pd.unique(matchesDF["NameMatch"])
 
@@ -32,10 +42,12 @@ periodDict = {}
 
 badCount = 0
 
+numObser = 2
+
 for name in unqNames:
     nameDf = name_cut(matchesDF, name)
     
-    numObser = 2
+    
     if nameDf.shape[0]<numObser:
         badCount+=1
         print(f"{badCount}. not enough points (<{numObser}) for {name}")
@@ -47,7 +59,7 @@ for name in unqNames:
     periodDict[name] = period
 
 
-print(periodDict)
+# print(periodDict)
 print(len(unqNames))
 
 fig, ax = plt.subplots()

### 2024-07-12 Notes/Notes.tex
fixed time offset and started progess report

@@
 \newunicodechar{ʻ}{\okina}
-
 \newcommand{\omuamua}{\okina Omuamua}
@@
 The Transiting Exoplanet Survey Satellite (TESS) \cite{Ricker2014} is a large area, high imaging cadence, space telescope. 
-TESS is tasked with staring at the same piece of sky for \qty{27}{\day} at a time (a sector), delivering \qtyproduct{96 x 24}{\degree} full frame images (FFIs) at regular intervals. 
+TESS is tasked with observing one piece of sky for \qty{27}{\day} at a time (a sector), delivering \qtyproduct{96 x 24}{\degree} full frame images (FFIs) at regular intervals. 
 These FFIs are built from stacked \qty{2}{\second} exposures, leveraging the short readout times of the 16 CCD cameras onboard. 
@@
 There have been attempts before to find and classify the asteroids in TESS data before by \citet{Pal2018, Pal2020}.
-This work aims to extend there study to more sectors, and a to use a different data reduction method, using the \texttt{TESSreduce} package \citep{Ridden-Harper2021}.
+This work aims to extend their study to more sectors, and to use a different data reduction method, using the \texttt{TESSreduce} package \citep{Ridden-Harper2021}.
 As part of a full sky transient survey using TESS, \texttt{TESSELLATE} (Ridden-Harper and Roxburgh et al., in prep), asteroids pop out as transient objects that are spikes in brightness of pixels for only a few frames.
@@
 To do this, catelogue matching is in order.
-Using the \texttt{KDTree} algorithm \citep{Maneewongvatana1999} as implemented in \texttt{SciPy} \citep{2020SciPy-NMeth} %TODO cite
+Using the \texttt{KDTree} algorithm \citep{Maneewongvatana1999} as implemented in \texttt{SciPy} \citep{2020SciPy-NMeth} %\\TODO cite
 , the RA and Dec values of the interpolated asteroid positions can be queried for each of the unknown detections in the small cut of a TESS sector of interest. 
-The smallest time gap between any match is \qty{0.1}{\day}, this is to stop any matches that are coincident spatialy but very far apart in time and is intentionally longer than one TESS FFI. 
+The largest time gap between any match is \qty{0.1}{\day}, this is to stop any matches that are coincident spatialy but very far apart in time and is intentionally longer than one TESS FFI. 
 Then, if multiple detection match to the same interpolated point, a cut is made by smallest distance, in coordinate space, and any distance must be smaller than \qty{0.01}{\degree} (\qty{36}{\arcsec}), which is within 2 TESS pixels.

### 2024-07-12 Progress Report/progressReportDraft.tex
fixed time offset and started progess report

\documentclass[12pt]{article}

\usepackage{natbib,hyperref,graphicx}
\usepackage[separate-uncertainty=true,multi-part-units=single]{siunitx}

\usepackage[margin=1.5cm, left=3cm,twoside]{geometry}

\DeclareSIUnit{\arcsec}{''}

\usepackage[utf8]{inputenc}
\usepackage{newunicodechar}
%\usepackage{libertine}

\DeclareRobustCommand{\okina}{%
  \raisebox{\dimexpr\fontcharht\font`A-\height}{%
    \scalebox{0.8}{`}%
  }%
}
\newunicodechar{ʻ}{\okina}
\newcommand{\omuamua}{\okina Omuamua }

\title{ASTR480 Progress Report}
\author{Brayden Leicester \\[1ex]
\small{Supervisors: Michele Bannister and Ryan Ridden-Harper}
}
\begin{document}
\maketitle

%*Aim / background summary

%To get all asteriods in TESS. 
This project aims to find and characterise the lightcurves of all the asteroids seen by the Transiting Exoplanet Survey Satellite (TESS). 
\section{Background}
    %Why TESS
TESS is a large area, high imaging cadence, space telescope  \citep{Ricker2014}. 
TESS is tasked with observing one piece of sky for \qty{27}{\day} at a time (a sector), delivering \qtyproduct{96 x 24}{\degree} full frame images (FFIs) at regular intervals. 
With the initial cadence for these full frame images set to \qty{30}{\minute}, the time resolution of TESS is unparalleled, with a Nyquist frequency of \qty{1}{\per\hour}, as the mission was extended the lenght of the FFIs has come down to \qty{10}{\minute} and then \qty{200}{\second}. 
This high time resolution and observation area does come at the cost of spatial resolution, as the pixels are each \qty{21}{\arcsec} square.
There have been attempts before to find and classify the asteroids in TESS data before by \citet{Pal2018, Pal2020}. This work aims to extend their study to more sectors, and to use a different data reduction method, the \texttt{TESSreduce} package \citep{Ridden-Harper2021}.
Because of the survey properties, TESS provides a selfconsitent was to measure the properties of asteroids over the full sky.   
Another beneficial part of this work is that as part of a full sky transient survey using TESS, \texttt{TESSELLATE} (Ridden-Harper and Roxburgh et al., in prep), asteroids are transient objects that spike the brightness of a pixel for only a few frames.
The goal of finding all the asteroids will allow for the removal of these spikes from the transient pipeline, as well as to understand the asteroid population better.

    %Why Asteroids
Asteroids are a key class of solar system objects. 
Understanding their rotation properties has long been of interest to astronomers \citep[e.g.][]{Weidenschilling1981,Harris1994}.

        %Omuamua
High amplitude variation has come to the forefront of questions about asteroid properties because of the first interstellar object (ISO) 1I/\omuamua \citep[see][for a review]{Bannister2019}. 
\omuamua was measured to have a rotation period of \qty{8.67(0.34)}{\hour} \citep{Belton2018} and seemed to be tumbling \citep[e.g.][]{Drahus2018,Fraser2018}. 
The peak to peak amplitude variation of 2.5 mag %TODO cite
on the double peaked light curve is of interest, as this is much higher than most asteroids. 
With the full sky survey of bright asteroids, we hope to find many asteroids with such a large amplitude variation, and to see just how rare \omuamua is.

\section{Work So Far}
%* Method / Initial Results

%Querry and interpolate

%Match to detections
Matching these interpolated positions to \texttt{TESSELLATE} detections is important to lower the unknown transient outputs of this pipeline. 
Using the \texttt{KDTree} algorithm \citep{Maneewongvatana1999} as implemented in \texttt{SciPy} \citep{2020SciPy-NMeth}, the right ascension (RA) and  declination (Dec) coordinates of the interpolated points and the detections can be compared and matched together. 
Filtering this \texttt{KDTree} output by not allowing the time between spatialy coincident matches to be longer than \qty{0.1}{\day}, \dots

%Lightcurves; detected VS forced interpolated

\section{Future Work}
%* Future work

%Periods and amplitueds for everything detected

%Run on server
The \texttt{TESSELLATE} pipeline has been running on the OzSTAR supercomputing facilities. 
After I am confident that all the parts of the asteriod detection and subsequent lightcurve analysis works as required, the same code can be refactored to work on OzSTAR and a largescale analysis of all of the processed TESS sectors can be run. 
Only after this has completed can the asteriod population statistics can be computed.

\section{Figures}
%*Figures:

%*Bib
\def\bibfont{\tiny}
\bibliographystyle{jphysicsB}
\bibliography{../Notes/bibfile.bib}
\end{document}

### 2024-07-12 Test Code/compareResults.py

@@ -102,7 +102,9 @@ detectedSourcesAll = pd.read_csv(f"../OzData/{sector}_{cam}_{ccd}_{cut}_detected
 #     indexs= interpRes.index[interpRes["Name"]==name]
 #     cutItrpDf = interpRes.loc[indexs]
 
-indexsOfInterest = detectedSourcesAll.index[detectedSourcesAll["Type"] == "0"]
+print(pd.unique(detectedSourcesAll["Type"]))
+
+indexsOfInterest = detectedSourcesAll.index[(detectedSourcesAll["Type"] == "0") | detectedSourcesAll["Type"] == "AST"] #TODO figure out what code is for asteroid.
 
 detectedSources = detectedSourcesAll.loc[indexsOfInterest]
 

### 2024-07-18 Progress Report/progressReportDraft.tex
progress on progress report, and some COM fluxes

@@
 \usepackage[margin=1.5cm, left=3cm,twoside]{geometry}
+
+\usepackage[para]{footmisc}
+
+\usepackage[compact]{titlesec}
+\titlespacing{\section}{0pt}{2ex}{1ex}
+\titlespacing{\subsection}{0pt}{1ex}{0ex}
+\titlespacing{\subsubsection}{0pt}{0.5ex}{0ex}
 
@@
 Asteroids are a key class of solar system objects. 
-Understanding their rotation properties has long been of interest to astronomers \citep[e.g.][]{Weidenschilling1981,Harris1994}.
-
-        %Omuamua
+Understanding their rotation properties has long been of interest to astronomers \citep[e.g.][]{Weidenschilling1981,Harris1994}. %Omuamu
 High amplitude variation has come to the forefront of questions about asteroid properties because of the first interstellar object (ISO) 1I/\omuamua \citep[see][for a review]{Bannister2019}. 
 \omuamua was measured to have a rotation period of \qty{8.67(0.34)}{\hour} \citep{Belton2018} and seemed to be tumbling \citep[e.g.][]{Drahus2018,Fraser2018}. 
-The peak to peak amplitude variation of 2.5 mag %TODO cite
-on the double peaked light curve is of interest, as this is much higher than most asteroids. 
+The peak to peak amplitude variation of 2.5 mag \citep{Meech2017} of the double peaked light curve is of interest, as this is much higher than most asteroids, and it implies a large axis ratio. 
 With the full sky survey of bright asteroids, we hope to find many asteroids with such a large amplitude variation, and to see just how rare \omuamua is.
@@
 %Querry and interpolate
+To check for asteroids in the TESS data, the positions of the asteroids with time are required.
+For most asteroids, their orbital elements are well known, so it is a matter of looking them up and cross-matching with transients in the TESS data.
+Python was used to make API calls to {Skybot}\footnote{\href{https://vo.imcce.fr/webservices/skybot/}{Skybot}} to get positions of asteroids in a cone search box in RA and Dec space.
+As TESS sectors are \qty{27}{\day} long, querying every \qty{12}{\hour} is manageable.
+These positions are still very sparsely spaced in time compared to the TESS data, so an interpolation is needed to bridge the gap.
+With TESS data coming in $\frac12\unit{\hour}$ chunks, 24 interpolated points are needed between each API call.
+This interpolation should be accurate, as asteroids move at close to a TESS pixel per TESS frame on average \citep{Pal2018,Pal2020}, and checking against a higher frequency query to {JPL Horizons}\footnote{\href{https://ssd.jpl.nasa.gov/horizons/}{JPL Horizons}} confirmed this.
+For the faster TESS data, more interpolated points are needed, but a smaller the change in position between each point.
 
@@
 Using the \texttt{KDTree} algorithm \citep{Maneewongvatana1999} as implemented in \texttt{SciPy} \citep{2020SciPy-NMeth}, the right ascension (RA) and  declination (Dec) coordinates of the interpolated points and the detections can be compared and matched together.
-Filtering this \texttt{KDTree} output by not allowing the time between spatialy coincident matches to be longer than \qty{0.1}{\day}, \dots
+Filtering this \texttt{KDTree} output by not allowing the time between spatially coincident matches to be longer than \qty{0.1}{\day} stops any accidental matches in position from non-asteroids.
 
 %Lightcurves; detected VS forced interpolated 
+There are two sets of points to take lightcurves from. 
+The matches from the detections, which already have a flux calculated, or the interpolated points themselves, which are more numerous but require forcing the photometry. 
+Not every interpolated point gets a match, due to a variety of reasons, %pass in front of stars, dip below limiting mag, bad pipeline, not extreme enough of a difference between frames 
+so a comparison between the two lightcurves is interesting. 
+There were some challenges getting the flux of the interpolated points, even when \texttt{TESSELLATE} had already reduced all the FFIs of interest, due to the timing of the TESS frames, but these were identified and corrected for.
 
@@
 %Periods and amplitueds for everything detected
+The next part of my analysis has to do with determining the periods and amplitudes of each asteroid's lightcurve.
+For this there are a few methods I can try, using the \texttt{Lightkurve} package built for period analysis of TESS (and Kepler) data of variable stars, or peel back a layer of abstraction and use the Lomb-Scargle periodogram as implemented by \texttt{Astropy}.
+Some trialling of both methods is needed, as preliminary testing reveals of interesting similarities and differences between the packages.
 
@@
 The \texttt{TESSELLATE} pipeline has been running on the OzSTAR supercomputing facilities. 
-After I am confident that all the parts of the asteriod detection and subsequent lightcurve analysis works as required, the same code can be refactored to work on OzSTAR and a largescale analysis of all of the processed TESS sectors can be run.
-Only after this has completed can the asteriod population statistics can be computed.
+After I am confident that all the parts of the asteroid detection and subsequent lightcurve analysis works as required, the same code can be refactored to work on OzSTAR and a large-scale analysis of all the processed TESS sectors can be run.
+Only after this has completed can the asteroid population statistics can be computed. 
+I will be looking for completeness of detections of asteroids, as well as accuracy of periods and amplitude variation.
 
@@
 
+%H-scaled pos plot of higher cam/less asteroids. 
+%Both Lightcurves (interp and detect)
+%? another one?
+
 %*Bib
+% \newgeometry{twocolumn,margin=1.5cm, left=3cm,twoside}
+% \newpage
 \def\bibfont{\tiny}

### 2024-07-18 Test Code/expectedPositionsSkyBot.py

@@ -124,7 +124,7 @@ def querySB(targetPos: list, qRad: float = 10.0, qLoc: str = "C57", numTimesteps
     result = pd.DataFrame()
     
     while len(result)<numTimesteps:
-        
+
         try:
             result = _Skybotquery(ra_i, dec_i, timeList.jd,
                           radius=qRad, location=qLoc, cache=True)
@@ -264,7 +264,7 @@ def plotExpectedPos(posDf: pd.DataFrame, timeList: npt.ArrayLike, targetPos: lis
     # ax.plot(cornersPix[0], cornersPix[1])
 
 
-    ax.text(-37, -10, s=scaleStr, fontsize=18)
+    if scaleAlpha: ax.text(-37, -10, s=scaleStr, fontsize=18)
 
     fig.tight_layout()
 
@@ -387,7 +387,7 @@ res, times = querySB(myTargetPos, magLim=magLim, numTimesteps=54, qRad=3.2)
 
 # res.to_csv(f"./querryResult_ra{myTargetPos[0]}_dec{myTargetPos[1]}_t{myTargetPos[2].mjd}_Mv{magLim}.csv")
 
-posFig = plotExpectedPos(res, times, myTargetPos, magLim=magLim, scaleAlpha=True)
+posFig = plotExpectedPos(res, times, myTargetPos, magLim=magLim, scaleAlpha=False)
 # posFig.savefig(f"./ExpectedPositionsPlot_ra{myTargetPos[0]}_dec{myTargetPos[1]}_t{myTargetPos[2].mjd}_Mv{magLim}.png")
 
 # eleFig = plotHorizons(unqNames, times[0], plotAEI=True)
@@ -429,7 +429,7 @@ for i, name in enumerate(unqNames):
 interpRes = pd.concat(dfsList) #puts evrything back together
 interpRes.reset_index(drop=True, inplace=True)
 # # interpRes.to_csv(f"./InterpolatedQuerryResult_ra{myTargetPos[0]}_dec{myTargetPos[1]}_t{myTargetPos[2].mjd}_Mv{magLim}.csv")
-interpRes.to_csv(f"./InterpolatedQuerryResult_{sector}_{cam}_{ccd}_{cut}.csv")
+#* interpRes.to_csv(f"./InterpolatedQuerryResult_{sector}_{cam}_{ccd}_{cut}.csv")
 
[…]

### 2024-07-25 Notes/Notes.tex
progress on progress report, and some COM fluxes
talk slides started, progress on periods, found edge cases in alldata

@@
 \newunicodechar{ʻ}{\okina}
-\newcommand{\omuamua}{\okina Omuamua}
+\newcommand{\omuamua}{\okina Oumuamua}
 
@@
 TESS is tasked with observing one piece of sky for \qty{27}{\day} at a time (a sector), delivering \qtyproduct{96 x 24}{\degree} full frame images (FFIs) at regular intervals. 
-These FFIs are built from stacked \qty{2}{\second} exposures, leveraging the short readout times of the 16 CCD cameras onboard. 
+These FFIs are built from stacked \qty{2}{\second} exposures, leveraging the short readout times of the 16 CCD cameras on-board. 
 With the initial cadence for these full frame images set to \qty{30}{\minute}, the time resolution of TESS is unparalleled. 
@@
 They can confuse pipelines looking for short time stellar transients such as flare stars and supernova. 
-As such, a way of filtering them out is required, as well as this filtering will get the properties of the lightcurves, such as the rotation periods and amplitude variation of these small solar system bodies.
-A full sky, self consistent catalogue of these properties is important for many reasons. 
+As such, a way of filtering them out is required, as well as this filtering will get the properties of the light curves, such as the rotation periods and amplitude variation of these small solar system bodies.
+A full sky, self-consistent catalogue of these properties is important for many reasons. 
 The asteroid population statistics are useful in their own right, understanding the orbits of near earth objects is useful for planetary defence, and these bodies \dots
@@
 
-Peak to peak LC of 2.5 mag %TODO cite
+Peak to peak LC of 2.5 mag \citep{Meech2017} %TODO cite
 
@@
 For most asteroids, their orbital elements are well known, so it is a matter of looking them up and cross-matching with transients in the TESS data.
-However, querying APIs for timesteps of \qty{30}{\min} or shorter is
-prohibitively expensive, especially when they rate limit their calls.
+However, querying APIs for timesteps of \qty{30}{\min} or shorter is prohibitively expensive, especially when they rate limit their calls.
 Python was used to make API calls to {Skybot}\footnote{\href{https://vo.imcce.fr/webservices/skybot/}{Skybot}} to get positions of asteroids in a cone search box in RA and Dec space.

### 2024-07-25 Progress Report/progressReportDraft.tex
talk slides started, progress on periods, found edge cases in alldata

--- 2024-07-12 Progress Report/progressReportDraft.tex
@@
 
-\usepackage[margin=1.5cm, left=3cm,twoside]{geometry}
+\usepackage[margin=1.5cm,twoside]{geometry}
+
+\usepackage[para]{footmisc}
+
+\usepackage[compact]{titlesec}
+% \titlespacing{\section}{0pt}{2ex}{1ex}
+% \titlespacing{\subsection}{0pt}{1ex}{0ex}
+% \titlespacing{\subsubsection}{0pt}{0.5ex}{0ex}
 
 \DeclareSIUnit{\arcsec}{''}
+\DeclareSIUnit{\mag}{\text{mag}}
 
@@
 \newunicodechar{ʻ}{\okina}
-\newcommand{\omuamua}{\okina Omuamua }
+\newcommand{\omuamua}{\okina Oumuamua }
 
@@
 
+\section{Background}
+
 %To get all asteriods in TESS. 
-This project aims to find and characterise the lightcurves of all the asteroids seen by the Transiting Exoplanet Survey Satellite (TESS). 
-\section{Background}
+This project aims to find and characterise the light curves of all the asteroids seen by the Transiting Exoplanet Survey Satellite (TESS). 
     %Why TESS
-TESS is a large area, high imaging cadence, space telescope  \citep{Ricker2014}. 
+TESS is a large area, high cadence imaging , space telescope  \citep{Ricker2014}. 
 TESS is tasked with observing one piece of sky for \qty{27}{\day} at a time (a sector), delivering \qtyproduct{96 x 24}{\degree} full frame images (FFIs) at regular intervals. 
-With the initial cadence for these full frame images set to \qty{30}{\minute}, the time resolution of TESS is unparalleled, with a Nyquist frequency of \qty{1}{\per\hour}, as the mission was extended the lenght of the FFIs has come down to \qty{10}{\minute} and then \qty{200}{\second}.
+With the initial cadence for these full frame images set to \qty{30}{\minute}, the time resolution of TESS is unparalleled, with a Nyquist frequency of \qty{1}{\per\hour}, as the mission was extended the length of the FFIs has come down to \qty{10}{\minute} and then \qty{200}{\second}.
 This high time resolution and observation area does come at the cost of spatial resolution, as the pixels are each \qty{21}{\arcsec} square.
@@
 This work aims to extend their study to more sectors, and to use a different data reduction method, the \texttt{TESSreduce} package \citep{Ridden-Harper2021}.
-Because of the survey properties, TESS provides a selfconsitent was to measure the properties of asteroids over the full sky.   
+%TODO expand here...
+Because of the survey properties, TESS provides a self-consistent way to measure the properties of asteroids over the full sky.   
 Another beneficial part of this work is that as part of a full sky transient survey using TESS, \texttt{TESSELLATE} (Ridden-Harper and Roxburgh et al., in prep), asteroids are transient objects that spike the brightness of a pixel for only a few frames.
@@
 Asteroids are a key class of solar system objects. 
-Understanding their rotation properties has long been of interest to astronomers \citep[e.g.][]{Weidenschilling1981,Harris1994}.
-
-        %Omuamua
-High amplitude variation has come to the forefront of questions about asteroid properties because of the first interstellar object (ISO) 1I/\omuamua \citep[see][for a review]{Bannister2019}. 
+Understanding their rotation properties has long been of interest to astronomers \citep[e.g.][]{Weidenschilling1981,Harris1994}. %Omuamu
+High amplitude variation has come to the forefront of questions about asteroid properties because of the first interstellar object, 1I/\omuamua \citep[see][for a review]{Bannister2019}. 
 \omuamua was measured to have a rotation period of \qty{8.67(0.34)}{\hour} \citep{Belton2018} and seemed to be tumbling \citep[e.g.][]{Drahus2018,Fraser2018}. 
-The peak to peak amplitude variation of 2.5 mag %TODO cite
-on the double peaked light curve is of interest, as this is much higher than most asteroids. 
+The peak to peak amplitude variation of \qty{2.5}{\mag} \citep{Meech2017} of the double peaked light curve is of interest, as this is much higher than most asteroids, and it implies a large axis ratio.
 With the full sky survey of bright asteroids, we hope to find many asteroids with such a large amplitude variation, and to see just how rare \omuamua is.
@@
 %Querry and interpolate
+To check for asteroids in the TESS data, the positions of the asteroids with time are required.
+For most asteroids, their orbital elements are well known, so it is a matter of looking them up and cross-matching with transients in the TESS data.
+Python was used to make API calls to {Skybot}\footnote{\href{https://vo.imcce.fr/webservices/skybot/}{Skybot: https://vo.imcce.fr/webservices/skybot/}} to get positions of asteroids in a cone search box in right ascension (RA) and  declination (Dec) space.
+As TESS sectors are \qty{27}{\day} long, querying every \qty{12}{\hour} is manageable.
+These positions are still very sparsely spaced in time compared to the TESS data, so an interpolation is needed to bridge the gap.
+With TESS data coming in $\frac12\unit{\hour}$ chunks, 24 interpolated points are needed between each API call.
+This interpolation should be accurate, as asteroids move at close to a TESS pixel per FFI on average \citep{Pal2018,Pal2020}, and checking against a higher frequency query to {JPL Horizons}\footnote{\href{https://ssd.jpl.nasa.gov/horizons/}{JPL Horizons: https://ssd.jpl.nasa.gov/horizons/}} confirmed this accuracy.
+For the shorter FFIs, the asteriods will be move fewer pixels per frame, which could cause slight mismatches between the interpolation and the detections. 
+Horizons was also used get properties of the individual asteroids, such as the absolute magnitude $H$. 
+For the faster TESS data, more interpolated points are needed, but a smaller the change in position between each point.
+These interpolated positions can be seen in \autoref{Fig:interpPos} for a cut from \texttt{TESSELLATE}. 
+There are a few interesting features, such as the asteroids are moving in the same direction, indicated by the colouring, they come in from low RA and Dec and tend to increase both coordinates as the month of the sector progresses.
+There is also a large size range in this slice of sky, ranging over \qty{5}{\mag} in absolute magnitude, which can be seen in the alpha (or transparency) of the tracks in \autoref{Fig:interpPos}.
 
@@
 Matching these interpolated positions to \texttt{TESSELLATE} detections is important to lower the unknown transient outputs of this pipeline. 
-Using the \texttt{KDTree} algorithm \citep{Maneewongvatana1999} as implemented in \texttt{SciPy} \citep{2020SciPy-NMeth}, the right ascension (RA) and  declination (Dec) coordinates of the interpolated points and the detections can be compared and matched together.
-Filtering this \texttt{KDTree} output by not allowing the time between spatialy coincident matches to be longer than \qty{0.1}{\day}, \dots
+Using the \texttt{KDTree} algorithm \citep{Maneewongvatana1999} as implemented in \texttt{SciPy} \citep{2020SciPy-NMeth}, the RA and Dec coordinates of the interpolated points and the detections can be compared and matched together.
+Filtering this \texttt{KDTree} output by restricting the time between spatially coincident matches to less than \qty{0.1}{\day} stops any accidental matches in position from non-asteroid detections.
 
-%Lightcurves; detected VS forced interpolated
+%light curves; detected VS forced interpolated 
+There are two sets of points to take light curves from. 
+The matches from the detections, which already have a flux calculated, and the interpolated points themselves, which are more numerous but require forcing the photometry. 
+There were some challenges getting the flux of the interpolated points, even when \texttt{TESSELLATE} has already reduced all the FFIs of interest, due to the timing of the TESS frames, but these were identified and corrected for.
+Not every interpolated point gets a match, due to a variety of reasons, %pass in front of stars, dip below limiting mag, bad pipeline, not extreme enough of a difference between frames 
+so a comparison between the two light curves is interesting. 
+\autoref{Fig:DifFlux} shows these two light curves for a chosen asteroid, ``Ruff''. 
+This was picked as it has a high number of points matched to detections.
+% There are more interpolated points than matched points, as not ever interpolated point gets matched to a detection.
+There seems to be a systematic offset between the fluxes, with the interpolated points having consistently lower flux, (here the means differ by 58 counts)
+Adjusting aperture to sum the fluxes over to the ``centre of mass'' of the asteroid in each frame does not alleviate this problem.
 
@@
 %Periods and amplitueds for everything detected
+The next part of my analysis has to do with determining the periods and amplitudes of each asteroid's light curve.
+For this there are a few methods I can try; using the \texttt{Lightkurve} \citep{Lightkurve2018} package built for period analysis of TESS (and Kepler) data of variable stars or peel back a layer of abstraction and use the Lomb-Scargle periodogram as implemented by \texttt{Astropy}\cite{Astropy2022}.
+Some trialling of both methods is needed, as preliminary testing reveals of interesting similarities and differences between the packages. 
+There will be differences in the period between the matched points and the interpolated points, not just because of the difference in average flux, but also from the larger number of points.
+This difference needs to be carefully thought through to understand what is the more likely period.
 
@@
 The \texttt{TESSELLATE} pipeline has been running on the OzSTAR supercomputing facilities. 
-After I am confident that all the parts of the asteriod detection and subsequent lightcurve analysis works as required, the same code can be refactored to work on OzSTAR and a largescale analysis of all of the processed TESS sectors can be run.
-Only after this has completed can the asteriod population statistics can be computed.
+After I am confident that all the parts of the asteroid detection and subsequent light curve analysis works as required, the same code can be refactored to work on OzSTAR and a large-scale analysis of all the processed TESS sectors can be run.
+Only after this has completed can the asteroid population statistics can be computed. 
+I will be looking for completeness of detections of asteroids, as well as accuracy of periods and amplitude variation.
 
-\section{Figures}
+% \section{Figures}
 %*Figures:
+\newgeometry{twocolumn,margin=1.5cm,twoside}
+
+\begin{figure}
+  \centering
+    \includegraphics[width =\columnwidth]{../Test Code/Testing Figures/interpPos\_22\_2\_3\_4.pdf}
+    \caption{The interpolated positions of asteroids in one cut of a TESS sector. 
+    The colours of the lines are time sequenced, as shown in the colour bar.
+    The alpha of the colours are scaled by the absolute magnitude $H$ of the asteroid, queried from JPL Horizons. 
+    Both celestial (RA and Dec) coordinates and ecliptic (ecLon and ecLat) coordinates axes are shown.
+    }
+    \label{Fig:interpPos}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width =\columnwidth]{../Test Code/Testing Figures/differentFluxes Ruff .pdf}
+  \caption{The light curves of the asteroid ``Ruff'', with the flux of the interpolated positions in blue circles and the flux from the matches to detected sources in orange diamonds.
+  The Flux axis is measured in counts, as calculated by \texttt{TESSreduce}, and the times are in Modified Julian Date.}
+  \label{Fig:DifFlux}
+\end{figure}
+
+% \section{Acknowledgements}
+% %OzSTAR asks for:
+% \small
+% This work was performed on the OzSTAR national facility at Swinburne University of Technology. 
+% The OzSTAR program receives funding in part from the Astronomy National Collaborative Research Infrastructure Strategy (NCRIS) allocation provided by the Australian Government, and from the Victorian Higher Education State Investment Fund (VHESIF) provided by the Victorian Government.
+
+%H-scaled pos plot of higher cam/less asteroids. 
+%Both light curves (interp and detect)
+%? another one?
 
 %*Bib
+% \newgeometry{twocolumn,margin=1.5cm, left=3cm}
+% \newpage
 \def\bibfont{\tiny}

### 2024-07-25 Test Code/compareResults.py

@@ -104,7 +104,9 @@ detectedSourcesAll = pd.read_csv(f"../OzData/{sector}_{cam}_{ccd}_{cut}_detected
 
 print(pd.unique(detectedSourcesAll["Type"]))
 
-indexsOfInterest = detectedSourcesAll.index[(detectedSourcesAll["Type"] == "0") | detectedSourcesAll["Type"] == "AST"] #TODO figure out what code is for asteroid.
+indexsOfInterest = detectedSourcesAll.index[(detectedSourcesAll["Type"] == "0")]
+
+indexsOfInterest = detectedSourcesAll.index[(detectedSourcesAll["Type"] == "0") | (detectedSourcesAll["Type"] == "Asteroid")] #// TODO figure out what code is for asteroid.
 
 detectedSources = detectedSourcesAll.loc[indexsOfInterest]
 
@@ -148,7 +150,7 @@ matches.to_csv(f"./{sector}_{cam}_{ccd}_{cut}_InterpAndDetect_Matches.csv")
 
 unqNames = pd.unique(interpRes["Name"])
 
-obsForDetect=80 #number of observations needed to count a detection
+obsForDetect=40 #number of observations needed to count a detection
 
 
 interpedMv = []
@@ -202,7 +204,7 @@ for name in unqNames:
         plt.scatter(nameCut["Time"], nameCut["flux"], label=name)
         plt.legend()
 
-    if name == " Ruff " or name ==" Lincoln ":
+    if name == " Ruff " or name ==" Lincoln " or name ==" 1999 JE82 ":
         nameCut.to_csv(f"./{name}Matches.csv")
 
 

### 2024-07-25 Test Code/expectedPositionsSkyBot.py

@@ -123,13 +123,17 @@ def querySB(targetPos: list, qRad: float = 10.0, qLoc: str = "C57", numTimesteps
     timeList = t_i + dt*np.arange(0, numTimesteps)
     result = pd.DataFrame()
     
-    while len(result)<numTimesteps:
-
+    while len(result)<numTimesteps: #so if query timesout, it will restart with the cache
         try:
             result = _Skybotquery(ra_i, dec_i, timeList.jd,
-                          radius=qRad, location=qLoc, cache=True)
+                          radius=qRad, location=qLoc, cache=True) #timeout throws error, so need to be in try/except to not close
         except:
-            continue
+            continue #restarts query with cache
+
+        if type(result) != "pandas.core.frame.DataFrame": 
+            #! need something for when no asteroids are in cut.
+            #! else everthing else breaks too. 
+            break #to get out of while
     
 
     brightResult = result.loc[result["Mv"] <= magLim].reset_index(drop=True)
@@ -142,7 +146,7 @@ def querySB(targetPos: list, qRad: float = 10.0, qLoc: str = "C57", numTimesteps
     return brightResult, timeList
 
 
-def plotExpectedPos(posDf: pd.DataFrame, timeList: npt.ArrayLike, targetPos: list, magLim: float = 20.0, scaleAlpha: bool = False, minLen: int = 0) -> plt.Figure:
+def plotExpectedPos(posDf: pd.DataFrame, timeList: npt.ArrayLike, targetPos: list, magLim: float = 20.0, scaleAlpha: bool = False, minLen: int = 0, saving=False) -> plt.Figure:
     """
     Takes the output of a query to SkyBot, sets up a WCS, and plots the objects position, coloured by time, in Ra/Dec and ecLon/ecLat space. 
 
@@ -160,6 +164,8 @@ def plotExpectedPos(posDf: pd.DataFrame, timeList: npt.ArrayLike, targetPos: lis
             If alpha of the symbols should be scaled to the average brigtness of the object
         minLen: int
[…]

### 2024-07-29 Progress Report/progressReportDraft.tex
talk done

@@
 \title{ASTR480 Progress Report}
-\author{Brayden Leicester \\[1ex]
+\author{Brayden Leicester \textit{61352063} \\[1ex]
 \small{Supervisors: Michele Bannister and Ryan Ridden-Harper}
@@
 
-%To get all asteriods in TESS. 
+%To get all asteroids in TESS. 
 This project aims to find and characterise the light curves of all the asteroids seen by the Transiting Exoplanet Survey Satellite (TESS). 
@@
 There have been attempts before to find and classify the asteroids in TESS data before by \citet{Pal2018, Pal2020}. 
-This work aims to extend their study to more sectors, and to use a different data reduction method, the \texttt{TESSreduce} package \citep{Ridden-Harper2021}.
-%TODO expand here...
+The first data release in \citet{Pal2020} catalogues nearly 10,000 asteroid light curves from the first year of TESS operation.
+This work aims to extend their study to more sectors, hence characterising more asteroids, and to use a different data reduction method, the \texttt{TESSreduce} package \citep{Ridden-Harper2021} which should increase the accuracy of the photometry due to the improved background subtraction process.
 Because of the survey properties, TESS provides a self-consistent way to measure the properties of asteroids over the full sky.   
@@
 This interpolation should be accurate, as asteroids move at close to a TESS pixel per FFI on average \citep{Pal2018,Pal2020}, and checking against a higher frequency query to {JPL Horizons}\footnote{\href{https://ssd.jpl.nasa.gov/horizons/}{JPL Horizons: https://ssd.jpl.nasa.gov/horizons/}} confirmed this accuracy.
-For the shorter FFIs, the asteriods will be move fewer pixels per frame, which could cause slight mismatches between the interpolation and the detections. 
+For the shorter FFIs, the asteroids will be move fewer pixels per frame, which could cause slight mismatches between the interpolation and the detections. 
 Horizons was also used get properties of the individual asteroids, such as the absolute magnitude $H$. 

### 2024-07-29 Test Code/expectedPositionsSkyBot.py

@@ -374,9 +374,9 @@ def plotHorizons(nameList: list[str], t_i:Time, t_f:Time| None = None,loc:str="5
     return fig
 
 sector = 22
-cam = 2
+cam = 1
 ccd = 3
-cut = 4
+cut = 8
 
 
 fname = f"../OzData/{sector}_{cam}_{ccd}_{cut}_wcs.fits"
@@ -404,8 +404,8 @@ posFig.savefig(f"./Testing Figures/posFig_{sector}_{cam}_{ccd}_{cut}.png")
 
 unqNames = list(pd.unique(res['Name']))
 
-distanceFig = plotHorizons(unqNames, times[0], t_f=times[-1], loc="500@-95", plotIRDel=True)
-distanceFig.savefig(f"./Testing Figures/delRHPlot_{sector}_{cam}_{ccd}_{cut}.png")
+# distanceFig = plotHorizons(unqNames, times[0], t_f=times[-1], loc="500@-95", plotIRDel=True)
+# distanceFig.savefig(f"./Testing Figures/delRHPlot_{sector}_{cam}_{ccd}_{cut}.png")
 # distanceFig.savefig(f"./DistancesVInclPlot_ra{myTargetPos[0]}_dec{myTargetPos[1]}_t{myTargetPos[2].mjd}_Mv{magLim}.png")
 
 

### 2024-07-31 Test Code/expectedPositionsSkyBot.py
talk done

@@ -430,6 +430,7 @@ for i, name in enumerate(unqNames):
     interpDecs = np.interp(x=interpTimes, xp=underSampledPos["epoch"], fp=underSampledPos["Dec"])
     interpDf = pd.DataFrame({"RA":interpRAs, "Dec":interpDecs, "epoch":interpTimes}) #make into DF
     #*concat with origonals, and then sorts and fills empty cols
+    #TODO remove origonal points... 
     concatedDF = pd.concat([underSampledPos, interpDf])
     concatedDF.sort_values(by=['epoch'], inplace=True)
     concatedDF.reset_index(drop=True, inplace=True)

### 2024-08-27 FinalReport/FinalReportDraft.tex
started on final report. Have many things to fix in code

--- 2024-07-29 Progress Report/progressReportDraft.tex
@@
 
-\usepackage[margin=1.5cm,twoside]{geometry}
+\usepackage[margin=1.5cm, left=3cm ,twoside]{geometry}
 
 \usepackage[para]{footmisc}
-
-\usepackage[compact]{titlesec}
-% \titlespacing{\section}{0pt}{2ex}{1ex}
-% \titlespacing{\subsection}{0pt}{1ex}{0ex}
-% \titlespacing{\subsubsection}{0pt}{0.5ex}{0ex}
 
@@
 \usepackage{newunicodechar}
-%\usepackage{libertine}
 
@@
 
-\title{ASTR480 Progress Report}
-\author{Brayden Leicester \textit{61352063} \\[1ex]
-\small{Supervisors: Michele Bannister and Ryan Ridden-Harper}
+\title{ASTR480 Final Report:\\
+\textit{Asteroids in TESS}
+}
+\author{Brayden Leicester \\ 
+\textit{61352063} \\
+[1ex]\small{Supervisors: Michele Bannister and Ryan Ridden-Harper}
 }
 \begin{document}
+\pagenumbering{gobble} %to blank pg # before Contents pg
+%TODO declaration page ????
+%\newpage
 \maketitle
+%*{ as per the report guidelines
+% \newpage
+% \section*{Abstract}
+% %*Background
 
-%*Aim / background summary
+% %*Methods
 
-\section{Background}
+% %*Main Results
+
+% %*Conclusion
+
+% \newpage
+% \pagenumbering{roman}
+% \tableofcontents
+% \newpage
+% \listoffigures
+% \newpage
+% \listoftables
+% \newpage
+%*}
+
+\pagenumbering{arabic}
+
+\section{Introduction}\label{Sec:Intro}
 
@@
 This project aims to find and characterise the light curves of all the asteroids seen by the Transiting Exoplanet Survey Satellite (TESS). 
+As such, a thorough discussion of both is required.
+
+\subsection{TESS}\label{SubSec:TESS}
+
     %Why TESS
-TESS is a large area, high cadence imaging , space telescope  \citep{Ricker2014}. 
+TESS is a large area, high cadence imaging, space telescope  \citep{Ricker2014}. 
 TESS is tasked with observing one piece of sky for \qty{27}{\day} at a time (a sector), delivering \qtyproduct{96 x 24}{\degree} full frame images (FFIs) at regular intervals. 
-With the initial cadence for these full frame images set to \qty{30}{\minute}, the time resolution of TESS is unparalleled, with a Nyquist frequency of \qty{1}{\per\hour}, as the mission was extended the length of the FFIs has come down to \qty{10}{\minute} and then \qty{200}{\second}.
+These FFIs are built from stacked \qty{2}{\second} exposures, leveraging the short readout times of the 16 CCD cameras on-board. 
+With the initial cadence for these full frame images set to \qty{30}{\minute}, the time resolution of TESS is unparalleled.
+A Nyquist frequency %TODO def, cite
+of \unit{\per\hour} is well sampled enough to characterise most variable stars, as well as find the orbital periods of exoplanets to a high precision. 
+As the mission was extended, after TESS had already mapped the entire sky, the length of the FFIs has decreased to \qty{10}{\minute} and then later to \qty{200}{\second}. 
+This decreased in the FFI exposure time comes with a complementary increase in the Nyquist frequency. 
 This high time resolution and observation area does come at the cost of spatial resolution, as the pixels are each \qty{21}{\arcsec} square.
+Of interest here is what such a high sampling rate can do for statistic on the asteroid population. 
+For bright asteroids the rotation periods should be able to be easily determined from this vast dataset. 
+The shortest FFIs will be able to accurately determine the rotation periods of all but the fastest rotating asteroids, most of which will be too dim to see in the TESS data.
+
 There have been attempts before to find and classify the asteroids in TESS data before by \citet{Pal2018, Pal2020}. 
 The first data release in \citet{Pal2020} catalogues nearly 10,000 asteroid light curves from the first year of TESS operation.
-This work aims to extend their study to more sectors, hence characterising more asteroids, and to use a different data reduction method, the \texttt{TESSreduce} package \citep{Ridden-Harper2021} which should increase the accuracy of the photometry due to the improved background subtraction process.
+%TODO the others
+
+This work aims to extend these studies to more sectors, hence characterising more asteroids, and to use a different data reduction method.
+Reducing the data with the \texttt{TESSreduce} package \citep{Ridden-Harper2021} should increase the accuracy of the photometry due to the improved background subtraction process.
 Because of the survey properties, TESS provides a self-consistent way to measure the properties of asteroids over the full sky.   
@@
 The goal of finding all the asteroids will allow for the removal of these spikes from the transient pipeline, as well as to understand the asteroid population better.  
+They can confuse pipelines looking for short time stellar transients such as flare stars and supernova. 
+As such, a way of filtering them out is required, as well as this filtering will get the properties of the light curves, such as the rotation periods and amplitude variation of these small solar system bodies.
+A full sky, self-consistent catalogue of these properties is important for many reasons. 
+The asteroid population statistics are useful in their own right, understanding the orbits of near earth objects is useful for planetary defence, and these bodies \dots
 
-    %Why Asteroids
+\subsection{Asteroids}\label{SubSec:Asteroid}
+
 Asteroids are a key class of solar system objects. 
-Understanding their rotation properties has long been of interest to astronomers \citep[e.g.][]{Weidenschilling1981,Harris1994}. %Omuamu
+Understanding their rotation properties has long been of interest to astronomers \citep[e.g.][]{Weidenschilling1981,Harris1994}. 
+%TODO asteroid elements
+
+The orbits of asteroids can be well characterised by only a few values. The semi-major axis, $a$, is the largest distance from the sun that the asteroid achieves on its elliptical orbit.
+Main belt asteroids have typical $a$s of \dots %TODO
+The eccentricity, $e$, of this orbit is the next parameter of interest.
+$e=0$ implies a perfectly circular orbit, $0<e<1$ are ellipses, $e=1$ give a parabolic orbit, and $e>1$ are unbound, hyperbolic orbits.
+Main belt asteroids have small $e$, in the range of \dots
+Comets have larger $e$, with some Oort cloud comets having $e\approx 1$. 
+Interstellar objects (ISOs) have $e>1$ as they are unbound.
+
+\subsubsection*{Interstellar Objects}
+%Omuamu
 High amplitude variation has come to the forefront of questions about asteroid properties because of the first interstellar object, 1I/\omuamua \citep[see][for a review]{Bannister2019}.
+\omuamua was determined to be spectroscopically red \citep{Fitzsimmons2017, Meech2017}, and having a photometric colour in the neutral end of the solar system range \citep{Bannister2017}.  
+1I was classed as asteroid due to lack of a coma, and no noticable activity %TODO cite, as opposed to 2I??
 \omuamua was measured to have a rotation period of \qty{8.67(0.34)}{\hour} \citep{Belton2018} and seemed to be tumbling \citep[e.g.][]{Drahus2018,Fraser2018}. 
-The peak to peak amplitude variation of \qty{2.5}{\mag} \citep{Meech2017} of the double peaked light curve is of interest, as this is much higher than most asteroids, and it implies a large axis ratio.
+Combing the tumbling with an elongation ratio of up to \qty{6(1)}{}:1 \citep{McNeill2018}, 1I is said to have a cigar shape \citep{Belton2018}.  
+The peak to peak amplitude variation of this ISO was \qty{2.5}{\mag} \citep{Meech2017} over its double peaked light curve.
+This is quite interesting, as it varies more than most asteroids. 
 With the full sky survey of bright asteroids, we hope to find many asteroids with such a large amplitude variation, and to see just how rare \omuamua is.
 
-\section{Work So Far}
-%* Method / Initial Results
+\section{Methods}\label{Sec:Meth}
 
-%Querry and interpolate
+\subsection{Querrying Databases}\label{SubSec:Querry}
+
 To check for asteroids in the TESS data, the positions of the asteroids with time are required.
@@
 These positions are still very sparsely spaced in time compared to the TESS data, so an interpolation is needed to bridge the gap.
+
+Knowing where the asteroids are is helpful for finding them in the archival TESS data, but knowing more about the individual asteroids is also useful for population statistics.
+{JPL Horizons}\footnote{\href{https://ssd.jpl.nasa.gov/horizons/}{JPL Horizons: https://ssd.jpl.nasa.gov/horizons/}} was used to obtain the orbital elements of each asteroid, $a$, $e$, and $i$, as well as an absolute magnitude $H$, which is a good proxy for size (as discussed above).
+%TODO ref
+
+\subsection{Interpolation}\label{SubSec:Interp}
+
 With TESS data coming in $\frac12\unit{\hour}$ chunks, 24 interpolated points are needed between each API call.
-This interpolation should be accurate, as asteroids move at close to a TESS pixel per FFI on average \citep{Pal2018,Pal2020}, and checking against a higher frequency query to {JPL Horizons}\footnote{\href{https://ssd.jpl.nasa.gov/horizons/}{JPL Horizons: https://ssd.jpl.nasa.gov/horizons/}} confirmed this accuracy.
+This interpolation should be accurate, as asteroids move at close to a TESS pixel per FFI on average \citep{Pal2018,Pal2020}, and checking against a higher frequency query to JPL Horizons confirmed this accuracy.
 For the shorter FFIs, the asteroids will be move fewer pixels per frame, which could cause slight mismatches between the interpolation and the detections. 
@@
 
+\subsection{Detection Matching}\label{SubSec:Match}
+
 %Match to detections
 Matching these interpolated positions to \texttt{TESSELLATE} detections is important to lower the unknown transient outputs of this pipeline. 
+Having interpolated their positions, the asteroids have a well sampled set of RA, Dec and time values of where they should be in the TESS data.
+They should show up in a pixel for a small number of frames, of order $\sim1$.
+This is the same as a lot of other transient events, a sharp rise in brightness and then disapering quickly again.
+The number of frames do change, type Ia supernova will brighten is a matter of a few hours and then dim for days, while stellar flares are of similar profile by a smaller max brightness and a corrsepondingly shorter decay time.
+Asteroids are very short, however detection pipelines are robust.
+These pipelines have already found the aforementioned supernova and stellar flares, the job of this work is to catch all the asteroids in the set of all the transients.
+To do this, catelogue matching is in order.
+
 Using the \texttt{KDTree} algorithm \citep{Maneewongvatana1999} as implemented in \texttt{SciPy} \citep{2020SciPy-NMeth}, the RA and Dec coordinates of the interpolated points and the detections can be compared and matched together.
 Filtering this \texttt{KDTree} output by restricting the time between spatially coincident matches to less than \qty{0.1}{\day} stops any accidental matches in position from non-asteroid detections. 
+A cut is then made by smallest distance, in coordinate space, and any distance must be smaller than \qty{0.01}{\degree} (\qty{36}{\arcsec}), which is within 2 TESS pixels. 
+Then, if multiple detection match to the same interpolated point, the detection with the smallest distance will be taken.
+
+\subsection{Ligthcurves and Periods}\label{SubSec:LCsandPs}
 
@@
 This was picked as it has a high number of points matched to detections.
-% There are more interpolated points than matched points, as not ever interpolated point gets matched to a detection.
+There are more interpolated points than matched points, as not ever interpolated point gets matched to a detection.
 There seems to be a systematic offset between the fluxes, with the interpolated points having consistently lower flux, (here the means differ by 58 counts)
 Adjusting aperture to sum the fluxes over to the ``centre of mass'' of the asteroid in each frame does not alleviate this problem.
-
-\section{Future Work}
-%* Future work
 
@@
 
+\subsection{Asteroid Statistics}\label{SubSec:StatMeth}
+
 %Run on server
@@
 
-% \section{Figures}
-%*Figures:
-\newgeometry{twocolumn,margin=1.5cm,twoside}
+\section{Results}\label{Sec:Res}
+\section{Discussion}\label{Sec:Disc}
+\section{Conclusion}\label{Sec:Conc}
+
+\section{Figures}
+%*Figures: Only here to keep them out of the way. Will be placed in text as it grows
 
@@
     \includegraphics[width =\columnwidth]{../Test Code/Testing Figures/interpPos\_22\_2\_3\_4.pdf}
-    \caption{The interpolated positions of asteroids in one cut of a TESS sector. 
+    \caption[Interpolated positions of asteroids]{The interpolated positions of asteroids in one cut of a TESS sector. 
     The colours of the lines are time sequenced, as shown in the colour bar.
@@
   \includegraphics[width =\columnwidth]{../Test Code/Testing Figures/differentFluxes Ruff .pdf}
-  \caption{The light curves of the asteroid ``Ruff'', with the flux of the interpolated positions in blue circles and the flux from the matches to detected sources in orange diamonds.
+  \caption[Light curves of Ruff]{The light curves of the asteroid ``Ruff'', with the flux of the interpolated positions in blue circles and the flux from the matches to detected sources in orange diamonds.
   The Flux axis is measured in counts, as calculated by \texttt{TESSreduce}, and the times are in Modified Julian Date.}
@@
 
-% \section{Acknowledgements}
-% %OzSTAR asks for:
-% \small
-% This work was performed on the OzSTAR national facility at Swinburne University of Technology. 
-% The OzSTAR program receives funding in part from the Astronomy National Collaborative Research Infrastructure Strategy (NCRIS) allocation provided by the Australian Government, and from the Victorian Higher Education State Investment Fund (VHESIF) provided by the Victorian Government.
+\newpage %! Page number above here must be <=25
 
-%H-scaled pos plot of higher cam/less asteroids. 
-%Both light curves (interp and detect)
-%? another one?
+\section*{Acknowledgements}
+%OzSTAR asks for:
+\small
+This work was performed on the OzSTAR national facility at Swinburne University of Technology. 
+The OzSTAR program receives funding in part from the Astronomy National Collaborative Research Infrastructure Strategy (NCRIS) allocation provided by the Australian Government, and from the Victorian Higher Education State Investment Fund (VHESIF) provided by the Victorian Government.
 
 %*Bib
-% \newgeometry{twocolumn,margin=1.5cm, left=3cm}
-% \newpage
-\def\bibfont{\tiny}
 \bibliographystyle{jphysicsB}
-\bibliography{../Notes/bibfile.bib}
+\bibliography{bibfile.bib}
 \end{document}

### 2024-08-27 Test Code/compareResults.py

@@ -29,12 +29,12 @@ def useKDTree(df1, df2, cols, maxDist:float=0.01,maxTimeSep:float=0.1,k:int=1):
 
     df2.rename(columns={'Unnamed: 0Match':"IDMatch"}, inplace=True, errors="raise")
 
-    #TODO compare time at indices returned
+    #// TODO compare time at indices returned
     timeInds = df2.index[np.abs((df2["Time"]-df2["TimeMatch"]))<maxTimeSep]
 
     df2=df2.loc[timeInds]
 
-    #TODO take only nearest match for each point
+    #// TODO take only nearest match for each point
     smallDistInds = df2.index[df2["distToMatch"]<maxDist]
     
     df2= df2.loc[smallDistInds]
@@ -69,7 +69,7 @@ def useKDTree(df1, df2, cols, maxDist:float=0.01,maxTimeSep:float=0.1,k:int=1):
 #     return cat1_id,cat2_id
 
 
-sector = 29
+sector = 22
 cam = 1
 ccd = 3
 cut = 7

### 2024-08-27 Test Code/expectedPositionsSkyBot.py

@@ -373,7 +373,7 @@ def plotHorizons(nameList: list[str], t_i:Time, t_f:Time| None = None,loc:str="5
 
     return fig
 
-sector = 29
+sector = 22
 cam = 1
 ccd = 3
 cut = 7
@@ -399,10 +399,12 @@ res, times = querySB(myTargetPos, magLim=magLim, numTimesteps=54, qRad=3.2)
 # posFig.savefig(f"./Testing Figures/posFig_{sector}_{cam}_{ccd}_{cut}.png")
 # posFig.savefig(f"./ExpectedPositionsPlot_ra{myTargetPos[0]}_dec{myTargetPos[1]}_t{myTargetPos[2].mjd}_Mv{magLim}.png")
 
-# eleFig = plotHorizons(unqNames, times[0], plotAEI=True)
-# eleFig.savefig(f"./OrbitalElementsPlot_ra{myTargetPos[0]}_dec{myTargetPos[1]}_t{myTargetPos[2].mjd}_Mv{magLim}.png")
-
 unqNames = list(pd.unique(res['Name']))
+
+
+eleFig = plotHorizons(unqNames, times[0], plotAEI=True)
+eleFig.savefig(f"./OrbitalElementsPlot_ra{myTargetPos[0]}_dec{myTargetPos[1]}_t{myTargetPos[2].mjd}_Mv{magLim}.png")
+
 
 # distanceFig = plotHorizons(unqNames, times[0], t_f=times[-1], loc="500@-95", plotIRDel=True)
 # distanceFig.savefig(f"./Testing Figures/delRHPlot_{sector}_{cam}_{ccd}_{cut}.png")
@@ -436,6 +438,7 @@ for i, name in enumerate(unqNames):
     else:
         interpPoints = 216  # 200 s ffi
     
+    #! This is all wrong, the deltaT was in days not in query lenghth, so I was geting each hour, not each 1/2 hr etc. 
 
 
     interpTimes = np.linspace(minTime,maxTime, int(interpPoints*deltaTime))#linspace to sample

### 2024-08-27 Test Code/findAndInterp.py

@@ -1,0 +1,291 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import numpy as np
+import numpy.typing as npt
+import matplotlib.pyplot as plt
+import matplotlib.colors as mplc
+import matplotlib.cm as mpcm
+import pandas as pd
+from astropy.coordinates import SkyCoord
+from astropy.time import Time
+import astropy.units as u
+from astropy import wcs
+from astropy.utils.data import download_file
+from astropy.io import fits
+from astroquery.jplhorizons import Horizons
+import itertools
+from tqdm import tqdm
+from scipy.spatial import KDTree
+
+
+plt.rcParams.update({
+    "font.size": 18,
+    "font.family": "serif",
+    "figure.autolayout": True,
+    "axes.grid": False,
+    # "xtick.minor.visible": True,
+    # "ytick.minor.visible": True,
+})
+
+
+
+def _Skybotquery(ra, dec, times, radius=10/60, location='C57', cache=False):
+    """Returns a list of asteroids/comets given a position and time.
+    This function relies on The Virtual Observatory Sky Body Tracker (SkyBot)
+    service which can be found at http://vo.imcce.fr/webservices/skybot/
+        Geert's magic code
+            Ryans Magic code
+
+    Parameters
+    ----------https://github.com/bray217/tessts.git
+        Right Ascension in degrees.
+    dec : float
+        Declination in degrees.
+    times : array of float
+        Times in Julian Date.
+    radius : float
+        Search radius in degrees.
+    location : str
+        Spacecraft location. Options include `'kepler'` and `'tess'`.
+    cache : bool
+        Whether to cache the search result. Default is True.
+    Returns
+    -------
+    result : `pandas.DataFrame`
+        DataFrame containing the list of known solar system objects at the
+        requested time and location.
+    """
+    url = 'http://vo.imcce.fr/webservices/skybot/skybotconesearch_query.php?'
+    url += '-mime=text&'
+    url += '-ra={}&'.format(ra)
+    url += '-dec={}&'.format(dec)
+    url += '-bd={}&'.format(radius)
+    url += '-loc={}&'.format(location)
+    # TODO add more things here?
+
+    df = None
+    times = np.atleast_1d(times)
+    for time in tqdm(times, desc='Querying for SSOs'):
[…]

### 2024-08-27 Test Code/makeLightcurve.py

@@ -8,7 +8,12 @@ from astropy import wcs
 from astropy.utils.data import download_file
 from astropy.io import fits
 
-# %matplotlib widget
+import warnings
+from astropy.utils.exceptions import AstropyWarning
+warnings.simplefilter('ignore', category=AstropyWarning)
+
+
+%matplotlib widget
 
 plt.rcParams.update({
     "font.size": 18,
@@ -33,18 +38,103 @@ def load_interpolations(sector, cam, ccd, cut, dirPath = "./"):
     interpDF.drop(columns="Unnamed: 0", inplace=True)
     return interpDF
 
-def load_fluxes(sector, cam, ccd, cut, dirPath = "../Ozdata/"):
+def load_fluxes(sector, cam, ccd, cut, dirPath = "../OzData/"):
     reduFlux = np.load(f"{dirPath}sector{sector}_cam{cam}_ccd{ccd}_cut{cut}_of16_ReducedFlux.npy")
     return reduFlux
 
-def load_times(sector, cam, ccd, cut, dirPath = "../Ozdata/"):
+def load_times(sector, cam, ccd, cut, dirPath = "../OzData/"):
     timesFromOz = np.load(f"{dirPath}sector{sector}_cam{cam}_ccd{ccd}_cut{cut}_of16_Times.npy")
     return timesFromOz
 
 def name_cut(df, name:str, colName:str="NameMatch"):
     "take a df and gives  only the values where the colName == name"
-    return df.loc[df.index[df[colName]==name]]
+    toReturn = df.loc[df.index[df[colName]==name]]
+    toReturn.reset_index(drop=True, inplace=True)
+    return toReturn
 
+def sum_fluxes(frames, ys, xs):
+    """Sums 3x3 box around all f,y,x, retuns list of fluxes in frame order"""
+
+    down = 1
+    up = 2
+    total = []
+
+    for f,y,x in zip(frames,ys,xs):   
+        #making sure everything is in bounds
+        ymin = y-down
+        ymax = y+up
+        xmin = x-down
+        xmax = x+up
+
+        if ymin < 0: ymin = 0
+        if ymax >= fluxBounds[1]: ymax = fluxBounds[1] -1
+        if xmin < 0: xmin = 0
+        if xmax >= fluxBounds[2]: xmax = fluxBounds[2]-1
+        total.append(np.sum(reducedFluxes[f,ymin:ymax,xmin:xmax]))
+
[…]

### 2024-08-27 Test Code/periodDetection.py

@@ -6,7 +6,7 @@ import lightkurve as lk
 import astropy.units as u
 from astropy.timeseries import LombScargle as lsp
 
-# %matplotlib widget 
+%matplotlib widget 
 
 plt.rcParams.update({
     "font.size": 18,
@@ -17,6 +17,8 @@ plt.rcParams.update({
     # "ytick.minor.visible": True,
 })
 
+#TODO Nyquist criterion
+#TODO LC database
 
 def detect_period_lk(times, fluxes, fluxErrs):
     """
@@ -33,31 +35,56 @@ def detect_period_ap(times, fluxes, plotting = False):
     """
     Calculates the periodogram with astropy and returns the period at the max power.
     """
-    lsper = lsp(times*u.day, fluxes)#.autopower(samples_per_peak=50)
-    freqs, powers = lsper.autopower(samples_per_peak=5)
+    #* Seems to be sec 27 and 56 when it changes
+    #12hr queries constant
+    if sector <27: 
+        nyquistP = 1*60*60*u.s #1/2hr ffi, so 1hr nyqist
+    elif sector>=27 and sector<56:
+        nyquistP = 20*60*u.s #1/6hr (10 min) ffi
+    else: 
+        nyquistP = 400*u.s   # 200 s ffi
+    
+
+    minFreq = 1/(2*(times.max()-times.min())*u.day).to(u.s)
+    maxFreq = 1/nyquistP
+
+    lsper = lsp((u.Quantity(times, u.day)).to(u.s), fluxes)
+    freqs, powers = lsper.autopower(samples_per_peak=5, minimum_frequency=minFreq, maximum_frequency=maxFreq)
     bestPow = np.max(powers)
     bestFreq = freqs[np.argmax(powers)]
-    t_fit=np.linspace(times.min(),times.max(),1000)*u.day
+
+    t_fit=(np.linspace(times.min(),times.max(),1000)*u.day).to(u.s)
     y_fit = lsper.model(t_fit, bestFreq)
     modelParams = lsper.model_parameters(bestFreq)
-    f_a_prob = lsper.false_alarm_probability(np.max(powers),samples_per_peak=50)
+    f_a_prob = lsper.false_alarm_probability(np.max(powers),samples_per_peak=50, minimum_frequency=minFreq, maximum_frequency=maxFreq)
 
     if plotting:
         fig, ax = plt.subplots(2, figsize = (8,10))
 
         ax[0].set_xscale("log")
[…]

### 2024-08-29 Test Code/findAndInterp.py
tried to get lcdb to work, and played around with mag

@@ -202,7 +202,7 @@ def find_asteroids(sector,cam,ccd,cut):
 
     for name in unqNames:
         nCut = name_cut(res, name, colName="Name")
-        avgMv = np.mean(nCut["Mv"])
+        avgMv = np.mean(nCut["Mv"]).round(3)
         num = nCut.at[0,"Num"]
         astrClass = nCut.at[0,"Class"]  
 

### 2024-08-29 Test Code/makeLightcurve.py

@@ -7,6 +7,7 @@ import astropy.units as u
 from astropy import wcs
 from astropy.utils.data import download_file
 from astropy.io import fits
+from astropy.stats import sigma_clip
 
 import warnings
 from astropy.utils.exceptions import AstropyWarning
@@ -79,35 +80,36 @@ def lightcurve_from_name(name):
     nameDf = name_cut(interpDF,name,"Name")
     
     #frameNums = [ np.argmin(np.abs(frameTimes -(time-timeOffset))) for time in nameDf["epoch"]]
-    frameNums = None
+    frameNums = []
     badIds = []
 
     for i, time in enumerate(nameDf["epoch"]):
         time = time-timeOffset
         
-        thisFrame = np.argmin(np.abs(frameTimes -time))
 
-        if frameNums==None:
-            frameNums = [thisFrame] #puts first one in
-        elif frameNums[-1] == thisFrame:
+
+        thisFrame = np.argmin(np.abs(frameTimes -time))
+        frameNums.append(thisFrame)
+
+        if ((time - frameTimes.min()) < 0) or ((time - frameTimes.max())>0):
+            
+            badIds.append(i) # makes sure they are in the right times
+            continue
+
+        try:
+            lastFrame = frameNums[-2]
+        except:
+            lastFrame = None
+
+        if lastFrame == thisFrame: #2nd to last frame, as have added this frame already
             oldTime = frameTimes[frameNums[-1]] - time
             newTime = frameTimes[thisFrame] - time
-
             #shorter dt gets taken if the go to the same frame
 
             if np.abs(newTime)>=np.abs(oldTime): #New is longer (or ==) dt, so keep old one
                 badIds.append(i)
             else:   #new dt is shorter, so use  new one
                 badIds.append((i-1))
-                frameNums[-1]=thisFrame 
-        else:
-            frameNums.append(thisFrame)
-    
-    
-    #remove bad rows from nameDF
-    nameDf.drop(index=badIds, inplace=True)
-    #add frames in
-    nameDf["Frames"] = frameNums
     
     targetWSC = fits.open(f"../OzData/{sector}_{cam}_{ccd}_{cut}_wcs.fits")[0]
[…]

### 2024-08-29 Test Code/periodDetection.py

@@ -45,7 +45,7 @@ def detect_period_ap(times, fluxes, plotting = False):
         nyquistP = 400*u.s   # 200 s ffi
     
 
-    minFreq = 1/(2*(times.max()-times.min())*u.day).to(u.s)
+    minFreq = 1/(0.1*(times.max()-times.min())*u.day).to(u.s)
     maxFreq = 1/nyquistP
 
     lsper = lsp((u.Quantity(times, u.day)).to(u.s), fluxes)
@@ -65,7 +65,7 @@ def detect_period_ap(times, fluxes, plotting = False):
         ax[0].set_yscale("log")
         ax[0].set(xlabel="Log Frequency", ylabel="Power")
         ax[0].plot(freqs, powers, c="k", label="Periodogram")
-        ax[0].scatter(bestFreq,bestPow, c="gold",marker="*", s=100, label=f"Best Frequency = {bestFreq.round(2)} \nFalse Alarm Probability = {f_a_prob.round(3)}")
+        ax[0].scatter(bestFreq,bestPow, c="gold",marker="*", s=100, label=f"Best Frequency = {bestFreq.round(7)} \nFalse Alarm Probability = {f_a_prob.round(3)}")
         ax[0].legend()
 
         ax[1].set(xlabel="Time [MJD]", ylabel="Flux")
@@ -139,9 +139,9 @@ def compute_periods_ap(posDF):
         flux = nameDf["Flux"].values
 
         period_ap, f_a_Prob, theta = detect_period_ap(time, flux)
-        periodList_ap.append([name, period_ap, period_ap/(60*60*24), f_a_Prob, theta])
+        periodList_ap.append([name, period_ap, period_ap/(60*60), period_ap/(60*60*24), f_a_Prob, theta])
 
-    apPeriods = pd.DataFrame(periodList_ap, columns=["Name", "Best Period", "Best Period [Day]",  "False Alarm Probability", "Model Parameters"])
+    apPeriods = pd.DataFrame(periodList_ap, columns=["Name", "Best Period [Seconds]", "Best Period [Hours]","Best Period [Days]",  "False Alarm Probability", "Model Parameters"])
 
     return apPeriods, badCount
 
@@ -158,6 +158,10 @@ cut = 7
 
 interpLcDF = load_interps(22,1,3,7)
 
+
[…]

### 2024-09-01 Test Code/findAndInterp.py
interpolate at frames, not linspace

@@ -29,6 +29,8 @@ plt.rcParams.update({
 
 
 
+timeOffset = 2400000.5
+
 def _Skybotquery(ra, dec, times, radius=10/60, location='C57', cache=False):
     """Returns a list of asteroids/comets given a position and time.
     This function relies on The Virtual Observatory Sky Body Tracker (SkyBot)
@@ -147,6 +149,12 @@ def querySB(targetPos: list, qRad: float = 10.0, qLoc: str = "C57", numTimesteps
 
     brightResult = result.loc[result["Mv"] <= magLim].reset_index(drop=True)
 
+    brightResult["Name"] = [name.strip() for name in brightResult["Name"]]
+
+    brightResult["MJD"] = brightResult["epoch"]-timeOffset
+
+    brightResult.drop(columns="epoch", inplace=True)
+
     coords = SkyCoord(
         brightResult["RA"], brightResult["Dec"], unit=(u.hourangle, u.deg))
     brightResult["RA"] = coords.ra.deg
@@ -241,22 +249,32 @@ def interplolation_of_pos(posDf, sector):
     unqNames = np.unique(posDf["Name"])
     dfsList = []
 
+
+    frameTimes = np.load(f"../OzData/sector{sector}_cam{cam}_ccd{ccd}_cut{cut}_of16_Times.npy")
+
     for name in unqNames:
         
         underSampledPos = name_cut(posDf, name, colName="Name")
 
-        underSampledPos["QueriedPoint"] = np.ones_like(underSampledPos["epoch"])
-        minTime = underSampledPos["epoch"].min()
-        maxTime = underSampledPos["epoch"].max()
+        underSampledPos["QueriedPoint"] = np.ones_like(underSampledPos["MJD"])
+        minTime = underSampledPos["MJD"].min()
+        maxTime = underSampledPos["MJD"].max()
         deltaTime = maxTime-minTime
+
+        #TODO interp times at frame times.
         interpTimes = np.linspace(minTime,maxTime, int(interpPoints*deltaTime))#linspace to sample
         #Ra and Dec samples
-        interpRAs = np.interp(x=interpTimes, xp=underSampledPos["epoch"], fp=underSampledPos["RA"])
-        interpDecs = np.interp(x=interpTimes, xp=underSampledPos["epoch"], fp=underSampledPos["Dec"])
[…]

### 2024-09-01 Test Code/makeLightcurve.py

@@ -61,6 +61,9 @@ def sum_fluxes(frames, ys, xs):
     total = []
 
     for f,y,x in zip(frames,ys,xs):   
+        
+        if int(f) != f: print("FUCK")
+        
         #making sure everything is in bounds
         ymin = y-down
         ymax = y+up
@@ -71,46 +74,43 @@ def sum_fluxes(frames, ys, xs):
         if ymax >= fluxBounds[1]: ymax = fluxBounds[1] -1
         if xmin < 0: xmin = 0
         if xmax >= fluxBounds[2]: xmax = fluxBounds[2]-1
-        total.append(np.sum(reducedFluxes[f,ymin:ymax,xmin:xmax]))
+        total.append(np.sum(reducedFluxes[int(f),ymin:ymax,xmin:xmax]))
 
     return total
 
 
 def lightcurve_from_name(name):
     nameDf = name_cut(interpDF,name,"Name")
-    
-    #frameNums = [ np.argmin(np.abs(frameTimes -(time-timeOffset))) for time in nameDf["epoch"]]
-    frameNums = []
+    frameNums = [] #*sidesteped by keeping frame ID in from interpolation
     badIds = []
 
-    for i, time in enumerate(nameDf["epoch"]):
-        time = time-timeOffset
+
+        # for i, time in enumerate(nameDf["MJD"]):
+        #     # time = time-timeOffset
+
+        #     thisFrame = np.argmin(np.abs(frameTimes -time))
+        #     frameNums.append(thisFrame)
+
+        #     if ((time - frameTimes.min()) < 0) or ((time - frameTimes.max())>0):
+                
+        #         badIds.append(i) # makes sure they are in the right times
+        #         continue
+
+        #     try:
+        #         lastFrame = frameNums[-2]
+        #     except:
+        #         lastFrame = None
+
+        #     if lastFrame == thisFrame: #2nd to last frame, as have added this frame already
+        #         oldTime = frameTimes[frameNums[-1]] - time
+        #         newTime = frameTimes[thisFrame] - time
+        #         #shorter dt gets taken if the go to the same frame
+
+        #         if np.abs(newTime)>=np.abs(oldTime): #New is longer (or ==) dt, so keep old one
+        #             badIds.append(i)
+        #         else:   #new dt is shorter, so use  new one
[…]

### 2024-09-01 Test Code/periodDetection.py

@@ -45,7 +45,7 @@ def detect_period_ap(times, fluxes, plotting = False):
         nyquistP = 400*u.s   # 200 s ffi
     
 
-    minFreq = 1/(0.1*(times.max()-times.min())*u.day).to(u.s)
+    minFreq = 1/(2*(times.max()-times.min())*u.day).to(u.s)
     maxFreq = 1/nyquistP
 
     lsper = lsp((u.Quantity(times, u.day)).to(u.s), fluxes)
@@ -68,12 +68,12 @@ def detect_period_ap(times, fluxes, plotting = False):
         ax[0].scatter(bestFreq,bestPow, c="gold",marker="*", s=100, label=f"Best Frequency = {bestFreq.round(7)} \nFalse Alarm Probability = {f_a_prob.round(3)}")
         ax[0].legend()
 
-        ax[1].set(xlabel="Time [MJD]", ylabel="Flux")
+        ax[1].set(xlabel="Time [JD]", ylabel="Flux")
         ax[1].scatter(times, fluxes, label = "Light curve", c="tab:orange", marker = "d", s=10)
         ax[1].plot(t_fit.to(u.day),y_fit, ls="--", c="k", label="Model")
 
 
-    return (1/bestFreq)/u.s, f_a_prob, modelParams
+    return (1/bestFreq)/u.s, f_a_prob, modelParams, bestPow
 
 
 def load_interps(sector, cam, ccd, cut):
@@ -112,7 +112,7 @@ def compute_periods_lk(posDF):
             badCount+=1
             # print(f"{badCount}. not enough points (<={numObser}) for {name}")
             continue
-        time = nameDf["epoch"].values
+        time = nameDf["MJD"].values
         flux = nameDf["Flux"].values
         fluxErr = np.ones_like(flux)*1e-4 #!Have to make up err for lk
         period_lk = detect_period_lk(time,flux,fluxErr)
@@ -135,19 +135,30 @@ def compute_periods_ap(posDF):
             badCount+=1
             # print(f"{badCount}. not enough points (<={numObser}) for {name}")
             continue
-        time = nameDf["epoch"].values
+        time = nameDf["MJD"].values
         flux = nameDf["Flux"].values
 
-        period_ap, f_a_Prob, theta = detect_period_ap(time, flux)
[…]

### 2024-09-04 Test Code/expectedPositionsSkyBot.py
TESS isn't on a up and down...

@@ -402,8 +402,8 @@ res, times = querySB(myTargetPos, magLim=magLim, numTimesteps=54, qRad=3.2)
 unqNames = list(pd.unique(res['Name']))
 
 
-eleFig = plotHorizons(unqNames, times[0], plotAEI=True)
-eleFig.savefig(f"./OrbitalElementsPlot_ra{myTargetPos[0]}_dec{myTargetPos[1]}_t{myTargetPos[2].mjd}_Mv{magLim}.png")
+# eleFig = plotHorizons(unqNames, times[0], plotAEI=True)
+# eleFig.savefig(f"./OrbitalElementsPlot_ra{myTargetPos[0]}_dec{myTargetPos[1]}_t{myTargetPos[2].mjd}_Mv{magLim}.png")
 
 
 # distanceFig = plotHorizons(unqNames, times[0], t_f=times[-1], loc="500@-95", plotIRDel=True)
@@ -441,7 +441,9 @@ for i, name in enumerate(unqNames):
     #! This is all wrong, the deltaT was in days not in query lenghth, so I was geting each hour, not each 1/2 hr etc. 
 
 
-    interpTimes = np.linspace(minTime,maxTime, int(interpPoints*deltaTime))#linspace to sample
+    #interpTimes = np.linspace(minTime,maxTime, int(interpPoints*deltaTime))#linspace to sample #! just use times of frames from cut 
+
+
     #Ra and Dec samples
     interpRAs = np.interp(x=interpTimes, xp=underSampledPos["epoch"], fp=underSampledPos["RA"])
     interpDecs = np.interp(x=interpTimes, xp=underSampledPos["epoch"], fp=underSampledPos["Dec"])

### 2024-09-04 Test Code/findAndInterp.py

@@ -61,7 +61,7 @@ def _Skybotquery(ra, dec, times, radius=10/60, location='C57', cache=False):
     url += '-mime=text&'
     url += '-ra={}&'.format(ra)
     url += '-dec={}&'.format(dec)
-    url += '-bd={}&'.format(radius)
+    url += '-rd={}&'.format(radius)
     url += '-loc={}&'.format(location)
     # TODO add more things here?
 
@@ -202,10 +202,10 @@ def get_properties_Horizons(asteroidsDf, time, loc:str="500@10")->pd.DataFrame:
 
 def find_asteroids(sector,cam,ccd,cut):
 
-    res, timeList = querySB(setupQuery(sector, cam, ccd, cut), numTimesteps=54, qRad=3.2)
+    res, timeList = querySB(setupQuery(sector, cam, ccd, cut), numTimesteps=54, qRad=3.05)
 
     unqNames = np.unique(res["Name"])
-
+    print(len(unqNames))
     propertiesList = []
 
     for name in unqNames:
@@ -224,7 +224,6 @@ def find_asteroids(sector,cam,ccd,cut):
     withEles.to_csv(f"{fname}.csv")
 
     return res
-
 
 
 def name_cut(df, name:str, colName:str="NameMatch"):
@@ -247,6 +246,7 @@ def interplolation_of_pos(posDf, sector):
     
 
     unqNames = np.unique(posDf["Name"])
+    print(len(unqNames))
     dfsList = []
 
 
@@ -261,13 +261,13 @@ def interplolation_of_pos(posDf, sector):
         maxTime = underSampledPos["MJD"].max()
         deltaTime = maxTime-minTime
 
-        #TODO interp times at frame times.
-        interpTimes = np.linspace(minTime,maxTime, int(interpPoints*deltaTime))#linspace to sample
-        #Ra and Dec samples
+        #// TODO interp times at frame times.
+        # interpTimes = np.linspace(minTime,maxTime, int(interpPoints*deltaTime))#linspace to sample
+        # #Ra and Dec samples
 
         frameIDs = np.where((frameTimes>=minTime) & (frameTimes<=maxTime))[0]
 
-        interpTimes = frameTimes[frameIDs]
+        interpTimes = frameTimes[frameIDs] #*gets IDs and times of frames all at once
 
         interpRAs = np.interp(x=interpTimes, xp=underSampledPos["MJD"], fp=underSampledPos["RA"])
[…]

### 2024-09-04 Test Code/makeLightcurve.py

@@ -8,6 +8,11 @@ from astropy import wcs
 from astropy.utils.data import download_file
 from astropy.io import fits
 from astropy.stats import sigma_clip
+from scipy import ndimage
+import numpy.typing as npt
+import matplotlib.colors as mplc
+import matplotlib.cm as mpcm
+import itertools
 
 import warnings
 from astropy.utils.exceptions import AstropyWarning
@@ -58,11 +63,12 @@ def sum_fluxes(frames, ys, xs):
 
     down = 1
     up = 2
+    checkUp = 4
+    checkDown = 3
     total = []
+    comTotal = []
 
     for f,y,x in zip(frames,ys,xs):   
-        
-        if int(f) != f: print("FUCK")
         
         #making sure everything is in bounds
         ymin = y-down
@@ -76,7 +82,20 @@ def sum_fluxes(frames, ys, xs):
         if xmax >= fluxBounds[2]: xmax = fluxBounds[2]-1
         total.append(np.sum(reducedFluxes[int(f),ymin:ymax,xmin:xmax]))
 
-    return total
+        ymin = y-checkDown
+        ymax = y+checkUp
+        xmin = x-checkDown
+        xmax = x+checkUp
+
+        if ymin < 0: ymin = 0
+        if ymax >= fluxBounds[1]: ymax = fluxBounds[1] -1
+        if xmin < 0: xmin = 0
+        if xmax >= fluxBounds[2]: xmax = fluxBounds[2]-1
+
+        comTotal.append(np.sum(reducedFluxes[int(f),ymin:ymax,xmin:xmax]))
+        
+
+    return total, comTotal
 
 
 def lightcurve_from_name(name):
@@ -118,6 +137,7 @@ def lightcurve_from_name(name):
 
     interpedX, interpedY  = w.all_world2pix(interpCoords.ra, interpCoords.dec,0)
 
+    #rounds to nearest whole number, which returns #.0 as a float. then int converts. if just int masked, it floors the number, not rounds it.
     interpedX = interpedX.round().astype(int)
     interpedY = interpedY.round().astype(int)
     
@@ -128,17 +148,20 @@ def lightcurve_from_name(name):
     # nameDf["Frames"] = frameNums
 
     #takes the flux sums, and 
-    sumedFluxes = sum_fluxes(nameDf["FrameIDs"], nameDf["Y"], nameDf["X"])
[…]

### 2024-09-04 Test Code/periodDetection.py

@@ -31,7 +31,7 @@ def detect_period_lk(times, fluxes, fluxErrs):
 
     return period.value
 
-def detect_period_ap(times, fluxes, plotting = False):
+def detect_period_ap(times, fluxes, plotting = False, knownFreq = None):
     """
     Calculates the periodogram with astropy and returns the period at the max power.
     """
@@ -49,7 +49,7 @@ def detect_period_ap(times, fluxes, plotting = False):
     maxFreq = 1/nyquistP
 
     lsper = lsp((u.Quantity(times, u.day)).to(u.s), fluxes)
-    freqs, powers = lsper.autopower(samples_per_peak=5, minimum_frequency=minFreq, maximum_frequency=maxFreq)
+    freqs, powers = lsper.autopower(minimum_frequency=minFreq, maximum_frequency=maxFreq)
     bestPow = np.max(powers)
     bestFreq = freqs[np.argmax(powers)]
 
@@ -66,6 +66,11 @@ def detect_period_ap(times, fluxes, plotting = False):
         ax[0].set(xlabel="Log Frequency", ylabel="Power")
         ax[0].plot(freqs, powers, c="k", label="Periodogram")
         ax[0].scatter(bestFreq,bestPow, c="gold",marker="*", s=100, label=f"Best Frequency = {bestFreq.round(7)} \nFalse Alarm Probability = {f_a_prob.round(3)}")
+
+        if knownFreq is not None:
+            ax[0].axvline(knownFreq, linestyle=":", c="tab:blue")
+            ax[0].axvline(2*knownFreq, linestyle=":", c="tab:purple")
+
         ax[0].legend()
 
         ax[1].set(xlabel="Time [JD]", ylabel="Flux")
@@ -136,7 +141,7 @@ def compute_periods_ap(posDF):
             # print(f"{badCount}. not enough points (<={numObser}) for {name}")
             continue
         time = nameDf["MJD"].values
-        flux = nameDf["Flux"].values
+        flux = nameDf["COM Flux"].values
 
         period_ap, f_a_Prob, theta, maxPow = detect_period_ap(time, flux)
         periodList_ap.append([name, period_ap, period_ap/(60*60), period_ap/(60*60*24), f_a_Prob, theta, maxPow])
[…]

### 2024-09-04 FinalReport/FinalReportDraft.tex
interpolate at frames, not linspace
no one was at iso chat

@@
 
-\subsection{Ligthcurves and Periods}\label{SubSec:LCsandPs}
+\subsection{Ligthcurves}\label{SubSec:Lightcurves}
 
@@
 
+%TODO data cleaning explaination
+
+\subsection*{Period detection}\label{SubSec:Periods}
+
 %Periods and amplitueds for everything detected
+
 The next part of my analysis has to do with determining the periods and amplitudes of each asteroid's light curve.
-For this there are a few methods I can try; using the \texttt{Lightkurve} \citep{Lightkurve2018} package built for period analysis of TESS (and Kepler) data of variable stars or peel back a layer of abstraction and use the Lomb-Scargle periodogram as implemented by \texttt{Astropy}\cite{Astropy2022}.
-Some trialling of both methods is needed, as preliminary testing reveals of interesting similarities and differences between the packages. 
+
+A widely used technique in determining the periods of astronomical data is the Lomb-Scargle periodogram \citep[\citet{Lomb1976,Scargle1982}, but see][for a review]{VanderPlas2018}.
+This method can efficently take discrete timeseries data of which the observations are unevenly spaced and calculate the frequencies present in the sample.
+TESS should have rather evenly spaced data, as the FFIs are produced on a set candance, but sometimes data will be missing.
+For example it may have been sigma clipped due to a spike in brightness that is unphysical for an asteroid.
+The Lomb-Scargle periodogram deals with this uneven sampling much better than either a fast fourier transform or a direct least squares fit.
+
+There are a few different implementation of a Lomb-Scargle periodogram avalible in world of astronomy python packages.
+I first tried using the \texttt{Lightkurve} \citep{Lightkurve2018} package built for period analysis of TESS (and Kepler) data of variable stars.
+\texttt{Lightkurve} was easy to use but didn't allow for fine tuning of the periodogram, or give any model of the curve it produced.
+Peeling back a layer of abstraction and useing the Lomb-Scargle periodogram as implemented by \texttt{Astropy}\citep{Astropy2022} was the next option.
+There were interesting similarities and differences between the packages.
+\texttt{Lightkurve} is based on the \texttt{Astropy} methods, but dose not give all of the functionallity, instead opting to simplify the process.
+\texttt{Astropy} gives a statstic on how confident it is in the detected frequency is, the false alarm probability.
+It also returns a model, in the form of
+
+\begin{equation}
+  \label{Eq:LCModel}
+  F(t;f,\vec{\theta}\,) = \theta_0 + \theta_1\cdot\sin{2\pi ft} +\theta_2\cdot\cos{2\pi ft}
+\end{equation}
+
+where the flux, $F$, is given in terms of time, $t$, the calculated frequency, f, and the model parameter $\vec{\theta} = [\theta_0, \theta_1,\theta_2]$.
+These parameters are useful in reconstructing the sinusoid without rerunning the periodogram.
+
 There will be differences in the period between the matched points and the interpolated points, not just because of the difference in average flux, but also from the larger number of points.

### 2024-09-05 FinalReport/FinalReportDraft.tex
some more background

@@
 
+\usepackage[super]{nth}
+
+\usepackage{amsmath,amssymb}
+
 \DeclareSIUnit{\arcsec}{''}
-\DeclareSIUnit{\mag}{\text{mag}}
+\DeclareSIUnit{\au}{\rm{AU}}
+\DeclareSIUnit{\mag}{\rm{mag}}
 
 \usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
 \usepackage{newunicodechar}
@@
 %*}
-
+\newpage
 \pagenumbering{arabic}
@@
 With the initial cadence for these full frame images set to \qty{30}{\minute}, the time resolution of TESS is unparalleled.
-A Nyquist frequency %TODO def, cite
-of \unit{\per\hour} is well sampled enough to characterise most variable stars, as well as find the orbital periods of exoplanets to a high precision.
+The highest frequency (shortest period) that can be found from time series data like this is at the Nyquist limit, twice the minimum time between images.  
+Due to the short cadence, the Nyquist frequency of TESS is quite high, \qty{1}{\per\hour}. 
+This is well sampled enough to characterise most variable stars, as well as find the orbital periods of exoplanets to a high precision.
 As the mission was extended, after TESS had already mapped the entire sky, the length of the FFIs has decreased to \qty{10}{\minute} and then later to \qty{200}{\second}.
@@
 
-There have been attempts before to find and classify the asteroids in TESS data before by \citet{Pal2018, Pal2020}.
-The first data release in \citet{Pal2020} catalogues nearly 10,000 asteroid light curves from the first year of TESS operation.
-%TODO the others
+There have been attempts before to find and classify the asteroids in TESS data before.
+\citet{Pal2018} proposes that TESS will be a good instrument for solar system object study, and simulate some detections down to \nth{20} visual magnitude, but that good photometry should only be expected to $V \lesssim 19$.
+The first data release in \citet{Pal2020} catalogues nearly 10,000 asteroid light curves from the first year of TESS operation, and they report that they triple the number of asteroids with accurate rotation periods.
+\citet{McNeill2023} also analyse TESS cycle 1, and detect almost 38,000 objects.
+They determine reliable rotation periods for about 3,500 asteroids in this sample, and show lack of reliability for objects with periods less than \qty{3}{\hour}. 
+In the overlap from the data from \citet{Pal2020}, \citeauthor{McNeill2023} find good agreement between the two sets of periods and amplitudes. 
+\citet{Gowanlock2024} uses TESS photometry as well as  observations by the Zwicky Transient Facility \citep[ZTF, ][]{Bellm2018} to get a longer baseline on mutually observed objects while combining ground and space based observations.
+There were not very many objects in both samples with known periods for comparison, with only 222 objects analysed, however they demonstrate that the technique is effective.
 
-This work aims to extend these studies to more sectors, hence characterising more asteroids, and to use a different data reduction method.
-Reducing the data with the \texttt{TESSreduce} package \citep{Ridden-Harper2021} should increase the accuracy of the photometry due to the improved background subtraction process.
+This work aims to extend these studies to more sectors, hence characterising more asteroids.
+The aim is to get a volume complete set of asteroid lightcurves and periods, by analysis of the whole sky as seen by TESS using data from more than just the first year of TESS's operation.
+I also employ a different data reduction method, reducing the data with the \texttt{TESSreduce} package \citep{Ridden-Harper2021} should increase the accuracy of the photometry due to the improved background subtraction process.
 Because of the survey properties, TESS provides a self-consistent way to measure the properties of asteroids over the full sky.
-Another beneficial part of this work is that as part of a full sky transient survey using TESS, \texttt{TESSELLATE} (Ridden-Harper and Roxburgh et al., in prep), asteroids are transient objects that spike the brightness of a pixel for only a few frames.
+By using the higher cadence FFIs of later TESS sectors, the $P\leq\qty{3}{\hour}$ limit on the accurate periods \citep[as found by ][]{McNeill2023} should be lowered.  
+Another beneficial part of this work is that as part of a full sky transient survey using TESS, \texttt{TESSELLATE} \citep{TESSELATE}.
+Asteroids can spike the brightness of a pixel for only a few frames while it is passing by.
+This can confuse pipelines looking for  transients such as flare stars and supernova, as they can look very similar, having the brightness change in one pixel for a short time.
 The goal of finding all the asteroids will allow for the removal of these spikes from the transient pipeline, as well as to understand the asteroid population better.
-They can confuse pipelines looking for short time stellar transients such as flare stars and supernova.
-As such, a way of filtering them out is required, as well as this filtering will get the properties of the light curves, such as the rotation periods and amplitude variation of these small solar system bodies.
-A full sky, self-consistent catalogue of these properties is important for many reasons.
-The asteroid population statistics are useful in their own right, understanding the orbits of near earth objects is useful for planetary defence, and these bodies \dots
+By filtering the asteroid detections out, while also self consistently determining rotation properties of a large sample of the population, more science can be done with the one action.
 
@@
 Asteroids are a key class of solar system objects.
-Understanding their rotation properties has long been of interest to astronomers \citep[e.g.][]{Weidenschilling1981,Harris1994}.
-%TODO asteroid elements
-
 The orbits of asteroids can be well characterised by only a few values. The semi-major axis, $a$, is the largest distance from the sun that the asteroid achieves on its elliptical orbit.
-Main belt asteroids have typical $a$s of \dots %TODO
+Main belt asteroids have typical $a$s of \qtyrange{2}{3.3}{\au} \citet{DeMeo2015}. 
+The bounds between the inner, middle and outer main belt at \qty{2.5}{\au} and \qty{2.82}{\au}, these separations are due to resonances with Jupiter (the 3:1 and 5:2 mean motion resonances mark the edges of middle main belt) and Saturn (the $\nu_6$ secular resonance cuts off the lowest $a$ edge of the inner belt) carving swathes where asteroids cannot be in stable orbits.
 The eccentricity, $e$, of this orbit is the next parameter of interest.
 $e=0$ implies a perfectly circular orbit, $0<e<1$ are ellipses, $e=1$ give a parabolic orbit, and $e>1$ are unbound, hyperbolic orbits.
-Main belt asteroids have small $e$, in the range of \dots
-Comets have larger $e$, with some Oort cloud comets having $e\approx 1$.
-Interstellar objects (ISOs) have $e>1$ as they are unbound.
+Main belt asteroids have small $e$, in the range of \qtyrange{0}{0.35}{} \citep{DeMeo2015}
+Comets have larger $e$, with some long period comets having $e\approx 1$, but the typical range is \qtyrange{0.2}{0.7}{} \citep{Lewis2012}.
+Interstellar objects (ISOs) have $e>1$ as they are on unbound hyperbolic trajectories.
+The inclination, $i$, of asteroids is also a key orbital element, it is the angle relative to the ecliptic plane and in principle asteroids can span the whole range. 
+Main belt asteroids have typical inclinations of \qtyrange{0}{30}{\degree} \citep{DeMeo2015}, so the TESS cameras closer to the ecliptic plane are expected to have the most asteroids in their fields.
+
+%TODO Abs mag H, and how that is size.
+
+%TODO asteroid families
+Asteroids can be grouped into families based on clustering of the proper orbital elements and similarities in their colour or spectra \citep{Nesvorny2015}. 
+These families are named for their largest member.
+
+Understanding the rotation properties of asteroids has long been of interest to astronomers \citep[e.g.][for early work into the limits of rotation period and the tumbling nature of some small bodies]{Weidenschilling1981,Harris1994}.
+This is still at the forefront of research, with more lightcurves being published year-on-year \citep{Harris2015}. 
+This will only increase with more large survey telescopes coming online soon, such as the Legacy Survey of Space and Time \citep[LSST][]{LSST2019}, the number of new asteroids that are well characterised will increase by more than 2 orders of magnitude on the Sloan Digital Sky Survey \citep[SDSS][]{York2000} which had $\sim 88000$ objects \citet{Parker2008}.
+LSST needs a dedicated pipeline for asteroid classification, the Solar System Notification Alert Processing System \citep[SNAPS][]{Trilling2023} has been developed using ZTF data and is can scale up handle volume of data coming from LSST.
+Work is being done to understand the biases of how  asymmetric shape high amplitude variations can lead to selection effects of such large sample of asteroids \citet{Levine2023}.
+For the smallest objects ingenious observation techniques are employed to measure rotation periods such as intentionally streaking the asteroid down a column of pixel on a detector, and making the lightcurve from the streak, as done by \citet{Bolin2023}.
+
+Because of such a rich history of study, many asteroids have a known rotation period. 
+These are catalogued by \citet{Warner2009} %TODO updated in citation
+in the Light Curve Database (LCDB).
+%TODO fig of period vs diameter.   
+In the October \nth{1} 2023 release, the LCDB has nearly 9000 objects with periods determined to a quality code $U\geq 2-$, which is the minimum reliability recommended by \citeauthor{Warner2009} for statistical analysis.
+This includes data from \citet{Pal2020}.
+Comparing the periods I find of asteroids that are already in the LCDB is key to making sure my methods are reliable, and that the periods for asteroids for which no data exists can be trusted.
 
@@
 \omuamua was determined to be spectroscopically red \citep{Fitzsimmons2017, Meech2017}, and having a photometric colour in the neutral end of the solar system range \citep{Bannister2017}.
-1I was classed as asteroid due to lack of a coma, and no noticable activity %TODO cite, as opposed to 2I??
+1I was classed as asteroid due to lack of a coma, and no noticeable activity. 
+This contrasts the second ISO, 2I/Borisov, which had many cometary characteristics \citep[see ][for a review]{Dorofeeva2023}. %\\TODO cite, as opposed to 2I??
 \omuamua was measured to have a rotation period of \qty{8.67(0.34)}{\hour} \citep{Belton2018} and seemed to be tumbling \citep[e.g.][]{Drahus2018,Fraser2018}.
@@
 With TESS data coming in $\frac12\unit{\hour}$ chunks, 24 interpolated points are needed between each API call.
-This interpolation should be accurate, as asteroids move at close to a TESS pixel per FFI on average \citep{Pal2018,Pal2020}, and checking against a higher frequency query to JPL Horizons confirmed this accuracy.
+This interpolation should be accurate, as asteroids move at close to a TESS pixel per FFI on average \citep{Pal2018}, and checking against a higher frequency query to JPL Horizons confirmed this accuracy.
 For the shorter FFIs, the asteroids will be move fewer pixels per frame, which could cause slight mismatches between the interpolation and the detections.
@@
 %TODO data cleaning explaination
+Following \citet{McNeill2023}, any fluxes more than $3\sigma$ from the mean value are sigma-clipped from the lightcurves.
 
@@
 A widely used technique in determining the periods of astronomical data is the Lomb-Scargle periodogram \citep[\citet{Lomb1976,Scargle1982}, but see][for a review]{VanderPlas2018}.
-This method can efficently take discrete timeseries data of which the observations are unevenly spaced and calculate the frequencies present in the sample.
-TESS should have rather evenly spaced data, as the FFIs are produced on a set candance, but sometimes data will be missing.
-For example it may have been sigma clipped due to a spike in brightness that is unphysical for an asteroid.
-The Lomb-Scargle periodogram deals with this uneven sampling much better than either a fast fourier transform or a direct least squares fit.
+This method can efficiently take discrete timeseries data of which the observations are unevenly spaced and calculate the frequencies present in the sample.
+TESS should have rather evenly spaced data, as the FFIs are produced on a set candence, but sometimes data will be missing.
+For example, it may have been sigma clipped due to a spike in brightness that is unphysical for an asteroid, or TESS sees the same asteroid in the same sector on either side of the day long mid-sector break it takes for data transfer.
+The Lomb-Scargle periodogram deals with this uneven sampling much better than either a fast Fourier transform or a direct least squares fit.
 
-There are a few different implementation of a Lomb-Scargle periodogram avalible in world of astronomy python packages.
+There are a few different implementations of a Lomb-Scargle periodogram available in world of astronomy python packages.
 I first tried using the \texttt{Lightkurve} \citep{Lightkurve2018} package built for period analysis of TESS (and Kepler) data of variable stars.
-\texttt{Lightkurve} was easy to use but didn't allow for fine tuning of the periodogram, or give any model of the curve it produced.
-Peeling back a layer of abstraction and useing the Lomb-Scargle periodogram as implemented by \texttt{Astropy}\citep{Astropy2022} was the next option.
+\texttt{Lightkurve} was easy to use but didn't allow for fine-tuning of the periodogram, or give any model of the curve it produced.
+Peeling back a layer of abstraction and using the Lomb-Scargle periodogram as implemented by \texttt{Astropy}\citep{Astropy2022} was the next option.
 There were interesting similarities and differences between the packages.
-\texttt{Lightkurve} is based on the \texttt{Astropy} methods, but dose not give all of the functionallity, instead opting to simplify the process.
-\texttt{Astropy} gives a statstic on how confident it is in the detected frequency is, the false alarm probability.
+\texttt{Lightkurve} is based on the \texttt{Astropy} methods, but does not give all the functionality, instead opting to simplify the process.
+\texttt{Astropy} gives a statistic on how confident it is in the detected frequency is, the false alarm probability.
 It also returns a model, in the form of

### 2024-09-06 FinalReport/FinalReportDraft.tex
cut with a bkg limit

@@
 
+%TODO YORP spin up
+
 Understanding the rotation properties of asteroids has long been of interest to astronomers \citep[e.g.][for early work into the limits of rotation period and the tumbling nature of some small bodies]{Weidenschilling1981,Harris1994}.
@@
 
-\subsection*{Period detection}\label{SubSec:Periods}
+\subsection{Period detection}\label{SubSec:Periods}
 
@@
 \texttt{Lightkurve} was easy to use but didn't allow for fine-tuning of the periodogram, or give any model of the curve it produced.
-Peeling back a layer of abstraction and using the Lomb-Scargle periodogram as implemented by \texttt{Astropy}\citep{Astropy2022} was the next option.
+The next option was peeling back a layer of abstraction and using the Lomb-Scargle periodogram as implemented by \texttt{Astropy}\citep[\citet{Astropy2022} but see][for the implementation]{Vanderplas2012,Vanderplas2015}.
 There were interesting similarities and differences between the packages.

### 2024-09-06 Test Code/findAndInterp.py

@@ -204,7 +204,58 @@ def find_asteroids(sector,cam,ccd,cut):
 
     res, timeList = querySB(setupQuery(sector, cam, ccd, cut), numTimesteps=54, qRad=3.05)
 
-    unqNames = np.unique(res["Name"])
+    print(len(np.unique(res["Name"])))
+
+    targetWSC = fits.open(f"../OzData/{sector}_{cam}_{ccd}_{cut}_wcs.fits")[0]
+    w = wcs.WCS(targetWSC.header)
+    
+    coords=SkyCoord(ra = res["RA"], dec = res["Dec"], unit="deg")
+
+    xCoord, yCoord  = w.all_world2pix(coords.ra, coords.dec,0)
+
+    #rounds to nearest whole number, which returns #.0 as a float. then int converts. if just int masked, it floors the number, not rounds it.
+    xCoord = xCoord.round().astype(int)
+    yCoord = yCoord.round().astype(int)
+    
+
+    #add X, Y, Fs in
+    res["X"] = xCoord
+    res["Y"] = yCoord
+
+    fluxBounds = 513
+
+    badIds = []
+    badIds+= (np.where((xCoord<-1) | (xCoord>= fluxBounds)| (yCoord<-1) | (yCoord>= fluxBounds))[0].tolist()) #! ugly af, but takes all out of bounds values in X or Y, and gets them into the badIDs
+    #give extra space for interploations to work with
+    res.drop(index=badIds, inplace=True)
+
+    interpRes = interplolation_of_pos(res, sector)
+
+    interpRes.drop(columns=["X", "Y"], inplace=True)
+
+    coords=SkyCoord(ra = interpRes["RA"], dec = interpRes["Dec"], unit="deg")
+
+    xCoord, yCoord  = w.all_world2pix(coords.ra, coords.dec,0)
+
+    #rounds to nearest whole number, which returns #.0 as a float. then int converts. if just int masked, it floors the number, not rounds it.
+    xCoord = xCoord.round().astype(int)
+    yCoord = yCoord.round().astype(int)
+    
+
+    #add X, Y, Fs in
+    interpRes["X"] = xCoord
+    interpRes["Y"] = yCoord
+
+    fluxBounds = 512
+
+    badIds = []
[…]

### 2024-09-06 Test Code/makeLightcurve.py

@@ -13,13 +13,17 @@ import numpy.typing as npt
 import matplotlib.colors as mplc
 import matplotlib.cm as mpcm
 import itertools
+from photutils.aperture import CircularAperture
+from photutils.aperture import aperture_photometry
+from astropy.stats import SigmaClip as astroSC
+from photutils.background import Background2D, MedianBackground
 
 import warnings
 from astropy.utils.exceptions import AstropyWarning
 warnings.simplefilter('ignore', category=AstropyWarning)
+warnings.simplefilter('ignore', category=RuntimeWarning)
 
-
-%matplotlib widget
+# %matplotlib widget
 
 plt.rcParams.update({
     "font.size": 18,
@@ -92,6 +96,39 @@ def sum_fluxes(frames, ys, xs):
         if xmin < 0: xmin = 0
         if xmax >= fluxBounds[2]: xmax = fluxBounds[2]-1
 
+
+        #???? comTotal.append(np.sum(reducedFluxes[int(f),ymin:ymax,xmin:xmax]))
+        
+
+        com = ndimage.center_of_mass(reducedFluxes[int(f),ymin:ymax,xmin:xmax])
+        comX = x-checkDown+com[1]
+        comY = y-checkDown+com[0]
+        
+        try:
+            comX = int(round(comX))
+            comY = int(round(comY))
+        except:
+            comTotal.append(np.nan)
+            continue
+        
+        photTable = aperture_photometry(reducedFluxes[int(f),:,:],CircularAperture((comX,comY), r=1.5))
+
+        comAppFlux = photTable["aperture_sum"][0]
+        comTotal.append(comAppFlux)
+        continue
+
+
+
+        ymin = comY-down
+        ymax = comY+up
+        xmin = comX-down
+        xmax = comX+up
+
+        if ymin < 0: ymin = 0
+        if ymax >= fluxBounds[1]: ymax = fluxBounds[1] -1
+        if xmin < 0: xmin = 0
+        if xmax >= fluxBounds[2]: xmax = fluxBounds[2]-1
+
         comTotal.append(np.sum(reducedFluxes[int(f),ymin:ymax,xmin:xmax]))
         
 
[…]

### 2024-09-06 Test Code/periodDetection.py

@@ -6,7 +6,7 @@ import lightkurve as lk
 import astropy.units as u
 from astropy.timeseries import LombScargle as lsp
 
-%matplotlib widget 
+# %matplotlib widget 
 
 plt.rcParams.update({
     "font.size": 18,
@@ -102,7 +102,7 @@ def name_cut(df, name:str, colName:str="NameMatch"):
     toReturn.reset_index(drop=True, inplace=True)
     return toReturn
 
-numObser = 3 #*any less and NaNs come up, or it completely breaks
+numObser = 3 #*any less than 3 and NaNs come up, or it completely breaks
 #TODO, what should it be for a reliable P
 
 def compute_periods_lk(posDF):
@@ -128,15 +128,16 @@ def compute_periods_lk(posDF):
     return lkPeriods, badCount
 
 def compute_periods_ap(posDF):
-    unqNames = pd.unique(interpLcDF["Name"])
+    unqNames = pd.unique(posDF["Name"])
     periodList_ap = []
     badCount = 0
 
     for name in unqNames:
-        nameDf = name_cut(interpLcDF, name, colName="Name")
+        nameDf = name_cut(posDF, name, colName="Name")
         
+        nameProperties = name_cut(astrData, name, colName="Name")
         
-        if nameDf.shape[0]<=numObser:
+        if nameDf.shape[0]<=numObser or nameProperties["Over Background Limit"][0] ==False:
             badCount+=1
             # print(f"{badCount}. not enough points (<={numObser}) for {name}")
             continue
@@ -174,6 +175,11 @@ cut = 7
 
 interpLcDF = load_interps(22,1,3,7)
 
+astrData = pd.read_csv(f"./asteroids_in_{sector}_{cam}_{ccd}_{cut}_properties.csv")
+
+astrData.drop(columns = ["Unnamed: 0"], inplace=True)
+
+
 unqNames = np.unique(interpLcDF["Name"])
 
 # lkPer, badCountlk = compute_periods_lk(interpLcDF)
@@ -183,18 +189,11 @@ apPer, badCountap = compute_periods_ap(interpLcDF)
 
 #Combine with what we already have for each asteroid
 
[…]

### 2024-09-10 Test Code/compareResults.py
deleted names with spaces replaced with no leading/trailing space

@@ -5,9 +5,9 @@ Currently faking random data.
 
 B Leicester 26/4/24
 
-Last edited 26/4/24
+Last edited 9/9/24
 '''
-
+ 
 import numpy as np
 import matplotlib.pyplot as plt
 import pandas as pd
@@ -27,12 +27,16 @@ def useKDTree(df1, df2, cols, maxDist:float=0.01,maxTimeSep:float=0.1,k:int=1):
     for c in fts:
         df2[f"{c}Match"]= df1[c].values[indices]
 
+
     df2.rename(columns={'Unnamed: 0Match':"IDMatch"}, inplace=True, errors="raise")
 
     #// TODO compare time at indices returned
     timeInds = df2.index[np.abs((df2["Time"]-df2["TimeMatch"]))<maxTimeSep]
 
+
     df2=df2.loc[timeInds]
+
+
 
     #// TODO take only nearest match for each point
     smallDistInds = df2.index[df2["distToMatch"]<maxDist]
@@ -110,15 +114,15 @@ indexsOfInterest = detectedSourcesAll.index[(detectedSourcesAll["Type"] == "0") 
 
 detectedSources = detectedSourcesAll.loc[indexsOfInterest]
 
-detectedSources["jd"] = detectedSources["mjd"] +2400000.5
+# detectedSources["jd"] = detectedSources["mjd"] +2400000.5
 detectedSources.reset_index(drop=True, inplace=True)
 
 colsToUse = ["RA", "Dec"]
 
-interpRes.rename(columns={"epoch":"Time"}, inplace=True)
+interpRes.rename(columns={"MJD":"Time"}, inplace=True)
 
 
-detectedSources.rename(columns={"ra":"RA", "dec":"Dec", "jd":"Time"}, inplace=True)
+detectedSources.rename(columns={"ra":"RA", "dec":"Dec", "mjd":"Time"}, inplace=True)
 
 matches = useKDTree(df1=interpRes.copy(deep=True), df2=detectedSources.copy(deep=True), cols=colsToUse, maxDist=21/3600, maxTimeSep=0.025)
 
@@ -150,17 +154,19 @@ matches.to_csv(f"./{sector}_{cam}_{ccd}_{cut}_InterpAndDetect_Matches.csv")
 
 unqNames = pd.unique(interpRes["Name"])
 
-obsForDetect=40 #number of observations needed to count a detection
+obsForDetect=3 #number of observations needed to count a detection
 
 
 interpedMv = []
[…]

### 2024-09-10 Test Code/expectedPositionsSkyBot.py

@@ -417,7 +417,7 @@ unqNames = list(pd.unique(res['Name']))
 #* interpolating data. needed so am no doing so many API querries.
 
 dfsList = []
-
+ 
 for i, name in enumerate(unqNames):
     indexs= res.index[res["Name"]==name]
     underSampledPos = res.loc[indexs]
@@ -442,7 +442,7 @@ for i, name in enumerate(unqNames):
 
 
     #interpTimes = np.linspace(minTime,maxTime, int(interpPoints*deltaTime))#linspace to sample #! just use times of frames from cut 
-
+    interpTimes = np.load(f"../OzData/sector{sector}_cam{cam}_ccd{ccd}_cut{cut}_of16_Times.npy")
 
     #Ra and Dec samples
     interpRAs = np.interp(x=interpTimes, xp=underSampledPos["epoch"], fp=underSampledPos["RA"])
@@ -464,7 +464,7 @@ for i, name in enumerate(unqNames):
 interpRes = pd.concat(dfsList) #puts evrything back together
 interpRes.reset_index(drop=True, inplace=True)
 # # interpRes.to_csv(f"./InterpolatedQuerryResult_ra{myTargetPos[0]}_dec{myTargetPos[1]}_t{myTargetPos[2].mjd}_Mv{magLim}.csv")
-interpRes.to_csv(f"./InterpolatedQuerryResult_{sector}_{cam}_{ccd}_{cut}.csv")
+# interpRes.to_csv(f"./InterpolatedQuerryResult_{sector}_{cam}_{ccd}_{cut}.csv")
 
 # posFig = plotExpectedPos(interpRes, times, myTargetPos, magLim=magLim, scaleAlpha=True, saving=True)
 # posFig.savefig(f"./InterpolatedExpectedPositionsPlot_HscaleOn_ra{myTargetPos[0]}_dec{myTargetPos[1]}_t{myTargetPos[2].mjd}_Mv{magLim}.png")

### 2024-09-10 Test Code/findAndInterp.py

@@ -275,9 +275,6 @@ def find_asteroids(sector,cam,ccd,cut):
     withEles.to_csv(f"{fname}.csv")
 
 
-
-
-
     return interpRes
 
 
@@ -351,7 +348,7 @@ def interplolation_of_pos(posDf, sector):
 
     #final clean
     interpRes.reset_index(drop=True, inplace=True)
-    interpRes.drop(columns=["Mv","Class"], inplace=True)
+    interpRes.drop(columns=["Class"], inplace=True)
     
 
 

### 2024-09-10 Test Code/makeLightcurve.py

@@ -23,7 +23,7 @@ from astropy.utils.exceptions import AstropyWarning
 warnings.simplefilter('ignore', category=AstropyWarning)
 warnings.simplefilter('ignore', category=RuntimeWarning)
 
-# %matplotlib widget
+%matplotlib widget
 
 plt.rcParams.update({
     "font.size": 18,
@@ -116,8 +116,6 @@ def sum_fluxes(frames, ys, xs):
         comAppFlux = photTable["aperture_sum"][0]
         comTotal.append(comAppFlux)
         continue
-
-
 
         ymin = comY-down
         ymax = comY+up
@@ -361,6 +359,45 @@ def setupQuery(sector,cam,ccd,cut,dirPath="../OzData/"):
     return myTargetPos
 
 
+def recoveredHist(astrProperties, bkgLim):
+    """Takes the mean and deviation of the fluxes, and compares to a background Limit. A histogram of those greater than the limit is made. If no asteroid is in a mag bin, then 'all' are recovered (i.e. 0/0=1 here)"""
+    
+    fig, ax = plt.subplots()
+
+    magMeans = astrProperties["Mv(mean)"].values
+
+    histRange = (13,20)
+    histBins = 14
+
+    vMagHist = np.histogram(magMeans, range=histRange, bins=histBins)
+
+    overBkgData = np.where((astrProperties["Mean COM Flux"]-astrProperties["STD COM Flux"])>=bkgLim)[0]
+
+    ax.set(xlabel="Mean V mag", ylabel="Fraction recovered")
+    #* Note recovered is when mean-std > bkg, 84% of data above noise floor
+    #* recoverd is 1 when no points in bin anyway. 
+    passedBkgLim = np.where((astrProperties["Mean COM Flux"]-astrProperties["STD COM Flux"])>=bkgLim, True, False)
+
+    # astrProperties["Over Background Limit"] = passedBkgLim
+
+    overBkgMags = magMeans[passedBkgLim]
+
+    overBkgHist = np.histogram(overBkgMags, range=histRange, bins=histBins)
+
+    normedHist = overBkgHist[0]/vMagHist[0]
+
[…]

### 2024-09-10 Test Code/periodDetection.py

@@ -6,7 +6,8 @@ import lightkurve as lk
 import astropy.units as u
 from astropy.timeseries import LombScargle as lsp
 
-# %matplotlib widget 
+%matplotlib widget 
+
 
 plt.rcParams.update({
     "font.size": 18,
@@ -68,12 +69,13 @@ def detect_period_ap(times, fluxes, plotting = False, knownFreq = None):
         ax[0].scatter(bestFreq,bestPow, c="gold",marker="*", s=100, label=f"Best Frequency = {bestFreq.round(7)} \nFalse Alarm Probability = {f_a_prob.round(3)}")
 
         if knownFreq is not None:
-            ax[0].axvline(knownFreq, linestyle=":", c="tab:blue")
-            ax[0].axvline(2*knownFreq, linestyle=":", c="tab:purple")
+            ax[0].axvline(knownFreq, linestyle=":", c="tab:blue", label = f"Known F={knownFreq.round(7)}")
+            ax[0].axvline(2*knownFreq, linestyle=":", c="tab:purple", label = f"Known F alias (2*)")
+            ax[0].axvline(0.5*knownFreq, linestyle=":", c="tab:red", label = f"Known F alias (0.5*)")
 
-        ax[0].legend()
-
-        ax[1].set(xlabel="Time [JD]", ylabel="Flux")
+        ax[0].legend(fontsize=10)
+        meanFlux = np.mean(fluxes)
+        ax[1].set(xlabel="Time [JD]", ylabel="Flux")#, ylim=(meanFlux-100, meanFlux+100))
         ax[1].scatter(times, fluxes, label = "Light curve", c="tab:orange", marker = "d", s=10)
         ax[1].plot(t_fit.to(u.day),y_fit, ls="--", c="k", label="Model")
 
@@ -137,7 +139,7 @@ def compute_periods_ap(posDF):
         
         nameProperties = name_cut(astrData, name, colName="Name")
         
-        if nameDf.shape[0]<=numObser or nameProperties["Over Background Limit"][0] ==False:
+        if nameDf.shape[0]<=numObser:# or nameProperties["Over Background Limit"][0] ==False:
             badCount+=1
             # print(f"{badCount}. not enough points (<={numObser}) for {name}")
             continue
[…]

### 2024-09-10 Test Code/ToPutOnOzstar/asteroidCatJob.sh
put some stuff on ozstar, it doesn't work

@@ -1,0 +1,12 @@
+#!/bin/bash
+#
+#SBATCH --job-name=TESS_S22_Asteroid_Cat
+#SBATCH --output=/fred/oz335/bleicester/joblogs/%A_Asteroid_Cat_Logs.txt
+#SBATCH --error=/fred/oz335/joblogs/bleicester/%A_asteroid_Cat_errors.txt
+#
+#SBATCH --ntasks=1
+#SBATCH --time=1-00:00:00
+#SBATCH --cpus-per-task=1
+#SBATCH --mem-per-cpu=4G
+
+python /fred/oz335/bleicester/Code/runAstrCat.py

### 2024-09-10 Test Code/ToPutOnOzstar/asteroidCatalouge.py

@@ -1,10 +1,5 @@
 import numpy as np
-import matplotlib.pyplot as plt
-import numpy as np
 import numpy.typing as npt
-import matplotlib.pyplot as plt
-import matplotlib.colors as mplc
-import matplotlib.cm as mpcm
 import pandas as pd
 from astropy.coordinates import SkyCoord
 from astropy.time import Time
@@ -13,23 +8,9 @@ from astropy import wcs
 from astropy.utils.data import download_file
 from astropy.io import fits
 from astroquery.jplhorizons import Horizons
-import itertools
 from tqdm import tqdm
-from scipy.spatial import KDTree
 
 
-plt.rcParams.update({
-    "font.size": 18,
-    "font.family": "serif",
-    "figure.autolayout": True,
-    "axes.grid": False,
-    # "xtick.minor.visible": True,
-    # "ytick.minor.visible": True,
-})
-
-
-
-timeOffset = 2400000.5
 
 def _Skybotquery(ra, dec, times, radius=10/60, location='C57', cache=False):
     """Returns a list of asteroids/comets given a position and time.
@@ -91,9 +72,9 @@ def _Skybotquery(ra, dec, times, radius=10/60, location='C57', cache=False):
     return df
 
 
-def setupQuery(sector,cam,ccd,cut,dirPath="../OzData/"):
+def _setupQuery(dirPath="../OzData/"):
 
-    fname = f"{dirPath}{sector}_{cam}_{ccd}_{cut}_wcs.fits"
+    fname = f"{dirPath}wcs.fits"
 
     targetWSC = fits.open(fname)[0]
 
@@ -105,7 +86,7 @@ def setupQuery(sector,cam,ccd,cut,dirPath="../OzData/"):
 
     return myTargetPos
 
-def querySB(targetPos: list, qRad: float = 10.0, qLoc: str = "C57", numTimesteps: int = 27, magLim: float = 20.0) -> tuple[pd.DataFrame, npt.ArrayLike]:
+def _querySB(targetPos: list, qRad: float = 10.0, qLoc: str = "C57", numTimesteps: int = 27, magLim: float = 20.0) -> tuple[pd.DataFrame, npt.ArrayLike]:
     """
     Queries SkyBot in a box of width qRad after making a list of the times to query at. returns this timelist after the query has been made and down selected to just the objects brighter than magLim
 
[…]

### 2024-09-10 Test Code/ToPutOnOzstar/runAstrCat.py

@@ -1,0 +1,19 @@
+from asteroidCatalouge import make_asteroid_cat
+import numpy as np
+
+thisSector = 22
+
+#*For Full runs
+# possibleCams = [1,2,3,4]
+# possibleCCDs = [1,2,3,4]
+# possibleCuts = [1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16]
+
+#For testing run
+possibleCams = [1]
+possibleCCDs = [3]
+possibleCuts = [7]
+
+for i in  possibleCams:
+    for j in possibleCCDs:
+        for k in possibleCuts:
+            make_asteroid_cat(thisSector,i,j,k)

### 2024-09-10 Test Code/makeLightcurve.py

@@ -465,6 +465,8 @@ bkgLim = np.mean(medBkgs)
 meanFluxes = []
 stdFluxes = []
 
+numObser = 3
+
 for i,name in enumerate(namesAfter):
     nCut = name_cut(totalLcDf,name,"Name")
     cflux = nCut["COM Flux"] 
@@ -474,7 +476,7 @@ for i,name in enumerate(namesAfter):
     sdev = np.std(cflux)
     stdFluxes.append(sdev)
 
-    if mean - sdev >= bkgLim :
+    if mean - sdev >= bkgLim and len(nCut.index)>=numObser:
         count+=1
         ax.errorbar(i,mean,sdev, markersize=10, capsize=3)
 

### 2024-09-10 Test Code/periodDetection.py

@@ -23,6 +23,8 @@ plt.rcParams.update({
 
 def detect_period_lk(times, fluxes, fluxErrs):
     """
+    DEPRICATED
+    
     Calculates the periodogram using LightKurve and returns the period at the max power.
     """
     lc =lk.LightCurve(data=None, time=times, flux=fluxes, flux_err=fluxErrs)
@@ -47,6 +49,8 @@ def detect_period_ap(times, fluxes, plotting = False, knownFreq = None):
     
 
     minFreq = 1/(2*(times.max()-times.min())*u.day).to(u.s)
+    # minFreq = 1e-5/u.s
+
     maxFreq = 1/nyquistP
 
     lsper = lsp((u.Quantity(times, u.day)).to(u.s), fluxes)
@@ -108,7 +112,7 @@ numObser = 3 #*any less than 3 and NaNs come up, or it completely breaks
 #TODO, what should it be for a reliable P
 
 def compute_periods_lk(posDF):
-
+    """DEPRECATED"""
     unqNames = pd.unique(posDF["Name"]) #!name col better
     periodList_lk = []
     badCount = 0
@@ -236,14 +240,6 @@ print("")
 # singleNameLSP(interpLcDF,maxPap["Name"])
 
 
-fig2, ax2 = plt.subplots()
-
-ax2.scatter(allData["Mean COM Flux"], allData["False Alarm Probability"])
-
-ax2.set(xlim=(-10,10))
-
-
-
 # minPlk = lkPer.loc[lkPer["Best Period"].idxmin()]
 minPap = apPer.loc[apPer["Best Period [Seconds]"].idxmin()]
 print("")
@@ -278,7 +274,7 @@ compedPs = pd.DataFrame(inLCDBList, columns=["Name", "Known Period", "Found Peri
 print(compedPs)
 
 
-trialName = "Bernoulli"
+trialName = "Ruff"
 
 try:
     knownFreq = 1/(compedPs.at[compedPs.index[compedPs["Name"]==trialName].values[0],"Known Period"]*60*60)
@@ -321,23 +317,38 @@ matchCut = name_cut(matchesDf, trialName)
 ax.scatter(matchCut["Time"],matchCut["flux"], label="Matched Flux",c="Pink", marker="^")
 
 
-ax.set_ylim(100,800)
+# ax.set_ylim(100,800)
 ax.legend()
 
 
[…]

### 2024-09-11 FinalReport/FinalReportDraft.tex
more writting at ISO chat

@@
 \DeclareSIUnit{\mag}{\rm{mag}}
+\DeclareSIUnit{\px}{\rm{px}}
 
@@
 By using the higher cadence FFIs of later TESS sectors, the $P\leq\qty{3}{\hour}$ limit on the accurate periods \citep[as found by ][]{McNeill2023} should be lowered.
-Another beneficial part of this work is that as part of a full sky transient survey using TESS, \texttt{TESSELLATE} \citep{TESSELATE}.
+Another beneficial part of this work is that as part of a full sky transient survey using TESS, \texttt{TESSELLATE} \citep{TESSELLATE}.
 Asteroids can spike the brightness of a pixel for only a few frames while it is passing by.
@@
 With TESS data coming in $\frac12\unit{\hour}$ chunks, 24 interpolated points are needed between each API call.
-This interpolation should be accurate, as asteroids move at close to a TESS pixel per FFI on average \citep{Pal2018}, and checking against a higher frequency query to JPL Horizons confirmed this accuracy.
-For the shorter FFIs, the asteroids will be move fewer pixels per frame, which could cause slight mismatches between the interpolation and the detections.
-Horizons was also used get properties of the individual asteroids, such as the absolute magnitude $H$.
-For the faster TESS data, more interpolated points are needed, but a smaller the change in position between each point.
+Because of a small change in position between each frame, \citep[$\sim \qty{1}{\px}$ per frame][]{Pal2018}, this interpolation should be accurate.
+Checking against a higher frequency query to the JPL Horizons ephemeris  confirmed this accuracy on a few asteroids. %TODO fig
+This is not a feasable way of getting all of the positions, as the queries are per name, not per area like the conesearch in Skybot.
+They are also rate limited, so while a few asteroids can be checked, filling the gaps bewteen the Skybot half day queries for all the asteroids would take far too long.
+For the faster TESS FFIs, more interpolated points are needed, but a smaller the change in position between each point will be.
+Forcing the interpolations to be at the same time as the TESS frames, which are known thanks to \texttt{TESSELLATE} saving them out, simplifies the matching to detections and forced photometry of these points.
 These interpolated positions can be seen in \autoref{Fig:interpPos} for a cut from \texttt{TESSELLATE}.
@@
 The number of frames do change, type Ia supernova will brighten is a matter of a few hours and then dim for days, while stellar flares are of similar profile by a smaller max brightness and a corrsepondingly shorter decay time.
-Asteroids are very short, however detection pipelines are robust.
-These pipelines have already found the aforementioned supernova and stellar flares, the job of this work is to catch all the asteroids in the set of all the transients.
+Asteroids are very short spikes, however these detection pipelines are robust. %? Clarinda begs to differ
+These pipelines have already found the aforementioned supernova, stellar flares and variable stars, by matching to catalogues of those types of objects (Gaia, ZTF etc.).
+The job of this work is to catch all the asteroids in the set of all the transients that are returned from the pipeline.
 To do this, catelogue matching is in order.
@@
 The matches from the detections, which already have a flux calculated, and the interpolated points themselves, which are more numerous but require forcing the photometry.
-There were some challenges getting the flux of the interpolated points, even when \texttt{TESSELLATE} has already reduced all the FFIs of interest, due to the timing of the TESS frames, but these were identified and corrected for.
-Not every interpolated point gets a match, due to a variety of reasons, %pass in front of stars, dip below limiting mag, bad pipeline, not extreme enough of a difference between frames 
-so a comparison between the two light curves is interesting.
+
+There were some challenges getting the flux of the interpolated points, even when \texttt{TESSELLATE} has already reduced all the FFIs of interest.
+The flux is summed in  $3\times3$ box centred on the integer $xy$ position returned from changing the world coordinates in RA and Dec to pixel coordinates using \texttt{Astropy}'s \texttt{SkyCoord} subpackage.
+%?
+cite
+
+The integer coordinates of the asteroids seemed to be leading the object in any given frame of TESS. %?FIG?
+This was decreasing the total flux of the asteroid, as the box used for sumation of the fluxes was hitting mostly sky.
+This also caused a ``sawtoothing'' effect, where the lighturve would jump from a minimum to a maximum when the integer coordinate snaped from being a pixel out to back on the asteroid.
+This both of these issues can be fixed by calculation of the centre of mass (COM) flux, the average brightness in $xy$ space on a $8\times8$ cut of the image centred on the interpolated position.
+I didn't take the highest pixel value in this box because there could be a rouge star, this would drag the COM around, but not by as much as snapping to that highest point.
+A \texttt{Photutils} %TODO cite 
+aperture photometry call to this COM with a radis of \qty{1.5}{\px} was enough to get fluxes of the same size, but with a far more smooth transition from maximum to minimum, as expected from rotating asteroids.
+
+Not every interpolated point gets a match, due to a variety of reasons, which culminate in there being not an extreme enough of a difference between frames.
+A comparison between the two light curves is interesting.
+
 \autoref{Fig:DifFlux} shows these two light curves for a chosen asteroid, ``Ruff''.
-This was picked as it has a high number of points matched to detections.
-There are more interpolated points than matched points, as not ever interpolated point gets matched to a detection.
+This was picked as it has a high number of points matched to detections, but of course still more interpolated points total.
 There seems to be a systematic offset between the fluxes, with the interpolated points having consistently lower flux, (here the means differ by 58 counts)
@@
 
+The completeness of the asteroid population in TESS as a function of their apparent magnitude can be determined.
+The apparent magnitude is one of the properties returned from the queries, and is in the V band.
+By taking a limit on the if an asteroid is detected \dots %TODO what is the limit
+Breaking the average V mag of these objects into \qty{0.5}{\mag} bins for both the sample returned from a query and the sample detected, a ratio of completeness can be calculated per bin.
+(Note: If there is no asteroids in a bin, none have a chance of being recovered, so the all did get recovered. I.e. $\frac{0}{0}=1$)
+
 \section{Results}\label{Sec:Res}

### 2024-09-17 FinalReport/FinalReportDraft.tex
fixed while hanging (!!!!!) also have report in proper format from MA

@@
-\documentclass[12pt]{article}
+\documentclass{UCreport}
 
-\usepackage{natbib,hyperref,graphicx}
+\usepackage[authoryear,comma,round]{natbib}
+\usepackage{hyperref,graphicx}
 \usepackage[separate-uncertainty=true,multi-part-units=single]{siunitx}
 
-\usepackage[margin=1.5cm, left=3cm ,twoside]{geometry}
+% \usepackage[margin=1.5cm, left=3cm ,twoside]{geometry}
 
@@
 
-\title{ASTR480 Final Report:\\
-\textit{Asteroids in TESS}
-}
-\author{Brayden Leicester \\ 
-\textit{61352063} \\
-[1ex]\small{Supervisors: Michele Bannister and Ryan Ridden-Harper}
-}
+% \title{ASTR480 Final Report:\\
+% \textit{Asteroids in TESS}
+% }
+% \author{Brayden Leicester \\ 
+% \textit{61352063} \\
+% [1ex]\small{Supervisors: Michele Bannister and Ryan Ridden-Harper}
+% }
 \begin{document}
-\pagenumbering{gobble} %to blank pg # before Contents pg
-%TODO declaration page ????
-%\newpage
-\maketitle
-%*{ as per the report guidelines
-% \newpage
-% \section*{Abstract}
+
+%----------- Report information ---------
+
+\logo{logos/UC.png}
+\school{\textbf{School of Physical and Chemical Sciences}}
+\course{\textbf{ASTR480}}
+\student{\textbf{Brayden Leicester}}
+\supervisor{Dr Michele Bannister and Dr Ryan Ridden-Harper}
+\ttitle{Asteroids in TESS} %title of the file
+
+%----------- Init -------------------
+
+\buildmargins % display margins
+\buildcover % create the front cover of the document
+
+%------------ Declaration ----------------
+\pagenumbering{roman}
+\addcontentsline{toc}{section}{Declaration}
+\declaration
+
+I certify that content of this report was mostly my own work.
+\\
+
+My supervisor helped by proofreading the report and offering feedback. \\
+
+Graduate students, X and Y, helped me understand ... \\
+
+The derivation in section 3 was taken from ... \\
+
+\vspace{2cm}
+\begin{centering}
+\textbf{\student}\par
+\end{centering}
+
+\newpage
+
+%------------ Abstract ----------------
+
+\addcontentsline{toc}{section}{Abstract}
+\abstract
+
+This is where the abstract goes. \\
+
+This is the next paragraph.
+
 % %*Background
@@
 
-% \newpage
-% \pagenumbering{roman}
-% \tableofcontents
-% \newpage
-% \listoffigures
-% \newpage
-% \listoftables
-% \newpage
-%*}
 \newpage
+%------------ Contents pages ----------------
+
+\toc % creates the table of contents
+
+\addcontentsline{toc}{section}{List of Figures}
+\listoffigures
+\newpage
+\addcontentsline{toc}{section}{List of Tables}
+\listoftables
+\newpage
+
+%------------ Report body ----------------
 \pagenumbering{arabic}
@@
 
-There have been attempts before to find and classify the asteroids in TESS data before.
+There have been attempts to find and classify the asteroids in TESS data before.
 \citet{Pal2018} proposes that TESS will be a good instrument for solar system object study, and simulate some detections down to \nth{20} visual magnitude, but that good photometry should only be expected to $V \lesssim 19$.
+The sentiment of using TESS in this way is shared by \citet{Wong2019}.
 The first data release in \citet{Pal2020} catalogues nearly 10,000 asteroid light curves from the first year of TESS operation, and they report that they triple the number of asteroids with accurate rotation periods.
 \citet{McNeill2023} also analyse TESS cycle 1, and detect almost 38,000 objects.
-They determine reliable rotation periods for about 3,500 asteroids in this sample, and show lack of reliability for objects with periods less than \qty{3}{\hour}.
+They determine reliable rotation periods for about 3,500 asteroids in this sample and show a lack of reliability for objects with periods less than \qty{3}{\hour}.
 In the overlap from the data from \citet{Pal2020}, \citeauthor{McNeill2023} find good agreement between the two sets of periods and amplitudes.
+The Minor Planet Center (MPC) gets regular updates from TESS thanks to the work of \citet{Woods2021} and their \texttt{LINEAR-TESS} program, which creates tracklets of objects over a day, and then stitches these together to form a track of each asteroid through an FFI.
 \citet{Gowanlock2024} uses TESS photometry as well as  observations by the Zwicky Transient Facility \citep[ZTF, ][]{Bellm2018} to get a longer baseline on mutually observed objects while combining ground and space based observations.
 There were not very many objects in both samples with known periods for comparison, with only 222 objects analysed, however they demonstrate that the technique is effective.
+Fainter and unknown solar system objects can be found by shift and stacking \citep{Holman2019, Payne2019,Rice2020} or taking a fast X-ray transform \citep{Nguyen2024} of the FFI data.
 
@@
 Because of such a rich history of study, many asteroids have a known rotation period.
-These are catalogued by \citet{Warner2009} %TODO updated in citation
+These are catalogued by \citet{Warner2009} %TODO updated in citation http://www.MinorPlanet.info/php/lcdb.php
 in the Light Curve Database (LCDB).
@@
 
+% \begin{table}
+%   \caption[Tab1]{My table}
+
+% \end{table}
+
 \subsection{Querrying Databases}\label{SubSec:Querry}
@@
 For most asteroids, their orbital elements are well known, so it is a matter of looking them up and cross-matching with transients in the TESS data.
-Python was used to make API calls to {Skybot}\footnote{\href{https://vo.imcce.fr/webservices/skybot/}{Skybot: https://vo.imcce.fr/webservices/skybot/}} to get positions of asteroids in a cone search box in right ascension (RA) and  declination (Dec) space.
+Python was used to make API calls to {Skybot}\footnote{\href{https://vo.imcce.fr/webservices/skybot/}{Skybot: https://vo.imcce.fr/webservices/skybot/}} \citep{Berthier2006} to get positions of asteroids in a cone search box in right ascension (RA) and  declination (Dec) space.
 As TESS sectors are \qty{27}{\day} long, querying every \qty{12}{\hour} is manageable.
@@
 Knowing where the asteroids are is helpful for finding them in the archival TESS data, but knowing more about the individual asteroids is also useful for population statistics.
-  {JPL Horizons}\footnote{\href{https://ssd.jpl.nasa.gov/horizons/}{JPL Horizons: https://ssd.jpl.nasa.gov/horizons/}} was used to obtain the orbital elements of each asteroid, $a$, $e$, and $i$, as well as an absolute magnitude $H$, which is a good proxy for size (as discussed above).
+The API of {JPL Horizons}\footnote{\href{https://ssd.jpl.nasa.gov/horizons/}{JPL Horizons: https://ssd.jpl.nasa.gov/horizons/}} was used to obtain the orbital elements of each asteroid, $a$, $e$, and $i$, as well as an absolute magnitude $H$, which is a good proxy for size (as discussed above).
 %TODO ref
@@
 \section*{Acknowledgements}
+\addcontentsline{toc}{section}{Acknowledgements}
 %OzSTAR asks for:
@@
 
+This research has made use of data and/or services provided by the International Astronomical Union's Minor Planet Center.
+
+This work has used data and/or services provided by NASA's Solar System Dynamics website, https://ssd.jpl.nasa.gov/. Specifically JPL Horizons: https://ssd.jpl.nasa.gov/horizons/
+
 %*Bib
-\bibliographystyle{jphysicsB}
+\addcontentsline{toc}{section}{References}
+% \bibliographystyle{jphysicsB}
 \bibliography{bibfile.bib}

### 2024-09-17 ReportTemplate/main.tex
fixed while hanging (!!!!!) also have report in proper format from MA

--- 2024-09-06 FinalReport/FinalReportDraft.tex
@@
-\documentclass[12pt]{article}
+\documentclass{UCreport}
+\usepackage{lipsum}
+\usepackage{gensymb}
+\usepackage{float}
+\usepackage{graphicx} % Required for inserting images
 
-\usepackage{natbib,hyperref,graphicx}
-\usepackage[separate-uncertainty=true,multi-part-units=single]{siunitx}
+%
+%
+%   To cite by author name, uncomment this line.
+\usepackage[authoryear,comma,round]{natbib} 
+%
+%
+%   To instead cite by number, uncomment this line.
+%\usepackage[square,numbers,comma,sort&compress]{natbib}
 
-\usepackage[margin=1.5cm, left=3cm ,twoside]{geometry}
+\begin{document}
 
-\usepackage[para]{footmisc}
+%----------- Report information ---------
 
-\usepackage[super]{nth}
+\logo{logos/UC.png}
+\school{\textbf{School of Physical and Chemical Sciences}}
+\course{\textbf{PHYS480}}
+\student{\textbf{My Name}}
+\supervisor{Dr Jane Doe}
+\ttitle{Title of the report} %title of the file
 
-\usepackage{amsmath,amssymb}
+%----------- Init -------------------
 
-\DeclareSIUnit{\arcsec}{''}
-\DeclareSIUnit{\au}{\rm{AU}}
-\DeclareSIUnit{\mag}{\rm{mag}}
+\buildmargins % display margins
+\buildcover % create the front cover of the document
 
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
-\usepackage{newunicodechar}
+%------------ Declaration ----------------
 
-\DeclareRobustCommand{\okina}{%
-  \raisebox{\dimexpr\fontcharht\font`A-\height}{%
-    \scalebox{0.8}{`}%
-  }%
-}
-\newunicodechar{ʻ}{\okina}
-\newcommand{\omuamua}{\okina Oumuamua }
+\pagenumbering{roman}
 
-\title{ASTR480 Final Report:\\
-\textit{Asteroids in TESS}
-}
-\author{Brayden Leicester \\ 
-\textit{61352063} \\
-[1ex]\small{Supervisors: Michele Bannister and Ryan Ridden-Harper}
-}
-\begin{document}
-\pagenumbering{gobble} %to blank pg # before Contents pg
-%TODO declaration page ????
-%\newpage
-\maketitle
-%*{ as per the report guidelines
-% \newpage
-% \section*{Abstract}
-% %*Background
+\addcontentsline{toc}{section}{Declaration}
+\declaration
 
-% %*Methods
+I certify that content of this report was mostly my own work.
+\\
 
-% %*Main Results
+My supervisor helped by proofreading the report and offering feedback. \\
 
-% %*Conclusion
+Graduate students, X and Y, helped me understand ... \\
 
-% \newpage
-% \pagenumbering{roman}
-% \tableofcontents
-% \newpage
-% \listoffigures
-% \newpage
-% \listoftables
-% \newpage
-%*}
+The derivation in section 3 was taken from ... \\
+
+\vspace{2cm}
+\begin{centering}
+\textbf{\student}\par
+\end{centering}
+
 \newpage
+
+%------------ Abstract ----------------
+
+\addcontentsline{toc}{section}{Abstract}
+\abstract
+
+This is where the abstract goes. \\
+
+This is the next paragraph.
+
+\newpage
+
+%------------ Contents pages ----------------
+
+\toc % creates the table of contents
+
+\addcontentsline{toc}{section}{List of Figures}
+\listoffigures
+\newpage
+\addcontentsline{toc}{section}{List of Tables}
+\listoftables
+\newpage
+
+%------------ Report body ----------------
+
 \pagenumbering{arabic}
 
-\section{Introduction}\label{Sec:Intro}
+\section{Introduction}
 
-%To get all asteroids in TESS. 
-This project aims to find and characterise the light curves of all the asteroids seen by the Transiting Exoplanet Survey Satellite (TESS).
-As such, a thorough discussion of both is required.
+\subsection{Part of introduction}
+This is a test.
 
-\subsection{TESS}\label{SubSec:TESS}
+\newpage
 
-%Why TESS
-TESS is a large area, high cadence imaging, space telescope  \citep{Ricker2014}.
-TESS is tasked with observing one piece of sky for \qty{27}{\day} at a time (a sector), delivering \qtyproduct{96 x 24}{\degree} full frame images (FFIs) at regular intervals.
-These FFIs are built from stacked \qty{2}{\second} exposures, leveraging the short readout times of the 16 CCD cameras on-board.
-With the initial cadence for these full frame images set to \qty{30}{\minute}, the time resolution of TESS is unparalleled.
-The highest frequency (shortest period) that can be found from time series data like this is at the Nyquist limit, twice the minimum time between images.  
-Due to the short cadence, the Nyquist frequency of TESS is quite high, \qty{1}{\per\hour}. 
-This is well sampled enough to characterise most variable stars, as well as find the orbital periods of exoplanets to a high precision.
-As the mission was extended, after TESS had already mapped the entire sky, the length of the FFIs has decreased to \qty{10}{\minute} and then later to \qty{200}{\second}.
-This decreased in the FFI exposure time comes with a complementary increase in the Nyquist frequency.
-This high time resolution and observation area does come at the cost of spatial resolution, as the pixels are each \qty{21}{\arcsec} square.
-Of interest here is what such a high sampling rate can do for statistic on the asteroid population.
-For bright asteroids the rotation periods should be able to be easily determined from this vast dataset.
-The shortest FFIs will be able to accurately determine the rotation periods of all but the fastest rotating asteroids, most of which will be too dim to see in the TESS data.
+\section{Next section}
 
-There have been attempts before to find and classify the asteroids in TESS data before.
-\citet{Pal2018} proposes that TESS will be a good instrument for solar system object study, and simulate some detections down to \nth{20} visual magnitude, but that good photometry should only be expected to $V \lesssim 19$.
-The first data release in \citet{Pal2020} catalogues nearly 10,000 asteroid light curves from the first year of TESS operation, and they report that they triple the number of asteroids with accurate rotation periods.
-\citet{McNeill2023} also analyse TESS cycle 1, and detect almost 38,000 objects.
-They determine reliable rotation periods for about 3,500 asteroids in this sample, and show lack of reliability for objects with periods less than \qty{3}{\hour}. 
-In the overlap from the data from \citet{Pal2020}, \citeauthor{McNeill2023} find good agreement between the two sets of periods and amplitudes. 
-\citet{Gowanlock2024} uses TESS photometry as well as  observations by the Zwicky Transient Facility \citep[ZTF, ][]{Bellm2018} to get a longer baseline on mutually observed objects while combining ground and space based observations.
-There were not very many objects in both samples with known periods for comparison, with only 222 objects analysed, however they demonstrate that the technique is effective.
+This is another test. A plot is shown in Fig.~\ref{fig:plot}.
 
-This work aims to extend these studies to more sectors, hence characterising more asteroids.
-The aim is to get a volume complete set of asteroid lightcurves and periods, by analysis of the whole sky as seen by TESS using data from more than just the first year of TESS's operation.
-I also employ a different data reduction method, reducing the data with the \texttt{TESSreduce} package \citep{Ridden-Harper2021} should increase the accuracy of the photometry due to the improved background subtraction process.
-Because of the survey properties, TESS provides a self-consistent way to measure the properties of asteroids over the full sky.
-By using the higher cadence FFIs of later TESS sectors, the $P\leq\qty{3}{\hour}$ limit on the accurate periods \citep[as found by ][]{McNeill2023} should be lowered.  
-Another beneficial part of this work is that as part of a full sky transient survey using TESS, \texttt{TESSELLATE} \citep{TESSELATE}.
-Asteroids can spike the brightness of a pixel for only a few frames while it is passing by.
-This can confuse pipelines looking for  transients such as flare stars and supernova, as they can look very similar, having the brightness change in one pixel for a short time.
-The goal of finding all the asteroids will allow for the removal of these spikes from the transient pipeline, as well as to understand the asteroid population better.
-By filtering the asteroid detections out, while also self consistently determining rotation properties of a large sample of the population, more science can be done with the one action.
-
-\subsection{Asteroids}\label{SubSec:Asteroid}
-
-Asteroids are a key class of solar system objects.
-The orbits of asteroids can be well characterised by only a few values. The semi-major axis, $a$, is the largest distance from the sun that the asteroid achieves on its elliptical orbit.
-Main belt asteroids have typical $a$s of \qtyrange{2}{3.3}{\au} \citet{DeMeo2015}. 
-The bounds between the inner, middle and outer main belt at \qty{2.5}{\au} and \qty{2.82}{\au}, these separations are due to resonances with Jupiter (the 3:1 and 5:2 mean motion resonances mark the edges of middle main belt) and Saturn (the $\nu_6$ secular resonance cuts off the lowest $a$ edge of the inner belt) carving swathes where asteroids cannot be in stable orbits.
-The eccentricity, $e$, of this orbit is the next parameter of interest.
-$e=0$ implies a perfectly circular orbit, $0<e<1$ are ellipses, $e=1$ give a parabolic orbit, and $e>1$ are unbound, hyperbolic orbits.
-Main belt asteroids have small $e$, in the range of \qtyrange{0}{0.35}{} \citep{DeMeo2015}
-Comets have larger $e$, with some long period comets having $e\approx 1$, but the typical range is \qtyrange{0.2}{0.7}{} \citep{Lewis2012}.
-Interstellar objects (ISOs) have $e>1$ as they are on unbound hyperbolic trajectories.
-The inclination, $i$, of asteroids is also a key orbital element, it is the angle relative to the ecliptic plane and in principle asteroids can span the whole range. 
-Main belt asteroids have typical inclinations of \qtyrange{0}{30}{\degree} \citep{DeMeo2015}, so the TESS cameras closer to the ecliptic plane are expected to have the most asteroids in their fields.
-
-%TODO Abs mag H, and how that is size.
-
-%TODO asteroid families
-Asteroids can be grouped into families based on clustering of the proper orbital elements and similarities in their colour or spectra \citep{Nesvorny2015}. 
-These families are named for their largest member.
-
-%TODO YORP spin up
-
-Understanding the rotation properties of asteroids has long been of interest to astronomers \citep[e.g.][for early work into the limits of rotation period and the tumbling nature of some small bodies]{Weidenschilling1981,Harris1994}.
-This is still at the forefront of research, with more lightcurves being published year-on-year \citep{Harris2015}. 
-This will only increase with more large survey telescopes coming online soon, such as the Legacy Survey of Space and Time \citep[LSST][]{LSST2019}, the number of new asteroids that are well characterised will increase by more than 2 orders of magnitude on the Sloan Digital Sky Survey \citep[SDSS][]{York2000} which had $\sim 88000$ objects \citet{Parker2008}.
-LSST needs a dedicated pipeline for asteroid classification, the Solar System Notification Alert Processing System \citep[SNAPS][]{Trilling2023} has been developed using ZTF data and is can scale up handle volume of data coming from LSST.
-Work is being done to understand the biases of how  asymmetric shape high amplitude variations can lead to selection effects of such large sample of asteroids \citet{Levine2023}.
-For the smallest objects ingenious observation techniques are employed to measure rotation periods such as intentionally streaking the asteroid down a column of pixel on a detector, and making the lightcurve from the streak, as done by \citet{Bolin2023}.
-
-Because of such a rich history of study, many asteroids have a known rotation period. 
-These are catalogued by \citet{Warner2009} %TODO updated in citation
-in the Light Curve Database (LCDB).
-%TODO fig of period vs diameter.   
-In the October \nth{1} 2023 release, the LCDB has nearly 9000 objects with periods determined to a quality code $U\geq 2-$, which is the minimum reliability recommended by \citeauthor{Warner2009} for statistical analysis.
-This includes data from \citet{Pal2020}.
-Comparing the periods I find of asteroids that are already in the LCDB is key to making sure my methods are reliable, and that the periods for asteroids for which no data exists can be trusted.
-
-\subsubsection*{Interstellar Objects}
-%Omuamu
-High amplitude variation has come to the forefront of questions about asteroid properties because of the first interstellar object, 1I/\omuamua \citep[see][for a review]{Bannister2019}.
-\omuamua was determined to be spectroscopically red \citep{Fitzsimmons2017, Meech2017}, and having a photometric colour in the neutral end of the solar system range \citep{Bannister2017}.
-1I was classed as asteroid due to lack of a coma, and no noticeable activity. 
-This contrasts the second ISO, 2I/Borisov, which had many cometary characteristics \citep[see ][for a review]{Dorofeeva2023}. %\\TODO cite, as opposed to 2I??
-\omuamua was measured to have a rotation period of \qty{8.67(0.34)}{\hour} \citep{Belton2018} and seemed to be tumbling \citep[e.g.][]{Drahus2018,Fraser2018}.
-Combing the tumbling with an elongation ratio of up to \qty{6(1)}{}:1 \citep{McNeill2018}, 1I is said to have a cigar shape \citep{Belton2018}.
-The peak to peak amplitude variation of this ISO was \qty{2.5}{\mag} \citep{Meech2017} over its double peaked light curve.
-This is quite interesting, as it varies more than most asteroids.
-With the full sky survey of bright asteroids, we hope to find many asteroids with such a large amplitude variation, and to see just how rare \omuamua is.
-
-\section{Methods}\label{Sec:Meth}
-
-\subsection{Querrying Databases}\label{SubSec:Querry}
-
-To check for asteroids in the TESS data, the positions of the asteroids with time are required.
-For most asteroids, their orbital elements are well known, so it is a matter of looking them up and cross-matching with transients in the TESS data.
-Python was used to make API calls to {Skybot}\footnote{\href{https://vo.imcce.fr/webservices/skybot/}{Skybot: https://vo.imcce.fr/webservices/skybot/}} to get positions of asteroids in a cone search box in right ascension (RA) and  declination (Dec) space.
-As TESS sectors are \qty{27}{\day} long, querying every \qty{12}{\hour} is manageable.
-These positions are still very sparsely spaced in time compared to the TESS data, so an interpolation is needed to bridge the gap.
-
-Knowing where the asteroids are is helpful for finding them in the archival TESS data, but knowing more about the individual asteroids is also useful for population statistics.
-{JPL Horizons}\footnote{\href{https://ssd.jpl.nasa.gov/horizons/}{JPL Horizons: https://ssd.jpl.nasa.gov/horizons/}} was used to obtain the orbital elements of each asteroid, $a$, $e$, and $i$, as well as an absolute magnitude $H$, which is a good proxy for size (as discussed above).
-%TODO ref
-
-\subsection{Interpolation}\label{SubSec:Interp}
-
-With TESS data coming in $\frac12\unit{\hour}$ chunks, 24 interpolated points are needed between each API call.
-This interpolation should be accurate, as asteroids move at close to a TESS pixel per FFI on average \citep{Pal2018}, and checking against a higher frequency query to JPL Horizons confirmed this accuracy.
-For the shorter FFIs, the asteroids will be move fewer pixels per frame, which could cause slight mismatches between the interpolation and the detections.
-Horizons was also used get properties of the individual asteroids, such as the absolute magnitude $H$.
-For the faster TESS data, more interpolated points are needed, but a smaller the change in position between each point.
-These interpolated positions can be seen in \autoref{Fig:interpPos} for a cut from \texttt{TESSELLATE}.
-There are a few interesting features, such as the asteroids are moving in the same direction, indicated by the colouring, they come in from low RA and Dec and tend to increase both coordinates as the month of the sector progresses.
-There is also a large size range in this slice of sky, ranging over \qty{5}{\mag} in absolute magnitude, which can be seen in the alpha (or transparency) of the tracks in \autoref{Fig:interpPos}.
-
-\subsection{Detection Matching}\label{SubSec:Match}
-
-%Match to detections
-Matching these interpolated positions to \texttt{TESSELLATE} detections is important to lower the unknown transient outputs of this pipeline.
-Having interpolated their positions, the asteroids have a well sampled set of RA, Dec and time values of where they should be in the TESS data.
-They should show up in a pixel for a small number of frames, of order $\sim1$.
-This is the same as a lot of other transient events, a sharp rise in brightness and then disapering quickly again.
-The number of frames do change, type Ia supernova will brighten is a matter of a few hours and then dim for days, while stellar flares are of similar profile by a smaller max brightness and a corrsepondingly shorter decay time.
-Asteroids are very short, however detection pipelines are robust.
-These pipelines have already found the aforementioned supernova and stellar flares, the job of this work is to catch all the asteroids in the set of all the transients.
-To do this, catelogue matching is in order.
-
-Using the \texttt{KDTree} algorithm \citep{Maneewongvatana1999} as implemented in \texttt{SciPy} \citep{2020SciPy-NMeth}, the RA and Dec coordinates of the interpolated points and the detections can be compared and matched together.
-Filtering this \texttt{KDTree} output by restricting the time between spatially coincident matches to less than \qty{0.1}{\day} stops any accidental matches in position from non-asteroid detections.
-A cut is then made by smallest distance, in coordinate space, and any distance must be smaller than \qty{0.01}{\degree} (\qty{36}{\arcsec}), which is within 2 TESS pixels.
-Then, if multiple detection match to the same interpolated point, the detection with the smallest distance will be taken.
-
-\subsection{Ligthcurves}\label{SubSec:Lightcurves}
-
-%light curves; detected VS forced interpolated 
-There are two sets of points to take light curves from.
-The matches from the detections, which already have a flux calculated, and the interpolated points themselves, which are more numerous but require forcing the photometry.
-There were some challenges getting the flux of the interpolated points, even when \texttt{TESSELLATE} has already reduced all the FFIs of interest, due to the timing of the TESS frames, but these were identified and corrected for.
-Not every interpolated point gets a match, due to a variety of reasons, %pass in front of stars, dip below limiting mag, bad pipeline, not extreme enough of a difference between frames 
-so a comparison between the two light curves is interesting.
-\autoref{Fig:DifFlux} shows these two light curves for a chosen asteroid, ``Ruff''.
-This was picked as it has a high number of points matched to detections.
-There are more interpolated points than matched points, as not ever interpolated point gets matched to a detection.
-There seems to be a systematic offset between the fluxes, with the interpolated points having consistently lower flux, (here the means differ by 58 counts)
-Adjusting aperture to sum the fluxes over to the ``centre of mass'' of the asteroid in each frame does not alleviate this problem.
-
-%TODO data cleaning explaination
-Following \citet{McNeill2023}, any fluxes more than $3\sigma$ from the mean value are sigma-clipped from the lightcurves.
-
-\subsection{Period detection}\label{SubSec:Periods}
-
-%Periods and amplitueds for everything detected
-
-The next part of my analysis has to do with determining the periods and amplitudes of each asteroid's light curve.
-
-A widely used technique in determining the periods of astronomical data is the Lomb-Scargle periodogram \citep[\citet{Lomb1976,Scargle1982}, but see][for a review]{VanderPlas2018}.
-This method can efficiently take discrete timeseries data of which the observations are unevenly spaced and calculate the frequencies present in the sample.
-TESS should have rather evenly spaced data, as the FFIs are produced on a set candence, but sometimes data will be missing.
-For example, it may have been sigma clipped due to a spike in brightness that is unphysical for an asteroid, or TESS sees the same asteroid in the same sector on either side of the day long mid-sector break it takes for data transfer.
-The Lomb-Scargle periodogram deals with this uneven sampling much better than either a fast Fourier transform or a direct least squares fit.
-
-There are a few different implementations of a Lomb-Scargle periodogram available in world of astronomy python packages.
-I first tried using the \texttt{Lightkurve} \citep{Lightkurve2018} package built for period analysis of TESS (and Kepler) data of variable stars.
-\texttt{Lightkurve} was easy to use but didn't allow for fine-tuning of the periodogram, or give any model of the curve it produced.
-The next option was peeling back a layer of abstraction and using the Lomb-Scargle periodogram as implemented by \texttt{Astropy}\citep[\citet{Astropy2022} but see][for the implementation]{Vanderplas2012,Vanderplas2015}.
-There were interesting similarities and differences between the packages.
-\texttt{Lightkurve} is based on the \texttt{Astropy} methods, but does not give all the functionality, instead opting to simplify the process.
-\texttt{Astropy} gives a statistic on how confident it is in the detected frequency is, the false alarm probability.
-It also returns a model, in the form of
-
-\begin{equation}
-  \label{Eq:LCModel}
-  F(t;f,\vec{\theta}\,) = \theta_0 + \theta_1\cdot\sin{2\pi ft} +\theta_2\cdot\cos{2\pi ft}
-\end{equation}
-
-where the flux, $F$, is given in terms of time, $t$, the calculated frequency, f, and the model parameter $\vec{\theta} = [\theta_0, \theta_1,\theta_2]$.
-These parameters are useful in reconstructing the sinusoid without rerunning the periodogram.
-
-There will be differences in the period between the matched points and the interpolated points, not just because of the difference in average flux, but also from the larger number of points.
-This difference needs to be carefully thought through to understand what is the more likely period.
-
-\subsection{Asteroid Statistics}\label{SubSec:StatMeth}
-
-%Run on server
-The \texttt{TESSELLATE} pipeline has been running on the OzSTAR supercomputing facilities.
-After I am confident that all the parts of the asteroid detection and subsequent light curve analysis works as required, the same code can be refactored to work on OzSTAR and a large-scale analysis of all the processed TESS sectors can be run.
-Only after this has completed can the asteroid population statistics can be computed.
-I will be looking for completeness of detections of asteroids, as well as accuracy of periods and amplitude variation.
-
-\section{Results}\label{Sec:Res}
-\section{Discussion}\label{Sec:Disc}
-\section{Conclusion}\label{Sec:Conc}
-
-\section{Figures}
-%*Figures: Only here to keep them out of the way. Will be placed in text as it grows
-
-\begin{figure}
+\begin{figure}[ht]
 \centering
-  \includegraphics[width =\columnwidth]{../Test Code/Testing Figures/interpPos\_22\_2\_3\_4.pdf}
-  \caption[Interpolated positions of asteroids]{The interpolated positions of asteroids in one cut of a TESS sector.
-    The colours of the lines are time sequenced, as shown in the colour bar.
-    The alpha of the colours are scaled by the absolute magnitude $H$ of the asteroid, queried from JPL Horizons.
-    Both celestial (RA and Dec) coordinates and ecliptic (ecLon and ecLat) coordinates axes are shown.
-  }
-  \label{Fig:interpPos}
+\includegraphics[width=0.8\textwidth]{figures/test_figure.png}
+\caption{This is the caption.}
+\label{fig:plot}
 \end{figure}
 
-\begin{figure}
-  \centering
-  \includegraphics[width =\columnwidth]{../Test Code/Testing Figures/differentFluxes Ruff .pdf}
-  \caption[Light curves of Ruff]{The light curves of the asteroid ``Ruff'', with the flux of the interpolated positions in blue circles and the flux from the matches to detected sources in orange diamonds.
-    The Flux axis is measured in counts, as calculated by \texttt{TESSreduce}, and the times are in Modified Julian Date.}
-  \label{Fig:DifFlux}
-\end{figure}
+As a citation test we can refer to \citet{Albrow2022} or use brackets (when using cite-by-name style) \citep{Hurley2005}.
 
-\newpage %! Page number above here must be <=25
+Here is a table that we can refer to as Table~\ref{table:table1}.
+
+\vspace{1cm}
+
+\begin{table}[ht]
+    \caption{Table captions go above.}
+    \label{table:table1}
+    \begin{tabular}{|c|cc|cc|}
+    \toprule
+    $AR$ & \multicolumn{2}{c}{$Re=110$} & \multicolumn{2}{c}{$Re=1400$} \\
+    & 1\textsuperscript{st} Revolution & 3\textsuperscript{rd} Revolution & 1\textsuperscript{st} Revolution & 3\textsuperscript{rd} Revolution \\
+    \hline
+    3 & 2.44\% & 6.74\% & 3.05\% & 10.4\% \\
+    5 & 1.52\% & 6.65\% & 3.88\% & 11.0\% \\
+    7 & 1.18\% & 2.34\% & 5.80\% & 7.11\% \\
+    \bottomrule
+    \end{tabular}
+\end{table}
+
+\section{Conclusions}
 
 \section*{Acknowledgements}
-%OzSTAR asks for:
-\small
-This work was performed on the OzSTAR national facility at Swinburne University of Technology.
-The OzSTAR program receives funding in part from the Astronomy National Collaborative Research Infrastructure Strategy (NCRIS) allocation provided by the Australian Government, and from the Victorian Higher Education State Investment Fund (VHESIF) provided by the Victorian Government.
 
-%*Bib
-\bibliographystyle{jphysicsB}
-\bibliography{bibfile.bib}
+\newpage
+
+%
+%  References section
+%
+
+\bibliography{references}
+
 \end{document}

### 2024-09-17 Test Code/ToPutOnOzstar/asteroidCatalouge.py

@@ -9,6 +9,7 @@ from astropy.utils.data import download_file
 from astropy.io import fits
 from astroquery.jplhorizons import Horizons
 from tqdm import tqdm
+import shutil
 
 
 
@@ -45,12 +46,13 @@ def _Skybotquery(ra, dec, times, radius=10/60, location='C57', cache=False):
     url += '-rd={}&'.format(radius)
     url += '-loc={}&'.format(location)
     # TODO add more things here?
-
+    responses = []
     df = None
     times = np.atleast_1d(times)
     for time in tqdm(times, desc='Querying for SSOs'):
         url_queried = url + 'EPOCH={}'.format(time)
         response = download_file(url_queried, cache=cache)
+        responses.append(response)
         if open(response).read(10) == '# Flag: -1':  # error code detected?
             raise IOError("SkyBot Solar System query failed.\n"
                           "URL used:\n" + url_queried + "\n"
@@ -69,7 +71,7 @@ def _Skybotquery(ra, dec, times, radius=10/60, location='C57', cache=False):
     if df is not None:
         # // ! should have inplace=True...
         df.reset_index(drop=True, inplace=True)
-    return df
+    return df, responses
 
 
 def _setupQuery(dirPath="../OzData/"):
@@ -117,7 +119,7 @@ def _querySB(targetPos: list, qRad: float = 10.0, qLoc: str = "C57", numTimestep
     
     while len(result)<numTimesteps: #so if query timesout, it will restart with the cache
         try:
-            result = _Skybotquery(ra_i, dec_i, timeList.jd,
+            result ,responses = _Skybotquery(ra_i, dec_i, timeList.jd,
                           radius=qRad, location=qLoc, cache=True) #timeout throws error, so need to be in try/except to not close
         except:
             continue #restarts query with cache
@@ -142,7 +144,7 @@ def _querySB(targetPos: list, qRad: float = 10.0, qLoc: str = "C57", numTimestep
     brightResult["RA"] = coords.ra.deg
     brightResult["Dec"] = coords.dec.deg
 
-    return brightResult, timeList
+    return brightResult, timeList, responses
 
 
 def _get_properties_Horizons(asteroidsDf, time, loc:str="500@10")->pd.DataFrame:
[…]

### 2024-09-17 Test Code/ToPutOnOzstar/runAstrCat.py

@@ -1,5 +1,7 @@
 from asteroidCatalouge import make_asteroid_cat
 import numpy as np
+from itertools import product
+
 
 thisSector = 22
 
@@ -13,7 +15,24 @@ possibleCams = [1]
 possibleCCDs = [3]
 possibleCuts = [7]
 
-for i in  possibleCams:
-    for j in possibleCCDs:
-        for k in possibleCuts:
-            make_asteroid_cat(thisSector,i,j,k)
+# for i in  possibleCams:
+#     for j in possibleCCDs:
+#         for k in possibleCuts:
+#             print(thisSector,i,j,k)
+#             
+
+
+from multiprocessing import Pool
+# import time
+
+# def thisPrint(sector,cam,ccd,cut):
+#     time.sleep(0.5)
+#     print(sector,cam,ccd,cut)
+
+if __name__ == '__main__':
+    with Pool(12) as p:
+        p.starmap(make_asteroid_cat, product([thisSector],possibleCams,possibleCCDs,possibleCuts))
+        # p.starmap(thisPrint, product([thisSector],possibleCams,possibleCCDs,possibleCuts))
+
+
+

### 2024-09-17 Test Code/findAndInterp.py

@@ -16,6 +16,7 @@ from astroquery.jplhorizons import Horizons
 import itertools
 from tqdm import tqdm
 from scipy.spatial import KDTree
+import shutil
 
 
 plt.rcParams.update({
@@ -67,9 +68,11 @@ def _Skybotquery(ra, dec, times, radius=10/60, location='C57', cache=False):
 
     df = None
     times = np.atleast_1d(times)
+    responses =[]
     for time in tqdm(times, desc='Querying for SSOs'):
         url_queried = url + 'EPOCH={}'.format(time)
         response = download_file(url_queried, cache=cache)
+        responses.append(response)
         if open(response).read(10) == '# Flag: -1':  # error code detected?
             raise IOError("SkyBot Solar System query failed.\n"
                           "URL used:\n" + url_queried + "\n"
@@ -88,7 +91,7 @@ def _Skybotquery(ra, dec, times, radius=10/60, location='C57', cache=False):
     if df is not None:
         # // ! should have inplace=True...
         df.reset_index(drop=True, inplace=True)
-    return df
+    return df, responses
 
 
 def setupQuery(sector,cam,ccd,cut,dirPath="../OzData/"):
@@ -132,19 +135,22 @@ def querySB(targetPos: list, qRad: float = 10.0, qLoc: str = "C57", numTimesteps
     dt = (27/numTimesteps)*u.day  # a dt in days
     # list of times constructed sequentially
     timeList = t_i + dt*np.arange(0, numTimesteps)
-    result = pd.DataFrame()
+    result= pd.DataFrame()
     
-    while len(result)<numTimesteps: #so if query timesout, it will restart with the cache
+    while len(result)==0: #so if query timesout, it will restart with the cache
         try:
-            result = _Skybotquery(ra_i, dec_i, timeList.jd,
+            result, responses = _Skybotquery(ra_i, dec_i, timeList.jd,
                           radius=qRad, location=qLoc, cache=True) #timeout throws error, so need to be in try/except to not close
-        except:
+        except Exception as e:
+            print(e)
             continue #restarts query with cache
 
-        if type(result) != "pandas.core.frame.DataFrame": 
[…]

### 2024-09-17 Test Code/periodDetection.py

@@ -6,6 +6,8 @@ import lightkurve as lk
 import astropy.units as u
 from astropy.timeseries import LombScargle as lsp
 
+import nifty_ls
+
 %matplotlib widget 
 
 
@@ -48,18 +50,21 @@ def detect_period_ap(times, fluxes, plotting = False, knownFreq = None):
         nyquistP = 400*u.s   # 200 s ffi
     
 
+
+    centredFlux = fluxes - np.mean(fluxes)
+
     minFreq = 1/(2*(times.max()-times.min())*u.day).to(u.s)
     # minFreq = 1e-5/u.s
 
     maxFreq = 1/nyquistP
 
-    lsper = lsp((u.Quantity(times, u.day)).to(u.s), fluxes)
-    freqs, powers = lsper.autopower(minimum_frequency=minFreq, maximum_frequency=maxFreq)
+    lsper = lsp((u.Quantity(times, u.day)).to(u.s), centredFlux)
+    freqs, powers = lsper.autopower(minimum_frequency=minFreq, maximum_frequency=maxFreq, method="fastnifty")
     bestPow = np.max(powers)
     bestFreq = freqs[np.argmax(powers)]
 
     t_fit=(np.linspace(times.min(),times.max(),1000)*u.day).to(u.s)
-    y_fit = lsper.model(t_fit, bestFreq)
+    y_fit = lsper.model(t_fit, bestFreq) + np.mean(fluxes)
     modelParams = lsper.model_parameters(bestFreq)
     f_a_prob = lsper.false_alarm_probability(np.max(powers),samples_per_peak=50, minimum_frequency=minFreq, maximum_frequency=maxFreq)
 
@@ -208,26 +213,37 @@ numPerap = len(apPer.index)
 # print(badCountlk, numPerlk)
 print(badCountap, numPerap)
 
-fig, ax = plt.subplots()
-# ax.errorbar(lkPer.index, lkPer["Best Period"], fmt=".", c="tab:blue", capsize = 2, label="Lightkurve")
-
-#! Not error in period, but level of uncertanty due to f_a_Prob
-ax.errorbar(apPer.index, apPer["Best Period [Days]"], apPer["False Alarm Probability"]*10, fmt=".",c="tab:orange", capsize=2, label="All")
-
-bigPowLim = 0.2
-
[…]

### 2024-09-19 Test Code/ToPutOnOzstar/makeLightcurve.py
trying with new access token

@@ -1,0 +1,160 @@
+import numpy as np
+import pandas as pd
+from astropy.coordinates import SkyCoord
+from astropy import wcs
+from astropy.io import fits
+from astropy.stats import sigma_clip
+from scipy import ndimage
+from photutils.aperture import CircularAperture
+from photutils.aperture import aperture_photometry
+
+import warnings
+from astropy.utils.exceptions import AstropyWarning
+warnings.simplefilter('ignore', category=AstropyWarning)
+warnings.simplefilter('ignore', category=RuntimeWarning)
+
+
+def load_matches(sector, cam, ccd, cut, dirPath = ""):
+    matchesDF = pd.read_csv(f"{dirPath}MatchesToDetections.csv")
+    matchesDF.drop(columns="Unnamed: 0", inplace=True)
+    return matchesDF
+
+def load_interpolations(sector, cam, ccd, cut, dirPath = ""):
+    interpDF = pd.read_csv(f"{dirPath}asteroid_interpolated_positions.csv")
+    interpDF.drop(columns="Unnamed: 0", inplace=True)
+    return interpDF
+
+def load_fluxes(sector, cam, ccd, cut, dirPath = ""):
+    reduFlux = np.load(f"{dirPath}sector{sector}_cam{cam}_ccd{ccd}_cut{cut}_of16_ReducedFlux.npy")
+    return reduFlux
+
+def load_times(sector, cam, ccd, cut, dirPath = ""):
+    timesFromOz = np.load(f"{dirPath}sector{sector}_cam{cam}_ccd{ccd}_cut{cut}_of16_Times.npy")
+    return timesFromOz
+
+def name_cut(df, name:str, colName:str="Name"):
+    "take a df and gives  only the values where the colName == name"
+    toReturn = df.loc[df.index[df[colName]==name]]
+    toReturn.reset_index(drop=True, inplace=True)
+    return toReturn
+
+def sum_fluxes(frames, ys, xs, fluxBounds:list,reducedFluxes):
+    """
+    1.5 px aperture photometry around the COM coord
+    
+    NO LONGER Sums 3x3 box around all f,y,x, retuns list of fluxes in frame order
+    
+    """
+
+    checkUp = 4
+    checkDown = 3
+
+    comTotal = []
+
+    for f,y,x in zip(frames,ys,xs):   
+        
+
+        ymin = y-checkDown
+        ymax = y+checkUp
+        xmin = x-checkDown
[…]

### 2024-09-19 Test Code/findAndInterp.py

@@ -183,6 +183,7 @@ def get_properties_Horizons(asteroidsDf, time, loc:str="500@10")->pd.DataFrame:
             e = elements["e"].value[0]
             i = elements["incl"].value[0]
             H = elements["H"].value[0]
+            #? phaseAng = elements["alpha"].value[0]
 
         except Exception as ex:
             print(f"{name} failed horizons check because:\n {ex}")
@@ -190,6 +191,7 @@ def get_properties_Horizons(asteroidsDf, time, loc:str="500@10")->pd.DataFrame:
             e = np.NaN
             i = np.NaN
             H = np.NaN
+            # phaseAng = np.NaN
 
         #TODO lcdb querry
         #lcdbQ = querryfunc()...
@@ -285,7 +287,7 @@ def find_asteroids(sector,cam,ccd,cut):
 
         asteroidPropertiesDf = pd.DataFrame(propertiesList,columns=["Num","Name","Mv(mean)","Class", "Number of Points"])
 
-        withEles = get_properties_Horizons(asteroidPropertiesDf, timeList[0])
+        withEles = get_properties_Horizons(asteroidPropertiesDf, timeList[0],"500@-95")
 
         fname = f"asteroids_in_{sector}_{cam}_{ccd}_{cut}_properties"
         withEles.to_csv(f"{fname}.csv")

### 2024-09-19 Test Code/makeLightcurve.py

@@ -101,8 +101,14 @@ def sum_fluxes(frames, ys, xs):
         
 
         com = ndimage.center_of_mass(reducedFluxes[int(f),ymin:ymax,xmin:xmax])
-        comX = x-checkDown+com[1]
-        comY = y-checkDown+com[0]
+        if (com[0]>=0) &(com[0]<=7) &(com[1]>=0) & (com[1]<=7): #make sure COM in 7x7 box
+            comX =x-checkDown+com[1]
+            comY = y-checkDown+com[0]
+        else:
+            #reset defaults
+            comX =x
+            comY = y
+        
         
         try:
             comX = int(round(comX))
@@ -117,17 +123,6 @@ def sum_fluxes(frames, ys, xs):
         comTotal.append(comAppFlux)
         continue
 
-        ymin = comY-down
-        ymax = comY+up
-        xmin = comX-down
-        xmax = comX+up
-
-        if ymin < 0: ymin = 0
-        if ymax >= fluxBounds[1]: ymax = fluxBounds[1] -1
-        if xmin < 0: xmin = 0
-        if xmax >= fluxBounds[2]: xmax = fluxBounds[2]-1
-
-        comTotal.append(np.sum(reducedFluxes[int(f),ymin:ymax,xmin:xmax]))
         
 
     return total, comTotal
@@ -419,7 +414,7 @@ print(len(unqNames))
 
 extendedDfList = []
 
-for name in np.unique(interpDF["Name"]):
+for name in unqNames:
     lcRes = lightcurve_from_name(name)
     extendedDfList.append(lcRes)   
 
@@ -494,20 +489,33 @@ astrData.to_csv(f"./asteroids_in_{sector}_{cam}_{ccd}_{cut}_properties.csv")
 
 print(count)
 
-plot_lc_from_name("Bernoulli")
+testName = "1991 RE6"
+
+plot_lc_from_name(testName)
 
 
 #* sawtooth trials
-trialCut = name_cut(totalLcDf,"Bernoulli", colName="Name")
+trialCut = name_cut(totalLcDf,testName, colName="Name")
 
-timeCut = trialCut.loc[np.where((trialCut["MJD"]>=58900.58) & (trialCut["MJD"]<=58901.11))]
[…]

### 2024-09-19 Test Code/periodDetection.py

@@ -54,7 +54,10 @@ def detect_period_ap(times, fluxes, plotting = False, knownFreq = None):
     centredFlux = fluxes - np.mean(fluxes)
 
     minFreq = 1/(2*(times.max()-times.min())*u.day).to(u.s)
-    # minFreq = 1e-5/u.s
+    
+    minFreq = 1e-5/u.s
+
+    # minFreq = 1/(1468800*u.s)
 
     maxFreq = 1/nyquistP
 
@@ -113,7 +116,7 @@ def name_cut(df, name:str, colName:str="NameMatch"):
     toReturn.reset_index(drop=True, inplace=True)
     return toReturn
 
-numObser = 3 #*any less than 3 and NaNs come up, or it completely breaks
+numObser = 50 #*any less than 3 and NaNs come up, or it completely breaks
 #TODO, what should it be for a reliable P
 
 def compute_periods_lk(posDF):
@@ -218,26 +221,27 @@ print(badCountap, numPerap)
 
 
 #* period vs index
-# fig, ax = plt.subplots()
-# # ax.errorbar(lkPer.index, lkPer["Best Period"], fmt=".", c="tab:blue", capsize = 2, label="Lightkurve")
+fig, ax = plt.subplots()
+# ax.errorbar(lkPer.index, lkPer["Best Period"], fmt=".", c="tab:blue", capsize = 2, label="Lightkurve")
 
-# #! Not error in period, but level of uncertanty due to f_a_Prob
-# ax.errorbar(apPer.index, apPer["Best Period [Days]"], apPer["False Alarm Probability"]*10, fmt=".",c="tab:orange", capsize=2, label="All")
+#! Not error in period, but level of uncertanty due to f_a_Prob
 
-# bigPowLim = 0.2
+ax.errorbar(apPer.index.values, apPer["Best Period [Days]"].values, apPer["False Alarm Probability"].values*0, fmt=".",c="tab:orange", capsize=2, label="All")
 
-# bigPows = apPer.iloc[apPer.index[apPer["Max Power"] > bigPowLim]]
+bigPowLim = 0.2
+
+bigPows = apPer.iloc[apPer.index[apPer["Max Power"] > bigPowLim]]
 
 
-# #?WHY DO I HAVE TO DO THIS????
-# ids = bigPows.index.values
-# peri = bigPows["Best Period [Days]"].values
[…]

### 2024-09-24 FinalReport/FinalReportDraft.tex
should have done this ages ago

@@
 \usepackage[separate-uncertainty=true,multi-part-units=single]{siunitx}
+
+\usepackage{caption}
+\usepackage[list=true,listformat=simple]{subcaption}
+\makeatletter
+\g@addto@macro\p@subfigure{.}
+\makeatother
 
@@
 \DeclareSIUnit{\arcsec}{''}
-\DeclareSIUnit{\au}{\rm{AU}}
-\DeclareSIUnit{\mag}{\rm{mag}}
-\DeclareSIUnit{\px}{\rm{px}}
+\DeclareSIUnit{\au}{\mathrm{AU}}
+\DeclareSIUnit{\mag}{\mathrm{mag}}
+\DeclareSIUnit{\px}{\mathrm{px}}
 
@@
 
-My supervisor helped by proofreading the report and offering feedback. \\
+My supervisors helped by proofreading the report and offering feedback. As well as providing useful insight into problems when I was stuck.\\
 
@@
 The derivation in section 3 was taken from ... \\
+
+\autoref{Fig:FreqVsDiam} has been constructed from data from the Light Curve Database (LCDB) \citet{Warner2009}, as cited in text.
+That data has being built suplimented with my own work in the corrseponding Figure %TODO
 
@@
 This project aims to find and characterise the light curves of all the asteroids seen by the Transiting Exoplanet Survey Satellite (TESS).
-As such, a thorough discussion of both is required.
+Combining knowledge of these asteroid's positions with the high cadence imaging of TESS should allow for the determination of rotation periods of these small bodies.
 
@@
 These families are named for their largest member.
+Families are thought to all originate from the same body that was destroyed by a collision earlier in the solar system's history.
 
 %TODO YORP spin up
+Asteroids can have their rotation properties changed with time, collisions can do this, but so can light from the sun. 
+The YORP effect (named for the discoverers Yarkovsky, O'Keefe, Radzievskii, and Paddack) is the process of tourquing asteroids of an asymmetric shape by both bombardment of photons and thermal (re-)emission \citep{Rubincam2000}.
 
@@
 %TODO fig of period vs diameter.   
-In the October \nth{1} 2023 release, the LCDB has nearly 9000 objects with periods determined to a quality code $U\geq 2-$, which is the minimum reliability recommended by \citeauthor{Warner2009} for statistical analysis.
-This includes data from \citet{Pal2020}.
+In the October \nth{1}, 2023 release, the LCDB has nearly 9000 objects with periods determined to a quality code $U\geq 2-$, which is the minimum reliability recommended by \citeauthor{Warner2009} for statistical analysis.
+This includes other TESS data from \citet{Pal2020} and \citet{Woods2021}.
+The frequencies against diameter plot of these asteroids in \autoref{Fig:FreqVsDiam} shows the collection of all the asteroids in the LCDB.
+The spin barrier \citep[first proposed by][]{Pravec2000} is seen in the red dashed lines as there are very few asteroids larger than $\qty{0.15}{\kilo\metre}$ rotating faster than $\sim \qty{10}{\per\day}$.
+It is thought that the YORP effect spins up asteroids so fast that they pass above this barrier and then no longer have the internal strength necessary to hold themselves together. 
+This , is shown  in \autoref{Fig:FreqVsDiam}. 
 Comparing the periods I find of asteroids that are already in the LCDB is key to making sure my methods are reliable, and that the periods for asteroids for which no data exists can be trusted.
+
+\begin{figure}[t!]
+  \centering
+  \includegraphics[width=0.7\textwidth]{../Test Code/Testing Figures/Diam-FreqPlot.pdf}
+  \caption[Asteroid frequencies vs diameter (LCDB)]{The $U\geq 2-$ rotation frequencies in \unit{\per\day} and the inverse as a period in \unit{\hour} againts the diameter of the body in the October \nth{1}, 2023 LCDB \citet{Warner2009}.
+  This is using the diameter of the asteroids from the LCDB $H$ to diameter conversion. The spin barrier of \citet{Pravec2000} is shown in red dashed lines}
+  \label{Fig:FreqVsDiam}
+\end{figure}
 
@@
 
-% \begin{table}
-%   \caption[Tab1]{My table}
+\subsection{Querrying Databases}\label{SubSec:Querry}
 
-% \end{table}
+\begin{figure}[h!]
+  \begin{subfigure}[t]{0.5\textwidth}
+    \centering
+    \includegraphics[width =\textwidth]{../Test Code/Testing Figures/interpPos\_22\_2\_3\_4.pdf}
+    \caption[Positions]{Positions of asteroids in a cut.}
+    \label{Fig:interpPos}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[t]{0.5\textwidth}
+    \centering
+    \includegraphics[width =\textwidth]{../Test Code/Testing Figures/1989 BC1PosCheck.pdf} %TODO one with more points???
+    \caption[Errors]{Error in interpolated positions}
+    \label{Fig:errPos}
+  \end{subfigure}
 
-\subsection{Querrying Databases}\label{SubSec:Querry}
+  \caption[Interpolated positions and error of asteroids]{
+  (a)
+  The interpolated positions of asteroids in one cut of a TESS sector.
+  The colours of the lines are time sequenced, as shown in the colour bar.
+  The alpha of the colours are scaled by the absolute magnitude $H$ of the asteroid, queried from JPL Horizons.
+  Both celestial (RA and Dec) and ecliptic (ecLon and ecLat) coordinates axes are shown.
+  The green $+$ is the centre of the cut. 
+  This cut was chosen because of how many asteroids there were, enough to see the shape of the box, but not so much that the lines are overwhelming to the eye.\\
+  (b)
+  The difference in the interpolated position of an asteroid from bulk queries to JPL Horizons, with interpolated coordinate being subtracted from the coordinate returned by Horizons.
+The colour bar is a different colour map to (a) as the range of time is shorter.
+The extent of the box is one TESS pixel, \qty{21}{\arcsec}.
+This is of asteroid ``1989 BC1'', but all others checked are similarly accurate (It is very expensive to check them all by brute force queries of Horizons).
+  }
+  \label{Fig:interpandErrPos}
+\end{figure}
 
@@
 For most asteroids, their orbital elements are well known, so it is a matter of looking them up and cross-matching with transients in the TESS data.
-Python was used to make API calls to {Skybot}\footnote{\href{https://vo.imcce.fr/webservices/skybot/}{Skybot: https://vo.imcce.fr/webservices/skybot/}} \citep{Berthier2006} to get positions of asteroids in a cone search box in right ascension (RA) and  declination (Dec) space.
+Python was used to make API calls to
+%{Skybot}\footnote{\href{https://vo.imcce.fr/webservices/skybot/}{Skybot: https://vo.imcce.fr/webservices/skybot/}} 
+Skybot \citep{Berthier2006} to get positions of asteroids in a cone search box in right ascension (RA) and  declination (Dec) space.
 As TESS sectors are \qty{27}{\day} long, querying every \qty{12}{\hour} is manageable.
@@
 Knowing where the asteroids are is helpful for finding them in the archival TESS data, but knowing more about the individual asteroids is also useful for population statistics.
-The API of {JPL Horizons}\footnote{\href{https://ssd.jpl.nasa.gov/horizons/}{JPL Horizons: https://ssd.jpl.nasa.gov/horizons/}} was used to obtain the orbital elements of each asteroid, $a$, $e$, and $i$, as well as an absolute magnitude $H$, which is a good proxy for size (as discussed above).
-%TODO ref
+The \texttt{astroquery} \citep{Ginsburg2019} package was use to query JPL Horizons% {JPL Horizons}\footnote{\href{https://ssd.jpl.nasa.gov/horizons/}{JPL Horizons: https://ssd.jpl.nasa.gov/horizons/}}
+was used to obtain the orbital elements of each asteroid, $a$, $e$, and $i$, as well as an absolute magnitude $H$, which is a good proxy for size (as discussed above). %TODO ref
 
@@
 With TESS data coming in $\frac12\unit{\hour}$ chunks, 24 interpolated points are needed between each API call.
-Because of a small change in position between each frame, \citep[$\sim \qty{1}{\px}$ per frame][]{Pal2018}, this interpolation should be accurate.
-Checking against a higher frequency query to the JPL Horizons ephemeris  confirmed this accuracy on a few asteroids. %TODO fig
-This is not a feasable way of getting all of the positions, as the queries are per name, not per area like the conesearch in Skybot.
-They are also rate limited, so while a few asteroids can be checked, filling the gaps bewteen the Skybot half day queries for all the asteroids would take far too long.
-For the faster TESS FFIs, more interpolated points are needed, but a smaller the change in position between each point will be.
-Forcing the interpolations to be at the same time as the TESS frames, which are known thanks to \texttt{TESSELLATE} saving them out, simplifies the matching to detections and forced photometry of these points.
 These interpolated positions can be seen in \autoref{Fig:interpPos} for a cut from \texttt{TESSELLATE}.
@@
 There is also a large size range in this slice of sky, ranging over \qty{5}{\mag} in absolute magnitude, which can be seen in the alpha (or transparency) of the tracks in \autoref{Fig:interpPos}.
+Because of a small change in position between each frame, \citep[$\sim \qty{1}{\px}$ per frame][]{Pal2018}, this interpolation should be accurate.
+The 
+Checking against a higher frequency query to the JPL Horizons ephemeris  confirmed this accuracy on a few asteroids, as shown in \autoref{Fig:errPos}. 
+This is not a feasible way of getting all the positions, as the queries are per name and time, not per area and time like the cone search in Skybot.
+They are also rate limited, so while a few asteroids can be checked, filling the gaps between the Skybot half day queries for all the asteroids would take far too long.
+For the faster TESS FFIs, more interpolated points are needed, but a smaller the change in position between each point will be.
+Forcing the interpolations to be at the same time as the TESS frames, which are known thanks to \texttt{TESSELLATE} saving them out, simplifies the matching to detections and forced photometry of these points.
 
@@
 Then, if multiple detection match to the same interpolated point, the detection with the smallest distance will be taken.
+An example of the matches found using this \texttt{KDTree} is shown in \autoref{Fig:RADecMatch}.
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width =\textwidth]{../Test Code/Testing Figures/DetectMatchPos.pdf}
+  \caption[Interpolated and Detected Positions]{The RA (right plot) and Dec (left plot) of the interpolated points (blue) and the matched \texttt{TESSELLATE} sources(orange) against the time.
+The gap is from the TESS downtime in the sector, when it was passing close to earth.}
+  \label{Fig:RADecMatch}
+\end{figure}
 
@@
   \centering
-  \includegraphics[width =\columnwidth]{../Test Code/Testing Figures/interpPos\_22\_2\_3\_4.pdf}
-  \caption[Interpolated positions of asteroids]{The interpolated positions of asteroids in one cut of a TESS sector.
-    The colours of the lines are time sequenced, as shown in the colour bar.
-    The alpha of the colours are scaled by the absolute magnitude $H$ of the asteroid, queried from JPL Horizons.
-    Both celestial (RA and Dec) coordinates and ecliptic (ecLon and ecLat) coordinates axes are shown.
-  }
-  \label{Fig:interpPos}
-\end{figure}
-
-\begin{figure}
-  \centering
   \includegraphics[width =\columnwidth]{../Test Code/Testing Figures/differentFluxes Ruff .pdf}
@@
 
-\newpage %! Page number above here must be <=25
+\newpage %* Page number above here must be <=25
 
@@
 
-This work has used data and/or services provided by NASA's Solar System Dynamics website, https://ssd.jpl.nasa.gov/. Specifically JPL Horizons: https://ssd.jpl.nasa.gov/horizons/
+This work has used data and/or services provided by NASA's Solar System Dynamics website, \url{https://ssd.jpl.nasa.gov/}. Specifically JPL Horizons: \url{https://ssd.jpl.nasa.gov/horizons/}.
 

### 2024-09-24 Test Code/findAndInterp.py

@@ -287,13 +287,13 @@ def find_asteroids(sector,cam,ccd,cut):
 
         asteroidPropertiesDf = pd.DataFrame(propertiesList,columns=["Num","Name","Mv(mean)","Class", "Number of Points"])
 
-        withEles = get_properties_Horizons(asteroidPropertiesDf, timeList[0],"500@-95")
+        withEles = get_properties_Horizons(asteroidPropertiesDf, timeList[0],"500@10")
 
         fname = f"asteroids_in_{sector}_{cam}_{ccd}_{cut}_properties"
         withEles.to_csv(f"{fname}.csv")
 
 
-        return interpRes, responses
+        return interpRes, responses, withEles
 
 
 def name_cut(df, name:str, colName:str="NameMatch"):
@@ -372,15 +372,148 @@ def interplolation_of_pos(posDf, sector):
     return interpRes
 
 
-# One that breaks [22, 4, 1, 7]
+    
+def plotExpectedPos(posDf: pd.DataFrame, timeList: npt.ArrayLike, targetPos: list, magLim: float = 20.0, scaleAlpha: bool = False, minLen: int = 0, saving=False, hsList = None) -> plt.Figure:
+    """
+    Takes the output of a query to SkyBot, sets up a WCS, and plots the objects position, coloured by time, in Ra/Dec and ecLon/ecLat space. 
+
+    Inputs
+        posDf: DataFrame
+            The positions of each object with the associated time. Should have structure of output of querySB  
+        timeList: arrayLike of Times
+            astropy.time array like of the times of queries, also in 'epoch' collumn in posDf, but makes sure that the whole queried time range is acounted for, instead of maybe nothing was found at one time.
+        targetPos: list of len=3, [float, float, Time]
+            list of initial ra (float), dec (float) and time (astropy.Time value), the center point of the query and the first time of observations 
+
+        magLim: float
+            Limiting magnitude of the query, default 20.0 
+        scaleAlpha: bool
+            If alpha of the symbols should be scaled to the average brigtness of the object
+        minLen: int
[…]

### 2024-09-24 Test Code/makeLightcurve.py

@@ -17,6 +17,7 @@ from photutils.aperture import CircularAperture
 from photutils.aperture import aperture_photometry
 from astropy.stats import SigmaClip as astroSC
 from photutils.background import Background2D, MedianBackground
+from astroquery.jplhorizons import Horizons
 
 import warnings
 from astropy.utils.exceptions import AstropyWarning
@@ -203,7 +204,7 @@ def plot_lc_from_name(name):
     ax.scatter(nCut["MJD"], nCut["COM Flux"])
     # ax.set_ylim(-200,200)
     
-def plotExpectedPos(posDf: pd.DataFrame, timeList: npt.ArrayLike, targetPos: list, magLim: float = 20.0, scaleAlpha: bool = False, minLen: int = 0, saving=False) -> plt.Figure:
+def plotExpectedPos(posDf: pd.DataFrame, timeList: npt.ArrayLike, targetPos: list, magLim: float = 20.0, scaleAlpha: bool = False, minLen: int = 0, saving=False, hsList = None) -> plt.Figure:
     """
     Takes the output of a query to SkyBot, sets up a WCS, and plots the objects position, coloured by time, in Ra/Dec and ecLon/ecLat space. 
 
@@ -270,19 +271,19 @@ def plotExpectedPos(posDf: pd.DataFrame, timeList: npt.ArrayLike, targetPos: lis
         scaleMult = 0.90  # changes how small alpha can get
         brightest = posDf["Mv"].min()  # * Mag. scale is backwards...
         deltaMag = magLim-brightest
-
         
-        hs = {}
+        if hsList == None:
+            hs = {}
+            for name in unqNames:
+                try:
+                    horizQ =Horizons(id = name, epochs = t_i.jd, location= "500@10")
+                    hs[name]=float(horizQ.elements()['H'])
+                except Exception as e:
+                    print(f"Name= {name} didn't work, error {str(e)}")#some names aren't in jpl?
+                    badNames.append(name)
 
-        for name in unqNames:
-            try:
-                horizQ =Horizons(id = name, epochs = t_i.jd, location= "500@10")
-                hs[name]=float(horizQ.elements()['H'])
[…]

### 2024-09-24 Test Code/periodDetection.py

@@ -55,11 +55,11 @@ def detect_period_ap(times, fluxes, plotting = False, knownFreq = None):
 
     minFreq = 1/(2*(times.max()-times.min())*u.day).to(u.s)
     
-    minFreq = 1e-5/u.s
+    # minFreq = 1e-5/u.s
 
     # minFreq = 1/(1468800*u.s)
 
-    maxFreq = 1/nyquistP
+    maxFreq = 1/nyquistP 
 
     lsper = lsp((u.Quantity(times, u.day)).to(u.s), centredFlux)
     freqs, powers = lsper.autopower(minimum_frequency=minFreq, maximum_frequency=maxFreq, method="fastnifty")
@@ -294,7 +294,7 @@ compedPs = pd.DataFrame(inLCDBList, columns=["Name", "Known Period", "Found Peri
 print(compedPs)
 
 
-trialName = "Lincoln"
+trialName = "2000 JA66"
 
 try:
     knownFreq = 1/(compedPs.at[compedPs.index[compedPs["Name"]==trialName].values[0],"Known Period"]*60*60)

### 2024-09-25 FinalReport/FinalReportDraft.tex
some changes to ap with xxy as well as com

@@
 The frequencies against diameter plot of these asteroids in \autoref{Fig:FreqVsDiam} shows the collection of all the asteroids in the LCDB.
-The spin barrier \citep[first proposed by][]{Pravec2000} is seen in the red dashed lines as there are very few asteroids larger than $\qty{0.15}{\kilo\metre}$ rotating faster than $\sim \qty{10}{\per\day}$.
-It is thought that the YORP effect spins up asteroids so fast that they pass above this barrier and then no longer have the internal strength necessary to hold themselves together. 
-This , is shown  in \autoref{Fig:FreqVsDiam}. 
+The spin barrier \citep{Pravec2000} is seen in the red dashed lines as there are very few asteroids larger than $\qty{0.15}{\kilo\metre}$ rotating faster than $\sim \qty{10}{\per\day}$.
+It is thought that the YORP effect spins up asteroids so fast that they pass above this barrier and then no longer have the internal strength necessary to hold themselves together, impling that all asteroids made to spin faster have torn themselves apart and that they have no tensile strength, i.e they are rubble piles instead of monoliths \citep{Harris1996}.
+This idea of the cohesive strength of small bodies is still explored today \citep[e.g.][for V-type asteroids]{Oszkiewicz2020}.
 Comparing the periods I find of asteroids that are already in the LCDB is key to making sure my methods are reliable, and that the periods for asteroids for which no data exists can be trusted.
@@
 Breaking the average V mag of these objects into \qty{0.5}{\mag} bins for both the sample returned from a query and the sample detected, a ratio of completeness can be calculated per bin.
-(Note: If there is no asteroids in a bin, none have a chance of being recovered, so the all did get recovered. I.e. $\frac{0}{0}=1$)
+(Note: If there are no asteroids in a bin, none have a chance of being recovered, so the all did get recovered. I.e. $\frac{0}{0}=1$)
 

### 2024-09-25 Test Code/ToPutOnOzstar/makeAsteroidLightcurve.py

@@ -1,0 +1,191 @@
+import numpy as np
+import pandas as pd
+from astropy.coordinates import SkyCoord
+from astropy import wcs
+from astropy.io import fits
+from astropy.stats import sigma_clip
+from scipy import ndimage
+from photutils.aperture import CircularAperture
+from photutils.aperture import aperture_photometry
+
+import warnings
+from astropy.utils.exceptions import AstropyWarning
+warnings.simplefilter('ignore', category=AstropyWarning)
+warnings.simplefilter('ignore', category=RuntimeWarning)
+
+
+def load_matches(sector, cam, ccd, cut, dirPath = ""):
+    matchesDF = pd.read_csv(f"{dirPath}MatchesToDetections.csv")
+    matchesDF.drop(columns="Unnamed: 0", inplace=True)
+    return matchesDF
+
+def load_interpolations(sector, cam, ccd, cut, dirPath = ""):
+    interpDF = pd.read_csv(f"{dirPath}asteroid_interpolated_positions.csv")
+    interpDF.drop(columns="Unnamed: 0", inplace=True)
+    return interpDF
+
+def load_fluxes(sector, cam, ccd, cut, dirPath = ""):
+    reduFlux = np.load(f"{dirPath}sector{sector}_cam{cam}_ccd{ccd}_cut{cut}_of16_ReducedFlux.npy")
+    return reduFlux
+
+def load_times(sector, cam, ccd, cut, dirPath = ""):
+    timesFromOz = np.load(f"{dirPath}sector{sector}_cam{cam}_ccd{ccd}_cut{cut}_of16_Times.npy")
+    return timesFromOz
+
+def name_cut(df, name:str, colName:str="Name"):
+    "take a df and gives  only the values where the colName == name"
+    toReturn = df.loc[df.index[df[colName]==name]]
+    toReturn.reset_index(drop=True, inplace=True)
+    return toReturn
+
+def sum_fluxes(frames, ys, xs, fluxBounds:list,reducedFluxes):
+    """
+    1.5 px aperture photometry around the COM coord
+    
+    Sums 3x3 box around all f,y,x, retuns list of fluxes in frame order
+    
+    """
+
+    down= 1
+    up = 2
+
+    checkUp = 3
+    checkDown = 2
+    total = []
+    comTotal = []
+
+    comCoordsList = []
+
+    for f,y,x in zip(frames,ys,xs):   
+        
+
[…]

### 2024-09-25 Test Code/ToPutOnOzstar/makeLightcurve.py

@@ -1,160 +1,0 @@
-import numpy as np
-import pandas as pd
-from astropy.coordinates import SkyCoord
-from astropy import wcs
-from astropy.io import fits
-from astropy.stats import sigma_clip
-from scipy import ndimage
-from photutils.aperture import CircularAperture
-from photutils.aperture import aperture_photometry
-
-import warnings
-from astropy.utils.exceptions import AstropyWarning
-warnings.simplefilter('ignore', category=AstropyWarning)
-warnings.simplefilter('ignore', category=RuntimeWarning)
-
-
-def load_matches(sector, cam, ccd, cut, dirPath = ""):
-    matchesDF = pd.read_csv(f"{dirPath}MatchesToDetections.csv")
-    matchesDF.drop(columns="Unnamed: 0", inplace=True)
-    return matchesDF
-
-def load_interpolations(sector, cam, ccd, cut, dirPath = ""):
-    interpDF = pd.read_csv(f"{dirPath}asteroid_interpolated_positions.csv")
-    interpDF.drop(columns="Unnamed: 0", inplace=True)
-    return interpDF
-
-def load_fluxes(sector, cam, ccd, cut, dirPath = ""):
-    reduFlux = np.load(f"{dirPath}sector{sector}_cam{cam}_ccd{ccd}_cut{cut}_of16_ReducedFlux.npy")
-    return reduFlux
-
-def load_times(sector, cam, ccd, cut, dirPath = ""):
-    timesFromOz = np.load(f"{dirPath}sector{sector}_cam{cam}_ccd{ccd}_cut{cut}_of16_Times.npy")
-    return timesFromOz
-
-def name_cut(df, name:str, colName:str="Name"):
-    "take a df and gives  only the values where the colName == name"
-    toReturn = df.loc[df.index[df[colName]==name]]
-    toReturn.reset_index(drop=True, inplace=True)
-    return toReturn
-
-def sum_fluxes(frames, ys, xs, fluxBounds:list,reducedFluxes):
-    """
-    1.5 px aperture photometry around the COM coord
-    
-    NO LONGER Sums 3x3 box around all f,y,x, retuns list of fluxes in frame order
-    
-    """
-
-    checkUp = 4
-    checkDown = 3
-
-    comTotal = []
-
-    for f,y,x in zip(frames,ys,xs):   
-        
-
-        ymin = y-checkDown
-        ymax = y+checkUp
-        xmin = x-checkDown
[…]

### 2024-09-25 Test Code/ToPutOnOzstar/runAstrLC.py

@@ -1,0 +1,23 @@
+from makeAsteroidLightcurve import makeLCs
+import numpy as np
+from itertools import product
+from multiprocessing import Pool
+
+thisSector = 22
+
+#*For Full runs
+# possibleCams = [1,2,3,4]
+# possibleCCDs = [1,2,3,4]
+# possibleCuts = [1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16]
+
+#For testing run
+possibleCams = [1]
+possibleCCDs = [3]
+possibleCuts = [7]        
+
+
+if __name__ == '__main__':
+    with Pool(12) as p:
+        p.starmap(makeLCs, product([thisSector],possibleCams,possibleCCDs,possibleCuts))
+        # p.starmap(thisPrint, product([thisSector],possibleCams,possibleCCDs,possibleCuts))
+

### 2024-09-25 Test Code/findAndInterp.py

@@ -437,7 +437,7 @@ def plotExpectedPos(posDf: pd.DataFrame, timeList: npt.ArrayLike, targetPos: lis
     badNames = []  # stops index problem later
     if scaleAlpha:
         # setup for possible alpha scaling
-        scaleMult = 0.85  # changes how small alpha can get
+        scaleMult = 0.95  # changes how small alpha can get
         if str(type(hsList)) == "<class 'NoneType'>": #doesn't like asking for None when it is a array
             hs = {}
             for name in unqNames:
@@ -468,9 +468,9 @@ def plotExpectedPos(posDf: pd.DataFrame, timeList: npt.ArrayLike, targetPos: lis
         # scales each objects alpha
 
         if scaleAlpha:
-            avgMag = posDf.loc[nameIDs]["Mv"].mean()
+            thisH = hsList[i]
             # scales the alpha of plotting to the brightness of the object, to give some idea of what might be detected
-            alpha = (1-scaleMult*((avgMag-brightest)/deltaMag))
+            alpha = (1-scaleMult*((thisH-brightest)/deltaMag))
         else:
             alpha = 1
 

### 2024-09-25 Test Code/makeLightcurve.py

@@ -68,8 +68,8 @@ def sum_fluxes(frames, ys, xs):
 
     down = 1
     up = 2
-    checkUp = 4
-    checkDown = 3
+    checkUp = 3
+    checkDown = 2
     total = []
     comTotal = []
 
@@ -85,7 +85,11 @@ def sum_fluxes(frames, ys, xs):
         if ymax >= fluxBounds[1]: ymax = fluxBounds[1] -1
         if xmin < 0: xmin = 0
         if xmax >= fluxBounds[2]: xmax = fluxBounds[2]-1
-        total.append(np.sum(reducedFluxes[int(f),ymin:ymax,xmin:xmax]))
+        
+        photTable = aperture_photometry(reducedFluxes[int(f),:,:],CircularAperture((x,y), r=1.5)) 
+        apFlux = photTable["aperture_sum"][0]
+        total.append(apFlux)
+
 
         ymin = y-checkDown
         ymax = y+checkUp
@@ -98,13 +102,13 @@ def sum_fluxes(frames, ys, xs):
         if xmax >= fluxBounds[2]: xmax = fluxBounds[2]-1
 
 
-        #???? comTotal.append(np.sum(reducedFluxes[int(f),ymin:ymax,xmin:xmax]))
-        
+        checkBox = reducedFluxes[int(f),ymin:ymax,xmin:xmax]
+        checkBox = np.where(checkBox>=0,checkBox,0)
 
-        com = ndimage.center_of_mass(reducedFluxes[int(f),ymin:ymax,xmin:xmax])
-        if (com[0]>=0) &(com[0]<=7) &(com[1]>=0) & (com[1]<=7): #make sure COM in 7x7 box
-            comX =x-checkDown+com[1]
-            comY = y-checkDown+com[0]
+        com = ndimage.center_of_mass(checkBox)
+        if (com[0]>=0) &(com[0]<=(checkUp+checkDown)) &(com[1]>=0) & (com[1]<=(checkUp+checkDown)): #make sure COM in 5x5 box
+            comX =x-checkDown+com[1] 
+            comY = y-checkDown+com[0] 
         else:
             #reset defaults
             comX =x
@@ -120,8 +124,8 @@ def sum_fluxes(frames, ys, xs):
         
[…]

### 2024-09-25 Test Code/periodDetection.py

@@ -294,7 +294,8 @@ compedPs = pd.DataFrame(inLCDBList, columns=["Name", "Known Period", "Found Peri
 print(compedPs)
 
 
-trialName = "2000 JA66"
+trialName = "2004 FQ5"
+
 
 try:
     knownFreq = 1/(compedPs.at[compedPs.index[compedPs["Name"]==trialName].values[0],"Known Period"]*60*60)

### 2024-09-26 FinalReport/FinalReportDraft.tex
ios chat work 26/9

@@
 An example of the matches found using this \texttt{KDTree} is shown in \autoref{Fig:RADecMatch}.
+It is clear that \texttt{TESSELLATE} is not detecting every frame that an asteroid is in.
+Detctions in \texttt{TESSELLATE} are done on a per pixel basis, and asteroids only cross each pixel for a short number of frames.
+Not every interpolated point gets a match, due to a variety of reasons, which culminate in there being not an extreme enough of a difference between frames at that pixel location.
+Instead there are bunches of freqntly matched points, and then patches where there are few matches detected at all.
+
+%TODO fig for agregate of all asteroids in a sector, to go in results
 
@@
 The matches from the detections, which already have a flux calculated, and the interpolated points themselves, which are more numerous but require forcing the photometry.
+There were some challenges getting the flux of the interpolated points, even when \texttt{TESSELLATE} has already reduced all the FFIs of interest.
 
-There were some challenges getting the flux of the interpolated points, even when \texttt{TESSELLATE} has already reduced all the FFIs of interest.
-The flux is summed in  $3\times3$ box centred on the integer $xy$ position returned from changing the world coordinates in RA and Dec to pixel coordinates using \texttt{Astropy}'s \texttt{SkyCoord} subpackage.
-%?
-cite
+Using the \texttt{SkyCoord} module of \texttt{Astropy}, the world coordinates, in RA and Dec, can be transformed into pixel coordinates, $x$ and $y$, using a world coordinates system (WCS) from TESSELLATE.
+\texttt{Photutils} \citep{Bradley2024} was then used to do aperture photometry with a \qty{1.5}{\px} radius to account for the spread in the TESS PRF. %?cite PRF
 
@@
 This also caused a ``sawtoothing'' effect, where the lighturve would jump from a minimum to a maximum when the integer coordinate snaped from being a pixel out to back on the asteroid.
-This both of these issues can be fixed by calculation of the centre of mass (COM) flux, the average brightness in $xy$ space on a $8\times8$ cut of the image centred on the interpolated position.
-I didn't take the highest pixel value in this box because there could be a rouge star, this would drag the COM around, but not by as much as snapping to that highest point.
-A \texttt{Photutils} %TODO cite 
-aperture photometry call to this COM with a radis of \qty{1.5}{\px} was enough to get fluxes of the same size, but with a far more smooth transition from maximum to minimum, as expected from rotating asteroids.
+This both of these issues can be fixed by calculation of the centre of mass (COM) flux, the average brightness in $xy$ space on a $5\times5$ cut of the image centred on the interpolated position.
+I didn't take the highest pixel value in this box because there could be a rouge star in the field that it snaps to, instead of the asteroid.
+A \texttt{Photutils} aperture photometry call to this COM with a radius of \qty{1.5}{\px} was enough to get fluxes of the same order, but with a far more smooth transition from maximum to minimum, as expected from rotating asteroids.
 
-Not every interpolated point gets a match, due to a variety of reasons, which culminate in there being not an extreme enough of a difference between frames.
-A comparison between the two light curves is interesting.
-
+A comparison between the two light curves is interesting, as one might expect only the brightest parts of the lightcurve will be matched.
 \autoref{Fig:DifFlux} shows these two light curves for a chosen asteroid, ``Ruff''.
 This was picked as it has a high number of points matched to detections, but of course still more interpolated points total.
-There seems to be a systematic offset between the fluxes, with the interpolated points having consistently lower flux, (here the means differ by 58 counts)
-Adjusting aperture to sum the fluxes over to the ``centre of mass'' of the asteroid in each frame does not alleviate this problem.
+There seems to be a systematic offset between the fluxes, with the interpolated points having consistently lower flux, (here the means differ by 58 counts)  %!check
 
 %TODO data cleaning explaination
-Following \citet{McNeill2023}, any fluxes more than $3\sigma$ from the mean value are sigma-clipped from the lightcurves.
+Following \citet{McNeill2023}, any fluxes more than $3\sigma$ from the mean value are sigma-clipped from the forced photometry lightcurves.
+This does introduce gaps into an otherwise regularly sampled lightcurve, but the contamination from background stars the object is passing in front of is too great.
 
@@
 These parameters are useful in reconstructing the sinusoid without rerunning the periodogram.
+The method used for the periodogram fits is the new \texttt{nifty-ls} package \citep{Garrison2024} that has can be implemented on its own, or by interfacing with \texttt{Astropy}.
+I use the later implementation as the speed up from the nonuniform fast Fourier transform is helpful on the large dataset from \texttt{TESSELLATE}, but the astropy interface is how I had everything set up to run on just one cut.
 
+I an using a frequency window of for the periodogram between the the inverse of twice the time the asteroid spends in the sector and the Nyquist frequency (1/TESS cadance).
+
+%???
 There will be differences in the period between the matched points and the interpolated points, not just because of the difference in average flux, but also from the larger number of points.
@@
 The \texttt{TESSELLATE} pipeline has been running on the OzSTAR supercomputing facilities.
-After I am confident that all the parts of the asteroid detection and subsequent light curve analysis works as required, the same code can be refactored to work on OzSTAR and a large-scale analysis of all the processed TESS sectors can be run.
-Only after this has completed can the asteroid population statistics can be computed.
-I will be looking for completeness of detections of asteroids, as well as accuracy of periods and amplitude variation.
+After I was confident that all the parts of the asteroid detection and subsequent light curve analysis worked as required on one cut of one camera of one ccd of one TESS sector, the same code was be refactored slightly to work on OzSTAR and a large-scale analysis of all the processed TESS sectors can be run.
+
+I will be looking for completeness of detections of asteroids, as well as accuracy of periods and amplitude variation by comparison to the LCDB.
 
@@
 Breaking the average V mag of these objects into \qty{0.5}{\mag} bins for both the sample returned from a query and the sample detected, a ratio of completeness can be calculated per bin.
-(Note: If there are no asteroids in a bin, none have a chance of being recovered, so the all did get recovered. I.e. $\frac{0}{0}=1$)
+(Note: If there are no asteroids in a bin, none have a chance of being recovered, so the all did get recovered.
+I.e.
+$\frac{0}{0}=1$, this is only an issue at the brightest magnitude where the fewest asteroids exist.)
 
@@
 
+This research made use of Photutils, an Astropy package for detection and photometry of astronomical sources \citep{Bradley2024}.
+
 %*Bib

### 2024-09-26 Test Code/findAndInterp.py

@@ -1,3 +1,4 @@
+import os
 import numpy as np
 import matplotlib.pyplot as plt
 import numpy as np
@@ -29,8 +30,8 @@ plt.rcParams.update({
 })
 
 
-
 timeOffset = 2400000.5
+
 
 def _Skybotquery(ra, dec, times, radius=10/60, location='C57', cache=False):
     """Returns a list of asteroids/comets given a position and time.
@@ -68,7 +69,7 @@ def _Skybotquery(ra, dec, times, radius=10/60, location='C57', cache=False):
 
     df = None
     times = np.atleast_1d(times)
-    responses =[]
+    responses = []
     for time in tqdm(times, desc='Querying for SSOs'):
         url_queried = url + 'EPOCH={}'.format(time)
         response = download_file(url_queried, cache=cache)
@@ -94,7 +95,7 @@ def _Skybotquery(ra, dec, times, radius=10/60, location='C57', cache=False):
     return df, responses
 
 
-def setupQuery(sector,cam,ccd,cut,dirPath="../OzData/"):
+def setupQuery(sector, cam, ccd, cut, dirPath="../OzData/"):
 
     fname = f"{dirPath}{sector}_{cam}_{ccd}_{cut}_wcs.fits"
 
@@ -107,6 +108,7 @@ def setupQuery(sector,cam,ccd,cut,dirPath="../OzData/"):
     myTargetPos = [targetRa_i, targetDec_i, targetTime_i]
 
     return myTargetPos
+
 
 def querySB(targetPos: list, qRad: float = 10.0, qLoc: str = "C57", numTimesteps: int = 27, magLim: float = 20.0) -> tuple[pd.DataFrame, npt.ArrayLike]:
     """
@@ -135,23 +137,22 @@ def querySB(targetPos: list, qRad: float = 10.0, qLoc: str = "C57", numTimesteps
     dt = (27/numTimesteps)*u.day  # a dt in days
     # list of times constructed sequentially
     timeList = t_i + dt*np.arange(0, numTimesteps)
-    result= pd.DataFrame()
-    
-    while len(result)==0: #so if query timesout, it will restart with the cache
+    result = pd.DataFrame()
+
+    while len(result) == 0:  # so if query timesout, it will restart with the cache
         try:
             result, responses = _Skybotquery(ra_i, dec_i, timeList.jd,
-                          radius=qRad, location=qLoc, cache=True) #timeout throws error, so need to be in try/except to not close
[…]

### 2024-09-30 Test Code/ToPutOnOzstar/asteriodPlotting.py
plotting, colating and matching for ozstar

@@ -1,0 +1,402 @@
+"""
+Plotting functions for asteroid analysis
+
+Import each function as needed, keeps other code for being a mess with plotting everywhere, and can be much more optional.
+
+All have some shared optional inputs
+
+    saving: bool
+        whether or not to save the file as a pdf, default is false
+    
+    savePath: str
+        /path/to/image/  the will be  called fName.pdf. trailing slash is needed
+    
+    nameChange: str
+        String to add to the file name. The fName of each function already has a decription of the plot, but if this wants to be changed this can be done here. Default is '', a blank str 
+
+BL 30/9/24
+"""
+
+
+import numpy as np
+import numpy.typing as npt
+import pandas as pd
+
+import matplotlib.pyplot as plt
+import matplotlib.colors as mplc
+import matplotlib.cm as mpcm
+
+import itertools
+from scipy import ndimage
+
+from astropy import wcs
+from astropy.coordinates import SkyCoord
+import astropy.units as u
+from astropy.io import fits
+from astropy.time import Time
+
+from astroquery.jplhorizons import Horizons
+from photutils.aperture import CircularAperture
+
+plt.rcParams.update({
+    "font.size": 18,
+    "font.family": "serif",
+    "figure.autolayout": True,
+    "axes.grid": False,
+    # "xtick.minor.visible": True,
+    # "ytick.minor.visible": True,
+})
+
+
+
+def _name_cut(df, name:str, colName:str="Name"):
+    "take a df and gives  only the values where the colName == name"
+    toReturn = df.loc[df.index[df[colName]==name]]
+    toReturn.reset_index(drop=True, inplace=True)
+    return toReturn
+
+
+
+def plotAEI(propDF:pd.DataFrame, saving:bool=False, savePath:str="../Figures/", nameChange=""):
+    ps=2
+    col="tab:blue"
+    fig, ax = plt.subplots(1,2, sharex=True, figsize = (12,6))
+    ax[0].set(xlabel=f"a [Au]", ylabel="e")
+    ax[1].set(xlabel=f"a [Au]", ylabel=f"i [^{{\circ}}]")
+    ax[0].scatter(propDF["a"], propDF["e"], c=col,s=ps)
+    ax[0].scatter(propDF["a"], propDF["i"], c=col,s=ps)
+    fig.tight_layout()
+
[…]

### 2024-09-30 Test Code/ToPutOnOzstar/colateAsteroidCuts.py

@@ -1,0 +1,58 @@
+import pandas as pd
+import os
+# from glob import glob #! slow recursivly
+
+def _concatDFList(dfList):
+    """Takes list of dataframes, returns concated DF with index reset"""
+    allDFs = pd.concat(dfList)
+    allDFs.reset_index(inplace=True, drop=True)
+    allDFs.rename(columns={"Unnamed: 0":"ID in Cut"}, inplace=True)
+    return allDFs
+
+def colate_cuts(sector):
+    """Take the data from a per cut level to a per sector level. This is ok as all analysis wants to be on a sector level. Does make large files though"""
+    # allPropPaths = glob(f"/fred/oz335/TESSdata/Sector{sector}/**/asteroid_properties_cat.csv", recursive=True) #!TOO SLOW
+
+    #* Faster to just itterate and check
+    possibleCams = [1,2,3,4]
+    possibleCCDs = [1,2,3,4]
+    possibleCuts = [1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16]
+
+    basePathsList = []
+
+    for cam in possibleCams:
+        for ccd in possibleCCDs:
+            for cut in possibleCuts:
+                thisPath = f"/fred/oz335/TESSdata/Sector{sector}/Cam{cam}/Ccd{ccd}/cut{cut}of16/"
+                basePathsList.append(thisPath)
+
+    propDfs = []
+    lcDfs = []
+    matchDfs = []
+
+    for path in basePathsList:
+        propPath = f"{path}asteroid_properties_cat.csv"
+        if os.path.exists(propPath):
+            propDfs.append(pd.read_csv(propPath))
+        
+        lcPath = f"{path}asteroid_interpolated_positions.csv"
+        if os.path.exists(lcPath):
+            lcDfs.append(pd.read_csv(lcDfs))
+        
+        matchPath = f"{path}Asteroid_InterpAndDetect_Matches.csv"
+        if os.path.exists(matchPath):
+            matchDfs.append(pd.read_csv(matchPath))
+        
+    allProps = _concatDFList(propDfs)
+    allLCs = _concatDFList(lcDfs)
+    allMatches = _concatDFList(matchDfs)
+
+
+    #TODO clean up duplicate props requery will be somewhere else
+
+
+    allProps.to_csv(f"/fred/oz335/bleicester/Data/{sector}_All_Properties.csv")
[…]

### 2024-09-30 Test Code/ToPutOnOzstar/makeAsteroidLightcurve.py

@@ -14,11 +14,6 @@ warnings.simplefilter('ignore', category=AstropyWarning)
 warnings.simplefilter('ignore', category=RuntimeWarning)
 
 
-def load_matches(sector, cam, ccd, cut, dirPath = ""):
-    matchesDF = pd.read_csv(f"{dirPath}MatchesToDetections.csv")
-    matchesDF.drop(columns="Unnamed: 0", inplace=True)
-    return matchesDF
-
 def load_interpolations(sector, cam, ccd, cut, dirPath = ""):
     interpDF = pd.read_csv(f"{dirPath}asteroid_interpolated_positions.csv")
     interpDF.drop(columns="Unnamed: 0", inplace=True)

### 2024-09-30 Test Code/ToPutOnOzstar/matchInterpToDetect.py

@@ -1,0 +1,116 @@
+import numpy as np
+import pandas as pd
+from scipy.spatial import KDTree
+
+
+def _load_lcs(sector, cam, ccd, cut, dirPath = "./"):
+    lcDF = pd.read_csv(f"{dirPath}InterpLC.csv")
+    #!need to not reset index lcDF.drop(columns="Unnamed: 0", inplace=True)
+    return lcDF
+
+def _load_fluxes(sector, cam, ccd, cut, dirPath = "../OzData/"):
+    reduFlux = np.load(f"{dirPath}sector{sector}_cam{cam}_ccd{ccd}_cut{cut}_of16_ReducedFlux.npy")
+    return reduFlux
+
+def _load_times(sector, cam, ccd, cut, dirPath = "../OzData/"):
+    timesFromOz = np.load(f"{dirPath}sector{sector}_cam{cam}_ccd{ccd}_cut{cut}_of16_Times.npy")
+    return timesFromOz
+
+def _load_detection(dirPath = "../OzData/"):
+    detectedSourcesAll = pd.read_csv(f"{dirPath}detected_sources.csv", usecols=["ra", "dec", "mjd", "flux", "Type"])
+
+    indexsOfInterest = detectedSourcesAll.index[(detectedSourcesAll["Type"] == "0")]
+
+    indexsOfInterest = detectedSourcesAll.index[(detectedSourcesAll["Type"] == "0") | (detectedSourcesAll["Type"] == "Asteroid")] #Just the unknown ones
+
+    detectedSources = detectedSourcesAll.loc[indexsOfInterest]
+
+
+    detectedSources.rename(columns={"ra":"RA", "dec":"Dec", "mjd":"MJD"}, inplace=True)
+
+    detectedSources.reset_index(drop=True, inplace=True)
+
+    return detectedSources
+
+def _load_properties(dirPath = "../OzData/"):
+    propDF = pd.read_csv(f"{dirPath}asteroid_properties_cat.csv")
+    propDF.drop(columns="Unnamed: 0", inplace=True)
+    return propDF
+
+def _name_cut(df, name:str, colName:str="NameMatch"):
+    "take a df and gives  only the values where the colName == name"
+    toReturn = df.loc[df.index[df[colName]==name]]
+    toReturn.reset_index(drop=True, inplace=True)
+    return toReturn
+
+def _use_KD_tree(df1, df2, cols, maxDist:float=0.01,maxTimeSep:float=0.1,k:int=1):
+    tree = KDTree(df1[cols].values)
[…]

### 2024-09-30 Test Code/compareResults.py

@@ -156,7 +156,6 @@ unqNames = pd.unique(interpRes["Name"])
 
 obsForDetect=3 #number of observations needed to count a detection
 
-
 interpedMv = []
 foundMv = []
 

### 2024-09-30 Test Code/findAndInterp.py

@@ -526,15 +526,17 @@ def plotExpectedPos(posDf: pd.DataFrame, timeList: npt.ArrayLike, targetPos: lis
 
 
 sector = 22
-cam = 2
+cam = 1
 ccd = 3
-cut = 4
+cut = 7
 
 
 resDf, responses, astrProps = find_asteroids(sector, cam, ccd, cut)
 # interpDf = interplolation_of_pos(resDf,sector) Interp now done inside of the finding, so lower horizons queries
 fname = f"InterpolatedQuerryResult_{sector}_{cam}_{ccd}_{cut}"
 resDf.to_csv(f"{fname}.csv")
+
+
 
 
 # TO DELETE CACHE
@@ -602,17 +604,17 @@ for name in unqNames:
     deltaDec = (decsActs - decsInterp)*60*60
     fig, ax = plt.subplots(1, figsize=(10, 8))
     # ax.set_title(name)
-    ax.set_xlabel("$\Delta$ RA [$''$]")
+    ax.set_xlabel("$\Delta$RA [$''$]")
     ax.set_xlim((-limAngle, limAngle))
-    ax.set_ylabel("$\Delta$ Dec [$''$]")
+    ax.set_ylabel("$\Delta$Dec [$''$]")
     ax.set_ylim((-limAngle, limAngle))
     # TODO axis with 'TESS px units'
 
     ax2 = ax.secondary_xaxis("top", functions=(arcstopx, pxtoarcs))
     ax3 = ax.secondary_yaxis("right", functions=(arcstopx, pxtoarcs))
 
-    ax2.set_xlabel("$\Delta$ RA [TESS px]")
-    ax3.set_ylabel("$\Delta$ Dec [TESS px]")
+    ax2.set_xlabel("$\Delta$RA [TESS px]")
+    ax3.set_ylabel("$\Delta$Dec [TESS px]")
 
     cmap = "winter"
     # The delta T of the data, and makes a norm for a cmap

### 2024-10-01 Test Code/ToPutOnOzstar/colateAsteroidCuts.py
requery and colation on oz

@@ -1,5 +1,10 @@
+"""
+Takes all the files on a per cut basis, and gets them together in a per sector basis. 
+"""
+
 import pandas as pd
 import os
+import numpy as np
 # from glob import glob #! slow recursivly
 
 def _concatDFList(dfList):
@@ -8,6 +13,23 @@ def _concatDFList(dfList):
     allDFs.reset_index(inplace=True, drop=True)
     allDFs.rename(columns={"Unnamed: 0":"ID in Cut"}, inplace=True)
     return allDFs
+
+def _dropNans(multiProp):
+    for prop in multiProp:
+            try:
+                propInt = int(prop)
+                return prop
+            except:
+                continue
+    
+    return prop
+
+def _name_cut(df, name:str, colName:str="NameMatch"):
+    "take a df and gives  only the values where the colName == name"
+    toReturn = df.loc[df.index[df[colName]==name]]
+    toReturn.reset_index(drop=True, inplace=True)
+    return toReturn
+
 
 def colate_cuts(sector):
     """Take the data from a per cut level to a per sector level. This is ok as all analysis wants to be on a sector level. Does make large files though"""
@@ -23,7 +45,7 @@ def colate_cuts(sector):
     for cam in possibleCams:
         for ccd in possibleCCDs:
             for cut in possibleCuts:
-                thisPath = f"/fred/oz335/TESSdata/Sector{sector}/Cam{cam}/Ccd{ccd}/cut{cut}of16/"
+                thisPath = f"/fred/oz335/TESSdata/Sector{sector}/Cam{cam}/Ccd{ccd}/Cut{cut}of16/"
                 basePathsList.append(thisPath)
 
     propDfs = []
@@ -50,9 +72,43 @@ def colate_cuts(sector):
 
     #TODO clean up duplicate props requery will be somewhere else
 
+    unqNames = np.unique(allProps["Name"])
+
+    nameLists = []
+
+    for name in unqNames:
+        
+        nCut = _name_cut(allProps,name)
+        totPoints = np.sum(nCut["Number of Points"]) 
+        newMean = (np.sum(nCut["Mv(mean)"]*nCut["Number of Points"])/totPoints).round(3)
+        
+        colapsedEles = []
+
+        for eleTitle in ["a", "e", "i", "H"]:
+            ele = pd.unique(nCut[eleTitle])
+            if len(ele) != 1:
[…]

### 2024-10-01 Test Code/ToPutOnOzstar/requeryHorizons.py

@@ -1,0 +1,54 @@
+"""Lots of queries timed out or overwhelmed Horizons, so asking again slower
+
+Has to be run on Trevor in OzStar
+"""
+
+from astroquery.jplhorizons import Horizons
+import pandas as pd
+from astropy.io import fits
+from astropy import wcs
+from astropy.time import Time
+import numpy as np
+
+def _get_wcs_time(sector):
+    basePath = f"/fred/oz335/TESSdata/Sector{sector}/Cam1/Ccd2/cut3of16/"
+    wcsFname = f"{basePath}wcs.fits"
+    targetWSC = fits.open(wcsFname)[0]
+    # w = wcs.WCS(targetWSC.header)
+    targetTime_i = Time(targetWSC.header["DATE-OBS"])
+    return targetTime_i
+
+
+def requery(sector):
+    allPropDF = pd.read_csv(f"/fred/oz335/bleicester/Data/{sector}_All_Properties.csv")
+
+    qTime = _get_wcs_time(sector)
+
+    for j in range(len(allPropDF.index)):
+        if str(allPropDF.at[j,"a"] == str(np.nan)):
+            try:
+                horizQ = Horizons(id=allPropDF.at[j,"Name"], epochs=qTime.jd,
+                                location="500@10")
+                elements = horizQ.elements()
+                # subscriting to get number instead of list of 1 value????
+                a = elements["a"].value[0]
+                e = elements["e"].value[0]
+                i = elements["incl"].value[0]
+                H = elements["H"].value[0]
+
+                allPropDF.at[j,"a"] = a
+                allPropDF.at[j,"e"] = e
+                allPropDF.at[j,"i"] = i
+                allPropDF.at[j,"H"] = H
+
+            except Exception as ex:
+                print(f"Query failed because {ex}")
+        else:
+            continue
+
+    allPropDF.to_csv(f"/fred/oz335/bleicester/Data/{sector}_All_Properties.csv")
+
+
+requery(29)
+
+    

### 2024-10-02 Test Code/ToPutOnOzstar/colateAsteroidCuts.py
added figs from OZstar

@@ -92,10 +92,10 @@ def colate_cuts(sector):
                 ele = ele[0]
             colapsedEles.append(ele)
 
-        nameList = [nCut.at[0,"Num"],nCut.at[0,"Name"],newMean,nCut.at[0,"Class"],totPoints, colapsedEles[0], colapsedEles[1], colapsedEles[2], colapsedEles[3]]
+        nameList = [nCut.at[0,"Num"],nCut.at[0,"Name"],newMean,nCut.at[0,"Class"],totPoints,nCut.at[0,"Num Matches"], colapsedEles[0], colapsedEles[1], colapsedEles[2], colapsedEles[3]]
         nameLists.append(nameList)
 
-    unqNamesPropDF = pd.DataFrame(nameLists, columns=["Num","Name","Mv(mean)","Class","Number of Points","a","e","i","H"])
+    unqNamesPropDF = pd.DataFrame(nameLists, columns=["Num","Name","Mv(mean)","Class","Number of Points","Num Matches","a","e","i","H"])
     
 
     allProps.to_csv(f"/fred/oz335/bleicester/Data/{sector}_All_Properties.csv")

### 2024-10-02 FinalReport/FinalReportDraft.tex
added figures to the report

@@
 
-A comparison between the two light curves is interesting, as one might expect only the brightest parts of the lightcurve will be matched.
-\autoref{Fig:DifFlux} shows these two light curves for a chosen asteroid, ``Ruff''.
-This was picked as it has a high number of points matched to detections, but of course still more interpolated points total.
-There seems to be a systematic offset between the fluxes, with the interpolated points having consistently lower flux, (here the means differ by 58 counts)  %!check
-
 %TODO data cleaning explaination
@@
 This does introduce gaps into an otherwise regularly sampled lightcurve, but the contamination from background stars the object is passing in front of is too great.
+
+\begin{figure}
+  \centering
+  \includegraphics[width =\textwidth]{../OzData/SingleBodyLC1991 VT.pdf}
+  \caption[Light curves of 1991 VT]{The light curves of the asteroid 1991 VT throughout its entrie journey across sector 29, with the flux of the interpolated positions in blue circles and the flux from the COM in orange diamonds, and the \texttt{TESSELLATE} flux for any matched points in the black squares.
+    The Flux axis is measured in counts, as calculated by \texttt{TESSreduce}, and the times are in modified Julian date (MJD).}
+  \label{Fig:FullSingleLC}
+\end{figure}
+A comparison between the three light curves is interesting, as one might expect only the brightest parts of the lightcurve will be matched.
+\autoref{Fig:FullSingleLC} shows these three light curves for a single body.
+The COM flux seems much more stable than the interperted flux on its own, and at a higher average flux count.
+This should result in more accurate periodograms as there is less overall curve to the data.
+The matched points are comparatavely sparce for this object, which is not the case for all.
+Many have no matches, and some have almost fully matched, this will be discussed more later.
 
@@
 This method can efficiently take discrete timeseries data of which the observations are unevenly spaced and calculate the frequencies present in the sample.
-TESS should have rather evenly spaced data, as the FFIs are produced on a set candence, but sometimes data will be missing.
+TESS should have rather evenly spaced data, as the FFIs are produced on a set cadence, but sometimes data will be missing.
 For example, it may have been sigma clipped due to a spike in brightness that is unphysical for an asteroid, or TESS sees the same asteroid in the same sector on either side of the day long mid-sector break it takes for data transfer.
@@
 
+%TODO single object periodogram
+%Maybe a bad one too
+
 \subsection{Asteroid Statistics}\label{SubSec:StatMeth}
@@
 \section{Results}\label{Sec:Res}
-\section{Discussion}\label{Sec:Disc}
-\section{Conclusion}\label{Sec:Conc}
 
-\section{Figures}
-%*Figures: Only here to keep them out of the way. Will be placed in text as it grows
+This is the results for TESS sector 29.
+This sector was chosen because it is early in the \qty{10}{\min} cadence of TESS FFIs, and had been fully run through TESSELLATE.
+Similar results can be found in any of the sectors, just the computaion time was a limiting factor.
+
+\subsection{Asteroid Properties}\label{SubSec:AstPropRes}
+
+The orbital element properties of the asteroids were returned by JPL Horizons queries.
+A plot of the semi-major axis against the other orbital elements, eccentricity and inclination, is shown in \autoref{Fig:AEI}.
+The disticntion between the resonances in the main belt are clearly seen. There are lines seperating the middle belt from the inner and outer belt.
+The Jupiter trojans are well defined, at just past \qty{5}{\au}, with a small range of eccentricity but a large spread in the inclination.
+Inbetween the outerbelt and the Jupiter trojans, sit some Mars crossers.
+A some near earth objects (NEOs) at smaller $a$ and a few more objects that are scattered with high $e$ and or $i$ complete the distrabution.
 
@@
   \centering
-  \includegraphics[width =\columnwidth]{../Test Code/Testing Figures/differentFluxes Ruff .pdf}
-  \caption[Light curves of Ruff]{The light curves of the asteroid ``Ruff'', with the flux of the interpolated positions in blue circles and the flux from the matches to detected sources in orange diamonds.
-    The Flux axis is measured in counts, as calculated by \texttt{TESSreduce}, and the times are in Modified Julian Date.}
-  \label{Fig:DifFlux}
+  \includegraphics[width=\textwidth]{../OzData/AEIplot29.pdf}
+  \caption[aei distribution]{a, e, i, plot of all the objects on sector 29}
+  \label{Fig:AEI}
 \end{figure}
+
+The number the belong to each class of objects is highlighted in \autoref{Fig:NumPerClass}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{../OzData/classesBar29.pdf}
+  \caption[Class distribution]{A distrabution of the number of asteroids per class that \texttt{SkyBot} knows the object is. }
+  \label{Fig:NumPerClass}
+\end{figure}
+
+\subsection{Detectability}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{../OzData/recoverdHistBkgLimof10.pdf}
+  \caption[Recovery plot]{A histogram of known average visual magnitude of each object against the whether or not the object has a lightcurve with an average flux value above 10 %!CHECK
+    counts for its COM flux.
+  }
+  \label{Fig:RecovOverBkg}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{../OzData/pointsMatchesNumberHistTEST.pdf}
+  \caption[short]{A histogram of the number of point in the lightcurve of each asteroid (blue) against the number of matches each asteroid gets in the \texttt{TESSELLATE} detections(orange dashed). }
+  \label{Fig:MatchInterpHists}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width = \textwidth]{../OzData/matchesHist29.pdf}
+  \caption[Magnitude of matches]{The visual magnitude of the asteroid against the fraction of asteroids in this bin with more than 3 %!CHECK
+    matches.
+  }
+  \label{Fig:MagofMatches}
+\end{figure}
+
+\section{Discussion}\label{Sec:Disc}
+\section{Conclusion}\label{Sec:Conc}
 

### 2024-10-03 FinalReport/FinalReportDraft.tex
more slight changes to the start of the results section

@@
 This sector was chosen because it is early in the \qty{10}{\min} cadence of TESS FFIs, and had been fully run through TESSELLATE.
-Similar results can be found in any of the sectors, just the computaion time was a limiting factor.
+Similar results can be found in any of the sectors, the computation time was a limiting factor in getting more sectors processed.
 
@@
 A plot of the semi-major axis against the other orbital elements, eccentricity and inclination, is shown in \autoref{Fig:AEI}.
-The disticntion between the resonances in the main belt are clearly seen. There are lines seperating the middle belt from the inner and outer belt.
-The Jupiter trojans are well defined, at just past \qty{5}{\au}, with a small range of eccentricity but a large spread in the inclination.
-Inbetween the outerbelt and the Jupiter trojans, sit some Mars crossers.
-A some near earth objects (NEOs) at smaller $a$ and a few more objects that are scattered with high $e$ and or $i$ complete the distrabution.
+The distinction between the resonances in the main belt are clearly seen. There are lines separating the middle belt from the inner and outer belt.
+The Jupiter Trojans are well-defined, at just past \qty{5}{\au}, with a small range of eccentricity but a large spread in the inclination.
+In between the outer belt and the Jupiter Trojans, sit some Mars crossers.
+A some near earth objects (NEOs) at smaller $a$ and a few more objects that are scattered with high $e$ and or $i$ complete the distribution.
+These are all objects in the field of the instrument during the sector. 
+Many of them will not be detected, as they will not be bright enough, this is seen below.
 
@@
   \includegraphics[width=\textwidth]{../OzData/AEIplot29.pdf}
-  \caption[aei distribution]{a, e, i, plot of all the objects on sector 29}
+  \caption[aei distribution]{
+    The $a$ against both $e$ and $i$ of all the objects that \texttt{SkyBot} found in Sector 29 with $V <\qty{20}{\mag}$. 
+    (Other than the comets and centaurs with poorly defined $a$ values)
+    The values of these proper elements were queried from JPL Horizons on 30/09/24.
+  }
   \label{Fig:AEI}
@@
 
-The number the belong to each class of objects is highlighted in \autoref{Fig:NumPerClass}
+The number of asteroids belong to each class of objects is highlighted in \autoref{Fig:NumPerClass}.
+As expected from the orbital elements in \autoref{Fig:AEI}, the majority of the minor planets fall in the three regions of the main belt.
+\dots
 
@@
   \includegraphics[width=\textwidth]{../OzData/classesBar29.pdf}
-  \caption[Class distribution]{A distrabution of the number of asteroids per class that \texttt{SkyBot} knows the object is. }
+  \caption[Class distribution]{A distribution of the number of asteroids per class that \texttt{SkyBot} found in the Sector. 
+  As with \autoref{Fig:AEI}, this is for all objects brighter than \nth{20}\unit{\mag}
+  For the class labels, MB is ``main belt'' and ``NEA'' is near earth asteroid.}
   \label{Fig:NumPerClass}
@@
   \includegraphics[width=\textwidth]{../OzData/recoverdHistBkgLimof10.pdf}
-  \caption[Recovery plot]{A histogram of known average visual magnitude of each object against the whether or not the object has a lightcurve with an average flux value above 10 %!CHECK
+  \caption[Recovery plot]{A histogram of known average visual magnitude of each object against the whether the object has a lightcurve with an average flux value above 10 %!CHECK
   counts for its COM flux.
@@
   \label{Fig:RecovOverBkg}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{../OzData/pointsMatchesNumberHistTEST.pdf}
-  \caption[short]{A histogram of the number of point in the lightcurve of each asteroid (blue) against the number of matches each asteroid gets in the \texttt{TESSELLATE} detections(orange dashed). }
-  \label{Fig:MatchInterpHists}
 \end{figure}
@@
 \end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{../OzData/pointsMatchesNumberHistTEST.pdf}
+  \caption[Number of observations or matches]{A histogram of the number of observations in the lightcurve of each asteroid (blue) against the number of matches each asteroid gets in the \texttt{TESSELLATE} detections(orange dashed).
+}
+  \label{Fig:MatchInterpHists}
+\end{figure}
+
+\subsection{Period Analysis Results}\label{SubSec:PerRes}
 

### 2024-10-04 Test Code/ToPutOnOzstar/matchInterpToDetect.py
got extra points out of lcs, and stated on what period detection to put on OZ

@@ -3,10 +3,6 @@ import pandas as pd
 from scipy.spatial import KDTree
 
 
-def _load_lcs(sector, cam, ccd, cut, dirPath = "./"):
-    lcDF = pd.read_csv(f"{dirPath}InterpLC.csv")
-    #!need to not reset index lcDF.drop(columns="Unnamed: 0", inplace=True)
-    return lcDF
 
 def _load_fluxes(sector, cam, ccd, cut, dirPath = "../OzData/"):
     reduFlux = np.load(f"{dirPath}sector{sector}_cam{cam}_ccd{ccd}_cut{cut}_of16_ReducedFlux.npy")

### 2024-10-04 Test Code/ToPutOnOzstar/periodAnalysis.py

@@ -1,0 +1,121 @@
+import numpy as np
+import pandas as pd 
+import astropy.units as u
+from astropy.timeseries import LombScargle as lsp
+import matplotlib as plt
+import nifty_ls
+
+def _name_cut(df, name:str, colName:str="Name"):
+    "take a df and gives  only the values where the colName == name"
+    toReturn = df.loc[df.index[df[colName]==name]]
+    toReturn.reset_index(drop=True, inplace=True)
+    return toReturn
+
+def _load_lcs(sector, dirPath = "../Data/"):
+    lcDF = pd.read_csv(f"{dirPath}{sector}_All_LCs.csv")
+    lcDF.drop(columns="Unnamed: 0", inplace=True)
+    return lcDF
+
+def _load_properties(sector, dirPath = "../Data/"):
+    propDF = pd.read_csv(f"{dirPath}{sector}_All_Properties.csv")
+    propDF.drop(columns="Unnamed: 0", inplace=True)
+    return propDF
+
+
+def detect_period_ap(times, fluxes, sector, plotting = False, knownFreq = None):
+    """
+    Calculates the periodogram with astropy and returns the period at the max power.
+    """
+    #* Seems to be sec 27 and 56 when it changes
+    #12hr queries constant
+    if sector <27: 
+        nyquistP = 1*60*60*u.s #1/2hr ffi, so 1hr nyqist
+    elif sector>=27 and sector<56:
+        nyquistP = 20*60*u.s #1/6hr (10 min) ffi
+    else: 
+        nyquistP = 400*u.s   # 200 s ffi
+    
+
+
+    centredFlux = fluxes - np.mean(fluxes)
+
+    minFreq = 1/(2*(times.max()-times.min())*u.day).to(u.s)
+    
+    # minFreq = 1e-5/u.s
+
+    # minFreq = 1/(1468800*u.s)
+
+    maxFreq = 1/nyquistP 
+
+    lsper = lsp((u.Quantity(times, u.day)).to(u.s), centredFlux)
+    freqs, powers = lsper.autopower(minimum_frequency=minFreq, maximum_frequency=maxFreq, method="fastnifty")
+    bestPow = np.max(powers)
+    bestFreq = freqs[np.argmax(powers)]
+
+    t_fit=(np.linspace(times.min(),times.max(),1000)*u.day).to(u.s)
+    y_fit = lsper.model(t_fit, bestFreq) + np.mean(fluxes)
+    modelParams = lsper.model_parameters(bestFreq)
[…]

### 2024-10-04 Test Code/periodDetection.py

@@ -20,8 +20,8 @@ plt.rcParams.update({
     # "ytick.minor.visible": True,
 })
 
-#TODO Nyquist criterion
-#TODO LC database
+#// TODO Nyquist criterion
+#// TODO LC database
 
 def detect_period_lk(times, fluxes, fluxErrs):
     """

### 2024-10-05 FinalReport/FinalReportDraft.tex
some more

@@
 The method used for the periodogram fits is the new \texttt{nifty-ls} package \citep{Garrison2024} that has can be implemented on its own, or by interfacing with \texttt{Astropy}.
-I use the later implementation as the speed up from the nonuniform fast Fourier transform is helpful on the large dataset from \texttt{TESSELLATE}, but the astropy interface is how I had everything set up to run on just one cut.
+I use the later implementation as the speed-up from the non-uniform fast Fourier transform is helpful on the large dataset from \texttt{TESSELLATE}, but the \texttt{Astropy} interface is how I had everything set up to run on just the subsection I was trialling my code on.
 
-I an using a frequency window of for the periodogram between the the inverse of twice the time the asteroid spends in the sector and the Nyquist frequency (1/TESS cadance).
+I am using a frequency window of for the periodogram between inverse of twice the time the asteroid spends in the sector and the Nyquist frequency (1/TESS cadence).
 
@@
 
-The number of asteroids belong to each class of objects is highlighted in \autoref{Fig:NumPerClass}.
+The number of asteroids belonging to each class of objects is highlighted in \autoref{Fig:NumPerClass}.
 As expected from the orbital elements in \autoref{Fig:AEI}, the majority of the minor planets fall in the three regions of the main belt.
-\dots
+There are some objects that are not asteroids, 20 comets were in the field and 1 centaur.
+While they have a V mag less than 20, these objects are often not bright enough to have lightcurves with high enough mean fluxes to be detected well by TESS.
 
@@
   \includegraphics[width=\textwidth]{../OzData/classesBar29.pdf}
-  \caption[Class distribution]{A distribution of the number of asteroids per class that \texttt{SkyBot} found in the Sector. 
+  \caption[Class distribution]{A distribution of the number of objects per class that \texttt{SkyBot} found in the Sector. 
   As with \autoref{Fig:AEI}, this is for all objects brighter than \nth{20}\unit{\mag}

### 2024-10-05 Test Code/ToPutOnOzstar/periodAnalysis.py

@@ -119,3 +119,6 @@ def period_analysis(sector):
 
     propDF.to_csv(f"../Data/{sector}Properties_and_Periods.csv")
 
+
+
+period_analysis(29)

### 2024-10-08 FinalReport/FinalReportDraft.tex
poster started, and work on period limts investigated

@@
 in the Light Curve Database (LCDB).
-%TODO fig of period vs diameter.   
+%// TODO fig of period vs diameter.   
 In the October \nth{1}, 2023 release, the LCDB has nearly 9000 objects with periods determined to a quality code $U\geq 2-$, which is the minimum reliability recommended by \citeauthor{Warner2009} for statistical analysis.
@@
 
-\begin{figure}[h!]
-  \begin{subfigure}[t]{0.5\textwidth}
-    \centering
-    \includegraphics[width =\textwidth]{../Test Code/Testing Figures/interpPos\_22\_2\_3\_4.pdf}
-    \caption[Positions]{Positions of asteroids in a cut.}
-    \label{Fig:interpPos}
-  \end{subfigure}
-  ~
-  \begin{subfigure}[t]{0.5\textwidth}
-    \centering
-    \includegraphics[width =\textwidth]{../Test Code/Testing Figures/1989 BC1PosCheck.pdf} %TODO one with more points???
-    \caption[Errors]{Error in interpolated positions}
-    \label{Fig:errPos}
-  \end{subfigure}
-
-  \caption[Interpolated positions and error of asteroids]{
-    (a)
-    The interpolated positions of asteroids in one cut of a TESS sector.
-    The colours of the lines are time sequenced, as shown in the colour bar.
-    The alpha of the colours are scaled by the absolute magnitude $H$ of the asteroid, queried from JPL Horizons.
-    Both celestial (RA and Dec) and ecliptic (ecLon and ecLat) coordinates axes are shown.
-    The green $+$ is the centre of the cut.
-    This cut was chosen because of how many asteroids there were, enough to see the shape of the box, but not so much that the lines are overwhelming to the eye.\\
-    (b)
-    The difference in the interpolated position of an asteroid from bulk queries to JPL Horizons, with interpolated coordinate being subtracted from the coordinate returned by Horizons.
-The colour bar is a different colour map to (a) as the range of time is shorter.
-The extent of the box is one TESS pixel, \qty{21}{\arcsec}.
-This is of asteroid ``1989 BC1'', but all others checked are similarly accurate (It is very expensive to check them all by brute force queries of Horizons).
-  }
-  \label{Fig:interpandErrPos}
-\end{figure}
-
 To check for asteroids in the TESS data, the positions of the asteroids with time are required.
@@
 
+\begin{figure}[h!]
+  \begin{subfigure}[t]{0.5\textwidth}
+    \centering
+    \includegraphics[width =\textwidth]{../Test Code/Testing Figures/interpPos\_22\_2\_3\_4.pdf}
+    \caption[Positions]{Positions of asteroids in a subsection.}
+    \label{Fig:interpPos}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[t]{0.5\textwidth}
+    \centering
+    \includegraphics[width =\textwidth]{../Test Code/Testing Figures/1989 BC1PosCheck.pdf} %TODO one with more points???
+    \caption[Errors]{Error in interpolated positions}
+    \label{Fig:errPos}
+  \end{subfigure}
+
+  \caption[Interpolated positions and error of asteroids]{
+    (a)
+    The interpolated positions of asteroids in one subsection of a TESS sector.
+    The colours of the lines are time sequenced, as shown in the colour bar.
+    The alpha of the colours are scaled by the absolute magnitude $H$ of the asteroid, queried from JPL Horizons.
+    Both celestial (RA and Dec) and ecliptic (ecLon and ecLat) coordinates axes are shown.
+    The green $+$ is the centre of the subsection.
+    This sunsection was chosen because of how many asteroids there were, enough to see the shape of the box, but not so much that the lines are overwhelming to the eye.\\
+    (b)
+    The difference in the interpolated position of an asteroid from bulk queries to JPL Horizons, with interpolated coordinate being subtracted from the coordinate returned by Horizons.
+The colour bar is a different colour map to (a) as the range of time is shorter.
+The extent of the box is one TESS pixel, \qty{21}{\arcsec}.
+This is of asteroid ``1989 BC1'', but all others checked are similarly accurate (It is very expensive to check them all by brute force queries of Horizons).
+  }
+  \label{Fig:interpandErrPos}
+\end{figure}
+
 With TESS data coming in $\frac12\unit{\hour}$ chunks, 24 interpolated points are needed between each API call.
-These interpolated positions can be seen in \autoref{Fig:interpPos} for a cut from \texttt{TESSELLATE}.
+These interpolated positions can be seen in \autoref{Fig:interpPos} for a subsection of sector from \texttt{TESSELLATE}.
 There are a few interesting features, such as the asteroids are moving in the same direction, indicated by the colouring, they come in from low RA and Dec and tend to increase both coordinates as the month of the sector progresses.
@@
 This was decreasing the total flux of the asteroid, as the box used for sumation of the fluxes was hitting mostly sky.
-This also caused a ``sawtoothing'' effect, where the lighturve would jump from a minimum to a maximum when the integer coordinate snaped from being a pixel out to back on the asteroid.
-This both of these issues can be fixed by calculation of the centre of mass (COM) flux, the average brightness in $xy$ space on a $5\times5$ cut of the image centred on the interpolated position.
+This also caused a ``sawtoothing'' effect, where the lighturve would jump from a minimum to a maximum when the integer coordinate snapped from being a pixel out to back on the asteroid.
+Both of these issues can be fixed by calculation of the centre of mass (COM), the average brightness in $xy$ space on a $5\times5$ \unit{\px} box of the image centred on the interpolated position.
 I didn't take the highest pixel value in this box because there could be a rouge star in the field that it snaps to, instead of the asteroid.
-A \texttt{Photutils} aperture photometry call to this COM with a radius of \qty{1.5}{\px} was enough to get fluxes of the same order, but with a far more smooth transition from maximum to minimum, as expected from rotating asteroids.
+A \texttt{Photutils} aperture photometry centred on the COM with the same radius of \qty{1.5}{\px} sorted some of the discrepancies in the lightcurve, with a far smoother transition from maximum to minimum, as expected from rotating asteroids.
+Hereafter, unless otherwise spesified, ``flux'' referes to this COM flux measurement of the asteroid.
 
 %TODO data cleaning explaination
+A few steps need to be taken to clean these COM lightcurves. 
 Following \citet{McNeill2023}, any fluxes more than $3\sigma$ from the mean value are sigma-clipped from the forced photometry lightcurves.
-This does introduce gaps into an otherwise regularly sampled lightcurve, but the contamination from background stars the object is passing in front of is too great.
+This does introduce gaps into an otherwise regularly sampled lightcurve, but the contamination from background stars the object moves through of is too great.
+The subsections made by \texttt{TESSELLATE} have a small overlap. 
+If an object moves between these, there will be a few frames where it is found in multiple of these subsections.
+This makes the lightcurve have two simultaneous observations of the asteroid, which clearly is not possible.
+At the very edge of the subsections, the asteroid can have a COM position that would have the aperture be summing over pixels that are outside the field, and thus the asteroid would have a lower flux value on the edge.
+The overlap is great enough that the decrease in flux should only be happening on one of the edges at a time.
+So, sorting by flux on a per asteroid basis, then removing the duplicate observation times, keeping only those with the highest flux, will clean the lightcurve.
 
@@
   \includegraphics[width =\textwidth]{../OzData/SingleBodyLC1991 VT.pdf}
-  \caption[Light curves of 1991 VT]{The light curves of the asteroid 1991 VT throughout its entrie journey across sector 29, with the flux of the interpolated positions in blue circles and the flux from the COM in orange diamonds, and the \texttt{TESSELLATE} flux for any matched points in the black squares.
+  \caption[Light curves Example]{The light curves of the asteroid 1991 VT throughout its entrie journey across sector 29, with the flux of the interpolated positions in blue circles and the flux from the COM in orange diamonds, and the \texttt{TESSELLATE} flux for any matched points in the black squares.
     The Flux axis is measured in counts, as calculated by \texttt{TESSreduce}, and the times are in modified Julian date (MJD).}
@@
 
-I am using a frequency window of for the periodogram between inverse of twice the time the asteroid spends in the sector and the Nyquist frequency (1/TESS cadence).
+I am using a frequency window of for the periodogram between inverse of twice the time the asteroid spends in the sector and the Nyquist frequency (1/TESS cadence). %TODO rework completely
 
@@
 There will be differences in the period between the matched points and the interpolated points, not just because of the difference in average flux, but also from the larger number of points. 
-This difference needs to be carefully thought through to understand what is the more likely period.
+The matches are often have fewer observations and are much sparcer in time that the forced photometry measurements (see the black squares in \autoref{Fig:FullSingleLC}), so would give worse periodograms.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{../OzData/PeriodogramUlysses.pdf}
+  \caption[Periodogram example]{\textit{Upper Panel}:The Periodogram of the COM flux of an asteroid, with the power of each frequency displayed, both axes are in log scale. 
+  The maximum power of the periodogram is highlighted with the gold star. 
+  This maximum is the calculated rotation frequency
+  The known frequency, from the LCDB \citep{Warner2009}, is shown in the blue dotted lines, and a factor of 2 alias each way is shown in the purple dotted lines ($2*$) and red dotted lines ($0.5*$). 
+  \textit{Lower Panel}: The COM flux lightcurve and the returned model (dashed black curve) of the lightcurve from the periodogram analysis.
+The frequency of this model is the same as the gold stared point in the upper panel.
+  }
+  \label{Fig:PeriodEx}
+\end{figure}
+
+An example of such a periodogram and the corresponding fit to the lightcurve is given in \autoref{Fig:PeriodEx}. 
+The periodogram calculated that this object had a frequency double that listed in the LCDB, as seen with the vertical dotted lines. 
+This is expected due to the double peaked nature of most asteroid lightcurves \citep{McNeill2023}.  
+The stared peak is by far the most prominent peak in the periodogram, as it is in log scale.
+Not all the asteroids have such well-defined maximums. 
+Many do not possess a signal strong enough to outperform the noise in the periodogram at low frequencies, which means many of the objects have a period that is consistent with the lowest frequency in the window (conversely the longest period).
+This behaviour can be accounted for and the dataset unbiased by removing any objects with periods within \qty{90}{\percent} of this maximum. 
+The window of \qty{17}{\day} was chosen to allow for this cut, and to have a TESS sector sample lightcurve with this implicit maximum period all but twice in the \qty{27}{\day} sector, assuming the asteroid was in the sector for its duration.
+%TODO discuss fit of model...
 
@@
 The \texttt{TESSELLATE} pipeline has been running on the OzSTAR supercomputing facilities.
-After I was confident that all the parts of the asteroid detection and subsequent light curve analysis worked as required on one cut of one camera of one ccd of one TESS sector, the same code was be refactored slightly to work on OzSTAR and a large-scale analysis of all the processed TESS sectors can be run.
+After I was confident that all the parts of the asteroid detection and subsequent light curve analysis worked as required on one subsection of one camera of one CCD of one TESS sector, the same code was refactored slightly to work on OzSTAR and a large-scale analysis of all the processed TESS sectors can be run.
 

### 2024-10-12 FinalReport/FinalReportDraft.tex
More on poster
sone for sat at uni

@@
 It is clear that \texttt{TESSELLATE} is not detecting every frame that an asteroid is in.
-Detctions in \texttt{TESSELLATE} are done on a per pixel basis, and asteroids only cross each pixel for a short number of frames.
+Detections in \texttt{TESSELLATE} are done on a per pixel basis, and asteroids only cross each pixel for a short number of frames.
 Not every interpolated point gets a match, due to a variety of reasons, which culminate in there being not an extreme enough of a difference between frames at that pixel location.
-Instead there are bunches of freqntly matched points, and then patches where there are few matches detected at all.
+Instead, there are bunches of frequently matched points, and then patches where there are few matches detected at all.
 
@@
   \includegraphics[width =\textwidth]{../Test Code/Testing Figures/DetectMatchPos.pdf}
-  \caption[Interpolated and Detected Positions]{The RA (right plot) and Dec (left plot) of the interpolated points (blue) and the matched \texttt{TESSELLATE} sources(orange) against the time.
-The gap is from the TESS downtime in the sector, when it was passing close to earth.}
+  \caption[Interpolated and Detected Positions]{The RA (right plot) and Dec (left plot) against the time of observation of the interpolated points (blue dots) or the matches to \texttt{TESSELLATE} detections (black squares).
+The gap at low times is from the TESS downtime in the sector, when it was passing close to earth.}
   \label{Fig:RADecMatch}
@@
 The integer coordinates of the asteroids seemed to be leading the object in any given frame of TESS. %?FIG?
-This was decreasing the total flux of the asteroid, as the box used for sumation of the fluxes was hitting mostly sky.
-This also caused a ``sawtoothing'' effect, where the lighturve would jump from a minimum to a maximum when the integer coordinate snapped from being a pixel out to back on the asteroid.
+This was decreasing the total flux of the asteroid, as the box used for aperture  was hitting mostly background.
+This also caused a ``sawtoothing'' effect, where the lightcurve would jump from a minimum to a maximum when the integer coordinate snapped from being a pixel out to back on the asteroid.
 Both of these issues can be fixed by calculation of the centre of mass (COM), the average brightness in $xy$ space on a $5\times5$ \unit{\px} box of the image centred on the interpolated position.
@@
 A \texttt{Photutils} aperture photometry centred on the COM with the same radius of \qty{1.5}{\px} sorted some of the discrepancies in the lightcurve, with a far smoother transition from maximum to minimum, as expected from rotating asteroids.
-Hereafter, unless otherwise spesified, ``flux'' referes to this COM flux measurement of the asteroid.
+Hereafter, unless otherwise specified, ``flux'' refers to this COM flux measurement of the asteroid.
 
@@
 \autoref{Fig:FullSingleLC} shows these three light curves for a single body.
-The COM flux seems much more stable than the interperted flux on its own, and at a higher average flux count.
+The COM flux seems much more stable than the interpolated positions flux on its own, and at a higher average flux count.
 This should result in more accurate periodograms as there is less overall curve to the data.
-The matched points are comparatavely sparce for this object, which is not the case for all.
+The matched points are comparatively sparse for this object, which is not the case for all.
 Many have no matches, and some have almost fully matched, this will be discussed more later.
+Lightcurves with an average flux $>10$ counts is lost in the random noise floor, so a minimum mean flux limit needs to be applied to ensure quality and reliability of any periods calculated.
 
@@
 A widely used technique in determining the periods of astronomical data is the Lomb-Scargle periodogram \citep[\citet{Lomb1976,Scargle1982}, but see][for a review]{VanderPlas2018}.
-This method can efficiently take discrete timeseries data of which the observations are unevenly spaced and calculate the frequencies present in the sample.
+This method can efficiently take discrete time-series data of which the observations are unevenly spaced and calculate the frequencies present in the sample.
 TESS should have rather evenly spaced data, as the FFIs are produced on a set cadence, but sometimes data will be missing.
 For example, it may have been sigma clipped due to a spike in brightness that is unphysical for an asteroid, or TESS sees the same asteroid in the same sector on either side of the day long mid-sector break it takes for data transfer.
-The Lomb-Scargle periodogram deals with this uneven sampling much better than either a fast Fourier transform or a direct least squares fit.
+The periodogram reduces to both a fast Fourier transform or a direct least squares fit in appropriate limits, but outperforms them markedly in unevenly sampled data.
 
@@
 
-I am using a frequency window of for the periodogram between inverse of twice the time the asteroid spends in the sector and the Nyquist frequency (1/TESS cadence). %TODO rework completely
+The minimum period returned by the periodograms will be set by the Nyquist limit of the data. 
+For the \qty{10}{\min} FFIs (the cadence analysed here), the Nyquist period is $\qty{20}{\min}=\frac13\unit{\hour}$, a threefold increase in time sensitivity for fast rotating asteroids.  
+Based on their experimentation with the \qty{30}{\min} cadence FFIs, \citet{McNeill2023} use a \qty{17}{\day} maximum period for their periodogram analysis.
+While a lack of signal in the rotation curve causes the calculated period to find this low frequency limit, such cases can be filtered out by keeping only periods less than \qty{90}{\percent} of this maximum.
+They recommended that periods less than \qty{3}{\hour} be interpreted with caution, due to lack of reliability in recovering injected sources.
+Because of the increase in FFI rate, this time will be droped to \qty{1}{\hour} for the minimum reliable period, as the lightcurve sampling on these timescales is equivalent.
 
@@
 There will be differences in the period between the matched points and the interpolated points, not just because of the difference in average flux, but also from the larger number of points. 
-The matches are often have fewer observations and are much sparcer in time that the forced photometry measurements (see the black squares in \autoref{Fig:FullSingleLC}), so would give worse periodograms.
+The matches are often having fewer observations and are much sparser in time that the forced photometry measurements (see the black squares in \autoref{Fig:FullSingleLC}), so would give worse periodograms.
+While the rotation period and other properties of the lightcurve will be calculated from the forced photometry on the COM positions, the matching is still beneficial to the \texttt{TESSELLATE} pipeline as a whole.
+Taking previously unidentified detections and assigning them to an asteroid category, the remaining unknown transients can be more efficiently searched for other interesting astrophysical phenomena.
 
-\begin{figure}
+\begin{figure}[!h]
   \centering
-  \includegraphics[width=\textwidth]{../OzData/PeriodogramUlysses.pdf}
+  \includegraphics[width=\textwidth]{../OzData/PeriodogramUlyssesUpdated.pdf}
   \caption[Periodogram example]{\textit{Upper Panel}:The Periodogram of the COM flux of an asteroid, with the power of each frequency displayed, both axes are in log scale. 
@@
 Not all the asteroids have such well-defined maximums. 
-Many do not possess a signal strong enough to outperform the noise in the periodogram at low frequencies, which means many of the objects have a period that is consistent with the lowest frequency in the window (conversely the longest period).
-This behaviour can be accounted for and the dataset unbiased by removing any objects with periods within \qty{90}{\percent} of this maximum. 
-The window of \qty{17}{\day} was chosen to allow for this cut, and to have a TESS sector sample lightcurve with this implicit maximum period all but twice in the \qty{27}{\day} sector, assuming the asteroid was in the sector for its duration.
+The peak power of the periodogram is also an important factor in its' quality, a peak power less than 0.1 will be almost indisdinguashable from the noise, so these periodograms  
 %TODO discuss fit of model...
@@
 
+The quality of the periodogram is also assessed. 
+From \citet{McNeill2023}, the window of the periodogram is
+
 \section{Results}\label{Sec:Res}
@@
 The Jupiter Trojans are well-defined, at just past \qty{5}{\au}, with a small range of eccentricity but a large spread in the inclination.
-In between the outer belt and the Jupiter Trojans, sit some Mars crossers.
-A some near earth objects (NEOs) at smaller $a$ and a few more objects that are scattered with high $e$ and or $i$ complete the distribution.
+A some Mars crossers and near earth objects (NEOs) at progressesivly smaller $a$ and a few more objects that are scattered with high $e$ and or $i$ complete the distribution.
 These are all objects in the field of the instrument during the sector. 
@@
   \centering
-  \includegraphics[width=\textwidth]{../OzData/recoverdHistBkgLimof10.pdf}
+  \includegraphics[width=\textwidth]{../OzData/recoverdHistBkgLimof10Atanerr.pdf}
   \caption[Recovery plot]{A histogram of known average visual magnitude of each object against the whether the object has a lightcurve with an average flux value above 10 %!CHECK

### 2024-10-14 FinalReport/FinalReportDraft.tex
fixed some figs and some writting

@@
 
-%TODO Abs mag H, and how that is size.
+% // below TODO Abs mag H, and how that is size.
 
@@
 The frequencies against diameter plot of these asteroids in \autoref{Fig:FreqVsDiam} shows the collection of all the asteroids in the LCDB.
+The diameter presented in the LCDB here calculated from the absolute magnitude $H$ of the bodies, assuming an albedo.
+\citeauthor{Warner2009} acknowledge that there are direct measurements of diameter for more asteroids in the present day, but continue to present the $H$ derived diameters for consistency.
 The spin barrier \citep{Pravec2000} is seen in the red dashed lines as there are very few asteroids larger than $\qty{0.15}{\kilo\metre}$ rotating faster than $\sim \qty{10}{\per\day}$.
@@
   \caption[Asteroid frequencies vs diameter (LCDB)]{The $U\geq 2-$ rotation frequencies in \unit{\per\day} and the inverse as a period in \unit{\hour} againts the diameter of the body in the October \nth{1}, 2023 LCDB \citet{Warner2009}.
-    This is using the diameter of the asteroids from the LCDB $H$ to diameter conversion. The spin barrier of \citet{Pravec2000} is shown in red dashed lines}
+  The axes are in log scale. 
+  This is using the diameter of the asteroids from the LCDB $H$ to diameter conversion. 
+  The spin barrier of \citet{Pravec2000} is shown in red dashed lines.
+  }
   \label{Fig:FreqVsDiam}
@@
 Skybot \citep{Berthier2006} to get positions of asteroids in a cone search box in right ascension (RA) and  declination (Dec) space.
-As TESS sectors are \qty{27}{\day} long, querying every \qty{12}{\hour} is manageable.
+The search was limited to only asteroids with an average visual magnitude $V\geq \qty{20}{\mag}$.
+As TESS sectors are \qty{27}{\day} long, so querying every \qty{12}{\hour} is both computationally feasible and allows for a good understanding of a bodies motion. %TODO cite numpy
 These positions are still very sparsely spaced in time compared to the TESS data, so an interpolation is needed to bridge the gap.
@@
 Knowing where the asteroids are is helpful for finding them in the archival TESS data, but knowing more about the individual asteroids is also useful for population statistics.
-The \texttt{astroquery} \citep{Ginsburg2019} package was use to query JPL Horizons% {JPL Horizons}\footnote{\href{https://ssd.jpl.nasa.gov/horizons/}{JPL Horizons: https://ssd.jpl.nasa.gov/horizons/}}
+The \texttt{astroquery} \citep{Ginsburg2019} package was used to query JPL Horizons% {JPL Horizons}\footnote{\href{https://ssd.jpl.nasa.gov/horizons/}{JPL Horizons: https://ssd.jpl.nasa.gov/horizons/}}
 was used to obtain the orbital elements of each asteroid, $a$, $e$, and $i$, as well as an absolute magnitude $H$, which is a good proxy for size (as discussed above). %TODO ref
@@
 They should show up in a pixel for a small number of frames, of order $\sim1$.
-This is the same as a lot of other transient events, a sharp rise in brightness and then disapering quickly again.
-The number of frames do change, type Ia supernova will brighten is a matter of a few hours and then dim for days, while stellar flares are of similar profile by a smaller max brightness and a corrsepondingly shorter decay time.
+This is the same as a lot of other transient events, a sharp rise in brightness and then disappearing quickly again.
+The number of frames do change, type Ia supernova will brighten is a matter of a few hours and then dim for days, while stellar flares are of similar profile by a smaller max brightness and a correspondingly shorter decay time.
 Asteroids are very short spikes, however these detection pipelines are robust. %? Clarinda begs to differ
@@
 The job of this work is to catch all the asteroids in the set of all the transients that are returned from the pipeline.
-To do this, catelogue matching is in order.
+To do this, catalogue matching is in order.
 
@@
 Many have no matches, and some have almost fully matched, this will be discussed more later.
-Lightcurves with an average flux $>10$ counts is lost in the random noise floor, so a minimum mean flux limit needs to be applied to ensure quality and reliability of any periods calculated.
+Lightcurves with an average flux $<10$ counts are lost in the random noise floor, so a minimum mean flux limit needs to be applied to ensure quality and reliability of any periods calculated.
 
@@
 
-\begin{figure}[!h]
+\begin{figure}[!t]
   \centering
@@
 The stared peak is by far the most prominent peak in the periodogram, as it is in log scale.
-Not all the asteroids have such well-defined maximums. 
-The peak power of the periodogram is also an important factor in its' quality, a peak power less than 0.1 will be almost indisdinguashable from the noise, so these periodograms  
-%TODO discuss fit of model...
+Not all the asteroids have such well-defined maximums, and the peak power of the periodogram is an important factor in its quality.
+A peak power less than 0.1 means there is little to no signal observed and it will be almost indistinguishable from the noise.
+The lower panel of \autoref{Fig:PeriodEx} shows the model given in \autoref{Eq:LCModel} fit to the COM flux lightcurve by \texttt{Astropy}.
+This looks to be a good fit, with the peaks and troughs lining up.
+The amplitude of the sinusoid appears to be slightly lower than the variation of the data, but it is consistent with the scatter.
+This model is successfully fit over a break in the lightcurve, and ignores the observations with a lower flux that did not quite get sigma clipped.
+This asteroid was chosen as an alias of the known frequency is recovered by this analysis, which is promising for those asteroids where no known frequency exists. 
+The double alias is expected here due to the double peaked nature of the lightcurve \citep{McNeill2023}.   
+%// TODO discuss fit of model...
 
-%TODO single object periodogram
+%// TODO single object periodogram
 %Maybe a bad one too
@@
 After I was confident that all the parts of the asteroid detection and subsequent light curve analysis worked as required on one subsection of one camera of one CCD of one TESS sector, the same code was refactored slightly to work on OzSTAR and a large-scale analysis of all the processed TESS sectors can be run.
+I will be looking for completeness of detections of asteroids, as well as accuracy of periods and amplitude variation.
 
-I will be looking for completeness of detections of asteroids, as well as accuracy of periods and amplitude variation by comparison to the LCDB.
+For accurate statistics, accurate data is needed. 
+So there must be checks on the quality of the values calculated. 
+These have mostly been discussed already, but for completeness I list them here.
+A lightcurve must have an average flux of 10 counts to be reliable.
+For an accurate periodogram, it is required that the lightcurve also contain more than 200 observations, which is very easy for TESS and its 10 minute cadance.
+The periodogram itself must have a peak power of at least 0.1, and the calculated period must be $\qty{1}{\hour}\leq P \leq \qty{367.2}{\hour}$, where the upper bound is \qty{90}{\percent} of the \qty{17}{\day} window suggested by \cite{McNeill2023}.
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\textwidth]{../OzData/LCBDcompLogScaleLineText1.pdf}
+  \caption[Known against calculated period]{A comparison of the calculated periods of the objects in the Sector to their known periods in the LCDB. 
+  Dotted lines of a constant relationship are plotted, coloured the same as \autoref{Fig:PeriodEx}.
+  Both axes are in log scale.
+  }
 
-The completeness of the asteroid population in TESS as a function of their apparent magnitude can be determined.
-The apparent magnitude is one of the properties returned from the queries, and is in the V band.
-By taking a limit on the if an asteroid is detected \dots %TODO what is the limit
-Breaking the average V mag of these objects into \qty{0.5}{\mag} bins for both the sample returned from a query and the sample detected, a ratio of completeness can be calculated per bin.
-(Note: If there are no asteroids in a bin, none have a chance of being recovered, so the all did get recovered.
-I.e.
-$\frac{0}{0}=1$, this is only an issue at the brightest magnitude where the fewest asteroids exist.)
-
-The quality of the periodogram is also assessed. 
-From \citet{McNeill2023}, the window of the periodogram is
+  \label{Fig:LCDBcomp}
+\end{figure}
+While some of the checks were chosen to agree with the past work of others, some were chosen to make sure the derived periods were in agreement with the known ones from the LCDB.
+A comparison between objects in both the sector and the LCDB is seen in \autoref{Fig:LCDBcomp}. 
+The $y=2x$ line is of interest here, the known periods being twice the detected period. 
+This line corresponds to the calculated frequency being at the double alias of the known frequency, as frequency and period are inveses of each other. 
+As seen in \autoref{Fig:PeriodEx} and expected from \citet{McNeill2023}, this line is what the periodogram should return. 
+Most of the asteroids the pass the cut line along or near this key relation. 
+Few asteroids on the line fail the checks, which is also reasuringing that the checks do what they are suposed to and filter out incorrect periods.
+The vertical line of grey points on the right hand side of \autoref{Fig:LCDBcomp} is on that maximum of the window of searched periods at $\qty{408}{\hour}$, and demonstrates why the \qty{90}{\percent} upper bound is necessary.
+Many asteroids across an order of magnitude of know periods have detected periods at the edge of the window, and shows that the periodogram will return this value when no signal is present.
 
@@
 A plot of the semi-major axis against the other orbital elements, eccentricity and inclination, is shown in \autoref{Fig:AEI}.
-The distinction between the resonances in the main belt are clearly seen. There are lines separating the middle belt from the inner and outer belt.
+Looking at the data as a whole, all objects in the field of the instrument during the sector, with $V \leq \qty{20}{\mag}$, are plotted.
+The exception being comets and centaurs, which have an $a$ (really $q$, the perihelion distance) too larger to see the detail of the majority of the objects. 
+The distinction between the resonances in the main belt can be seen.
+There are vertical strips separating the middle belt from the inner and outer belt.
 The Jupiter Trojans are well-defined, at just past \qty{5}{\au}, with a small range of eccentricity but a large spread in the inclination.
-A some Mars crossers and near earth objects (NEOs) at progressesivly smaller $a$ and a few more objects that are scattered with high $e$ and or $i$ complete the distribution.
-These are all objects in the field of the instrument during the sector. 
-Many of them will not be detected, as they will not be bright enough, this is seen below.
+A some Mars crossers and near earth asteroids (NEAs) at progressively smaller $a$ and a few more objects that are scattered with high $e$ and/or $i$ complete the distribution.
+Many objects have failed the quality checks on their lightcurve and periodogram, but the ones that pass have a similar distribution in orbital parameters to the sample as a whole.
+In total, 309 objects passed all the quality checks, out of a total sample of 5664 small bodies in the sector.
 
@@
   \centering
-  \includegraphics[width=\textwidth]{../OzData/AEIplot29.pdf}
+  \includegraphics[width=\textwidth]{../OzData/AEIplotqualCut.pdf}
   \caption[aei distribution]{
-    The $a$ against both $e$ and $i$ of all the objects that \texttt{SkyBot} found in Sector 29 with $V <\qty{20}{\mag}$. 
-    (Other than the comets and centaurs with poorly defined $a$ values)
+    The $a$ against both $e$ and $i$ of all the objects that \texttt{SkyBot} found in Sector 29 with $V <\qty{20}{\mag}$(Other than the comets and centaurs with poorly defined $a$ values).
+    The grey crosses are those objects that did not pass the quality checks on the lightcurve and periodogram, while the green is those that did. 
     The values of these proper elements were queried from JPL Horizons on 30/09/24.
@@
 There are some objects that are not asteroids, 20 comets were in the field and 1 centaur.
-While they have a V mag less than 20, these objects are often not bright enough to have lightcurves with high enough mean fluxes to be detected well by TESS.
+While they have a V mag less than 20, these objects are often not bright enough to have lightcurves with high enough mean fluxes to be detected well by TESS, as seen by the green bars.
+Around \qty{5}{\percent} of any class have lightcurves and periodograms that pass the quality checks. 
+The highest fraction is the $\frac29$ of the Amor family of NEAs, however with only 9 total objects the sample is too small to be representative of the detectability of this class as a whole.
 
@@
   \centering
-  \includegraphics[width=\textwidth]{../OzData/classesBar29.pdf}
-  \caption[Class distribution]{A distribution of the number of objects per class that \texttt{SkyBot} found in the Sector. 
-  As with \autoref{Fig:AEI}, this is for all objects brighter than \nth{20}\unit{\mag}
+  \includegraphics[width=\textwidth]{../OzData/classesBarqualCut.pdf}
+  \caption[Class distribution]{A bar chart distribution of the total number of objects per class that \texttt{SkyBot} found in the Sector (blue) and the same distribution of those that passed all the quality checks (green).
+  The numbers above each class (coloured the same) correspond to the height of the bar, as some classes have too few objects to be seen.
+  As with \autoref{Fig:AEI}, this is for all objects brighter than \nth{20}\unit{\mag}.
   For the class labels, MB is ``main belt'' and ``NEA'' is near earth asteroid.}
@@
   \centering
-  \includegraphics[width=\textwidth]{../OzData/recoverdHistBkgLimof10Atanerr.pdf}
+  \includegraphics[width=\textwidth]{../OzData/recoverdHistBkgLimof10AtanBothLegendFixed.pdf}
   \caption[Recovery plot]{A histogram of known average visual magnitude of each object against the whether the object has a lightcurve with an average flux value above 10 %!CHECK
-  counts for its COM flux.
+  counts for its COM flux (blue triangles) or passes all of the checks on the lightcurve and periodogram (green dots).
+  A fit of an $\arctan$ function is provided in dotted lines of the same colour. 
   }
-  \label{Fig:RecovOverBkg}
+  \label{Fig:RecovPassorFail}
 \end{figure}
 
-\begin{figure}
+% \begin{figure}
+%   \centering
+%   \includegraphics[width = \textwidth]{../OzData/matchesHist29.pdf}
+%   \caption[Magnitude of matches]{The visual magnitude of the asteroid against the fraction of asteroids in this bin with more than 3 %!CHECK
+%     matches.
+%   }
+%   \label{Fig:MagofMatches}
+% \end{figure}
+
+\begin{figure} %TODO fix.
   \centering
-  \includegraphics[width = \textwidth]{../OzData/matchesHist29.pdf}
-  \caption[Magnitude of matches]{The visual magnitude of the asteroid against the fraction of asteroids in this bin with more than 3 %!CHECK
-    matches.
-  }
-  \label{Fig:MagofMatches}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{../OzData/pointsMatchesNumberHistTEST.pdf}
-  \caption[Number of observations or matches]{A histogram of the number of observations in the lightcurve of each asteroid (blue) against the number of matches each asteroid gets in the \texttt{TESSELLATE} detections(orange dashed).
-}
+  \includegraphics[width=\textwidth]{../OzData/pointsMatchesNumberHistdouble1.pdf}
+  \caption[Number of observations or matches]{\textit{Left}: A histogram of the number of observations in the lightcurve of each asteroid (orange solid line) in 100 observation bins.
+\textit{Right}: The number of matches each asteroid gets in the \texttt{TESSELLATE} detections (black dashed)in 100 match bins.
+  Both are limited to 600 asteroids on the y-axis, which pales in comparison to the number of asteroids with $\leq$ 100 matches and yet is too large for detail to be seen in any higher bins. }
   \label{Fig:MatchInterpHists}
@@
 
+%TODO amp vs period
+
+\begin{figure}
+  \centering
+  \includegraphics[width = \textwidth]{../OzData/QualPervsAmp1.pdf}
+  \caption[Amplitude of Variation against Rotaion Period]{The amplitude of variation in the model from the period analysis against the calculated rotation period.
+  This is for all 309 %!CHECK
+  objects that pass the quality checks.
+  }
+  \label{Fig:perVSamp}
+\end{figure}
+
 \section{Discussion}\label{Sec:Disc}
+
+\subsection{Comparison to other work }%\citet{McNeill2023}}
+
+%TODO redo fig 1??
+
+%! IF Matching works both to M23 and LCDB !!!
+
+\subsection{Future work}
+
+quality checks aren't great
+
+run on rest of TESSELLATE data
+
 \section{Conclusion}\label{Sec:Conc}
+
+It works ok
+
+but could be better
+
+summerise findings
 

### 2024-10-14 FinalReport/FinalReportDraft.tex
Send report draft to supervisors

@@
 These have mostly been discussed already, but for completeness I list them here.
-A lightcurve must have an average flux of 10 counts to be reliable.
-For an accurate periodogram, it is required that the lightcurve also contain more than 200 observations, which is very easy for TESS and its 10 minute cadance.
+A lightcurve must have an mean flux of at least 10 counts to be reliable.
+For an accurate periodogram, it is required that the lightcurve also contain more than 200 observations, which is very easy for TESS and its \qty{10}{\minute} cadence.
 The periodogram itself must have a peak power of at least 0.1, and the calculated period must be $\qty{1}{\hour}\leq P \leq \qty{367.2}{\hour}$, where the upper bound is \qty{90}{\percent} of the \qty{17}{\day} window suggested by \cite{McNeill2023}.
@@
 Most of the asteroids the pass the cut line along or near this key relation. 
-Few asteroids on the line fail the checks, which is also reasuringing that the checks do what they are suposed to and filter out incorrect periods.
-The vertical line of grey points on the right hand side of \autoref{Fig:LCDBcomp} is on that maximum of the window of searched periods at $\qty{408}{\hour}$, and demonstrates why the \qty{90}{\percent} upper bound is necessary.
-Many asteroids across an order of magnitude of know periods have detected periods at the edge of the window, and shows that the periodogram will return this value when no signal is present.
+Few asteroids on the line fail the checks, which is also reassuring that the checks do what they are supposed to and filter out incorrect periods.
+The vertical line of grey points on the right-hand side of \autoref{Fig:LCDBcomp} is on that maximum of the window of searched periods at $\qty{408}{\hour}$, and demonstrates why the \qty{90}{\percent} upper bound is necessary.
+Many asteroids, across an order of magnitude of know periods, have detected periods at the edge of the window which shows that the periodogram will return this value when no signal is present.  
+There are about a dozen asteroids that pass the checks with large detected periods when compared to their known period.
+These green points in the lower right of \autoref{Fig:LCDBcomp} speak to the checks not being perfect.
+Some experimentation was done on the limiting values, and the ones presented here achieved the highest recovery of asteroid periods, while minimising the number that were incorrectly recovered. 
+The idea of doubleing the detected period of all the asteroids and calling that number the rotation period has had some doubt cast on it, as there are a couple of points recovered nicely on the $y=x$ line, indicating that the true period is found.
+Either the LCDB has not accounted for the alias, which is unlikely as I only took the periods with high quality codes as they suggest \citep{Warner2009}, or the rotation signal sometimes returns the true period.
+The latter implies the lightcurve is not double peaked, as is normally expected.
 

### 2024-10-15 FinalReport/FinalReportDraft.tex
done for last tues

@@
 
-\subsection{TESS}\label{SubSec:TESS}
-
-%Why TESS
-TESS is a large area, high cadence imaging, space telescope  \citep{Ricker2014}.
-TESS is tasked with observing one piece of sky for \qty{27}{\day} at a time (a sector), delivering \qtyproduct{96 x 24}{\degree} full frame images (FFIs) at regular intervals.
-These FFIs are built from stacked \qty{2}{\second} exposures, leveraging the short readout times of the 16 CCD cameras on-board.
-With the initial cadence for these full frame images set to \qty{30}{\minute}, the time resolution of TESS is unparalleled.
-The highest frequency (shortest period) that can be found from time series data like this is at the Nyquist limit, twice the minimum time between images.
-Due to the short cadence, the Nyquist frequency of TESS is quite high, \qty{1}{\per\hour}.
-This is well sampled enough to characterise most variable stars, as well as find the orbital periods of exoplanets to a high precision.
-As the mission was extended, after TESS had already mapped the entire sky, the length of the FFIs has decreased to \qty{10}{\minute} and then later to \qty{200}{\second}.
-This decreased in the FFI exposure time comes with a complementary increase in the Nyquist frequency.
-This high time resolution and observation area does come at the cost of spatial resolution, as the pixels are each \qty{21}{\arcsec} square.
-Of interest here is what such a high sampling rate can do for statistic on the asteroid population.
-For bright asteroids the rotation periods should be able to be easily determined from this vast dataset.
-The shortest FFIs will be able to accurately determine the rotation periods of all but the fastest rotating asteroids, most of which will be too dim to see in the TESS data.
-
-There have been attempts to find and classify the asteroids in TESS data before.
-\citet{Pal2018} proposes that TESS will be a good instrument for solar system object study, and simulate some detections down to \nth{20} visual magnitude, but that good photometry should only be expected to $V \lesssim 19$.
-The sentiment of using TESS in this way is shared by \citet{Wong2019}.
-The first data release in \citet{Pal2020} catalogues nearly 10,000 asteroid light curves from the first year of TESS operation, and they report that they triple the number of asteroids with accurate rotation periods.
-\citet{McNeill2023} also analyse TESS cycle 1, and detect almost 38,000 objects.
-They determine reliable rotation periods for about 3,500 asteroids in this sample and show a lack of reliability for objects with periods less than \qty{3}{\hour}.
-In the overlap from the data from \citet{Pal2020}, \citeauthor{McNeill2023} find good agreement between the two sets of periods and amplitudes.
-The Minor Planet Center (MPC) gets regular updates from TESS thanks to the work of \citet{Woods2021} and their \texttt{LINEAR-TESS} program, which creates tracklets of objects over a day, and then stitches these together to form a track of each asteroid through an FFI.
-\citet{Gowanlock2024} uses TESS photometry as well as  observations by the Zwicky Transient Facility \citep[ZTF, ][]{Bellm2018} to get a longer baseline on mutually observed objects while combining ground and space based observations.
-There were not very many objects in both samples with known periods for comparison, with only 222 objects analysed, however they demonstrate that the technique is effective.
-Fainter and unknown solar system objects can be found by shift and stacking \citep{Holman2019, Payne2019,Rice2020} or taking a fast X-ray transform \citep{Nguyen2024} of the FFI data.
-
-This work aims to extend these studies to more sectors, hence characterising more asteroids.
-The aim is to get a volume complete set of asteroid lightcurves and periods, by analysis of the whole sky as seen by TESS using data from more than just the first year of TESS's operation.
-I also employ a different data reduction method, reducing the data with the \texttt{TESSreduce} package \citep{Ridden-Harper2021} should increase the accuracy of the photometry due to the improved background subtraction process.
-Because of the survey properties, TESS provides a self-consistent way to measure the properties of asteroids over the full sky.
-By using the higher cadence FFIs of later TESS sectors, the $P\leq\qty{3}{\hour}$ limit on the accurate periods \citep[as found by ][]{McNeill2023} should be lowered.
-Another beneficial part of this work is that as part of a full sky transient survey using TESS, \texttt{TESSELLATE} \citep{TESSELLATE}.
-Asteroids can spike the brightness of a pixel for only a few frames while it is passing by.
-This can confuse pipelines looking for  transients such as flare stars and supernova, as they can look very similar, having the brightness change in one pixel for a short time.
-The goal of finding all the asteroids will allow for the removal of these spikes from the transient pipeline, as well as to understand the asteroid population better.
-By filtering the asteroid detections out, while also self consistently determining rotation properties of a large sample of the population, more science can be done with the one action.
-
 \subsection{Asteroids}\label{SubSec:Asteroid}
@@
   \includegraphics[width=0.7\textwidth]{../Test Code/Testing Figures/Diam-FreqPlot.pdf}
-  \caption[Asteroid frequencies vs diameter (LCDB)]{The $U\geq 2-$ rotation frequencies in \unit{\per\day} and the inverse as a period in \unit{\hour} againts the diameter of the body in the October \nth{1}, 2023 LCDB \citet{Warner2009}.
-  The axes are in log scale. 
-  This is using the diameter of the asteroids from the LCDB $H$ to diameter conversion. 
-  The spin barrier of \citet{Pravec2000} is shown in red dashed lines.
+  \caption[Asteroid frequencies vs diameter (LCDB)]{
+  Rotaion rate of asteroids by their size, recreated from the  October \nth{1}, 2023 LCDB after \citet{Warner2009}.   
   }
@@
 With the full sky survey of bright asteroids, we hope to find many asteroids with such a large amplitude variation, and to see just how rare \omuamua is.
+
+\subsection{TESS}\label{SubSec:TESS}
+
+%Why TESS
+TESS is a large area, high cadence imaging, space telescope  \citep{Ricker2014}.
+TESS is tasked with observing one piece of sky for \qty{27}{\day} at a time (a sector), delivering \qtyproduct{96 x 24}{\degree} full frame images (FFIs) at regular intervals.
+These FFIs are built from stacked \qty{2}{\second} exposures, leveraging the short readout times of the 16 CCD cameras on-board.
+With the initial cadence for these full frame images set to \qty{30}{\minute}, the time resolution of TESS is unparalleled.
+The highest frequency (shortest period) that can be found from time series data like this is at the Nyquist limit, twice the minimum time between images.
+Due to the short cadence, the Nyquist frequency of TESS is quite high, \qty{1}{\per\hour}.
+This is well sampled enough to characterise most variable stars, as well as find the orbital periods of exoplanets to a high precision.
+As the mission was extended, after TESS had already mapped the entire sky, the length of the FFIs has decreased to \qty{10}{\minute} and then later to \qty{200}{\second}.
+This decreased in the FFI exposure time comes with a complementary increase in the Nyquist frequency.
+This high time resolution and observation area does come at the cost of spatial resolution, as the pixels are each \qty{21}{\arcsec} square.
+Of interest here is what such a high sampling rate can do for statistic on the asteroid population.
+For bright asteroids the rotation periods should be able to be easily determined from this vast dataset.
+The shortest FFIs will be able to accurately determine the rotation periods of all but the fastest rotating asteroids, most of which will be too dim to see in the TESS data.
+
+There have been attempts to find and classify the asteroids in TESS data before.
+\citet{Pal2018} proposes that TESS will be a good instrument for solar system object study, and simulate some detections down to \nth{20} visual magnitude, but that good photometry should only be expected to $V \lesssim 19$.
+The sentiment of using TESS in this way is shared by \citet{Wong2019}.
+The first data release in \citet{Pal2020} catalogues nearly 10,000 asteroid light curves from the first year of TESS operation, and they report that they triple the number of asteroids with accurate rotation periods.
+\citet{McNeill2023} also analyse TESS cycle 1, and detect almost 38,000 objects.
+They determine reliable rotation periods for about 3,500 asteroids in this sample and show a lack of reliability for objects with periods less than \qty{3}{\hour}.
+In the overlap from the data from \citet{Pal2020}, \citeauthor{McNeill2023} find good agreement between the two sets of periods and amplitudes.
+The Minor Planet Center (MPC) gets regular updates from TESS thanks to the work of \citet{Woods2021} and their \texttt{LINEAR-TESS} program, which creates tracklets of objects over a day, and then stitches these together to form a track of each asteroid through an FFI.
+\citet{Gowanlock2024} uses TESS photometry as well as  observations by the Zwicky Transient Facility \citep[ZTF, ][]{Bellm2018} to get a longer baseline on mutually observed objects while combining ground and space based observations.
+There were not very many objects in both samples with known periods for comparison, with only 222 objects analysed, however they demonstrate that the technique is effective.
+Fainter and unknown solar system objects can be found by shift and stacking \citep{Holman2019, Payne2019,Rice2020} or taking a fast X-ray transform \citep{Nguyen2024} of the FFI data.
+
+This work aims to extend these studies to more sectors, hence characterising more asteroids.
+The aim is to get a volume complete set of asteroid lightcurves and periods, by analysis of the whole sky as seen by TESS using data from more than just the first year of TESS's operation.
+I also employ a different data reduction method, reducing the data with the \texttt{TESSreduce} package \citep{Ridden-Harper2021} should increase the accuracy of the photometry due to the improved background subtraction process.
+Because of the survey properties, TESS provides a self-consistent way to measure the properties of asteroids over the full sky.
+By using the higher cadence FFIs of later TESS sectors, the $P\leq\qty{3}{\hour}$ limit on the accurate periods \citep[as found by ][]{McNeill2023} should be lowered.
+Another beneficial part of this work is that as part of a full sky transient survey using TESS, \texttt{TESSELLATE} \citep{TESSELLATE}.
+Asteroids can spike the brightness of a pixel for only a few frames while it is passing by.
+This can confuse pipelines looking for  transients such as flare stars and supernova, as they can look very similar, having the brightness change in one pixel for a short time.
+The goal of finding all the asteroids will allow for the removal of these spikes from the transient pipeline, as well as to understand the asteroid population better.
+By filtering the asteroid detections out, while also self consistently determining rotation properties of a large sample of the population, more science can be done with the one action.
 
@@
   \centering
-  \includegraphics[width =\textwidth]{../OzData/SingleBodyLC1991 VT.pdf}
-  \caption[Light curves Example]{The light curves of the asteroid 1991 VT throughout its entrie journey across sector 29, with the flux of the interpolated positions in blue circles and the flux from the COM in orange diamonds, and the \texttt{TESSELLATE} flux for any matched points in the black squares.
+  \includegraphics[width =\textwidth]{../OzData/SingleBodyLCUlyssesCorrectedAgain.pdf}
+  \caption[Light curves Example]{The light curves of the asteroid (5254) Ulysses throughout its entire journey across sector 29, with the flux of the interpolated positions in blue circles and the flux from the COM in orange diamonds, and the \texttt{TESSELLATE} flux for any matched points in the black squares.
     The Flux axis is measured in counts, as calculated by \texttt{TESSreduce}, and the times are in modified Julian date (MJD).}

### 2024-10-16 FinalReport/FinalReportDraft.tex
some homework

@@
 Because of a small change in position between each frame, \citep[$\sim \qty{1}{\px}$ per frame][]{Pal2018}, this interpolation should be accurate.
-The
+
 Checking against a higher frequency query to the JPL Horizons ephemeris  confirmed this accuracy on a few asteroids, as shown in \autoref{Fig:errPos}.
@@
 Both of these issues can be fixed by calculation of the centre of mass (COM), the average brightness in $xy$ space on a $5\times5$ \unit{\px} box of the image centred on the interpolated position.
-I didn't take the highest pixel value in this box because there could be a rouge star in the field that it snaps to, instead of the asteroid.
+I did not take the highest pixel value in this box because there could be a rouge star in the field that it snaps to, instead of the asteroid.
 A \texttt{Photutils} aperture photometry centred on the COM with the same radius of \qty{1.5}{\px} sorted some of the discrepancies in the lightcurve, with a far smoother transition from maximum to minimum, as expected from rotating asteroids.
@@
 I first tried using the \texttt{Lightkurve} \citep{Lightkurve2018} package built for period analysis of TESS (and Kepler) data of variable stars.
-\texttt{Lightkurve} was easy to use but didn't allow for fine-tuning of the periodogram, or give any model of the curve it produced.
+\texttt{Lightkurve} was easy to use but did not allow for fine-tuning of the periodogram, or give any model of the curve it produced.
 The next option was peeling back a layer of abstraction and using the Lomb-Scargle periodogram as implemented by \texttt{Astropy}\citep[\citet{Astropy2022} but see][for the implementation]{Vanderplas2012,Vanderplas2015}.
@@
 
+\begin{figure} %? // TODO fix.
+  \centering
+  \includegraphics[width=\textwidth]{../OzData/pointsMatchesNumberHistdouble1.pdf}
+  \caption[Number of observations or matches]{\textit{Left}: A histogram of the number of observations in the lightcurve of each asteroid (orange solid line) in 100 observation bins.
+\textit{Right}: The number of matches each asteroid gets in the \texttt{TESSELLATE} detections (black dashed).
+  }
+  \label{Fig:MatchInterpHists}
+\end{figure}
+
+The number of forced photometry observations in a lightcurve is more evenly distributed than the number of matches, as shown in \autoref{Fig:MatchInterpHists}.
+The quality cut of 200 observations in a lightcurve only cuts out $\sim 200$ asteroids.
+Most asteroids are well sampled, with the mode of the distribution in the 2900 to 3000 observation bin, which means the asteroids are in a large fraction of the about 3500 frames in a sector.
+
+The matches have a very different distribution to the observations.
+While both plots are limited to 600 asteroids on the y-axis, the matches histogram far exceeds this in the first bin, $\leq$ 100 matches, as indicated by the text and arrow.
+Yet the y-limit is too large for detail to be seen in any higher bin.
+There are only 102 asteroids matched more than 100 times in the \texttt{TESSELLATE} detections.
+
 \begin{figure}
@@
   \includegraphics[width=\textwidth]{../OzData/recoverdHistBkgLimof10AtanBothLegendFixed.pdf}
-  \caption[Recovery plot]{A histogram of known average visual magnitude of each object against the whether the object has a lightcurve with an average flux value above 10 %!CHECK
-  counts for its COM flux (blue triangles) or passes all of the checks on the lightcurve and periodogram (green dots).
+  \caption[Recovery plot]{The recovery fraction against the known visual magnitude of each object against the whether the
+    object exceeds only the averge flux requirement  of the quality checks (blue triangles) or passes all of the quality checks on the lightcurve and periodogram (green dots).
     A fit of an $\arctan$ function is provided in dotted lines of the same colour.
@@
 
-% \begin{figure}
-%   \centering
-%   \includegraphics[width = \textwidth]{../OzData/matchesHist29.pdf}
-%   \caption[Magnitude of matches]{The visual magnitude of the asteroid against the fraction of asteroids in this bin with more than 3 %!CHECK
-%     matches.
-%   }
-%   \label{Fig:MagofMatches}
-% \end{figure}
+The fits of to the fraction recoverd in \autoref{Fig:RecovPassorFail} take the form
+\begin{equation} \label{Eq:Atan}
+  y= 0.5 -\frac{1}{\pi} \arctan(\alpha \cdot (m-m_{1/2}))
+\end{equation}
+where $y$ is the recoved fraction, $\alpha$ is how the steep the transition from 1 down to 0 is,  $m$ the magnitude and $m_{1/2}$ is the shift along the magnitude axis.
+The constant of 0.5 in \autoref{Eq:Atan} raises the point of inflection to be at $\frac 1/2$, and the $\frac{-1}{\pi}$ makes the function take the expected maximum at bright magnitudes at \qty{100}{\percent} recoved, down to no dim asteroids being detected.
+The $m_{1/2}$ parameter dictates the limiting magnitude, as dimmer than this value less than half of the asteroids are recoverd.
+The fit is preformed on all bins (of \qty{0.5}{\mag}) dimmer than the first bin with more than one asteroid in it, to stop extremely bright asteroids from having too much influence.
 
-\begin{figure} %TODO fix.
-  \centering
-  \includegraphics[width=\textwidth]{../OzData/pointsMatchesNumberHistdouble1.pdf}
-  \caption[Number of observations or matches]{\textit{Left}: A histogram of the number of observations in the lightcurve of each asteroid (orange solid line) in 100 observation bins.
-\textit{Right}: The number of matches each asteroid gets in the \texttt{TESSELLATE} detections (black dashed)in 100 match bins.
-  Both are limited to 600 asteroids on the y-axis, which pales in comparison to the number of asteroids with $\leq$ 100 matches and yet is too large for detail to be seen in any higher bins. }
-  \label{Fig:MatchInterpHists}
-\end{figure}
+Only taking the minimum average flux of 10 counts requirement, blue in \autoref{Fig:RecovPassorFail}, the limiting magnitude is \qty{17.78(0.01)}{\mag}.
+When all the quality checks are imposed, this gets over a \unit{\mag} brighter, to \qty{16.46(0.37)}{\mag}.
+It is expected that a higher threshold for quality will make the limiting magnitude brighter, but is value is so uncertain because of the scatter in the recovery fraction at brighter magnitudes, and the shallower decrease in efficiency.
 
@@
 
+The limiting magnitude of the recovery fraction demonstrated in \autoref{Fig:RecovPassorFail} agrees well with other tests of \texttt{TESSELLATE}.
+Concurrent work by another Honours student \citep{MontillaHons} achieves a similar recovery fraction detection efficiency of $\sim$\nth{17} \unit{\mag}.
+She has been preforming injection/recovery tests across many sectors, and finding the limiting magnitude is consistent across sectors and cadence of the FFIs.
+This is promising, as it means that asteroids will be detected to a similar depth as they are here for any further sectors processed, something I could not check myself.
+
 \subsection{Future work}
 
-quality checks aren't great
+%quality checks aren't great
+It is clear from the outliers in \autoref{Fig:perVSamp} and the clearly incorrect periods of know objects in \autoref{Fig:LCDBcomp} that the quality checks have room for improvement.
+Some of these checks were  taken to agree with the how others analyse similar data, like the maximum period from \citet{McNeill2023}.
+Others were more empherically derived, such as the minimum average flux value.
+There is a delecate balance between detecting as many known periods as possible and maintating strict enough cuts to keep out the junk.
 
-run on rest of TESSELLATE data
+The periodogram noise at the low frequency end is quite strong throughout the range of rotation periods. 
+Currently it is responsible for a lot of undetected rotaion rates, due to the \qty{90}{\percent} of the \qty{17}{\day} window check, and this noise region having a larger power than any signal.
+Modelling this excess to try and remove the noise and boost the signal from the rotaion could be done to improve the maximum power of the periodogram. 
+There are a few asteroids with known longer rotaion periods than this check allows for, but one TESS sector will not be enough to determine those accurately.
+
+Linking lightcurves of asteroids between sectors was not considered here, as only one sector has been analysed.
+Any one asteroid will be in multiple sectors, as TESS has surveyed that whole sky more than once. 
+The different FFI cadence, as well as the large temporal gap between these observations makes it non-trivial to recover the rotation periods.
+The different depths that the shortest cadence FFIs have \citep{MontillaHons}
+
+%run on rest of TESSELLATE data
 

### 2024-10-17 FinalReport/FinalReportDraft.tex
report intro restucture

@@
 This project aims to find and characterise the light curves of all the asteroids seen by the Transiting Exoplanet Survey Satellite (TESS).
-Combining knowledge of these asteroid's positions with the high cadence imaging of TESS should allow for the determination of rotation periods of these small bodies.
+Combining knowledge of these asteroids positions with the high cadence imaging of TESS should allow for the determination of rotation periods of these small bodies, and the amplitude of the variation.
 
@@
 
-Asteroids are a key class of solar system objects.
-The orbits of asteroids can be well characterised by only a few values. The semi-major axis, $a$, is the largest distance from the sun that the asteroid achieves on its elliptical orbit.
-Main belt asteroids have typical $a$s of \qtyrange{2}{3.3}{\au} \citet{DeMeo2015}.
-The bounds between the inner, middle and outer main belt at \qty{2.5}{\au} and \qty{2.82}{\au}, these separations are due to resonances with Jupiter (the 3:1 and 5:2 mean motion resonances mark the edges of middle main belt) and Saturn (the $\nu_6$ secular resonance cuts off the lowest $a$ edge of the inner belt) carving swathes where asteroids cannot be in stable orbits.
-The eccentricity, $e$, of this orbit is the next parameter of interest.
+Asteroids are a key class of solar system objects. Much is known about them already, but there is always more to learn. 
+The orbits of asteroids can be well characterised by only a few values, the proper elements of the object.
+These are the lenght of the long axis of elliptical orbit, the semi-major axis, $a$, how elliptical the orbit is, the eccentricity, $e$, and the angle to the ecliptic plane, the inclination,$i$. 
+Typical values for main belt asteroids are an $a$ of \qtyrange{2}{3.3}{\au}, an $e$ in \qtyrange{0}{0.35}{} and an $i$ around \qtyrange{0}{30}{\degree} \citep{DeMeo2015}.
+Comets have undefined $a$ and larger $e$, with some long period comets having $e\approx 1$, but the typical range is \qtyrange{0.2}{0.7}{} \citep{Lewis2012}.
+Interstellar objects (ISOs) have $e>1$ as they are on unbound hyperbolic trajectories \citep{McGlynn1989}.
+
+% The semi-major axis, $a$, is the largest distance from the sun that the asteroid achieves on its elliptical orbit.
+% Main belt asteroids have a typical $a$ of \qtyrange{2}{3.3}{\au} \citep{DeMeo2015}.
+% The bounds between the inner, middle and outer main belt at \qty{2.5}{\au} and \qty{2.82}{\au}, these separations are due to resonances with Jupiter (the 3:1 and 5:2 mean motion resonances mark the edges of middle main belt) and Saturn (the $\nu_6$ secular resonance cuts off the lowest $a$ edge of the inner belt) carving swathes where asteroids cannot be in stable orbits.
+
+% The eccentricity, $e$, of this orbit is the next proper element.
 $e=0$ implies a perfectly circular orbit, $0<e<1$ are ellipses, $e=1$ give a parabolic orbit, and $e>1$ are unbound, hyperbolic orbits.
-Main belt asteroids have small $e$, in the range of \qtyrange{0}{0.35}{} \citep{DeMeo2015}
-Comets have larger $e$, with some long period comets having $e\approx 1$, but the typical range is \qtyrange{0.2}{0.7}{} \citep{Lewis2012}.
-Interstellar objects (ISOs) have $e>1$ as they are on unbound hyperbolic trajectories.
-The inclination, $i$, of asteroids is also a key orbital element, it is the angle relative to the ecliptic plane and in principle asteroids can span the whole range.
-Main belt asteroids have typical inclinations of \qtyrange{0}{30}{\degree} \citep{DeMeo2015}, so the TESS cameras closer to the ecliptic plane are expected to have the most asteroids in their fields.
+% Main belt asteroids have small $e$, in the range of  \citep{DeMeo2015}
+% Comets have larger $e$, with some long period comets having $e\approx 1$, but the typical range is \qtyrange{0.2}{0.7}{} \citep{Lewis2012}.
+% Interstellar objects (ISOs) have $e>1$ as they are on unbound hyperbolic trajectories.
 
-% // below TODO Abs mag H, and how that is size.
+% The inclination, $i$, of asteroids is also a key orbital element, it is the angle relative to the ecliptic plane.
+% In principle asteroids can span the whole range of inclinations, though they rarely do.
+% Main belt asteroids have typical inclinations of \qtyrange{0}{30}{\degree} \citep{DeMeo2015}, so the TESS cameras closer to the ecliptic plane are expected to have the most asteroids in their fields.
 
-%TODO asteroid families
 Asteroids can be grouped into families based on clustering of the proper orbital elements and similarities in their colour or spectra \citep{Nesvorny2015}.
@@
 
-%TODO YORP spin up
+%TODO Rotation of asteroids
+As an asteroid rotates, the amount of sunlight reflected changes with time.
+This produces a lightcurve that is often sinusoidal in appearance.
+These lightcurves are normally ``double-peaked'', having two maximums for one full rotation of the body, due to symmetries in the shape of the object.
+The shape of the asteroid has a large effect on the scale of variation in the lightcurve \citep{Durech2015}, the more elongated the object is, the larger the variation.
+The rate of rotation is limited by the size of the asteroid \citep{Pravec2000}, as the internal strength of the body is put under too much strain if large asteroid rotate with short periods.
+
+Understanding the rotation properties of asteroids has long been of interest to astronomers \citep[e.g.][for early work into the limits of rotation period and the tumbling nature of some small bodies]{Weidenschilling1981,Harris1994}.
+This is still at the forefront of research, with more lightcurves being published year-on-year \citep{Harris2015}.
+For the smallest objects ingenious observation techniques are employed to measure rotation periods such as intentionally streaking the asteroid down a column of pixel on a detector, and making the lightcurve from the streak, as done by \citet{Bolin2023}.
+
 Asteroids can have their rotation properties changed with time, collisions can do this, but so can light from the sun.
@@
 
-Understanding the rotation properties of asteroids has long been of interest to astronomers \citep[e.g.][for early work into the limits of rotation period and the tumbling nature of some small bodies]{Weidenschilling1981,Harris1994}.
-This is still at the forefront of research, with more lightcurves being published year-on-year \citep{Harris2015}.
-This will only increase with more large survey telescopes coming online soon, such as the Legacy Survey of Space and Time \citep[LSST][]{LSST2019}, the number of new asteroids that are well characterised will increase by more than 2 orders of magnitude on the Sloan Digital Sky Survey \citep[SDSS][]{York2000} which had $\sim 88000$ objects \citet{Parker2008}.
+The data on solar system bodies will only increase with more large survey telescopes coming online soon, such as the Legacy Survey of Space and Time \citep[LSST][]{LSST2019}.
+The number of new asteroids that are well characterised will increase by more than 2 orders of magnitude on the Sloan Digital Sky Survey \citep[SDSS][]{York2000} which had $\sim 88000$ objects \citet{Parker2008}.
 LSST needs a dedicated pipeline for asteroid classification, the Solar System Notification Alert Processing System \citep[SNAPS][]{Trilling2023} has been developed using ZTF data and is can scale up handle volume of data coming from LSST.
 Work is being done to understand the biases of how  asymmetric shape high amplitude variations can lead to selection effects of such large sample of asteroids \citet{Levine2023}.
-For the smallest objects ingenious observation techniques are employed to measure rotation periods such as intentionally streaking the asteroid down a column of pixel on a detector, and making the lightcurve from the streak, as done by \citet{Bolin2023}.
 
@@
 in the Light Curve Database (LCDB).
-%// TODO fig of period vs diameter.   
 In the October \nth{1}, 2023 release, the LCDB has nearly 9000 objects with periods determined to a quality code $U\geq 2-$, which is the minimum reliability recommended by \citeauthor{Warner2009} for statistical analysis.
@@
 \citeauthor{Warner2009} acknowledge that there are direct measurements of diameter for more asteroids in the present day, but continue to present the $H$ derived diameters for consistency.
-The spin barrier \citep{Pravec2000} is seen in the red dashed lines as there are very few asteroids larger than $\qty{0.15}{\kilo\metre}$ rotating faster than $\sim \qty{10}{\per\day}$.
-It is thought that the YORP effect spins up asteroids so fast that they pass above this barrier and then no longer have the internal strength necessary to hold themselves together, impling that all asteroids made to spin faster have torn themselves apart and that they have no tensile strength, i.e they are rubble piles instead of monoliths \citep{Harris1996}.
+
+The asteroid spin barrier \citep{Pravec2000} is seen in the red dashed lines in \autoref{Fig:FreqVsDiam}.
+There are very few asteroids larger than $\qty{0.15}{\kilo\metre}$ rotating faster than $\sim \qty{10}{\per\day}$.
+It is thought that the YORP effect, as wel as collisons, spin up asteroids so fast that they pass above this barrier and then no longer have the internal strength necessary to hold themselves together.
+This implies that all asteroids made to spin faster have torn themselves apart and that they have no tensile strength, i.e they are rubble piles instead of monoliths \citep{Harris1996}. 
 This idea of the cohesive strength of small bodies is still explored today \citep[e.g.][for V-type asteroids]{Oszkiewicz2020}.
+
 Comparing the periods I find of asteroids that are already in the LCDB is key to making sure my methods are reliable, and that the periods for asteroids for which no data exists can be trusted.
@@
 1I was classed as asteroid due to lack of a coma, and no noticeable activity.
-This contrasts the second ISO, 2I/Borisov, which had many cometary characteristics \citep[see ][for a review]{Dorofeeva2023}. %\\TODO cite, as opposed to 2I??
+This contrasts the second ISO, 2I/Borisov, which had many cometary characteristics \citep[see ][for a review]{Dorofeeva2023}.
+
 \omuamua was measured to have a rotation period of \qty{8.67(0.34)}{\hour} \citep{Belton2018} and seemed to be tumbling \citep[e.g.][]{Drahus2018,Fraser2018}.
 Combing the tumbling with an elongation ratio of up to \qty{6(1)}{}:1 \citep{McNeill2018}, 1I is said to have a cigar shape \citep{Belton2018}.
-The peak to peak amplitude variation of this ISO was \qty{2.5}{\mag} \citep{Meech2017} over its double peaked light curve.
-This is quite interesting, as it varies more than most asteroids.
-With the full sky survey of bright asteroids, we hope to find many asteroids with such a large amplitude variation, and to see just how rare \omuamua is.
+The peak to peak amplitude variation of this ISO was \qty{2.5}{\mag} \citep{Meech2017} over its double peaked lightcurve, which is interesting as it is more variation than most asteroids more than most asteroids \citep[as seen in the LCDB of][]{Warner2009}.
+With the full sky survey of bright asteroids, I hope to and to see just how rare \omuamua is, by finding many asteroids with large amplitude variations.
 
@@
 TESS is a large area, high cadence imaging, space telescope  \citep{Ricker2014}.
-TESS is tasked with observing one piece of sky for \qty{27}{\day} at a time (a sector), delivering \qtyproduct{96 x 24}{\degree} full frame images (FFIs) at regular intervals.
-These FFIs are built from stacked \qty{2}{\second} exposures, leveraging the short readout times of the 16 CCD cameras on-board.
+TESS is tasked with observing one piece of sky for \qty{27}{\day} at a time , which defines a sector.
+The telescope delivers \qtyproduct{96 x 24}{\degree} full frame images (FFIs) at regular intervals.
+These FFIs are built from stacked \qty{2}{\second} exposures, leveraging the short readout times of the 16 charged-coupled devices (CCD) spread over four cameras on-board.
 With the initial cadence for these full frame images set to \qty{30}{\minute}, the time resolution of TESS is unparalleled.
+As the mission was extended, after TESS had already mapped the entire sky, the length of the FFIs has decreased to \qty{10}{\minute} and then later to \qty{200}{\second}.
+This high time resolution and observation area does come at the cost of spatial resolution, as the pixels are each \qty{21}{\arcsec} square.
+
+Of interest here is what such a high sampling rate can do for statistics on the asteroid population.
+For bright asteroids the rotation periods should be able to be easily determined from this vast dataset, due to the high sampling of the lightcurve.
+The shortest FFIs will be able to accurately determine the rotation periods of all but the fastest rotating asteroids, most of which will be too dim to see in the TESS data.
+
 The highest frequency (shortest period) that can be found from time series data like this is at the Nyquist limit, twice the minimum time between images.
@@
 This is well sampled enough to characterise most variable stars, as well as find the orbital periods of exoplanets to a high precision.
-As the mission was extended, after TESS had already mapped the entire sky, the length of the FFIs has decreased to \qty{10}{\minute} and then later to \qty{200}{\second}.
-This decreased in the FFI exposure time comes with a complementary increase in the Nyquist frequency.
-This high time resolution and observation area does come at the cost of spatial resolution, as the pixels are each \qty{21}{\arcsec} square.
-Of interest here is what such a high sampling rate can do for statistic on the asteroid population.
-For bright asteroids the rotation periods should be able to be easily determined from this vast dataset.
-The shortest FFIs will be able to accurately determine the rotation periods of all but the fastest rotating asteroids, most of which will be too dim to see in the TESS data.
+The decreased in the FFI exposure time in later sectors comes with a complementary increase in the Nyquist frequency.
 
-There have been attempts to find and classify the asteroids in TESS data before.
-\citet{Pal2018} proposes that TESS will be a good instrument for solar system object study, and simulate some detections down to \nth{20} visual magnitude, but that good photometry should only be expected to $V \lesssim 19$.
-The sentiment of using TESS in this way is shared by \citet{Wong2019}.
+Previous work has been conducted to find and classify all the asteroids in TESS data.
+\citet{Pal2018} proposes that TESS will be a good instrument for solar system object study, and simulate some detections down to \nth{20} visual magnitude, but that good photometry should only be expected to $V \lesssim 19$, a sentiment shared by \citet{Wong2019}.
 The first data release in \citet{Pal2020} catalogues nearly 10,000 asteroid light curves from the first year of TESS operation, and they report that they triple the number of asteroids with accurate rotation periods.
@@
 The Minor Planet Center (MPC) gets regular updates from TESS thanks to the work of \citet{Woods2021} and their \texttt{LINEAR-TESS} program, which creates tracklets of objects over a day, and then stitches these together to form a track of each asteroid through an FFI.
+
+There is ongoing analysis of subsets of the asteroid population in TESS.
 \citet{Gowanlock2024} uses TESS photometry as well as  observations by the Zwicky Transient Facility \citep[ZTF, ][]{Bellm2018} to get a longer baseline on mutually observed objects while combining ground and space based observations.
@@
 The aim is to get a volume complete set of asteroid lightcurves and periods, by analysis of the whole sky as seen by TESS using data from more than just the first year of TESS's operation.
+This has not been achieved, as only one sectors analysis is presented here, but the methods are scalable.
+Because of the survey properties, TESS provides a self-consistent way to measure the properties of asteroids over the full sky.
+By using the higher cadence FFIs of later TESS sectors, the $P\leq\qty{3}{\hour}$ limit on the accurate periods \citep[as found by ][]{McNeill2023} can be lowered.
 I also employ a different data reduction method, reducing the data with the \texttt{TESSreduce} package \citep{Ridden-Harper2021} should increase the accuracy of the photometry due to the improved background subtraction process.
-Because of the survey properties, TESS provides a self-consistent way to measure the properties of asteroids over the full sky.
-By using the higher cadence FFIs of later TESS sectors, the $P\leq\qty{3}{\hour}$ limit on the accurate periods \citep[as found by ][]{McNeill2023} should be lowered.
+
 Another beneficial part of this work is that as part of a full sky transient survey using TESS, \texttt{TESSELLATE} \citep{TESSELLATE}.
@@
 This can confuse pipelines looking for  transients such as flare stars and supernova, as they can look very similar, having the brightness change in one pixel for a short time.
-The goal of finding all the asteroids will allow for the removal of these spikes from the transient pipeline, as well as to understand the asteroid population better.
-By filtering the asteroid detections out, while also self consistently determining rotation properties of a large sample of the population, more science can be done with the one action.
+Knowing where the asteroids are will allow for the removal of these spikes from the transient pipeline, as well as to understand the asteroid population better.
+By filtering the asteroid detections out, while also self consistently determining rotation properties of a large sample of the population, more science can be done with the one set of analyses.
 
 \section{Methods}\label{Sec:Meth}
+
+The \texttt{TESSELLATE} pipeline has been running on the OzSTAR supercomputing facilities.
+I built the asteroid detection and subsequent lightcurve and periodogram analysis, to be described below, by testing on one subsection of one CCD of one camera  of one TESS sector.
+Once I was confident this worked as requird, the same code was refactored slightly to work on OzSTAR and a large-scale analysis of all the processed TESS sectors can be run. 
+I will be looking for completeness of detections of asteroids, as well as accuracy of periods and amplitude variation.
+
+For this work, I present analysis of TESS sector 29\footnote{\url{https://tess.mit.edu/observations/sector-29/}}, which was observed between \nth{26} August and \nth{22} September 2020.
+This was the third TESS sector to be using \qty{10}{\min} cadence FFIs, which has had less analysis of the asteroids preformed.
+It was chosen mainly because of this higher cadence, it was the only sector of this cadence that had been processed by \texttt{TESSELLATE} with enough time for my analysis to be preformed. 
+What is presented for this section is scalable, on a per sector basis, to any sector.
 
@@
 An example of the matches found using this \texttt{KDTree} is shown in \autoref{Fig:RADecMatch}.
+
 It is clear that \texttt{TESSELLATE} is not detecting every frame that an asteroid is in.
@@
 
-Using the \texttt{SkyCoord} module of \texttt{Astropy}, the world coordinates, in RA and Dec, can be transformed into pixel coordinates, $x$ and $y$, using a world coordinates system (WCS) from TESSELLATE.
-\texttt{Photutils} \citep{Bradley2024} was then used to do aperture photometry with a \qty{1.5}{\px} radius to account for the spread in the TESS PRF. %?cite PRF
+The flux is calculated with \texttt{Photutils} \citep{Bradley2024} aperture photometry with a \qty{1.5}{\px} radius.
+This aperture is placed on the centre of mass (COM) position of a $5\times5$ \unit{\px} box around the integer $xy$ coordinate of the asteroid.  
+The integer coordinates are found with the \texttt{SkyCoord} module of \texttt{Astropy}, the world coordinates, in RA and Dec, can be transformed into pixel coordinates, $x$ and $y$, using a world coordinates system (WCS) from TESSELLATE.
 
-The integer coordinates of the asteroids seemed to be leading the object in any given frame of TESS. %?FIG?
-This was decreasing the total flux of the asteroid, as the box used for aperture  was hitting mostly background.
+The integer coordinates  were not used themselves because the asteroids seemed to be leading the object in any given frame of TESS. %?FIG?
+This was decreasing the total flux of the asteroid, as the aperture  was hitting mostly background.
 This also caused a ``sawtoothing'' effect, where the lightcurve would jump from a minimum to a maximum when the integer coordinate snapped from being a pixel out to back on the asteroid.
-Both of these issues can be fixed by calculation of the centre of mass (COM), the average brightness in $xy$ space on a $5\times5$ \unit{\px} box of the image centred on the interpolated position.
-I did not take the highest pixel value in this box because there could be a rouge star in the field that it snaps to, instead of the asteroid.
-A \texttt{Photutils} aperture photometry centred on the COM with the same radius of \qty{1.5}{\px} sorted some of the discrepancies in the lightcurve, with a far smoother transition from maximum to minimum, as expected from rotating asteroids.
+
+I did not take the highest pixel value in a box to make this correction because there could be a rouge star in the field that it detects as the maximum, instead of the asteroid.
+The COM technique would still be affected by such contamination, but the magnitude would be less. 
+The photometry on the COM aperture did fix the sawtoothing, as there is subpixel accuracy in the COM calculation.  
+There was a far smoother transition from maximum to minimum, as expected from rotating asteroids. 
 Hereafter, unless otherwise specified, ``flux'' refers to this COM flux measurement of the asteroid.
@@
 Following \citet{McNeill2023}, any fluxes more than $3\sigma$ from the mean value are sigma-clipped from the forced photometry lightcurves.
-This does introduce gaps into an otherwise regularly sampled lightcurve, but the contamination from background stars the object moves through of is too great.
+The contamination from background stars the object passes is too great to allow in the lightcurve.
+This does introduce gaps into an otherwise regularly sampled lightcurve, but the Lomb-Scargle periodogram is designed to deal with poorly sampled data.
+
 The subsections made by \texttt{TESSELLATE} have a small overlap.
@@
 
-The next part of my analysis has to do with determining the periods and amplitudes of each asteroid's light curve.
+The next part of my analysis involves determining the periods and amplitudes of each asteroid's light curve.
 
@@
 
-\subsection{Asteroid Statistics}\label{SubSec:StatMeth}
+\subsection{Periodogram Quality Checks}\label{SubSec:QualCheck}
 
 %Run on server
-The \texttt{TESSELLATE} pipeline has been running on the OzSTAR supercomputing facilities.
-After I was confident that all the parts of the asteroid detection and subsequent light curve analysis worked as required on one subsection of one camera of one CCD of one TESS sector, the same code was refactored slightly to work on OzSTAR and a large-scale analysis of all the processed TESS sectors can be run.
-I will be looking for completeness of detections of asteroids, as well as accuracy of periods and amplitude variation.
 
@@
 Looking at the data as a whole, all objects in the field of the instrument during the sector, with $V \leq \qty{20}{\mag}$, are plotted.
-The exception being comets and centaurs, which have an $a$ (really $q$, the perihelion distance) too larger to see the detail of the majority of the objects.
+The exception being centaurs, which have an $a$ too larger to see the detail of the majority of the objects and comets, which have an undefined $a$.
 The distinction between the resonances in the main belt can be seen.
@@
 The different FFI cadence, as well as the large temporal gap between these observations makes it non-trivial to recover the rotation periods.
-The different depths that the shortest cadence FFIs have \citep{MontillaHons}
+%?? The different depths that the shortest cadence FFIs have \citep{MontillaHons}
 

### 2024-10-17 FinalReport/FinalReportDraft.tex
Thursday nigh writing done

@@
 Combing the tumbling with an elongation ratio of up to \qty{6(1)}{}:1 \citep{McNeill2018}, 1I is said to have a cigar shape \citep{Belton2018}.
-The peak to peak amplitude variation of this ISO was \qty{2.5}{\mag} \citep{Meech2017} over its double peaked lightcurve, which is interesting as it is more variation than most asteroids more than most asteroids \citep[as seen in the LCDB of][]{Warner2009}.
+The peak to peak amplitude variation of this ISO was \qty{2.5}{\mag} \citep{Meech2017} over its double-peaked lightcurve, which is interesting as it is more variation than most asteroids more than most asteroids \citep[as seen in the LCDB of][]{Warner2009}.
 With the full sky survey of bright asteroids, I hope to and to see just how rare \omuamua is, by finding many asteroids with large amplitude variations.
@@
 
-\subsection{Querrying Databases}\label{SubSec:Querry}
+\subsection{Querrying Databases for Asteroid Positions and Properties}\label{SubSec:Querry}
 
@@
 
-\subsection{Interpolation}\label{SubSec:Interp}
+\subsection{Interpolation of the Positions to Match TESS Cadence}\label{SubSec:Interp}
 
@@
 
-With TESS data coming in $\frac12\unit{\hour}$ chunks, 24 interpolated points are needed between each API call.
+With TESS data coming in $\qty{10}{\minute}$ chunks, 72 interpolated points are needed between each API call.
+A linear interpolation between the querried RA and Dec is preformed at the times of the TESS frames.
+\texttt{TESSELLATE} saves out these time-frame pairs when doing its dataprocessing, allowing for the time coordinate of any position to be exactly at a frame time.
+
 These interpolated positions can be seen in \autoref{Fig:interpPos} for a subsection of sector from \texttt{TESSELLATE}.
@@
 There is also a large size range in this slice of sky, ranging over \qty{5}{\mag} in absolute magnitude, which can be seen in the alpha (or transparency) of the tracks in \autoref{Fig:interpPos}.
-Because of a small change in position between each frame, \citep[$\sim \qty{1}{\px}$ per frame][]{Pal2018}, this interpolation should be accurate.
+Because of a small change in position between each frame, \citep[$\sim \qty{1}{\px}$ per \qty{30}{min} frame][]{Pal2018}, this interpolation should be accurate.
 
@@
 
-\subsection{Detection Matching}\label{SubSec:Match}
+\subsection{Matching to \texttt{TESSELLATE} Detections}\label{SubSec:Match}
 
-%Match to detections
 Matching these interpolated positions to \texttt{TESSELLATE} detections is important to lower the unknown transient outputs of this pipeline.
@@
 The number of frames do change, type Ia supernova will brighten is a matter of a few hours and then dim for days, while stellar flares are of similar profile by a smaller max brightness and a correspondingly shorter decay time.
-Asteroids are very short spikes, however these detection pipelines are robust. %? Clarinda begs to differ
-These pipelines have already found the aforementioned supernova, stellar flares and variable stars, by matching to catalogues of those types of objects (Gaia, ZTF etc.).
+
+Asteroids are very short spikes, however these detection pipelines are robust.
+These pipelines have already found the aforementioned supernova, stellar flares and variable stars, by matching to catalogues of those types of objects (Gaia, ZTF etc.). %? cite gaia
 The job of this work is to catch all the asteroids in the set of all the transients that are returned from the pipeline.
@@
 Then, if multiple detection match to the same interpolated point, the detection with the smallest distance will be taken.
+
 An example of the matches found using this \texttt{KDTree} is shown in \autoref{Fig:RADecMatch}.
-
 It is clear that \texttt{TESSELLATE} is not detecting every frame that an asteroid is in.
@@
 Instead, there are bunches of frequently matched points, and then patches where there are few matches detected at all.
-
-%TODO fig for agregate of all asteroids in a sector, to go in results
 
@@
 
-\subsection{Ligthcurves}\label{SubSec:Lightcurves}
+\subsection{Constructing Asteroid Ligthcurves}\label{SubSec:Lightcurves}
 
@@
 This aperture is placed on the centre of mass (COM) position of a $5\times5$ \unit{\px} box around the integer $xy$ coordinate of the asteroid.
+The COM was calculate using \texttt{Scipy} and the \texttt{ndimage.center\_of\_mass}.
 The integer coordinates are found with the \texttt{SkyCoord} module of \texttt{Astropy}, the world coordinates, in RA and Dec, can be transformed into pixel coordinates, $x$ and $y$, using a world coordinates system (WCS) from TESSELLATE.
@@
   \includegraphics[width =\textwidth]{../OzData/SingleBodyLCUlyssesCorrectedAgain.pdf}
-  \caption[Light curves Example]{The light curves of the asteroid (5254) Ulysses throughout its entire journey across sector 29, with the flux of the interpolated positions in blue circles and the flux from the COM in orange diamonds, and the \texttt{TESSELLATE} flux for any matched points in the black squares.
-    The Flux axis is measured in counts, as calculated by \texttt{TESSreduce}, and the times are in modified Julian date (MJD).}
+  \caption[Light curves Example]{The light curves of the asteroid (5254) Ulysses throughout its entire journey across sector 29, with the flux of the interpolated positions in blue circles and the flux from the COM in orange diamonds, and the \texttt{TESSELLATE} flux for any matched points in the black squares.}
   \label{Fig:FullSingleLC}
 \end{figure}
-A comparison between the three light curves is interesting, as one might expect only the brightest parts of the lightcurve will be matched.
+
+A comparison between the three light curves is interesting.
 \autoref{Fig:FullSingleLC} shows these three light curves for a single body.
@@
 This should result in more accurate periodograms as there is less overall curve to the data.
+
 The matched points are comparatively sparse for this object, which is not the case for all.
-Many have no matches, and some have almost fully matched, this will be discussed more later.
+Many have no matches, and some have almost fully matched, this will be discussed more later. %ref
 Lightcurves with an average flux $<10$ counts are lost in the random noise floor, so a minimum mean flux limit needs to be applied to ensure quality and reliability of any periods calculated.
 
-\subsection{Period detection}\label{SubSec:Periods}
+\subsection{Calculation of Rotation Periods}\label{SubSec:Periods}
 
 %Periods and amplitueds for everything detected
-
 The next part of my analysis involves determining the periods and amplitudes of each asteroid's light curve.
+The rotation period and other properties of the lightcurve will be calculated from the forced photometry on the COM positions.
+the matching is still beneficial to the \texttt{TESSELLATE} pipeline as a whole.
+Taking previously unidentified detections and assigning them to an asteroid category, the remaining unknown transients can be more efficiently searched for other interesting astrophysical phenomena.
 
@@
 There are a few different implementations of a Lomb-Scargle periodogram available in world of astronomy python packages.
+I have used the Lomb-Scargle periodogram as implemented by \texttt{Astropy}\citep[\citet{Astropy2022} but see][for the implementation]{Vanderplas2012,Vanderplas2015}.
+
 I first tried using the \texttt{Lightkurve} \citep{Lightkurve2018} package built for period analysis of TESS (and Kepler) data of variable stars.
 \texttt{Lightkurve} was easy to use but did not allow for fine-tuning of the periodogram, or give any model of the curve it produced.
-The next option was peeling back a layer of abstraction and using the Lomb-Scargle periodogram as implemented by \texttt{Astropy}\citep[\citet{Astropy2022} but see][for the implementation]{Vanderplas2012,Vanderplas2015}.
 There were interesting similarities and differences between the packages.
@@
 \texttt{Astropy} gives a statistic on how confident it is in the detected frequency is, the false alarm probability.
-It also returns a model, in the form of
+
+\texttt{Astropy} also returns a model, in the form of
 
@@
 
-where the flux, $F$, is given in terms of time, $t$, the calculated frequency, f, and the model parameter $\vec{\theta} = [\theta_0, \theta_1,\theta_2]$.
+where the flux, $F$, is given in terms of time, $t$, the calculated frequency, $f$, and the model parameter $\vec{\theta} = [\theta_0, \theta_1,\theta_2]$.
 These parameters are useful in reconstructing the sinusoid without rerunning the periodogram.
+
 The method used for the periodogram fits is the new \texttt{nifty-ls} package \citep{Garrison2024} that has can be implemented on its own, or by interfacing with \texttt{Astropy}.
-I use the later implementation as the speed-up from the non-uniform fast Fourier transform is helpful on the large dataset from \texttt{TESSELLATE}, but the \texttt{Astropy} interface is how I had everything set up to run on just the subsection I was trialling my code on.
+I use the later implementation as the speed-up from the non-uniform fast Fourier transform is helpful.
+The large dataset from \texttt{TESSELLATE} requires this calculation efficiency but the \texttt{Astropy} interface is how I was trialling my code on smaller datasets.
 
@@
 For the \qty{10}{\min} FFIs (the cadence analysed here), the Nyquist period is $\qty{20}{\min}=\frac13\unit{\hour}$, a threefold increase in time sensitivity for fast rotating asteroids.
-Based on their experimentation with the \qty{30}{\min} cadence FFIs, \citet{McNeill2023} use a \qty{17}{\day} maximum period for their periodogram analysis.
-While a lack of signal in the rotation curve causes the calculated period to find this low frequency limit, such cases can be filtered out by keeping only periods less than \qty{90}{\percent} of this maximum.
-They recommended that periods less than \qty{3}{\hour} be interpreted with caution, due to lack of reliability in recovering injected sources.
+Based on their experimentation with the \qty{30}{\min} cadence FFIs, \citet{McNeill2023} find that periods less than \qty{3}{\hour} be interpreted with caution, due to lack of reliability in recovering injected sources.
 Because of the increase in FFI rate, this time will be droped to \qty{1}{\hour} for the minimum reliable period, as the lightcurve sampling on these timescales is equivalent.
 
-%???
-There will be differences in the period between the matched points and the interpolated points, not just because of the difference in average flux, but also from the larger number of points.
-The matches are often having fewer observations and are much sparser in time that the forced photometry measurements (see the black squares in \autoref{Fig:FullSingleLC}), so would give worse periodograms.
-While the rotation period and other properties of the lightcurve will be calculated from the forced photometry on the COM positions, the matching is still beneficial to the \texttt{TESSELLATE} pipeline as a whole.
-Taking previously unidentified detections and assigning them to an asteroid category, the remaining unknown transients can be more efficiently searched for other interesting astrophysical phenomena.
+A \qty{17}{\day} maximum period for  periodogram analysis follows \citeauthor{McNeill2023}.
+A lack of signal in the rotation curve causes the calculated period to default this low frequency limit.
+Such cases can be filtered out by keeping only periods less than \qty{90}{\percent} of this maximum.
 
@@
   \includegraphics[width=\textwidth]{../OzData/PeriodogramUlyssesUpdated.pdf}
-  \caption[Periodogram example]{\textit{Upper Panel}:The Periodogram of the COM flux of an asteroid, with the power of each frequency displayed, both axes are in log scale.
+  \caption[Periodogram example]{\textit{Upper Panel}:The Periodogram of the COM flux of an asteroid, with the power of each frequency displayed.
     The maximum power of the periodogram is highlighted with the gold star.
-    This maximum is the calculated rotation frequency
+    This maximum is the calculated rotation frequency.
     The known frequency, from the LCDB \citep{Warner2009}, is shown in the blue dotted lines, and a factor of 2 alias each way is shown in the purple dotted lines ($2*$) and red dotted lines ($0.5*$).
@@
 The periodogram calculated that this object had a frequency double that listed in the LCDB, as seen with the vertical dotted lines.
-This is expected due to the double peaked nature of most asteroid lightcurves \citep{McNeill2023}.
-The stared peak is by far the most prominent peak in the periodogram, as it is in log scale.
+This is expected due to the double-peaked nature of most asteroid lightcurves \citep{McNeill2023}.
+This asteroid was highlighted as an alias of the known frequency is recovered by this analysis.
+This is promising for those asteroids where no known frequency exists.
+
+The stared peak in \autoref{Fig:PeriodEx} is by far the most prominent peak in the periodogram, as it is in log scale.
 Not all the asteroids have such well-defined maximums, and the peak power of the periodogram is an important factor in its quality.
 A peak power less than 0.1 means there is little to no signal observed and it will be almost indistinguishable from the noise.
+
 The lower panel of \autoref{Fig:PeriodEx} shows the model given in \autoref{Eq:LCModel} fit to the COM flux lightcurve by \texttt{Astropy}.
@@
 This model is successfully fit over a break in the lightcurve, and ignores the observations with a lower flux that did not quite get sigma clipped.
-This asteroid was chosen as an alias of the known frequency is recovered by this analysis, which is promising for those asteroids where no known frequency exists.
-The double alias is expected here due to the double peaked nature of the lightcurve \citep{McNeill2023}.
-%// TODO discuss fit of model...
 
-%// TODO single object periodogram
-%Maybe a bad one too
+The amplitude of the variation is calculated from this model fit.
+This is measured in magnitudes, and as such is found from the standard
+
+\begin{equation} \label{Eq:DelMag}
+  \Delta m = -2.5 \log_{10}(\frac{F_{min}}{F_{max}})
+\end{equation}
+
+with $\Delta m$ being the variation amplitude.
+The $F_{min}$ and $F_{max}$ terms in \autoref{Eq:DelMag} and the minimum and maximum flux of the model, respectivly.
+These are measured in counts.
 
 \subsection{Periodogram Quality Checks}\label{SubSec:QualCheck}
-
-%Run on server
 
@@
 A lightcurve must have an mean flux of at least 10 counts to be reliable.
-For an accurate periodogram, it is required that the lightcurve also contain more than 200 observations, which is very easy for TESS and its \qty{10}{\minute} cadence.
-The periodogram itself must have a peak power of at least 0.1, and the calculated period must be $\qty{1}{\hour}\leq P \leq \qty{367.2}{\hour}$, where the upper bound is \qty{90}{\percent} of the \qty{17}{\day} window suggested by \cite{McNeill2023}.
+For an accurate periodogram, it is required that the lightcurve also contain more than 200 observations, which is very easy for TESS and its \qty{10}{\minute} cadence, as the asteroid only has to be in the sector for \qty{33.3}{\hour}.
+The periodogram itself must have a peak power of at least 0.1 to be distingused from the noise
+The calculated period must be $\qty{1}{\hour}\leq P \leq \qty{367.2}{\hour}$, where the upper bound is \qty{90}{\percent} of the \qty{17}{\day} window suggested by \cite{McNeill2023}.
+
+\section{Results}\label{Sec:Res}
+
+Here I present the results for TESS sector 29.
+% This sector was chosen because it is early in the \qty{10}{\min} cadence of TESS FFIs, and had been fully run through TESSELLATE.
+Similar analysis can be preformed, and the same figures made, for any of the sectors.
+The computation time was a limiting factor in getting more sectors processed.
+
+\subsection{Quality Check Verification}
+
 \begin{figure}[t]
@@
     Dotted lines of a constant relationship are plotted, coloured the same as \autoref{Fig:PeriodEx}.
-    Both axes are in log scale.
   }
-
   \label{Fig:LCDBcomp}
 \end{figure}
-While some of the checks were chosen to agree with the past work of others, some were chosen to make sure the derived periods were in agreement with the known ones from the LCDB.
+
+While some of the quality checks were chosen to agree with the past work of others, some were chosen to make sure the derived periods were in agreement with the known ones from the LCDB.
 A comparison between objects in both the sector and the LCDB is seen in \autoref{Fig:LCDBcomp}.
@@
 This line corresponds to the calculated frequency being at the double alias of the known frequency, as frequency and period are inveses of each other.
-As seen in \autoref{Fig:PeriodEx} and expected from \citet{McNeill2023}, this line is what the periodogram should return.
+As seen in \autoref{Fig:PeriodEx} and expected from \citet{McNeill2023}, this line is what the periodogram should return due to double-peaked lightcurves.
 Most of the asteroids the pass the cut line along or near this key relation.
@@
 Some experimentation was done on the limiting values, and the ones presented here achieved the highest recovery of asteroid periods, while minimising the number that were incorrectly recovered.
+
 The idea of doubleing the detected period of all the asteroids and calling that number the rotation period has had some doubt cast on it, as there are a couple of points recovered nicely on the $y=x$ line, indicating that the true period is found.
 Either the LCDB has not accounted for the alias, which is unlikely as I only took the periods with high quality codes as they suggest \citep{Warner2009}, or the rotation signal sometimes returns the true period.
-The latter implies the lightcurve is not double peaked, as is normally expected.
+The latter implies the lightcurve is not double-peaked, as is normally expected.
+On the whole, there are enough periods that pass the checks with incorrect periods that these extra erronious periods will not significantly change the fraction of reliable detections.
 
-\section{Results}\label{Sec:Res}
-
-This is the results for TESS sector 29.
-This sector was chosen because it is early in the \qty{10}{\min} cadence of TESS FFIs, and had been fully run through TESSELLATE.
-Similar results can be found in any of the sectors, the computation time was a limiting factor in getting more sectors processed.
-
-\subsection{Asteroid Properties}\label{SubSec:AstPropRes}
+\subsection{Asteroid Properties}\label{SubSec:AstPropRes} %TODO fix name
 

### 2024-10-19 FinalReport/FinalReportDraft.tex
?final? Figures added from Ozstar

@@
 
-\autoref{Fig:FreqVsDiam} has been constructed from data from the Light Curve Database (LCDB) \citet{Warner2009}, as cited in text.
-That data has being built suplimented with my own work in the corrseponding Figure %TODO
+\autoref{Fig:FreqVsDiam} has been recreated after data from the Light Curve Database (LCDB) \citet{Warner2009}, as cited in text.
+That data has been supplemented with my own work in the corresponding \autoref{Fig:FreqDiamUpdate}.
 
@@
 
-Asteroids can have their rotation properties changed with time, collisions can do this, but so can light from the sun.
+Asteroids can have their rotation properties changed with time, collisions can do this, and so can light from the sun.
 The YORP effect (named for the discoverers Yarkovsky, O'Keefe, Radzievskii, and Paddack) is the process of tourquing asteroids of an asymmetric shape by both bombardment of photons and thermal (re-)emission \citep{Rubincam2000}.
@@
 The number of new asteroids that are well characterised will increase by more than 2 orders of magnitude on the Sloan Digital Sky Survey \citep[SDSS][]{York2000} which had $\sim 88000$ objects \citet{Parker2008}.
-LSST needs a dedicated pipeline for asteroid classification, the Solar System Notification Alert Processing System \citep[SNAPS][]{Trilling2023} has been developed using ZTF data and is can scale up handle volume of data coming from LSST.
+LSST needs a dedicated pipeline for asteroid classification, the Solar System Notification Alert Processing System \citep[SNAPS][]{Trilling2023} has been developed using Zwicky Transient Facility \citep[ZTF, ][]{Bellm2018} data and can scale up handle volume of data coming from LSST.
 Work is being done to understand the biases of how  asymmetric shape high amplitude variations can lead to selection effects of such large sample of asteroids \citet{Levine2023}.
@@
 Because of such a rich history of study, many asteroids have a known rotation period.
-These are catalogued by \citet{Warner2009} %TODO updated in citation http://www.MinorPlanet.info/php/lcdb.php
-in the Light Curve Database (LCDB).
+These are catalogued by \citet{Warner2009} in the Light Curve Database (LCDB).
 In the October \nth{1}, 2023 release, the LCDB has nearly 9000 objects with periods determined to a quality code $U\geq 2-$, which is the minimum reliability recommended by \citeauthor{Warner2009} for statistical analysis.
 This includes other TESS data from \citet{Pal2020} and \citet{Woods2021}.
-The frequencies against diameter plot of these asteroids in \autoref{Fig:FreqVsDiam} shows the collection of all the asteroids in the LCDB.
+The frequencies against diameter plot of these asteroids in \autoref{Fig:FreqVsDiam} shows the collection of all the asteroids in the LCDB with $U\geq 2-$.
 The diameter presented in the LCDB here calculated from the absolute magnitude $H$ of the bodies, assuming an albedo.
@@
 There are very few asteroids larger than $\qty{0.15}{\kilo\metre}$ rotating faster than $\sim \qty{10}{\per\day}$.
-It is thought that the YORP effect, as wel as collisons, spin up asteroids so fast that they pass above this barrier and then no longer have the internal strength necessary to hold themselves together.
+It is thought that the YORP effect, as well as collisions, spin up asteroids so fast that they pass above this barrier and then no longer have the internal strength necessary to hold themselves together.
 This implies that all asteroids made to spin faster have torn themselves apart and that they have no tensile strength, i.e they are rubble piles instead of monoliths \citep{Harris1996}.
@@
 
-Comparing the periods I find of asteroids that are already in the LCDB is key to making sure my methods are reliable, and that the periods for asteroids for which no data exists can be trusted.
+Comparing the periods I calculate of asteroids that are already in the LCDB is key to making sure my methods are reliable, and that the periods for asteroids for which no data exists can be trusted.
 
-\begin{figure}[t!]
+\begin{figure}[h]
   \centering
-  \includegraphics[width=0.7\textwidth]{../Test Code/Testing Figures/Diam-FreqPlot.pdf}
+  \includegraphics[width=0.7\textwidth]{./Figures/Diam-FreqPlot.pdf}
   \caption[Asteroid frequencies vs diameter (LCDB)]{
-    Rotaion rate of asteroids by their size, recreated from the  October \nth{1}, 2023 LCDB after \citet{Warner2009}.
+    Rotation rate of asteroids by their size, recreated from the  October \nth{1}, 2023 LCDB after \citet{Warner2009}.
   }
@@
 There is ongoing analysis of subsets of the asteroid population in TESS.
-\citet{Gowanlock2024} uses TESS photometry as well as  observations by the Zwicky Transient Facility \citep[ZTF, ][]{Bellm2018} to get a longer baseline on mutually observed objects while combining ground and space based observations.
+\citet{Gowanlock2024} uses TESS photometry as well as  observations by ZTF to get a longer baseline on mutually observed objects while combining ground and space based observations.
 There were not very many objects in both samples with known periods for comparison, with only 222 objects analysed, however they demonstrate that the technique is effective.
@@
 This work aims to extend these studies to more sectors, hence characterising more asteroids.
-The aim is to get a volume complete set of asteroid lightcurves and periods, by analysis of the whole sky as seen by TESS using data from more than just the first year of TESS's operation.
-This has not been achieved, as only one sectors analysis is presented here, but the methods are scalable.
+The goal is to get a volume complete set of asteroid lightcurves and periods, by analysis of the whole sky as seen by TESS using data from more than just the first year of TESS's operation.
+This has not been achieved, as the analysis of only one sector is presented here, but the methods are scalable.
 Because of the survey properties, TESS provides a self-consistent way to measure the properties of asteroids over the full sky.
 By using the higher cadence FFIs of later TESS sectors, the $P\leq\qty{3}{\hour}$ limit on the accurate periods \citep[as found by ][]{McNeill2023} can be lowered.
-I also employ a different data reduction method, reducing the data with the \texttt{TESSreduce} package \citep{Ridden-Harper2021} should increase the accuracy of the photometry due to the improved background subtraction process.
+I also employ a different data reduction method, using reduced data from the \texttt{TESSreduce} package \citep{Ridden-Harper2021}.
+This should increase the accuracy of the photometry due to the improved background subtraction process.
 
-Another beneficial part of this work is that as part of a full sky transient survey using TESS, \texttt{TESSELLATE} \citep{TESSELLATE}.
+This work is part of a full sky transient survey using TESS, the TESS Extensive Lightcurve Logging and Analysis of Transient Events4 (\texttt{TESSELLATE}) pipeline \citep{TESSELLATE}.
+To see what \texttt{TESSELLATE} detects and to help progress the pipeline, one can participate in the Cosmic Cataclysim Zooniverse project\footnote{\href{https://www.zooniverse.org/projects/cheerfuluser/cosmic-cataclysms}{https://www.zooniverse.org/projects/cheerfuluser/cosmic-cataclysms}}.
 Asteroids can spike the brightness of a pixel for only a few frames while it is passing by.
-This can confuse pipelines looking for  transients such as flare stars and supernova, as they can look very similar, having the brightness change in one pixel for a short time.
+This can confuse the pipeline looking for transients such as flare stars and supernova, as they can look very similar, having the brightness change in one pixel for a short time.
 Knowing where the asteroids are will allow for the removal of these spikes from the transient pipeline, as well as to understand the asteroid population better.
@@
 The \texttt{TESSELLATE} pipeline has been running on the OzSTAR supercomputing facilities.
-I built the asteroid detection and subsequent lightcurve and periodogram analysis, to be described below, by testing on one subsection of one CCD of one camera  of one TESS sector.
-Once I was confident this worked as requird, the same code was refactored slightly to work on OzSTAR and a large-scale analysis of all the processed TESS sectors can be run.
-I will be looking for completeness of detections of asteroids, as well as accuracy of periods and amplitude variation.
+I built this asteroid detection and subsequent lightcurve and periodogram analysis program by testing on one subsection of one CCD of one camera  of one TESS sector.
+The code written can be accessed at my ASTR480 GitHub repo \footnote{\href{https://github.com/ble61/ASTR480}{https://github.com/ble61/ASTR480}}. 
+Because of the large number of asteroids there are too many lightcurves to look at, so the code has to run with no human decisions. 
+Once I was confident this worked as required, the same code was refactored slightly to work on OzSTAR and a large-scale analysis of a processed TESS sector was run.
+% I will be looking for completeness of detections of asteroids, as well as accuracy of periods and amplitude variation.
 
@@
 It was chosen mainly because of this higher cadence, it was the only sector of this cadence that had been processed by \texttt{TESSELLATE} with enough time for my analysis to be preformed.
-What is presented for this section is scalable, on a per sector basis, to any sector.
+What is presented for this section is scalable, on a per-sector basis, to any sector that has been reduced.
 
@@
 To check for asteroids in the TESS data, the positions of the asteroids with time are required.
-For most asteroids, their orbital elements are well known, so it is a matter of looking them up and cross-matching with transients in the TESS data.
-Python was used to make API calls to
-%{Skybot}\footnote{\href{https://vo.imcce.fr/webservices/skybot/}{Skybot: https://vo.imcce.fr/webservices/skybot/}} 
-Skybot \citep{Berthier2006} to get positions of asteroids in a cone search box in right ascension (RA) and  declination (Dec) space.
+For asteroids bright enough to be in TESS, their ephemeris is well known, so I needed to access their positions and then cross match them with transients in the TESS data.
+
+\texttt{SkyBoT} \citep{Berthier2006} was queried to get positions of asteroids in a cone search.
+A box in right ascension (RA) and  declination (Dec) space.
 The search was limited to only asteroids with an average visual magnitude $V\geq \qty{20}{\mag}$.
-As TESS sectors are \qty{27}{\day} long, so querying every \qty{12}{\hour} is both computationally feasible and allows for a good understanding of a bodies motion. %TODO cite numpy
-These positions are still very sparsely spaced in time compared to the TESS data, so an interpolation is needed to bridge the gap.
+As TESS sectors are \qty{27}{\day} long, so querying every \qty{12}{\hour} is both computationally feasible and allows for a good understanding of a bodies motion.
+
+These positions are still very sparsely spaced in time compared to the TESS data. 
+An interpolation is needed to bridge the gap, this is preformed by \texttt{numpy} \citep{Harris2020}.
 
 Knowing where the asteroids are is helpful for finding them in the archival TESS data, but knowing more about the individual asteroids is also useful for population statistics.
-The \texttt{astroquery} \citep{Ginsburg2019} package was used to query JPL Horizons% {JPL Horizons}\footnote{\href{https://ssd.jpl.nasa.gov/horizons/}{JPL Horizons: https://ssd.jpl.nasa.gov/horizons/}}
-was used to obtain the orbital elements of each asteroid, $a$, $e$, and $i$, as well as an absolute magnitude $H$, which is a good proxy for size (as discussed above). %TODO ref
+The \texttt{astroquery} \citep{Ginsburg2019} package was used to query \texttt{JPL Horizons} was used to obtain the orbital elements of each asteroid, $a$, $e$, and $i$, as well as an absolute magnitude $H$, which is a good proxy for size (as discussed above).
+%TODO ref
 
@@
 
-\begin{figure}[h!]
-  \begin{subfigure}[t]{0.5\textwidth}
+\begin{figure}[t!]
   \centering
-    \includegraphics[width =\textwidth]{../Test Code/Testing Figures/interpPos\_22\_2\_3\_4.pdf}
-    \caption[Positions]{Positions of asteroids in a subsection.}
-    \label{Fig:interpPos}
-  \end{subfigure}
-  ~
-  \begin{subfigure}[t]{0.5\textwidth}
-    \centering
-    \includegraphics[width =\textwidth]{../Test Code/Testing Figures/1989 BC1PosCheck.pdf} %TODO one with more points???
-    \caption[Errors]{Error in interpolated positions}
-    \label{Fig:errPos}
-  \end{subfigure}
-
-  \caption[Interpolated positions and error of asteroids]{
-    (a)
+  \includegraphics[width =0.8 \textwidth]{./Figures/interpPos\_29\_1\_2\_2.pdf}
+  \caption[Interpolated Positions of Asteroids]{
     The interpolated positions of asteroids in one subsection of a TESS sector.
     The colours of the lines are time sequenced, as shown in the colour bar.
-    The alpha of the colours are scaled by the absolute magnitude $H$ of the asteroid, queried from JPL Horizons.
-    Both celestial (RA and Dec) and ecliptic (ecLon and ecLat) coordinates axes are shown.
-    The green $+$ is the centre of the subsection.
-    This sunsection was chosen because of how many asteroids there were, enough to see the shape of the box, but not so much that the lines are overwhelming to the eye.\\
-    (b)
-    The difference in the interpolated position of an asteroid from bulk queries to JPL Horizons, with interpolated coordinate being subtracted from the coordinate returned by Horizons.
-The colour bar is a different colour map to (a) as the range of time is shorter.
-The extent of the box is one TESS pixel, \qty{21}{\arcsec}.
-This is of asteroid ``1989 BC1'', but all others checked are similarly accurate (It is very expensive to check them all by brute force queries of Horizons).
-  }
-  \label{Fig:interpandErrPos}
+    The alpha of the colours are scaled by the absolute magnitude $H$ of the asteroid.
+    Both celestial coordinates (RA and Dec) and ecliptic longitude and latitude  axes are shown.
+    The green $+$ is the centre of the subsection.}
+    \label{Fig:interpPos}
+\end{figure}
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width =0.8 \textwidth]{./Figures/1990 KC1PosCheck.pdf} %TODO one with more points???
+  \caption[Interpolated Position Verification]{
+  The difference in the interpolated position of an asteroid from queries to \texttt{JPL Horizons}. 
+  The change in position in both arcseconds (\unit{\arcsec}) and TESS pixels are shown.  
+  }\label{Fig:errPos}
 \end{figure}
@@
 With TESS data coming in $\qty{10}{\minute}$ chunks, 72 interpolated points are needed between each API call.
-A linear interpolation between the querried RA and Dec is preformed at the times of the TESS frames.
-\texttt{TESSELLATE} saves out these time-frame pairs when doing its dataprocessing, allowing for the time coordinate of any position to be exactly at a frame time.
+A linear interpolation between the queried RA and Dec is preformed at the times of the TESS frames.
+\texttt{TESSELLATE} saves out these time-frame pairs as part of its analysis, allowing for the time coordinate of any position to be exactly at a frame time.
+This will also simplify the detection matching, as the any difference in the time will mean the detection is in a different frame.
 
@@
 There is also a large size range in this slice of sky, ranging over \qty{5}{\mag} in absolute magnitude, which can be seen in the alpha (or transparency) of the tracks in \autoref{Fig:interpPos}.
+
 Because of a small change in position between each frame, \citep[$\sim \qty{1}{\px}$ per \qty{30}{min} frame][]{Pal2018}, this interpolation should be accurate.
+Checking against a higher frequency query to the \texttt{JPL Horizons} ephemeris  confirmed this accuracy on a few asteroids, as shown in \autoref{Fig:errPos}.
+This is of asteroid ``(8654) 1990 KC1'', the largest difference is only a third of a TESS pixel in RA, so the interpolations are accurate enough for these measurements. 
+All others checked are similarly accurate.
+This is not a feasible way of getting all the positions, as the queries are per name and time, not per area and time like the cone search in \texttt{SkyBoT}.
+They are also rate limited, so while a few asteroids can be checked, filling the gaps between the \texttt{SkyBoT} half day queries for all the asteroids would be too expensive.
 
-Checking against a higher frequency query to the JPL Horizons ephemeris  confirmed this accuracy on a few asteroids, as shown in \autoref{Fig:errPos}.
-This is not a feasible way of getting all the positions, as the queries are per name and time, not per area and time like the cone search in Skybot.
-They are also rate limited, so while a few asteroids can be checked, filling the gaps between the Skybot half day queries for all the asteroids would take far too long.
-For the faster TESS FFIs, more interpolated points are needed, but a smaller the change in position between each point will be.
-Forcing the interpolations to be at the same time as the TESS frames, which are known thanks to \texttt{TESSELLATE} saving them out, simplifies the matching to detections and forced photometry of these points.
-
-\subsection{Matching to \texttt{TESSELLATE} Detections}\label{SubSec:Match}
+\subsection{Matching Asteroids to \texttt{TESSELLATE} Detections}\label{SubSec:Match}
 
@@
 Having interpolated their positions, the asteroids have a well sampled set of RA, Dec and time values of where they should be in the TESS data.
-They should show up in a pixel for a small number of frames, of order $\sim1$.
+They should show up in a single pixel in $xy$ space for a small number of frames, of order $\sim1$.
 This is the same as a lot of other transient events, a sharp rise in brightness and then disappearing quickly again.
-The number of frames do change, type Ia supernova will brighten is a matter of a few hours and then dim for days, while stellar flares are of similar profile by a smaller max brightness and a correspondingly shorter decay time.
+The number of frames for events is variable, type Ia supernova will brighten is a matter of a few hours and then dim for days, while stellar flares are of similar profile by a smaller max brightness and a correspondingly shorter decay time.
 
-Asteroids are very short spikes, however these detection pipelines are robust.
+Asteroids are very short events, however these detection pipelines are robust and do pick them up.
 These pipelines have already found the aforementioned supernova, stellar flares and variable stars, by matching to catalogues of those types of objects (Gaia, ZTF etc.). %? cite gaia
-The job of this work is to catch all the asteroids in the set of all the transients that are returned from the pipeline.
+A Goal of this work is to catch all the asteroids in the set of all the transients that are returned from the pipeline.
 To do this, catalogue matching is in order.
@@
 Filtering this \texttt{KDTree} output by restricting the time between spatially coincident matches to less than \qty{0.1}{\day} stops any accidental matches in position from non-asteroid detections.
-A cut is then made by smallest distance, in coordinate space, and any distance must be smaller than \qty{0.01}{\degree} (\qty{36}{\arcsec}), which is within 2 TESS pixels.
-Then, if multiple detection match to the same interpolated point, the detection with the smallest distance will be taken.
+A cut is then made by distance, which must be smaller than \qty{0.01}{\degree} (\qty{36}{\arcsec}), which is within 1.7 TESS pixels.
+Then, if multiple detection match to the same interpolated point, the detection with the smallest distance to the match is taken.
 
@@
 Detections in \texttt{TESSELLATE} are done on a per pixel basis, and asteroids only cross each pixel for a short number of frames.
-Not every interpolated point gets a match, due to a variety of reasons, which culminate in there being not an extreme enough of a difference between frames at that pixel location.
-Instead, there are bunches of frequently matched points, and then patches where there are few matches detected at all.
+The residuals of \autoref{Fig:RADecMatch} are all within a TESS pixel, shown on the right hand axis of each of the lower pannels.
+
+There are strings of frequently matched points, and then patches where there are few matches detected at all.
+Not every interpolated point gets a match, due to a variety of reasons.
+These all culminate in too small of a change in flux of the pixel from normal brightness and the frames where the asteroid is there.
+The change has to be quite extreme for \texttt{TESSELLATE} to detect it.
 
@@
   \centering
-  \includegraphics[width =\textwidth]{../Test Code/Testing Figures/DetectMatchPos.pdf}
-  \caption[Interpolated and Detected Positions]{The RA (right plot) and Dec (left plot) against the time of observation of the interpolated points (blue dots) or the matches to \texttt{TESSELLATE} detections (black squares).
-The gap at low times is from the TESS downtime in the sector, when it was passing close to earth.}
+  \includegraphics[width =\textwidth]{./Figures/DetectMatchPosUlysses.pdf}
+  \caption[Interpolated and Detected Positions]{The RA (right) and Dec (left) of interpolated position (blue dots) or the matches to \texttt{TESSELLATE} detections (pink hexagons) against the time of observation.
+The lower panels show the residuals of the difference between the positions for points that occur at the same time.}
   \label{Fig:RADecMatch}
@@
 
-\subsection{Constructing Asteroid Ligthcurves}\label{SubSec:Lightcurves}
+\subsection{Photometry and the Construction of Asteroid Lightcurves}\label{SubSec:Lightcurves}
 
@@
 A few steps need to be taken to clean these COM lightcurves.
-Following \citet{McNeill2023}, any fluxes more than $3\sigma$ from the mean value are sigma-clipped from the forced photometry lightcurves.
+Following \citet{McNeill2023}, any fluxes more than $3\sigma$ from the mean value are sigma-clipped recursivly from the forced photometry lightcurves, until convergence is achieved.
 The contamination from background stars the object passes is too great to allow in the lightcurve.
 This does introduce gaps into an otherwise regularly sampled lightcurve, but the Lomb-Scargle periodogram is designed to deal with poorly sampled data.
+Lightcurves with an average flux $<10$ counts are lost in the random noise floor, so a minimum mean flux limit needs to be applied to ensure quality and reliability of any periods calculated.
 
-The subsections made by \texttt{TESSELLATE} have a small overlap.
+The subsections of the FFIs made by \texttt{TESSELLATE} have a small overlap.
 If an object moves between these, there will be a few frames where it is found in multiple of these subsections.
@@
   \centering
-  \includegraphics[width =\textwidth]{../OzData/SingleBodyLCUlyssesCorrectedAgain.pdf}
-  \caption[Light curves Example]{The light curves of the asteroid (5254) Ulysses throughout its entire journey across sector 29, with the flux of the interpolated positions in blue circles and the flux from the COM in orange diamonds, and the \texttt{TESSELLATE} flux for any matched points in the black squares.}
+  \includegraphics[width =\textwidth]{./Figures/SingleBodyLCUlyssesgapSpan.pdf}
+  \caption[Light curves Example]{The light curves of the asteroid (5254) Ulysses throughout its entire journey across sector 29. 
+  The flux of the interpolated positions in blue circles,the \texttt{TESSELLATE} flux for any matched points in the pink hexagons, and the COM flux in orange diamonds.
+  The mid-sector gap is highlighted by the grey box.}
   \label{Fig:FullSingleLC}
@@
 \autoref{Fig:FullSingleLC} shows these three light curves for a single body.
-The COM flux seems much more stable than the interpolated positions flux on its own, and at a higher average flux count.
-This should result in more accurate periodograms as there is less overall curve to the data.
+The COM flux seems much more stable than the interpolated positions flux, and has a higher average flux count.
+This should result in more accurate periodograms as there is less overall variation in the data caused by the misplaced aperture.
 
 The matched points are comparatively sparse for this object, which is not the case for all.
-Many have no matches, and some have almost fully matched, this will be discussed more later. %ref
-Lightcurves with an average flux $<10$ counts are lost in the random noise floor, so a minimum mean flux limit needs to be applied to ensure quality and reliability of any periods calculated.
+Many have no matches, and some have almost fully matched, as discussed above (\autoref{SubSec:Match}) this will be discussed more later in \autoref{SubSec:Detect}.
 
-\subsection{Calculation of Rotation Periods}\label{SubSec:Periods}
+\subsection{Calculation of Asteroid Rotation Periods}\label{SubSec:Periods}
 
@@
 I have used the Lomb-Scargle periodogram as implemented by \texttt{Astropy}\citep[\citet{Astropy2022} but see][for the implementation]{Vanderplas2012,Vanderplas2015}.
-
-I first tried using the \texttt{Lightkurve} \citep{Lightkurve2018} package built for period analysis of TESS (and Kepler) data of variable stars.
-\texttt{Lightkurve} was easy to use but did not allow for fine-tuning of the periodogram, or give any model of the curve it produced.
-There were interesting similarities and differences between the packages.
-\texttt{Lightkurve} is based on the \texttt{Astropy} methods, but does not give all the functionality, instead opting to simplify the process.
-\texttt{Astropy} gives a statistic on how confident it is in the detected frequency is, the false alarm probability.
+The times are converted from MJD into seconds before the periodogram is calculated, so the frequency is calculated in \unit{\hertz}. 
+The flux values have their mean subtracted so the change in flux being analysed is about 0 counts.
+\texttt{Astropy} has ways of doing this internally, but I opted to do it explicitly.
 
 \texttt{Astropy} also returns a model, in the form of
-
 \begin{equation}
@@
 \end{equation}
-
 where the flux, $F$, is given in terms of time, $t$, the calculated frequency, $f$, and the model parameter $\vec{\theta} = [\theta_0, \theta_1,\theta_2]$.
 These parameters are useful in reconstructing the sinusoid without rerunning the periodogram.
+\texttt{Astropy} gives a statistic on how confident it is in the detected frequency is, the false alarm probability.
 
@@
 
+I originally tried using the \texttt{Lightkurve} \citep{Lightkurve2018} package built for period analysis of TESS (and Kepler) data of variable stars.
+There were interesting similarities and differences between the packages.
+\texttt{Lightkurve} was easy to use but did not allow for fine-tuning of the periodogram, or give any model of the best-fitting curve it produced.
+\texttt{Lightkurve} is based on the \texttt{Astropy} methods, but does not give all the functionality, instead opting to simplify the process.
+
 \begin{figure}[!t]
   \centering
-  \includegraphics[width=\textwidth]{../OzData/PeriodogramUlyssesUpdated.pdf}
-  \caption[Periodogram example]{\textit{Upper Panel}:The Periodogram of the COM flux of an asteroid, with the power of each frequency displayed.
+  \includegraphics[width=\textwidth]{./Figures/PeriodogramUlyssesResid.pdf}
+  \caption[Periodogram example]{\textit{Upper Panel}:The Periodogram of the COM flux of (5254) Ulysses.
     The maximum power of the periodogram is highlighted with the gold star.
-    This maximum is the calculated rotation frequency.
     The known frequency, from the LCDB \citep{Warner2009}, is shown in the blue dotted lines, and a factor of 2 alias each way is shown in the purple dotted lines ($2*$) and red dotted lines ($0.5*$).
-    \textit{Lower Panel}: The COM flux lightcurve and the returned model (dashed black curve) of the lightcurve from the periodogram analysis.
+    \textit{Middle Panel}: The COM flux lightcurve and the returned model (dashed black curve) of the lightcurve from the periodogram analysis.
     The frequency of this model is the same as the gold stared point in the upper panel.
+    \textit{Lower Panel}: The residuals of the model fit.
   }
@@
 
-An example of such a periodogram and the corresponding fit to the lightcurve is given in \autoref{Fig:PeriodEx}.
+An example of such an \texttt{Astropy} Lomb-Scargle periodogram and the corresponding fit to the lightcurve is given \autoref{Fig:PeriodEx}, along with the fit residuals.
+The periodogram is limited below at a power of $10^{-3}$ to highlight only the regions with large power. 
 The periodogram calculated that this object had a frequency double that listed in the LCDB, as seen with the vertical dotted lines.
@@
 This asteroid was highlighted as an alias of the known frequency is recovered by this analysis.
-This is promising for those asteroids where no known frequency exists.
+To account for this, the detected periods will be doubled for comparison to published works.
 
-The stared peak in \autoref{Fig:PeriodEx} is by far the most prominent peak in the periodogram, as it is in log scale.
+The stared peak in \autoref{Fig:PeriodEx} is  the most prominent peak in the periodogram, more that is seems at first glance due to the $\log$ scale.
 Not all the asteroids have such well-defined maximums, and the peak power of the periodogram is an important factor in its quality.
@@
 
-The lower panel of \autoref{Fig:PeriodEx} shows the model given in \autoref{Eq:LCModel} fit to the COM flux lightcurve by \texttt{Astropy}.
+The middle panel of \autoref{Fig:PeriodEx} shows the model given in \autoref{Eq:LCModel} fit to the COM flux lightcurve by \texttt{Astropy}.
 This looks to be a good fit, with the peaks and troughs lining up.
@@
 This is measured in magnitudes, and as such is found from the standard
-
 \begin{equation} \label{Eq:DelMag}
@@
 \end{equation}
-
 with $\Delta m$ being the variation amplitude.
@@
 
+The lower panel in \autoref{Fig:PeriodEx} shows the residuals of the model subtracted from the data.
+For the majority of the observations, the residuals are close to the 0 line, indicating a high quality fit.
+There are places where there is a significant deviation from 0, corresponding to places in the lightcurve that look out of place to the eye. 
+These large residuals could be removed by decreasing the tollerance on the sigma clipping, but that could cut true signal in lightcurves with a large variation amplitude.
+
 \subsection{Periodogram Quality Checks}\label{SubSec:QualCheck}
 
-For accurate statistics, accurate data is needed.
+For accurate statistics, reliable data is needed.
 So there must be checks on the quality of the values calculated.
-These have mostly been discussed already, but for completeness I list them here.
-A lightcurve must have an mean flux of at least 10 counts to be reliable.
+Some of these have been discussed already, but for completeness I list again here.
+A lightcurve must have a mean flux of at least 10 counts to be reliable.
 For an accurate periodogram, it is required that the lightcurve also contain more than 200 observations, which is very easy for TESS and its \qty{10}{\minute} cadence, as the asteroid only has to be in the sector for \qty{33.3}{\hour}.
-The periodogram itself must have a peak power of at least 0.1 to be distingused from the noise
+The periodogram itself must have a peak power of at least 0.1 to be distinguished from the noise
 The calculated period must be $\qty{1}{\hour}\leq P \leq \qty{367.2}{\hour}$, where the upper bound is \qty{90}{\percent} of the \qty{17}{\day} window suggested by \cite{McNeill2023}.
@@
 Here I present the results for TESS sector 29.
-% This sector was chosen because it is early in the \qty{10}{\min} cadence of TESS FFIs, and had been fully run through TESSELLATE.
 Similar analysis can be preformed, and the same figures made, for any of the sectors.
 The computation time was a limiting factor in getting more sectors processed.
+There were 5664 objects detected in the sector with $V\leq \qty{20}{\mag}$, of which 374 passed all the quality checks imposed on their lightcurve and periodogram.
 
-\subsection{Quality Check Verification}
+\subsection{Quality Check Verification}\label{SubSec:QualCheckVer}
 
@@
   \centering
-  \includegraphics[width=\textwidth]{../OzData/LCBDcompLogScaleLineText1.pdf}
-  \caption[Known against calculated period]{A comparison of the calculated periods of the objects in the Sector to their known periods in the LCDB.
+  \includegraphics[width=\textwidth]{./Figures/LCBDcompFinal.pdf}
+  \caption[LCDB known period against detected period]{A comparison of the detected periods of the objects in the sector to their known periods in the LCDB \citep{Warner2009}, coloured by the mean flux of the object.
+  Crosses failed the quality checks, while circles passed them.
+  Names of asteroids are referring to the data point to the right of the text.
     Dotted lines of a constant relationship are plotted, coloured the same as \autoref{Fig:PeriodEx}.
@@
 
-While some of the quality checks were chosen to agree with the past work of others, some were chosen to make sure the derived periods were in agreement with the known ones from the LCDB.
+While some quality checks were chosen to agree with the past work of others, some were chosen to make sure the derived periods were in agreement with the known ones from the LCDB.
 A comparison between objects in both the sector and the LCDB is seen in \autoref{Fig:LCDBcomp}.
-The $y=2x$ line is of interest here, the known periods being twice the detected period.
-This line corresponds to the calculated frequency being at the double alias of the known frequency, as frequency and period are inveses of each other.
-As seen in \autoref{Fig:PeriodEx} and expected from \citet{McNeill2023}, this line is what the periodogram should return due to double-peaked lightcurves.
+Of the 121 objects in both samples, 59 of them passed all the checks.
+These are included in the total 374 asteroids that pass  all the quality checks in sector 29.
+This means over \qty{80}{\percent} of the asteroids that pass all the checks do not have known periods.
+
+\autoref{Fig:LCDBcomp} is coloured by the flux of the asteroid to give an idea of what quality check the object failed. 
+If a cross is bright yellow, the periodogram checks would have been responsible, as the average flux is well above the necessary 10 counts.
+The one bright asteroid that failed the checks in the top left is right on the boundary of the \qty{1}{\hour} lower limit, indicating that this bound may be slightly too lenient.
+The vertical line of points on the right-hand side of \autoref{Fig:LCDBcomp} is on that maximum of the window of searched periods at $\qty{408}{\hour}$, and demonstrates why the \qty{90}{\percent} upper bound is necessary.
+Many asteroids, across an order of magnitude of known periods and across the range of mean fluxes, have detected periods at the edge of the window, showing that the periodogram will return this value when no signal is present.
+
+The largest asteroid, (1) Ceres, is highlighted in \autoref{Fig:LCDBcomp}. 
+It is not recovered, as it is almost too bright for TESS, and the flux bleeds into too many pixels for a rotation period to be accurately obtained from the \qty{1.5}{\px} aperture used.
+An exception to the methods is not made here, no reanalysis is attempted.
+This object is the main reason for the arrow on the colour bar indicating higher mean flux values without a change in colour.
+(1) Ceres has a mean flux is over 100000 counts which, if included, completely washes out the gradient.
+200 counts was chosen as a replacement value because of the range of detail that can be identified.
+
+The brightness of those asteroids that passed the checks is also useful to see.
+If a circle is dark blue, it may only just be above the minimum requirements for average flux.
+The bright circles in the lower right of \autoref{Fig:LCDBcomp} are cause for concern, as these are bright asteroids that are passing the checks with an incorrect period. 
+A range of brightness pass the checks, indicating sensitivity to a wide range of distances and sizes.
+These are the two competing effects that determine the average flux of an object, those that are further away need to be larger to reflect the same amount of light to the detector.
+
+The $y=2x$ line of \autoref{Fig:LCDBcomp} is of interest here, the known periods being twice the detected period.
+This line corresponds to the calculated frequency being an alias of double the known frequency, as frequency and period are inverses of each other.
+This is expected \citep{McNeill2023}, this line is what the periodogram should return due to double-peaked lightcurves.
+As seen in \autoref{Fig:PeriodEx}, (5254) Ulysses falls on this line, highlighted by text in \autoref{Fig:LCDBcomp}.
 Most of the asteroids the pass the cut line along or near this key relation.
 Few asteroids on the line fail the checks, which is also reassuring that the checks do what they are supposed to and filter out incorrect periods.
-The vertical line of grey points on the right-hand side of \autoref{Fig:LCDBcomp} is on that maximum of the window of searched periods at $\qty{408}{\hour}$, and demonstrates why the \qty{90}{\percent} upper bound is necessary.
-Many asteroids, across an order of magnitude of know periods, have detected periods at the edge of the window which shows that the periodogram will return this value when no signal is present.
+
 There are about a dozen asteroids that pass the checks with large detected periods when compared to their known period. 
-These green points in the lower right of \autoref{Fig:LCDBcomp} speak to the checks not being perfect.
-Some experimentation was done on the limiting values, and the ones presented here achieved the highest recovery of asteroid periods, while minimising the number that were incorrectly recovered.
+These points in the lower right of \autoref{Fig:LCDBcomp} imply the checks are not perfect.
+Some experimentation was done on the limiting values, and the ones presented here achieved the highest recovery of accurate asteroid periods, while minimising the number that were incorrectly recovered.
 
-The idea of doubleing the detected period of all the asteroids and calling that number the rotation period has had some doubt cast on it, as there are a couple of points recovered nicely on the $y=x$ line, indicating that the true period is found.
+The idea of doubling the detected period of all the asteroids and calling that number the rotation period has had some doubt cast on it, as there are a couple of points recovered nicely on the $y=x$ line, indicating that the true period is found.
 Either the LCDB has not accounted for the alias, which is unlikely as I only took the periods with high quality codes as they suggest \citep{Warner2009}, or the rotation signal sometimes returns the true period.
 The latter implies the lightcurve is not double-peaked, as is normally expected.
-On the whole, there are enough periods that pass the checks with incorrect periods that these extra erronious periods will not significantly change the fraction of reliable detections.
+On the whole, there are enough periods that pass the checks with incorrect periods that these extra erroneous periods will not significantly change the fraction of reliable detections.
 
@@
 
-The orbital element properties of the asteroids were returned by JPL Horizons queries.
-A plot of the semi-major axis against the other orbital elements, eccentricity and inclination, is shown in \autoref{Fig:AEI}.
+A plot of the semi-major axis against the other proper elements, eccentricity and inclination, is shown in \autoref{Fig:AEI}.
 Looking at the data as a whole, all objects in the field of the instrument during the sector, with $V \leq \qty{20}{\mag}$, are plotted.
-The exception being centaurs, which have an $a$ too larger to see the detail of the majority of the objects and comets, which have an undefined $a$.
+The exception being centaurs, which have an $a$ too larger for the detail in the distribution of the majority of the objects to be seen, and comets, which have an undefined $a$.
+The 374 objects that passed all the quality checks have a similar distribution of orbital parameters to the total sample of 5664 small bodies in the sector. 
+This shows just how beneficial TESS is, and why its full sky coverage is so important to solarsystem science.
+
 The distinction between the resonances in the main belt can be seen.
@@
 Many objects have failed the quality checks on their lightcurve and periodogram, but the ones that pass have a similar distribution in orbital parameters to the sample as a whole.
-In total, 309 objects passed all the quality checks, out of a total sample of 5664 small bodies in the sector.
 
@@
   \centering
-  \includegraphics[width=\textwidth]{../OzData/AEIplotqualCut.pdf}
+  \includegraphics[width=\textwidth]{./Figures/AEIplotqualCut.pdf}
   \caption[aei distribution]{
-    The $a$ against both $e$ and $i$ of all the objects that \texttt{SkyBot} found in Sector 29 with $V <\qty{20}{\mag}$(Other than the comets and centaurs with poorly defined $a$ values).
-    The grey crosses are those objects that did not pass the quality checks on the lightcurve and periodogram, while the green is those that did.
-    The values of these proper elements were queried from JPL Horizons on 30/09/24.
+    The $a$ against both $e$ and $i$ of the objects that \texttt{SkyBoT} found in Sector 29 with $V <\qty{20}{\mag}$.
+    The orange crosses are those objects that did not pass the quality checks on the lightcurve and periodogram, while the green is those that did.
+    The values of these proper elements were queried from \texttt{JPL Horizons} on 30/09/24.
   }
@@
 There are some objects that are not asteroids, 20 comets were in the field and 1 centaur.
-While they have a V mag less than 20, these objects are often not bright enough to have lightcurves with high enough mean fluxes to be detected well by TESS, as seen by the green bars.
-Around \qty{5}{\percent} of any class have lightcurves and periodograms that pass the quality checks.
-The highest fraction is the $\frac29$ of the Amor family of NEAs, however with only 9 total objects the sample is too small to be representative of the detectability of this class as a whole.
+While they have a V mag less than 20, these objects are often not bright enough to have lightcurves with high enough mean fluxes to be detected well by TESS.
+As seen by the green bars in \autoref{Fig:NumPerClass}, around \qty{5}{\percent} of any class have lightcurves and periodograms that pass the quality checks, and this is relatively class independent.
+The highest fraction is the detection of 1 object in the Amor family of NEAs.
+However, with only 9 total objects in the dataset, the sample is too small to make claims on the detectability of this class as a whole.
 
@@
   \centering
-  \includegraphics[width=\textwidth]{../OzData/classesBarqualCut.pdf}
-  \caption[Class distribution]{A bar chart distribution of the total number of objects per class that \texttt{SkyBot} found in the Sector (blue) and the same distribution of those that passed all the quality checks (green).
+  \includegraphics[width=\textwidth]{./Figures/classesBarqualCut.pdf}
+  \caption[Class distribution]{A bar chart distribution of the total number of objects per class that \texttt{SkyBoT} found in the Sector (orange) and the same distribution of those that passed all the quality checks (green).
     The numbers above each class (coloured the same) correspond to the height of the bar, as some classes have too few objects to be seen.
@@
 
-\subsection{Detectability}
+\subsection{Asteroid Detectability in TESS} \label{SubSec:Detect}
 
-\begin{figure} %? // TODO fix.
+\begin{figure}
   \centering
-  \includegraphics[width=\textwidth]{../OzData/pointsMatchesNumberHistdouble1.pdf}
+  \includegraphics[width=\textwidth]{./Figures/pointsMatchesNumberHistdoubleChangedbound.pdf}
   \caption[Number of observations or matches]{\textit{Left}: A histogram of the number of observations in the lightcurve of each asteroid (orange solid line) in 100 observation bins.
@@
 The matches have a very different distribution to the observations.
-While both plots are limited to 600 asteroids on the y-axis, the matches histogram far exceeds this in the first bin, $\leq$ 100 matches, as indicated by the text and arrow.
-Yet the y-limit is too large for detail to be seen in any higher bin.
-There are only 102 asteroids matched more than 100 times in the \texttt{TESSELLATE} detections.
+While both plots are limited to 500 asteroids on the y-axis, the matches' histogram in the right plot far exceeds this in the first bin, $\leq$ 100 matches, as indicated by the text and arrow.
+Yet the y-limit is too large for detail to be seen in any higher bin of this plot.
+There are only a total of 101 asteroids, out of the 5664 in the sector, that are matched more than 100 times to \texttt{TESSELLATE} detections.
+So using the matches as lightcurves for periodogram analysis would mean most fail the minimum observations check, which limits the sample before any periods are calculated.
 
@@
   \centering
-  \includegraphics[width=\textwidth]{../OzData/recoverdHistBkgLimof10AtanBothLegendFixed.pdf}
-  \caption[Recovery plot]{The recovery fraction against the known visual magnitude of each object against the whether the
-    object exceeds only the averge flux requirement  of the quality checks (blue triangles) or passes all of the quality checks on the lightcurve and periodogram (green dots).
+  \includegraphics[width=\textwidth]{./Figures/recoverdHistBkgLimof10AtanBothLegendFixed.pdf}
+  \caption[Visual Magnitude Recovery]{The recovery fraction against the known visual magnitude of each object in \qty{0.25}{\mag} bins.
+  Coloured by whether the object exceeds only the mean flux requirement of the quality checks (blue triangles) or passes all the quality checks on the lightcurve and periodogram (green circles).
     A fit of an $\arctan$ function is provided in dotted lines of the same colour.
@@
 
-The fits of to the fraction recoverd in \autoref{Fig:RecovPassorFail} take the form
+It is clear for \autoref{Fig:RecovPassorFail} that the mean flux check prohibits most of the asteroids dimmer than \nth{18} \unit{\mag}  from being detected. 
+The other checks cut down in the middle of the brightness range of objects rather evenly.
+The brightest bins, which only have a handful of asteroids in them are not reliable data points, as the values they can take on are too discrete.
+(1) Ceres has $V<\qty{8}{\mag}$ and is not recovered, as discussed in \autoref{SubSec:QualCheckVer}, and is the only object in this magnitude bin, so a 0 recovery fraction is returned.
+
+The fits of to the fraction recovered in \autoref{Fig:RecovPassorFail} take the form
 \begin{equation} \label{Eq:Atan}
@@
 \end{equation}
-where $y$ is the recoved fraction, $\alpha$ is how the steep the transition from 1 down to 0 is,  $m$ the magnitude and $m_{1/2}$ is the shift along the magnitude axis.
-The constant of 0.5 in \autoref{Eq:Atan} raises the point of inflection to be at $\frac 1/2$, and the $\frac{-1}{\pi}$ makes the function take the expected maximum at bright magnitudes at \qty{100}{\percent} recoved, down to no dim asteroids being detected.
-The $m_{1/2}$ parameter dictates the limiting magnitude, as dimmer than this value less than half of the asteroids are recoverd.
-The fit is preformed on all bins (of \qty{0.5}{\mag}) dimmer than the first bin with more than one asteroid in it, to stop extremely bright asteroids from having too much influence.
+where $y$ is the recovered fraction, $\alpha$ is how the steep the transition from 1 down to 0 is,  $m$ the magnitude and $m_{1/2}$ is the shift along the magnitude axis.
+The constant of 0.5 in \autoref{Eq:Atan} raises the point of inflection to be at $\frac 1/2$, and the $\frac{-1}{\pi}$ makes the function take the expected maximum at bright magnitudes at \qty{100}{\percent} recovered, down to no dim asteroids being detected.
+The $m_{1/2}$ parameter dictates the limiting magnitude, as dimmer than this value less than half of the asteroids are recovered.
+The fit is preformed on all bins (of \qty{0.5}{\mag}) dimmer than the first bin with more than one asteroid in it, to stop the discrete values from extremely bright asteroids from having too much influence.
 
-Only taking the minimum average flux of 10 counts requirement, blue in \autoref{Fig:RecovPassorFail}, the limiting magnitude is \qty{17.78(0.01)}{\mag}.
+There is a significant change in $m_{1/2}$ between the recovery curves in \autoref{Fig:RecovPassorFail}.
+When only the minimum average flux of 10 counts requirement is applied, the limiting magnitude is \qty{17.78(0.01)}{\mag}.
 When all the quality checks are imposed, this gets over a \unit{\mag} brighter, to \qty{16.46(0.37)}{\mag}.
-It is expected that a higher threshold for quality will make the limiting magnitude brighter, but is value is so uncertain because of the scatter in the recovery fraction at brighter magnitudes, and the shallower decrease in efficiency.
+It is expected that a higher threshold for quality will make the limiting magnitude brighter.
+The $m_{1/2}$ for the full suite of quality checks is rather  uncertain because of the scatter in the recovery fraction at brighter magnitudes, and the shallower decrease in efficiency.
 
-\subsection{Period Analysis Results}\label{SubSec:PerRes}
+\subsection{Rotation Period Analysis Results}\label{SubSec:PerRes}
 
-%TODO amp vs period
+For the 374 asteroids that pass the quality checks, relable rotation periods are obtained.
+\autoref{Fig:PerDistr} shows the distribution of these rotation rates. 
+The period plotted is double the detected period, to account for the double peaked nature of most lightcurves.
+The mode of the distribution is a period less that \qty{1}{\day}, by a factor of 10 over periods in the \qtyrange{1}{2}{\day} bin, the \nth{2} most populated.
 
@@
   \centering
-  \includegraphics[width = \textwidth]{../OzData/QualPervsAmp1.pdf}
-  \caption[Amplitude of Variation against Rotaion Period]{The amplitude of variation in the model from the period analysis against the calculated rotation period.
-    This is for all 309 %!CHECK
+  \includegraphics[width=\textwidth]{./Figures/perDistr.pdf}
+  \caption[Period Distribution]{The distribution of rotation periods in \qty{24}{\hour} (\qty{1}{\day}) bins for all the asteroids that passed the quality checks.}
+  \label{Fig:PerDistr}
+\end{figure}
+
+The distribution of the amplitude of the lightcurves is given in \autoref{Fig:ampDistr}.
+The peak of the distrabution at $\frac13$ \unit{\mag} and is right skewed.
+The high $\Delta m$ values are from lightcurves of dubious quality. 
+Manually checking the analysis, it is clear that these points above $\Delta m =2$ have all been affected by background contamination, and have untrustworth photometry.
+Some examples are given in the appendix for asteroids with known periods.
+
+It is unclear how to avoid such contamination.
+For some, analysis of the residuals could show a pattern that leads to a period and possible fit of a model, but this is not the case for all of them.
+For only one sector worth of data, it was managable for me to preform a manual check on these few outliers
+But this will not be the case for all the whole dataset \texttt{TESSELLATE} will analyse.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{./Figures/delMagDistr.pdf}
+  \caption[Amplitude Variation Distribution]{The distribution of amplitude variation for all the asteroids that passed the quality checks. 
+  The data are binned in steps of $\frac{1}{8}$ \unit{\mag}.
+  }
+  \label{Fig:ampDistr}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width = \textwidth]{./Figures/QualPervsAmp1I.pdf}
+  \caption[Amplitude of Variation against Rotation Period]{The amplitude of variation in the model from the period analysis against the calculated rotation period.
+    This is for all 374 %!CHECK
     objects that pass the quality checks.
@@
 
-%TODO redo fig 1??
+%TODO redo fig 1
 
-%! IF Matching works both to M23 and LCDB !!!
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{./Figures/Diam-FreqPlotThisWork.pdf}
+  \caption{This work's addition to \autoref{Fig:FreqVsDiam} and the LCDB of \citet{Warner2009}. The diameters are from NEOWISE where available.}
+  \label{Fig:FreqDiamUpdate}
+\end{figure}
+
+\autoref{Fig:FreqDiamUpdate} shows how my rotation periods fit with those in the LCDB of \cite{Warner2009} from \autoref{Fig:FreqVsDiam}.
+It is clear that TESS struggles to determine periods for asteroids with a diameter $\lesssim \qty{1}{\kilo \metre}$. 
+This is most likely due to these bodies being too dim for the telescope.  
+The spin barrier of \citet{Pravec2000} is not violated by any of the periods I have calculated.
+
+269 of the asteroids that pass the quality checks have diameters in the near-Earth object wide-field infrared survey explorer (NEOWISE) V2.0 database \citep{Masiero2011, NEOWISE2019}.
+It is only this sub-sample of the 374 asteroids that are plotted in \autoref{Fig:FreqDiamUpdate}.
+The diameters in the LCBD are not calculated using NEOWISE data, so there could be discrepancies in those asteroids that I have calculated periods for that are also in the LCDB. 
+The NEOWISE data is self-consistent and encompasses more asteroids that the diameters from just the LCDB, so it was preferred for the over plotting of the rotation data from this work.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{./Figures/M23andP20Comp.pdf}
+  \caption[Comparison to periods from other work]{Comparison of the rotation periods I have calculated to the publicly available data from \citet{McNeill2023} (left, pink crosses) and \citet{Pal2020} (right, purple pluses).
+The asteroids that are in all three are highlighted with black rings around the symbols.}
+\end{figure}
+
+18 in \citet{McNeill2023}
+13 in \citet{Pal2020}
 
 The limiting magnitude of the recovery fraction demonstrated in \autoref{Fig:RecovPassorFail} agrees well with other tests of \texttt{TESSELLATE}.
-Concurrent work by another Honours student \citep{MontillaHons} achieves a similar recovery fraction detection efficiency of $\sim$\nth{17} \unit{\mag}.
-She has been preforming injection/recovery tests across many sectors, and finding the limiting magnitude is consistent across sectors and cadence of the FFIs.
+Concurrent work by another Honours' student \citep{MontillaHons} achieves a similar recovery fraction detection efficiency of $\sim$\nth{17} \unit{\mag}.
+They have been preforming injection/recovery tests across many TESS sectors, and they find that the limiting magnitude is consistent across sectors
+The cadence of the FFIs also seems to have no effect on the depth of the images.
 This is promising, as it means that asteroids will be detected to a similar depth as they are here for any further sectors processed, something I could not check myself.
@@
 The periodogram noise at the low frequency end is quite strong throughout the range of rotation periods.
-Currently it is responsible for a lot of undetected rotaion rates, due to the \qty{90}{\percent} of the \qty{17}{\day} window check, and this noise region having a larger power than any signal.
-Modelling this excess to try and remove the noise and boost the signal from the rotaion could be done to improve the maximum power of the periodogram.
-There are a few asteroids with known longer rotaion periods than this check allows for, but one TESS sector will not be enough to determine those accurately.
+Currently, it is responsible for a lot of undetected rotation rates, due to the \qty{90}{\percent} of the \qty{17}{\day} window check, and this noise region having a larger power than any signal.
+Modelling this excess to try and remove the noise and boost the signal from the rotation could be done to improve the maximum power of the periodogram.
+There are a few asteroids with known longer rotation periods than this check allows for, but one TESS sector will not be enough to determine those accurately.
 
 Linking lightcurves of asteroids between sectors was not considered here, as only one sector has been analysed.
-Any one asteroid will be in multiple sectors, as TESS has surveyed that whole sky more than once.
+An asteroid will be in multiple sectors, as TESS has surveyed that whole sky more than once.
 The different FFI cadence, as well as the large temporal gap between these observations makes it non-trivial to recover the rotation periods.
-%?? The different depths that the shortest cadence FFIs have \citep{MontillaHons}
+One could calculate the rotation period in each sector, and then average.
+This technique once again biases against long rotation periods, which combining sectors should help with, not hinder.
+
+Lomb-Scargle periodograms can deal with large gaps in the time data, but would require the flux values to have the same normilisation.
+\texttt{TESSELLATE} is self consistent in the data reduction for a sector, but comparing fluxes between sectors may have some difficulties.
+Assuming the mean flux of the asteroid should be constant, subtracting the mean and
+
+There are also phase and viewing angle effects on the lightcurve that are important to correct for with large time gaps. 
+These were not considered for the single sector analysis because
 
@@
 
-This work has used data and/or services provided by NASA's Solar System Dynamics website, \url{https://ssd.jpl.nasa.gov/}. Specifically JPL Horizons: \url{https://ssd.jpl.nasa.gov/horizons/}.
+This work has used data and/or services provided by NASA's Solar System Dynamics website, \url{https://ssd.jpl.nasa.gov/}. Specifically \texttt{JPL Horizons}: \url{https://ssd.jpl.nasa.gov/horizons/}.
+
+The figures seen in this work have been plotted by \texttt{Matplotlib} \citep{Hunter2007}, specifically Version 3.9.2 \citep{PLT3.9.2}
 

### 2024-10-19 Test Code/findAndInterp.py

@@ -605,9 +605,9 @@ for name in unqNames:
     fig, ax = plt.subplots(1, figsize=(10, 8))
     # ax.set_title(name)
     ax.set_xlabel("$\Delta$RA [$''$]")
-    ax.set_xlim((-limAngle, limAngle))
+    # ax.set_xlim((-limAngle, limAngle))
     ax.set_ylabel("$\Delta$Dec [$''$]")
-    ax.set_ylim((-limAngle, limAngle))
+    # ax.set_ylim((-limAngle, limAngle))
     # TODO axis with 'TESS px units'
 
     ax2 = ax.secondary_xaxis("top", functions=(arcstopx, pxtoarcs))
@@ -625,4 +625,4 @@ for name in unqNames:
     clb.ax.ticklabel_format(useOffset=False)
 
     ax.scatter(deltaRa, deltaDec, c=nameCut["MJD"], cmap=cmap, norm=cnorm, s=5)
-    fig.savefig(f"./Testing Figures/{name}PosCheck.pdf")
+    # fig.savefig(f"./Testing Figures/{name}PosCheck.pdf")

### 2024-10-19 FinalReport/FinalReportDraft.tex
comp Table made

@@
 \usepackage{hyperref,graphicx}
+\usepackage{datatool,booktabs,pgfplotstable}
 \usepackage[separate-uncertainty=true,multi-part-units=single]{siunitx}
@@
 \makeatother
-
-% \usepackage[margin=1.5cm, left=3cm ,twoside]{geometry}
 
@@
 
-% \title{ASTR480 Final Report:\\
-% \textit{Asteroids in TESS}
-% }
-% \author{Brayden Leicester \\ 
-% \textit{61352063} \\
-% [1ex]\small{Supervisors: Michele Bannister and Ryan Ridden-Harper}
-% }
+\pgfplotsset{compat=1.18}
+\renewcommand*\dtldisplaystarttab{\toprule}
+\renewcommand*\dtldisplayendtab{\tabularnewline\bottomrule}
+\renewcommand*\dtldisplayafterhead{\midrule}
+\pgfplotstableset{
+    every head row/.style={before row=\toprule, after row=\midrule},
+    every last row/.style={after row=\bottomrule}
+}
+
 \begin{document}
@@
 Detections in \texttt{TESSELLATE} are done on a per pixel basis, and asteroids only cross each pixel for a short number of frames.
-The residuals of \autoref{Fig:RADecMatch} are all within a TESS pixel, shown on the right hand axis of each of the lower pannels.
+The residuals of \autoref{Fig:RADecMatch} are all within a TESS pixel, shown on the right hand axis of each of the lower panels.
 
@@
 
-For the 374 asteroids that pass the quality checks, relable rotation periods are obtained.
-\autoref{Fig:PerDistr} shows the distribution of these rotation rates. 
-The period plotted is double the detected period, to account for the double peaked nature of most lightcurves.
-The mode of the distribution is a period less that \qty{1}{\day}, by a factor of 10 over periods in the \qtyrange{1}{2}{\day} bin, the \nth{2} most populated.
-
-\begin{figure}
+\begin{figure}[h]
   \centering
@@
 
-The distribution of the amplitude of the lightcurves is given in \autoref{Fig:ampDistr}.
-The peak of the distrabution at $\frac13$ \unit{\mag} and is right skewed.
-The high $\Delta m$ values are from lightcurves of dubious quality. 
-Manually checking the analysis, it is clear that these points above $\Delta m =2$ have all been affected by background contamination, and have untrustworth photometry.
-Some examples are given in the appendix for asteroids with known periods.
+For the 374 asteroids that pass the quality checks, relable rotation periods are obtained.
+\autoref{Fig:PerDistr} shows the distribution of these rotation rates.
+The period plotted is double the detected period, to account for the double peaked nature of most lightcurves.
+The mode of the distribution is a period less that \qty{1}{\day}, by a factor of 10 over periods in the \qtyrange{1}{2}{\day} bin, the \nth{2} most populated.
 
-It is unclear how to avoid such contamination.
-For some, analysis of the residuals could show a pattern that leads to a period and possible fit of a model, but this is not the case for all of them.
-For only one sector worth of data, it was managable for me to preform a manual check on these few outliers
-But this will not be the case for all the whole dataset \texttt{TESSELLATE} will analyse.
-
-\begin{figure}
+\begin{figure}[!hb]
   \centering
@@
 
-\begin{figure}
+The distribution of the amplitude of the lightcurves is given in \autoref{Fig:ampDistr}.
+The peak of the distribution at $\frac13$ \unit{\mag} and is right skewed.
+The high $\Delta m$ values are from lightcurves of dubious quality.
+Manually checking the analysis, it is clear that these points above $\Delta m =2$ have all been affected by background contamination, and have untrustworth photometry.
+Some examples are given in the appendix for asteroids with known periods.
+
+It is unclear how to avoid such contamination.
+For some objects, analysis of the residuals could show a pattern that leads to a period and possible fit of a model, but this is not the case for all of them.
+For only one sector worth of data, it was managable for me to preform a manual check on these few outliers
+But this will not be the case for all the whole dataset \texttt{TESSELLATE} will analyse.
+
+\begin{figure}[h]
   \centering
@@
 
+Comparing the rotation period to the $\Delta m$ is insightful and is shown in \autoref{Fig:perVSamp}.
+All 374 of the objects that pass the quality checks are plotted here, and no trend is observed.
+The variation is quite constant accros the range of rotation periods, although there are less objects at longer periods.
+
+The modes of each of the distributions in \autoref{Fig:PerDistr} and \autoref{Fig:ampDistr} do not fully overlap, as indicated by \autoref{Fig:perVSamp}.
+Those asteroids with rotation periods $<\qty{24}{\hour}$ have a range of rotation periods.
+Converly, the objects with a $\Delta m \approx 0.3$ span the entire period range.
+
+The position in these axes of \omuamua is shown in \autoref{Fig:perVSamp} by the black caltrop and is highlighted with text.
+1I's rotation period of \qty{8.67(0.34)}{\hour} from \citet{Belton2018} and a $\Delta m$ of \qty{2.5}{\mag} from \citet{Meech2017} are used.
+There are some asteroids with a larger $\Delta m$ but, as discussed, these are not trustworthy.
+
 \section{Discussion}\label{Sec:Disc}
 
-\subsection{Comparison to other work }%\citet{McNeill2023}}
-
-%TODO redo fig 1
+\subsection{Comparison to Literature Data}\label{SubSec:LitComp}
 
@@
 
-\autoref{Fig:FreqDiamUpdate} shows how my rotation periods fit with those in the LCDB of \cite{Warner2009} from \autoref{Fig:FreqVsDiam}.
-It is clear that TESS struggles to determine periods for asteroids with a diameter $\lesssim \qty{1}{\kilo \metre}$. 
-This is most likely due to these bodies being too dim for the telescope.  
-The spin barrier of \citet{Pravec2000} is not violated by any of the periods I have calculated.
-
+\autoref{Fig:FreqDiamUpdate} displays my calculated rotation periods with those in the LCDB of \citet{Warner2009} (\autoref{Fig:FreqVsDiam}).
 269 of the asteroids that pass the quality checks have diameters in the near-Earth object wide-field infrared survey explorer (NEOWISE) V2.0 database \citep{Masiero2011, NEOWISE2019}.
@@
 
-\begin{figure}
+It is clear that TESS struggles to determine periods for asteroids with a diameter $\lesssim \qty{1}{\kilo \metre}$.
+This is most likely due to these bodies being too dim for the telescope.
+Those that are seen must not pass the quality checks I have imposed.
+
+The spin barrier of \citet{Pravec2000} is not violated by any of the periods calculated.
+The quality checks to do exclued this, as the spin barrier is at $\sim \qty{2.4}{\hour}$, and I only impose the period must be $>\qty{1}{\hour}$.
+This cadence of FFI has the ability to probe the forbidden region, to find more objects that violate the spin barrier.
+Unfortunalty, none were found in sector 29.
+
+\begin{figure}[!t]
   \centering
   \includegraphics[width=\textwidth]{./Figures/M23andP20Comp.pdf}
-  \caption[Comparison to periods from other work]{Comparison of the rotation periods I have calculated to the publicly available data from \citet{McNeill2023} (left, pink crosses) and \citet{Pal2020} (right, purple pluses).
-The asteroids that are in all three are highlighted with black rings around the symbols.}
+  \caption[Comparison to TESS periods from the Literature]{Comparison of the rotation periods I have calculated to the publicly available data from \citet{McNeill2023} (left, pink crosses) and \citet{Pal2020} (right, purple pluses).
+The asteroids that are in all three are highlighted with black rings around the symbols.
+    A $y=x$ line is plotted to guide the eye.}
+  \label{Fig:TESSLit}
 \end{figure}
 
-18 in \citet{McNeill2023}
-13 in \citet{Pal2020}
+As mentioned in \autoref{SubSec:Asteroid}, this is not the first work done analysing asteroids in TESS data.
+This work has 18 objects in common with \citet{McNeill2023} and an overlap of 13 asteroids with the data of \citet{Pal2020}, shown in \autoref{Fig:TESSLit}.
+There is agreement in the derivered periods for some of the objects in both datasets, as seen by the points on or near the $y=x$ line in each subplot.
+These datasets are in broad, but not complete, agreement for the asteroids they both measure \citep[see the comparison in ][]{McNeill2023}, so it is not surpising that this work is not in complete agreement.
+
+It does not appear that my calculated periods are biases towards longer or shorter values.
+Both in of these comparison datasets, and in the check against the LCDB in \autoref{Fig:LCDBcomp}, there are periods on either extreme.
+Sometimes the periods found here are much shorter than those found in the external data, and othertimes much longer, but this is in about equal proportions.
+The sample size is really too small to make any statistical statments, once again more sectors need to be analysed.
+
+There are 8 objects that appear in all three datasets.
+Two of them lie almost directly on top of each other in the \citeauthor{Pal2020} data on the right panels\autoref{Fig:TESSLit}, but there is a difference on the left hand side in the \citeauthor{McNeill2023} data.
+This intersection is shown again in \autoref{Tab:AllPers}, where it is more clear how well some of these data line up.
+Both (13364) 1998 UK20 and (16151) 1999 XF230 have the same rotation period measured in all three analysies.
+My period for (36240) 1999 VN44 is two orders of magnitude larger than either of the two previos studies, so there must have been an issue with the data for this object in sector 29.
+
+I seem to calculate similar periods to \citeauthor{Pal2020} more frequently than with \citeauthor{McNeill2023}, but again this is too small of a sample size.
+I have only analysesed a sector of data, they both measure periods for the same 13 sectors. %!Check
+When more data has been run through both \texttt{TESSELLATE} and my pipelines, then the number of intersecting objects will increase.
+
+\begin{table}[!t]
+  \centering
+  \caption[TESS Literature Comparison Table]{The 8 asteroids have rotation periods in this work rounded to two decimal places, as well as \citet{Pal2020} and \citet{McNeill2023}.
+All periods are measured in hours (\unit{\hour})} \label{Tab:AllPers}
+  \pgfplotstabletypeset[
+    col sep=&,
+    string type,
+    columns/a/.style={column name=(MPC designation) Asteroid Name},
+    columns/b/.style={column name=This work},
+    columns/c/.style={column name=\citeauthor{McNeill2023}},
+    columns/d/.style={column name=\citeauthor{Pal2020}}
+  ]{tabDat.dat}
+\end{table}
 

### 2024-10-20 FinalReport/FinalReportDraft.tex
Almost done

@@
 \makeatother
+
+\usepackage{chngcntr}
 
@@
 \newunicodechar{ʻ}{\okina}
-\newcommand{\omuamua}{\okina Oumuamua }
+\newcommand{\omuamua}{\okina Oumuamua } %space needed to get next word spaced 
+\newcommand{\omuamuans}{\okina Oumuamua} %no space after
 
@@
 \student{\textbf{Brayden Leicester}}
-\supervisor{Dr Michele Bannister and Dr Ryan Ridden-Harper}
+\supervisor{Dr Michele Bannister \\ and Dr Ryan Ridden-Harper}
 \ttitle{Asteroids in TESS} %title of the file
@@
 
-Graduate students, X and Y, helped me understand ... \\
+\autoref{Fig:FreqVsDiam} has been recreated after data from the Light Curve Database (LCDB) \citet{Warner2009}, as cited in text.
+That data has been supplemented with my own work in the corresponding \autoref{Fig:FreqDiamUpdate}.\\
 
-The derivation in section 3 was taken from ... \\
+\autoref{Eq:LCModel} was adapted from the \texttt{Astropy} documentation
+\citep{Astropy2022} ( \url{https://docs.astropy.org/en/stable/api/astropy.timeseries.LombScargle.html#astropy.timeseries.LombScargle.model_parameters}) for how the model is fit to the data. \\
 
-\autoref{Fig:FreqVsDiam} has been recreated after data from the Light Curve Database (LCDB) \citet{Warner2009}, as cited in text.
-That data has been supplemented with my own work in the corresponding \autoref{Fig:FreqDiamUpdate}.
+The TESS data was reduced for me by \texttt{TESSreduce} as part of the \texttt{TESSELLATE} pipeline, but the photometry is my own.
+All the analysis of the photometry was also done by me.
 
@@
 
-This is where the abstract goes. \\
-
-This is the next paragraph.
-
 % %*Background
+The Transiting Exoplanet Survey Satellite (TESS) is a wealth of high-cadence photometric data. 
+Asteroids move across the TESS field, and their intrinsic rotation varies their brightness periodically. 
+Here, I present an analysis of the asteroids in TESS sector 29, as an addition to the \texttt{TESSELLATE} pipeline's analysis.
 
 % %*Methods
+After obtaining the positions of the asteroids in \qty{12}{\hour} intervals, they are interpolated to match the cadence of the TESS images (\qty{10}{\min}).
+Aperture photometry of the asteroids is performed to construct a lightcurve.
+The rotation periods and amplitudes are calculated by a Lomb-Scargle periodogram.
+Quality checks are applied to properties of both the lightcurve and the periodogram of each object, to ensure accurate quantities are calculated. 
+Comparing derived values for objects with known periods, it is found that \qty{70}{\percent} of the detected periods are reliable.
 
 % %*Main Results
+5664 asteroids are found in the sector, of which 374 passed all the quality checks.
+\qty{80}{\percent} of these have no known period. 
+The distribution of periods and amplitudes broadly agrees with other analyses of TESS data.
+While some asteroids are calculated to have a large amplitude of variation, manual checking determined this to be an artefact of data contamination.
+It is shown that more than half of the objects with a visual magnitude above \qty{16.36(0.73)}{\mag} have reliable periods detected.
 
 % %*Conclusion
+While only one sector is analysed here, the methods are shown to be robust and can now be run on any TESS sector.
 
@@
 This project aims to find and characterise the light curves of all the asteroids seen by the Transiting Exoplanet Survey Satellite (TESS).
-Combining knowledge of these asteroids positions with the high cadence imaging of TESS should allow for the determination of rotation periods of these small bodies, and the amplitude of the variation.
+Combining knowledge of these asteroids positions with the high-cadence imaging of TESS should allow for the determination of rotation periods of these small bodies, and the amplitude of the variation in their lightcurves.
 
@@
 Asteroids are a key class of solar system objects. Much is known about them already, but there is always more to learn.
-The orbits of asteroids can be well characterised by only a few values, the proper elements of the object.
-These are the lenght of the long axis of elliptical orbit, the semi-major axis, $a$, how elliptical the orbit is, the eccentricity, $e$, and the angle to the ecliptic plane, the inclination,$i$.
+Three values characterise the orbits of asteroids, the proper elements of the object.
+These are the length of the long axis of the orbit, the semi-major axis, $a$, how elliptical the orbit is, the eccentricity, $e$, and the angle to the ecliptic plane, the inclination, $i$.
 Typical values for main belt asteroids are an $a$ of \qtyrange{2}{3.3}{\au}, an $e$ in \qtyrange{0}{0.35}{} and an $i$ around \qtyrange{0}{30}{\degree} \citep{DeMeo2015}.
-Comets have undefined $a$ and larger $e$, with some long period comets having $e\approx 1$, but the typical range is \qtyrange{0.2}{0.7}{} \citep{Lewis2012}.
+Comets have undefined $a$ and larger $e$, with some long-period comets having $e\approx 1$, but the typical range is \qtyrange{0.2}{0.7}{} \citep{Lewis2012}.
 Interstellar objects (ISOs) have $e>1$ as they are on unbound hyperbolic trajectories \citep{McGlynn1989}.
@@
 
-Asteroids can be grouped into families based on clustering of the proper orbital elements and similarities in their colour or spectra \citep{Nesvorny2015}.
+Asteroids can be grouped into families based on  the  clustering of the proper orbital elements and similarities in their colour or spectra \citep{Nesvorny2015}.
 These families are named for their largest member.
@@
 This produces a lightcurve that is often sinusoidal in appearance.
-These lightcurves are normally ``double-peaked'', having two maximums for one full rotation of the body, due to symmetries in the shape of the object.
+These lightcurves are normally ``double-peaked'', with two maximums for one full rotation, due to symmetries in the object's shape. 
 The shape of the asteroid has a large effect on the scale of variation in the lightcurve \citep{Durech2015}, the more elongated the object is, the larger the variation.
-The rate of rotation is limited by the size of the asteroid \citep{Pravec2000}, as the internal strength of the body is put under too much strain if large asteroid rotate with short periods.
+The rate of rotation is limited by the size of the asteroid \citep{Pravec2000}, as the body's internal strength is placed under too much strain if large
+asteroids rotate with short periods.
 
@@
 This is still at the forefront of research, with more lightcurves being published year-on-year \citep{Harris2015}.
-For the smallest objects ingenious observation techniques are employed to measure rotation periods such as intentionally streaking the asteroid down a column of pixel on a detector, and making the lightcurve from the streak, as done by \citet{Bolin2023}.
+For the smallest objects ingenious observation techniques are employed to measure rotation periods such as intentionally streaking the asteroid down a column of pixel on a detector, and making a lightcurve from the streak, as done by \citet{Bolin2023}.
 
 Asteroids can have their rotation properties changed with time, collisions can do this, and so can light from the sun.
-The YORP effect (named for the discoverers Yarkovsky, O'Keefe, Radzievskii, and Paddack) is the process of tourquing asteroids of an asymmetric shape by both bombardment of photons and thermal (re-)emission \citep{Rubincam2000}.
+The YORP effect (named for the discoverers Yarkovsky, O'Keefe, Radzievskii, and Paddack) is the process of torquing asteroids of an asymmetric shape by both bombardment of photons and thermal (re-)emission \citep{Rubincam2000}.
 
-The data on solar system bodies will only increase with more large survey telescopes coming online soon, such as the Legacy Survey of Space and Time \citep[LSST][]{LSST2019}.
+The long cadence data on solar system bodies will only increase with more large survey telescopes coming online soon, such as the Legacy Survey of Space and Time \citep[LSST][]{LSST2019}.
 The number of new asteroids that are well characterised will increase by more than 2 orders of magnitude on the Sloan Digital Sky Survey \citep[SDSS][]{York2000} which had $\sim 88000$ objects \citet{Parker2008}.
-LSST needs a dedicated pipeline for asteroid classification, the Solar System Notification Alert Processing System \citep[SNAPS][]{Trilling2023} has been developed using Zwicky Transient Facility \citep[ZTF, ][]{Bellm2018} data and can scale up handle volume of data coming from LSST.
-Work is being done to understand the biases of how  asymmetric shape high amplitude variations can lead to selection effects of such large sample of asteroids \citet{Levine2023}.
+LSST needs a dedicated pipeline for asteroid classification, the Solar System Notification Alert Processing System \citep[SNAPS][]{Trilling2023} has been developed using Zwicky Transient Facility \citep[ZTF, ][]{Bellm2018} data and can scale up to handle the volume of data from LSST.
+Work is being done to understand the biases of how  asymmetric shape high amplitude variations can lead to selection effects of such a large sample of asteroids \citet{Levine2023}.
 
@@
 This includes other TESS data from \citet{Pal2020} and \citet{Woods2021}.
-The frequencies against diameter plot of these asteroids in \autoref{Fig:FreqVsDiam} shows the collection of all the asteroids in the LCDB with $U\geq 2-$.
-The diameter presented in the LCDB here calculated from the absolute magnitude $H$ of the bodies, assuming an albedo.
-\citeauthor{Warner2009} acknowledge that there are direct measurements of diameter for more asteroids in the present day, but continue to present the $H$ derived diameters for consistency.
+A frequencies and diameter plot of these asteroids in \autoref{Fig:FreqVsDiam} shows the collection of all the asteroids in the LCDB with $U\geq 2-$.
+The diameter presented in the LCDB here is calculated from the absolute magnitude, $H$, of the bodies, assuming an albedo.
+\citeauthor{Warner2009} acknowledge that there are direct measurements of diameter for more asteroids in the present day, but continue to present the $H$-derived diameters for consistency.
 
@@
 It is thought that the YORP effect, as well as collisions, spin up asteroids so fast that they pass above this barrier and then no longer have the internal strength necessary to hold themselves together.
-This implies that all asteroids made to spin faster have torn themselves apart and that they have no tensile strength, i.e they are rubble piles instead of monoliths \citep{Harris1996}.
+This implies that all asteroids made to spin faster have torn themselves apart because they have no tensile strength, i.e. they are rubble piles instead of monoliths \citep{Harris1996}.
 This idea of the cohesive strength of small bodies is still explored today \citep[e.g.][for V-type asteroids]{Oszkiewicz2020}.
 
-Comparing the periods I calculate of asteroids that are already in the LCDB is key to making sure my methods are reliable, and that the periods for asteroids for which no data exists can be trusted.
+Comparing the periods I calculate for asteroids already in the LCDB is an important sanity check of my methods, and ensures the periods calculated for asteroids for which no data exists are reliable.
 
@@
   \includegraphics[width=0.7\textwidth]{./Figures/Diam-FreqPlot.pdf}
-  \caption[Asteroid frequencies vs diameter (LCDB)]{
+  \caption[Asteroid frequencies against diameter (LCDB)]{
     Rotation rate of asteroids by their size, recreated from the  October \nth{1}, 2023 LCDB after \citet{Warner2009}.
@@
 \subsubsection*{Interstellar Objects}
-%Omuamu
 High amplitude variation has come to the forefront of questions about asteroid properties because of the first interstellar object, 1I/\omuamua \citep[see][for a review]{Bannister2019}.
 \omuamua was determined to be spectroscopically red \citep{Fitzsimmons2017, Meech2017}, and having a photometric colour in the neutral end of the solar system range \citep{Bannister2017}.
-1I was classed as asteroid due to lack of a coma, and no noticeable activity.
+1I was classed as an asteroid due to lack of a coma, and no noticeable activity.
 This contrasts the second ISO, 2I/Borisov, which had many cometary characteristics \citep[see ][for a review]{Dorofeeva2023}.
@@
 The peak to peak amplitude variation of this ISO was \qty{2.5}{\mag} \citep{Meech2017} over its double-peaked lightcurve, which is interesting as it is more variation than most asteroids more than most asteroids \citep[as seen in the LCDB of][]{Warner2009}.
-With the full sky survey of bright asteroids, I hope to and to see just how rare \omuamua is, by finding many asteroids with large amplitude variations.
+With the full sky survey of bright asteroids, I aim to see how rare \omuamua is, by finding asteroids with large amplitude variations.
 
@@
 %Why TESS
-TESS is a large area, high cadence imaging, space telescope  \citep{Ricker2014}.
-TESS is tasked with observing one piece of sky for \qty{27}{\day} at a time , which defines a sector.
-The telescope delivers \qtyproduct{96 x 24}{\degree} full frame images (FFIs) at regular intervals.
-These FFIs are built from stacked \qty{2}{\second} exposures, leveraging the short readout times of the 16 charged-coupled devices (CCD) spread over four cameras on-board.
+TESS is a large area, high-cadence imaging, space telescope  \citep{Ricker2014}.
+TESS is tasked with observing one piece of sky for \qty{27}{\day} at a time, which defines a sector.
+The telescope delivers \qtyproduct{96 x 24}{\degree} full frame images (FFIs).
+These FFIs are built from stacked \qty{2}{\second} exposures, leveraging the short readout times of the 16 charged-coupled devices (CCDs) spread over four cameras.
 With the initial cadence for these full frame images set to \qty{30}{\minute}, the time resolution of TESS is unparalleled.
-As the mission was extended, after TESS had already mapped the entire sky, the length of the FFIs has decreased to \qty{10}{\minute} and then later to \qty{200}{\second}.
+As the mission was extended after TESS had already mapped the entire sky, the length of the FFIs decreased to \qty{10}{\minute} and then later to \qty{200}{\second}.
 This high time resolution and observation area does come at the cost of spatial resolution, as the pixels are each \qty{21}{\arcsec} square.
@@
 Of interest here is what such a high sampling rate can do for statistics on the asteroid population.
-For bright asteroids the rotation periods should be able to be easily determined from this vast dataset, due to the high sampling of the lightcurve.
+Rotation periods of bright asteroids can be easily determined from this vast dataset, due to the high sampling of the lightcurve.
 The shortest FFIs will be able to accurately determine the rotation periods of all but the fastest rotating asteroids, most of which will be too dim to see in the TESS data.
 
-The highest frequency (shortest period) that can be found from time series data like this is at the Nyquist limit, twice the minimum time between images.
+The highest frequency (the shortest period) that can be found from time series data is the Nyquist limit, twice the minimum time between images \citep{VanderPlas2018}. 
 Due to the short cadence, the Nyquist frequency of TESS is quite high, \qty{1}{\per\hour}.
-This is well sampled enough to characterise most variable stars, as well as find the orbital periods of exoplanets to a high precision.
-The decreased in the FFI exposure time in later sectors comes with a complementary increase in the Nyquist frequency.
+This is well sampled enough to characterise most variable stars and determine the orbital periods of exoplanets to high precision.
+The decrease in the FFI exposure time in later sectors comes with a complementary increase in the Nyquist frequency.
 
 Previous work has been conducted to find and classify all the asteroids in TESS data.
-\citet{Pal2018} proposes that TESS will be a good instrument for solar system object study, and simulate some detections down to \nth{20} visual magnitude, but that good photometry should only be expected to $V \lesssim 19$, a sentiment shared by \citet{Wong2019}.
-The first data release in \citet{Pal2020} catalogues nearly 10,000 asteroid light curves from the first year of TESS operation, and they report that they triple the number of asteroids with accurate rotation periods.
+\citet{Pal2018} propose that TESS will be a good instrument for solar system object study, a sentiment shared by \citet{Wong2019}.
+\citeauthor{Pal2018} simulate some detections down to \nth{20} visual magnitude and find good photometry should only be expected to $V \lesssim 19$.
+The first data release in \citet{Pal2020} catalogues nearly 10,000 asteroid lightcurves from the first year of TESS operation (cycle 1) with the \qty{30}{\min} cadence, and they report that they triple the number of asteroids with accurate rotation periods.
+The Minor Planet Center (MPC) gets regular updates from TESS because of the work of \citet{Woods2021} and their \texttt{LINEAR-TESS} program, which creates tracklets of objects over a day, and then stitches these together to form a track of each asteroid through an FFI.
+
 \citet{McNeill2023} also analyse TESS cycle 1, and detect almost 38,000 objects.
@@
 In the overlap from the data from \citet{Pal2020}, \citeauthor{McNeill2023} find good agreement between the two sets of periods and amplitudes.
-The Minor Planet Center (MPC) gets regular updates from TESS thanks to the work of \citet{Woods2021} and their \texttt{LINEAR-TESS} program, which creates tracklets of objects over a day, and then stitches these together to form a track of each asteroid through an FFI.
 
 There is ongoing analysis of subsets of the asteroid population in TESS.
-\citet{Gowanlock2024} uses TESS photometry as well as  observations by ZTF to get a longer baseline on mutually observed objects while combining ground and space based observations.
-There were not very many objects in both samples with known periods for comparison, with only 222 objects analysed, however they demonstrate that the technique is effective.
+\citet{Gowanlock2024} use TESS photometry as well as  observations by ZTF to get a longer baseline on mutually observed objects while combining ground and space-based observations.
+There were a limited number of objects in both samples with known periods for comparison, with only 222 objects analysed. However, they demonstrate that the technique is effective.
 Fainter and unknown solar system objects can be found by shift and stacking \citep{Holman2019, Payne2019,Rice2020} or taking a fast X-ray transform \citep{Nguyen2024} of the FFI data.
 
-This work aims to extend these studies to more sectors, hence characterising more asteroids.
+This work aims to extend these other TESS studies to more sectors and examine the higher cadence FFIs, hence characterising more asteroids.
 The goal is to get a volume complete set of asteroid lightcurves and periods, by analysis of the whole sky as seen by TESS using data from more than just the first year of TESS's operation.
 This has not been achieved, as the analysis of only one sector is presented here, but the methods are scalable.
+
+\section{Methods}\label{Sec:Meth}
+
 Because of the survey properties, TESS provides a self-consistent way to measure the properties of asteroids over the full sky.
@@
 
-This work is part of a full sky transient survey using TESS, the TESS Extensive Lightcurve Logging and Analysis of Transient Events4 (\texttt{TESSELLATE}) pipeline \citep{TESSELLATE}.
-To see what \texttt{TESSELLATE} detects and to help progress the pipeline, one can participate in the Cosmic Cataclysim Zooniverse project\footnote{\href{https://www.zooniverse.org/projects/cheerfuluser/cosmic-cataclysms}{https://www.zooniverse.org/projects/cheerfuluser/cosmic-cataclysms}}.
+This work is part of a full sky transient survey using TESS, the TESS Extensive Lightcurve Logging and Analysis of Transient Events (\texttt{TESSELLATE}) pipeline \citep{TESSELLATE}.
+To see what \texttt{TESSELLATE} detects and to help progress the pipeline, one can participate in the Cosmic Cataclysms Zooniverse project\footnote{\href{https://www.zooniverse.org/projects/cheerfuluser/cosmic-cataclysms}{https://www.zooniverse.org/projects/cheerfuluser/cosmic-cataclysms}}.
+
 Asteroids can spike the brightness of a pixel for only a few frames while it is passing by.
-This can confuse the pipeline looking for transients such as flare stars and supernova, as they can look very similar, having the brightness change in one pixel for a short time.
-Knowing where the asteroids are will allow for the removal of these spikes from the transient pipeline, as well as to understand the asteroid population better.
-By filtering the asteroid detections out, while also self consistently determining rotation properties of a large sample of the population, more science can be done with the one set of analyses.
-
-\section{Methods}\label{Sec:Meth}
+When the pipeline is looking for transients such as flare stars and supernovae, asteroids can look very similar.
+There is a brightness change in one pixel for a short time, as the object moves through the field.
+Finding the asteroids will allow for these spikes to be filtered from the transient pipeline, while also gaining a better understanding of the asteroid population.
+By filtering the asteroid detections out, while also self-consistently determining the rotation properties of a large sample of the population, more science can be done with one set of analyses.
 
 The \texttt{TESSELLATE} pipeline has been running on the OzSTAR supercomputing facilities.
-I built this asteroid detection and subsequent lightcurve and periodogram analysis program by testing on one subsection of one CCD of one camera  of one TESS sector.
+I built this asteroid detection and subsequent lightcurve and periodogram analysis program by testing one subsection of one CCD of one camera  of one TESS sector.
 The code written can be accessed at my ASTR480 GitHub repo \footnote{\href{https://github.com/ble61/ASTR480}{https://github.com/ble61/ASTR480}}.
@@
 Once I was confident this worked as required, the same code was refactored slightly to work on OzSTAR and a large-scale analysis of a processed TESS sector was run.
-% I will be looking for completeness of detections of asteroids, as well as accuracy of periods and amplitude variation.
+I have been looking for completeness of detections of asteroids in \texttt{TESSELLATE}, as well as accuracy of periods and amplitude variation calculated.
 
-For this work, I present analysis of TESS sector 29\footnote{\url{https://tess.mit.edu/observations/sector-29/}}, which was observed between \nth{26} August and \nth{22} September 2020.
-This was the third TESS sector to be using \qty{10}{\min} cadence FFIs, which has had less analysis of the asteroids preformed.
-It was chosen mainly because of this higher cadence, it was the only sector of this cadence that had been processed by \texttt{TESSELLATE} with enough time for my analysis to be preformed.
-What is presented for this section is scalable, on a per-sector basis, to any sector that has been reduced.
+Here, I present an analysis of TESS sector 29\footnote{\url{https://tess.mit.edu/observations/sector-29/}}, which was observed between \nth{26} August and \nth{22} September 2020.
+Sector 29 was the third TESS sector employing \qty{10}{\min} FFIs and this cadence has been focused on less in the literature. 
+It was chosen because it was the only sector of this cadence that had been processed by TESSELLATE with enough time for my analysis to be performed.
+However, what is presented in this section is scalable
+on a per-sector basis to any sector that has been reduced.
 
-\subsection{Querrying Databases for Asteroid Positions and Properties}\label{SubSec:Querry}
+\subsection{Querying Databases for Asteroid Positions and Properties}\label{SubSec:Querry}
 
 To check for asteroids in the TESS data, the positions of the asteroids with time are required.
-For asteroids bright enough to be in TESS, their ephemeris is well known, so I needed to access their positions and then cross match them with transients in the TESS data.
+For asteroids bright enough to be in TESS, their ephemeris is well known, so I needed to access their positions and then cross-match them with transients in the TESS data.
 
-\texttt{SkyBoT} \citep{Berthier2006} was queried to get positions of asteroids in a cone search.
-A box in right ascension (RA) and  declination (Dec) space.
-The search was limited to only asteroids with an average visual magnitude $V\geq \qty{20}{\mag}$.
-As TESS sectors are \qty{27}{\day} long, so querying every \qty{12}{\hour} is both computationally feasible and allows for a good understanding of a bodies motion.
+The sky body tracker, \texttt{SkyBoT} \citep{Berthier2006}, was queried to get the positions of asteroids.
+A box in right ascension (RA) and declination (Dec) space was drawn on the sky, centred in the middle of each subsection made by \texttt{TESSELLATE} and a cone search was executed by \texttt{SkyBoT}.
+The search was limited to only asteroids with an average visual magnitude $V\geq \qty{20}{\mag}$ because of the analysis of \citet{Pal2018}.
+As TESS sectors are \qty{27}{\day} long, querying every \qty{12}{\hour} is computationally feasible and also allows for a good understanding of a body's motion.
 
-These positions are still very sparsely spaced in time compared to the TESS data.
-An interpolation is needed to bridge the gap, this is preformed by \texttt{numpy} \citep{Harris2020}.
-
-Knowing where the asteroids are is helpful for finding them in the archival TESS data, but knowing more about the individual asteroids is also useful for population statistics.
-The \texttt{astroquery} \citep{Ginsburg2019} package was used to query \texttt{JPL Horizons} was used to obtain the orbital elements of each asteroid, $a$, $e$, and $i$, as well as an absolute magnitude $H$, which is a good proxy for size (as discussed above).
-%TODO ref
+Knowing where the asteroids were helped to find them in the archival TESS data, but knowing more about the individual asteroids is also useful for population statistics.
+The \texttt{Astroquery} \citep{Ginsburg2019} package was used to query \texttt{JPL Horizons} was used to obtain the orbital elements of each asteroid, $a$, $e$, and $i$, as well as an absolute magnitude $H$.
+The proper elements do not change with the asteroid's positions in its orbit, so only one query is needed per object.
 
@@
   \includegraphics[width =0.8 \textwidth]{./Figures/interpPos\_29\_1\_2\_2.pdf}
-  \caption[Interpolated Positions of Asteroids]{
+  \caption[Interpolated Positions of Asteroids on Sky]{
     The interpolated positions of asteroids in one subsection of a TESS sector.
@@
 
-With TESS data coming in $\qty{10}{\minute}$ chunks, 72 interpolated points are needed between each API call.
-A linear interpolation between the queried RA and Dec is preformed at the times of the TESS frames.
+With TESS data coming in $\qty{10}{\minute}$ chunks, these positions are still very sparsely spaced in time compared to the TESS data.
+72 interpolated points are needed between each \texttt{SkyBoT} query.
+A linear interpolation between the queried RA and Dec is performed by \texttt{NumPy} \citep{Harris2020} at the times of the TESS frames.
 \texttt{TESSELLATE} saves out these time-frame pairs as part of its analysis, allowing for the time coordinate of any position to be exactly at a frame time.
@@
 These interpolated positions can be seen in \autoref{Fig:interpPos} for a subsection of sector from \texttt{TESSELLATE}.
-There are a few interesting features, such as the asteroids are moving in the same direction, indicated by the colouring, they come in from low RA and Dec and tend to increase both coordinates as the month of the sector progresses.
-There is also a large size range in this slice of sky, ranging over \qty{5}{\mag} in absolute magnitude, which can be seen in the alpha (or transparency) of the tracks in \autoref{Fig:interpPos}.
+There are a few interesting features, such as the asteroids are moving in the same direction, indicated by the colouring with time.
+They tend to decrease their RA and increase their Dec as the month of the sector progresses, generally moving from the bottom right to the top left of the figure.
+There is also a large size range in this slice of sky, ranging over \qty{6}{\mag} in absolute magnitude, which can be seen in the alpha (or transparency) of the tracks in \autoref{Fig:interpPos}.
 
-Because of a small change in position between each frame, \citep[$\sim \qty{1}{\px}$ per \qty{30}{min} frame][]{Pal2018}, this interpolation should be accurate.
-Checking against a higher frequency query to the \texttt{JPL Horizons} ephemeris  confirmed this accuracy on a few asteroids, as shown in \autoref{Fig:errPos}.
-This is of asteroid ``(8654) 1990 KC1'', the largest difference is only a third of a TESS pixel in RA, so the interpolations are accurate enough for these measurements.
-All others checked are similarly accurate.
-This is not a feasible way of getting all the positions, as the queries are per name and time, not per area and time like the cone search in \texttt{SkyBoT}.
-They are also rate limited, so while a few asteroids can be checked, filling the gaps between the \texttt{SkyBoT} half day queries for all the asteroids would be too expensive.
+Because of a small change in position between each frame, \citep[$\sim \qty{1}{\px}$ per \qty{30}{min} FFI][]{Pal2018}, this interpolation should be accurate.
+Checking this prediction against a higher frequency query to the \texttt{JPL Horizons} ephemeris confirmed this accuracy on a few asteroids, as shown in \autoref{Fig:errPos}.
+This is of the asteroid (8654) 1990 KC1, but all other objects checked are just as accurate.
+The largest difference is only a third of a TESS pixel in RA, so the interpolations are accurate enough for these measurements.
+
+The high frequency queries are not a feasible way of getting the positions of all the asteroids.
+The \texttt{JPL Horizons} queries are specific to the asteroid name and time of observation, so each asteroid needs as separate query for each TESS frame.
+The cone search \texttt{SkyBoT} query would still be needed, to determine what asteroids are in the field to begin with.
+All other objects checked with \texttt{Horizons} are similarly accurate to \autoref{Fig:errPos}, so it is also not needed for reliable positions.
+The queries are also rate limited, so while a few asteroids can be checked, filling the gaps between the \texttt{SkyBoT} half day queries for all the asteroids would be too expensive.
 
@@
 Matching these interpolated positions to \texttt{TESSELLATE} detections is important to lower the unknown transient outputs of this pipeline.
-Having interpolated their positions, the asteroids have a well sampled set of RA, Dec and time values of where they should be in the TESS data.
-They should show up in a single pixel in $xy$ space for a small number of frames, of order $\sim1$.
+Taking previously unidentified detections and assigning them to an asteroid category, the remaining unknown transients can be more efficiently searched for other interesting astrophysical phenomena.
+Having interpolated their positions, the asteroids have a well-sampled set of RA, Dec and time values of where they should be in the TESS data.
+They should show up in a single pixel in $xy$ space for a few frames, of order $\sim1$.
 This is the same as a lot of other transient events, a sharp rise in brightness and then disappearing quickly again.
-The number of frames for events is variable, type Ia supernova will brighten is a matter of a few hours and then dim for days, while stellar flares are of similar profile by a smaller max brightness and a correspondingly shorter decay time.
 
+The number of frames for events is variable \citep{TESSELLATE}.
+For example, type Ia supernova will brighten is a matter of a few hours and then dim for days, while stellar flares are of a similar profile but have a smaller maximum brightness and a correspondingly shorter decay time.
 Asteroids are very short events, however these detection pipelines are robust and do pick them up.
-These pipelines have already found the aforementioned supernova, stellar flares and variable stars, by matching to catalogues of those types of objects (Gaia, ZTF etc.). %? cite gaia
-A Goal of this work is to catch all the asteroids in the set of all the transients that are returned from the pipeline.
-To do this, catalogue matching is in order.
+A goal of this work is to catch all the asteroids in the set of all the transients that are returned from the pipeline.
+To achieve this, catalogue matching is in order.
 
@@
 A cut is then made by distance, which must be smaller than \qty{0.01}{\degree} (\qty{36}{\arcsec}), which is within 1.7 TESS pixels.
-Then, if multiple detection match to the same interpolated point, the detection with the smallest distance to the match is taken.
+Then, if multiple detections match the same interpolated point, the detection with the smallest distance to the match is taken.
 
-An example of the matches found using this \texttt{KDTree} is shown in \autoref{Fig:RADecMatch}.
-It is clear that \texttt{TESSELLATE} is not detecting every frame that an asteroid is in.
-Detections in \texttt{TESSELLATE} are done on a per pixel basis, and asteroids only cross each pixel for a short number of frames.
-The residuals of \autoref{Fig:RADecMatch} are all within a TESS pixel, shown on the right hand axis of each of the lower panels.
+An example of the matches found using this \texttt{KDTree} is shown in \autoref{Fig:RADecMatch}, and appear to be coinciding with the asteroid both spatially and temporally.
+\texttt{TESSELLATE} is not detecting every frame that an asteroid is in.
+Detections in \texttt{TESSELLATE} are done on a per-pixel basis, and asteroids only cross each pixel for a short number of frames.
+The residuals of \autoref{Fig:RADecMatch} are all within a TESS pixel, shown on the right-hand axis of each of the lower panels.
+The small residuals mean the filtering done after the \texttt{KDTree} was run worked as intended.
 
@@
 These all culminate in too small of a change in flux of the pixel from normal brightness and the frames where the asteroid is there.
-The change has to be quite extreme for \texttt{TESSELLATE} to detect it.
+The change has to be large for \texttt{TESSELLATE} to detect it, so only when the asteroid will only be matched when it is brightest.
 
@@
   \includegraphics[width =\textwidth]{./Figures/DetectMatchPosUlysses.pdf}
-  \caption[Interpolated and Detected Positions]{The RA (right) and Dec (left) of interpolated position (blue dots) or the matches to \texttt{TESSELLATE} detections (pink hexagons) against the time of observation.
+  \caption[Interpolated and \texttt{TESSELLATE} Detected Positions Comparison]{The RA (right) and Dec (left) of interpolated position (blue dots) or the matches to \texttt{TESSELLATE} detections (pink hexagons) against the time of observation.
 The lower panels show the residuals of the difference between the positions for points that occur at the same time.}
@@
 The matches from the detections, which already have a flux calculated, and the interpolated points themselves, which are more numerous but require forcing the photometry.
-There were some challenges getting the flux of the interpolated points, even when \texttt{TESSELLATE} has already reduced all the FFIs of interest.
+\texttt{TESSELLATE} had already reduced all the FFIs using \texttt{TESSreduce}, I did not do any of the data reduction.
 
-The flux is calculated with \texttt{Photutils} \citep{Bradley2024} aperture photometry with a \qty{1.5}{\px} radius.
-This aperture is placed on the centre of mass (COM) position of a $5\times5$ \unit{\px} box around the integer $xy$ coordinate of the asteroid.
-The COM was calculate using \texttt{Scipy} and the \texttt{ndimage.center\_of\_mass}.
-The integer coordinates are found with the \texttt{SkyCoord} module of \texttt{Astropy}, the world coordinates, in RA and Dec, can be transformed into pixel coordinates, $x$ and $y$, using a world coordinates system (WCS) from TESSELLATE.
+The flux for the forced photometry is calculated with \texttt{Photutils} \citep{Bradley2024} aperture photometry.
+A \qty{1.5}{\px} radius aperture used, as this is the standard size used by the rest of \texttt{TESSELLATE}.
+It is placed on the centre of mass (COM) position of a $5\times5$ \unit{\px} box around the integer $xy$ coordinate of the asteroid.
+The COM was calculated using \texttt{Scipy} and the \texttt{ndimage.center\_of\_mass} routine.
+The integer coordinates are found with the \texttt{SkyCoord} module of \texttt{Astropy}, the world coordinates, in RA and Dec, can be transformed into pixel coordinates, $x$ and $y$, using a world coordinates system (WCS) from \texttt{TESSELLATE}.
 
-The integer coordinates were not used themselves because the asteroids seemed to be leading the object in any given frame of TESS. %?FIG?
+The integer coordinates from the interpolated positions were not used themselves because the asteroids seemed to be leading the object in any given frame of TESS.
 This was decreasing the total flux of the asteroid, as the aperture  was hitting mostly background.
@@
 
-I did not take the highest pixel value in a box to make this correction because there could be a rouge star in the field that it detects as the maximum, instead of the asteroid.
-The COM technique would still be affected by such contamination, but the magnitude would be less.
-The photometry on the COM aperture did fix the sawtoothing, as there is subpixel accuracy in the COM calculation.
+I did not take the highest pixel value in a box to correct this effect because there could be a rouge star in the field that it detects as the maximum, instead of the asteroid.
+The COM technique would still be affected by such contamination, but the total change in position would be less.
+The photometry on the COM aperture did fix the sawtoothing, as there is sub-pixel accuracy in the COM calculation.
 There was a far smoother transition from maximum to minimum, as expected from rotating asteroids.
@@
 %TODO data cleaning explaination
-A few steps need to be taken to clean these COM lightcurves.
-Following \citet{McNeill2023}, any fluxes more than $3\sigma$ from the mean value are sigma-clipped recursivly from the forced photometry lightcurves, until convergence is achieved.
+A few steps need to be taken to clean these lightcurves.
+Following \citet{McNeill2023}, any fluxes more than $3\sigma$ from the mean value are sigma-clipped recursively from the forced photometry lightcurves, until convergence is achieved.
 The contamination from background stars the object passes is too great to allow in the lightcurve.
-This does introduce gaps into an otherwise regularly sampled lightcurve, but the Lomb-Scargle periodogram is designed to deal with poorly sampled data.
+This does introduce gaps into an otherwise regularly sampled lightcurve, but this will not be a problem for the rotation period analysis.
 Lightcurves with an average flux $<10$ counts are lost in the random noise floor, so a minimum mean flux limit needs to be applied to ensure quality and reliability of any periods calculated.
@@
 If an object moves between these, there will be a few frames where it is found in multiple of these subsections.
-This makes the lightcurve have two simultaneous observations of the asteroid, which clearly is not possible.
-At the very edge of the subsections, the asteroid can have a COM position that would have the aperture be summing over pixels that are outside the field, and thus the asteroid would have a lower flux value on the edge.
+This produces lightcurves that have two simultaneous observations of the asteroid, which clearly is not possible.
+At the very edge of the subsections, the asteroid can have a position that would have the aperture be summing over pixels that are outside the field, and thus the asteroid would have a lower flux value on the edge.
 The overlap is great enough that the decrease in flux should only be happening on one of the edges at a time.
-So, sorting by flux on a per asteroid basis, then removing the duplicate observation times, keeping only those with the highest flux, will clean the lightcurve.
+By removing the duplicate observation times, keeping only those with the highest flux to filter out the edge cases, lightcurve is cleaned.
 
@@
   \includegraphics[width =\textwidth]{./Figures/SingleBodyLCUlyssesgapSpan.pdf}
-  \caption[Light curves Example]{The light curves of the asteroid (5254) Ulysses throughout its entire journey across sector 29.
-    The flux of the interpolated positions in blue circles,the \texttt{TESSELLATE} flux for any matched points in the pink hexagons, and the COM flux in orange diamonds.
-    The mid-sector gap is highlighted by the grey box.}
+  \caption[Lightcurve Example]{The light curves of the asteroid (5254) Ulysses throughout its entire journey across sector 29.
+    The flux of the interpolated positions in blue circles, the \texttt{TESSELLATE} flux for any matched points in the pink hexagons, and the COM flux in orange diamonds.
+    The mid-sector gap is highlighted by the grey shaded box.}
   \label{Fig:FullSingleLC}
@@
 A comparison between the three light curves is interesting.
-\autoref{Fig:FullSingleLC} shows these three light curves for a single body.
-The COM flux seems much more stable than the interpolated positions flux, and has a higher average flux count.
-This should result in more accurate periodograms as there is less overall variation in the data caused by the misplaced aperture.
+\autoref{Fig:FullSingleLC} shows these three light curves for a single asteroid, (5254) Ulysses.
+The COM flux is much more stable than the interpolated positions flux, and has a higher average flux count.
+This should result in a more accurate period analysis as there is less overall variation in the data caused by the misplaced aperture.
+All the change in flux is instead due to the rotation of the asteroid.
 
-The matched points are comparatively sparse for this object, which is not the case for all.
-Many have no matches, and some have almost fully matched, as discussed above (\autoref{SubSec:Match}) this will be discussed more later in \autoref{SubSec:Detect}.
+The matched points are comparatively sparse for this object, as discussed above (\autoref{SubSec:Match}, \autoref{Fig:RADecMatch} is also (5254) Ulysses).
+This is not the case for all asteroids.
+Many have no matches, and some have almost fully matched.
+The match rates for the sample in sector 29 will be discussed more later in \autoref{SubSec:Detect}.
 
@@
 The rotation period and other properties of the lightcurve will be calculated from the forced photometry on the COM positions. 
-the matching is still beneficial to the \texttt{TESSELLATE} pipeline as a whole.
-Taking previously unidentified detections and assigning them to an asteroid category, the remaining unknown transients can be more efficiently searched for other interesting astrophysical phenomena.
+The matching is still beneficial to the \texttt{TESSELLATE} pipeline as a whole, but is too sparse for this rotation analysis.
 
@@
 For example, it may have been sigma clipped due to a spike in brightness that is unphysical for an asteroid, or TESS sees the same asteroid in the same sector on either side of the day long mid-sector break it takes for data transfer.
-The periodogram reduces to both a fast Fourier transform or a direct least squares fit in appropriate limits, but outperforms them markedly in unevenly sampled data.
+The periodogram reduces to both a fast Fourier transform or a direct least squares fit in appropriate limits, but outperforms them markedly in unevenly sampled data \citep{VanderPlas2018}.
 
@@
 I have used the Lomb-Scargle periodogram as implemented by \texttt{Astropy}\citep[\citet{Astropy2022} but see][for the implementation]{Vanderplas2012,Vanderplas2015}.
-The times are converted from MJD into seconds before the periodogram is calculated, so the frequency is calculated in \unit{\hertz}.
+The times are converted from days to seconds before the periodogram is calculated, so the frequency is calculated in \unit{\hertz}.
 The flux values have their mean subtracted so the change in flux being analysed is about 0 counts.
@@
 \end{equation}
-where the flux, $F$, is given in terms of time, $t$, the calculated frequency, $f$, and the model parameter $\vec{\theta} = [\theta_0, \theta_1,\theta_2]$.
-These parameters are useful in reconstructing the sinusoid without rerunning the periodogram.
+where the flux, $F$, is given in terms of time, $t$, the calculated frequency, $f$, and the model parameters $\vec{\theta} = [\theta_0, \theta_1,\theta_2]$.
+These parameters are useful in reconstructing the model without rerunning the periodogram.
 \texttt{Astropy} gives a statistic on how confident it is in the detected frequency is, the false alarm probability.
 
-The method used for the periodogram fits is the new \texttt{nifty-ls} package \citep{Garrison2024} that has can be implemented on its own, or by interfacing with \texttt{Astropy}.
+The chosen method used for the periodogram fits is the new \texttt{nifty-ls} package \citep{Garrison2024} that has can be implemented on its own, or by interfacing with \texttt{Astropy}.
 I use the later implementation as the speed-up from the non-uniform fast Fourier transform is helpful.
@@
 The minimum period returned by the periodograms will be set by the Nyquist limit of the data.
-For the \qty{10}{\min} FFIs (the cadence analysed here), the Nyquist period is $\qty{20}{\min}=\frac13\unit{\hour}$, a threefold increase in time sensitivity for fast rotating asteroids.
+For the \qty{10}{\min} FFIs (the cadence analysed here), the Nyquist period is $\qty{20}{\min}=\frac13\unit{\hour}$, a threefold increase in time sensitivity from the longer FFIs
+This will improve the sensitivity of TESS to fast rotating asteroids.
 Based on their experimentation with the \qty{30}{\min} cadence FFIs, \citet{McNeill2023} find that periods less than \qty{3}{\hour} be interpreted with caution, due to lack of reliability in recovering injected sources.
-Because of the increase in FFI rate, this time will be droped to \qty{1}{\hour} for the minimum reliable period, as the lightcurve sampling on these timescales is equivalent.
+Because of the increase in FFI rate, this time will be decreased to \qty{1}{\hour} for the minimum reliable period, as the lightcurve sampling on these timescales is equivalent.
 
-A \qty{17}{\day} maximum period for  periodogram analysis follows \citeauthor{McNeill2023}.
-A lack of signal in the rotation curve causes the calculated period to default this low frequency limit.
-Such cases can be filtered out by keeping only periods less than \qty{90}{\percent} of this maximum.
+A \qty{17}{\day} maximum period for periodogram analysis is recommended by \citeauthor{McNeill2023}.
+If there is a lack of signal in the rotation curve, the calculated period to defaults this long period limit.
+Such cases can be filtered out by keeping only periods less than \qty{90}{\percent} of this maximum, and removing all others from further analysis.
 
-I originally tried using the \texttt{Lightkurve} \citep{Lightkurve2018} package built for period analysis of TESS (and Kepler) data of variable stars.
-There were interesting similarities and differences between the packages.
+For the periodograms, I originally tried using the \texttt{Lightkurve} \citep{Lightkurve2018} package built for period analysis of TESS (and Kepler) data of variable stars.
+There were interesting similarities and differences between the implementations.
 \texttt{Lightkurve} was easy to use but did not allow for fine-tuning of the periodogram, or give any model of the best-fitting curve it produced.
@@
 
-An example of such an \texttt{Astropy} Lomb-Scargle periodogram and the corresponding fit to the lightcurve is given \autoref{Fig:PeriodEx}, along with the fit residuals.
-The periodogram is limited below at a power of $10^{-3}$ to highlight only the regions with large power.
-The periodogram calculated that this object had a frequency double that listed in the LCDB, as seen with the vertical dotted lines.
+An example of such an \texttt{Astropy} Lomb-Scargle periodogram and the corresponding model fit to the lightcurve is given in \autoref{Fig:PeriodEx}, along with the residuals of the fit.
+It is calculated that (5254) Ulysses had a frequency double that listed in the LCDB, as seen with the vertical dotted lines.
 This is expected due to the double-peaked nature of most asteroid lightcurves \citep{McNeill2023}.
 This asteroid was highlighted as an alias of the known frequency is recovered by this analysis.
-To account for this, the detected periods will be doubled for comparison to published works.
+To account for this, the detected periods will be doubled to obtain the true rotation period of the object.
 
-The stared peak in \autoref{Fig:PeriodEx} is  the most prominent peak in the periodogram, more that is seems at first glance due to the $\log$ scale.
+The peak highlighted with a gold star in \autoref{Fig:PeriodEx} is the most prominent peak in the periodogram.
+The periodogram is limited below at a power of $10^{-3}$ to highlight only the regions with large power.
 Not all the asteroids have such well-defined maximums, and the peak power of the periodogram is an important factor in its quality.
-A peak power less than 0.1 means there is little to no signal observed and it will be almost indistinguishable from the noise.
+A peak power less than 0.1 means there is little to no signal observed, and it will be almost indistinguishable from the noise.
 
@@
 The amplitude of the sinusoid appears to be slightly lower than the variation of the data, but it is consistent with the scatter.
-This model is successfully fit over a break in the lightcurve, and ignores the observations with a lower flux that did not quite get sigma clipped.
+This model is successfully fit over the mid-sector break in the lightcurve, and ignores the observations with a lower flux that did not quite get sigma clipped as they are within $3\sigma$ of the mean flux.
+
+The lower panel in \autoref{Fig:PeriodEx} shows the residuals of the model subtracted from the data.
+For the majority of the observations, the residuals are close to the 0 line, indicating a high quality fit.
+There are places where there is a significant deviation from 0, corresponding to places in the lightcurve that look out of place to the eye.
+These large residuals could be removed by decreasing the tolerance on the sigma clipping, but that could cut true signal in lightcurves with a large variation amplitude.
 
@@
 with $\Delta m$ being the variation amplitude.
-The $F_{min}$ and $F_{max}$ terms in \autoref{Eq:DelMag} and the minimum and maximum flux of the model, respectivly.
+The $F_{min}$ and $F_{max}$ terms in \autoref{Eq:DelMag} and the minimum and maximum flux of the model, respectively.
 These are measured in counts.
-
-The lower panel in \autoref{Fig:PeriodEx} shows the residuals of the model subtracted from the data.
-For the majority of the observations, the residuals are close to the 0 line, indicating a high quality fit.
-There are places where there is a significant deviation from 0, corresponding to places in the lightcurve that look out of place to the eye.
-These large residuals could be removed by decreasing the tollerance on the sigma clipping, but that could cut true signal in lightcurves with a large variation amplitude.
 
@@
 So there must be checks on the quality of the values calculated.
-Some of these have been discussed already, but for completeness I list again here.
+Some of these have been discussed already, but for completeness I list them again here.
 A lightcurve must have a mean flux of at least 10 counts to be reliable.
-For an accurate periodogram, it is required that the lightcurve also contain more than 200 observations, which is very easy for TESS and its \qty{10}{\minute} cadence, as the asteroid only has to be in the sector for \qty{33.3}{\hour}.
+For an accurate periodogram, it is required that the lightcurve also contain more than 200 observations.
+This is very easy for TESS and its \qty{10}{\minute} cadence, as the asteroid only has to be in the sector for \qty{33.3}{\hour}.
 The periodogram itself must have a peak power of at least 0.1 to be distinguished from the noise
-The calculated period must be $\qty{1}{\hour}\leq P \leq \qty{367.2}{\hour}$, where the upper bound is \qty{90}{\percent} of the \qty{17}{\day} window suggested by \cite{McNeill2023}.
+The calculated period must be $\qty{1}{\hour}\leq P \leq \qty{367.2}{\hour}$, where the upper bound is \qty{90}{\percent} of the \qty{17}{\day} window suggested by \cite{McNeill2023} is taken from their recommendation and adjusted for the shorter FFI times.
 
@@
   \includegraphics[width=\textwidth]{./Figures/LCBDcompFinal.pdf}
-  \caption[LCDB known period against detected period]{A comparison of the detected periods of the objects in the sector to their known periods in the LCDB \citep{Warner2009}, coloured by the mean flux of the object.
+  \caption[Detected Period Compared to Known Period (LCDB)]{A comparison of the detected periods of the objects in the sector to their known periods in the LCDB \citep{Warner2009}, coloured by the mean flux of the object.
     Crosses failed the quality checks, while circles passed them.
     Names of asteroids are referring to the data point to the right of the text.
-    Dotted lines of a constant relationship are plotted, coloured the same as \autoref{Fig:PeriodEx}.
+    Dotted lines of a constant relationship are plotted, coloured to show the same relationship to the known value as \autoref{Fig:PeriodEx}.
   }
@@
 This means over \qty{80}{\percent} of the asteroids that pass all the checks do not have known periods.
+The periods plotted are the output of the periodogram, they have not yet been doubled to account for the double-peaked lightcurve.
 
@@
 If a cross is bright yellow, the periodogram checks would have been responsible, as the average flux is well above the necessary 10 counts.
-The one bright asteroid that failed the checks in the top left is right on the boundary of the \qty{1}{\hour} lower limit, indicating that this bound may be slightly too lenient.
-The vertical line of points on the right-hand side of \autoref{Fig:LCDBcomp} is on that maximum of the window of searched periods at $\qty{408}{\hour}$, and demonstrates why the \qty{90}{\percent} upper bound is necessary.
+The one bright asteroid that failed the checks in the top left is on the boundary of the \qty{1}{\hour} lower limit, indicating that this bound may be slightly too lenient.
+The vertical line of crosses on the right-hand side of \autoref{Fig:LCDBcomp} highlights the maximum window of searched periods at $\qty{408}{\hour}$, and demonstrates why the \qty{90}{\percent} upper bound is necessary.
 Many asteroids, across an order of magnitude of known periods and across the range of mean fluxes, have detected periods at the edge of the window, showing that the periodogram will return this value when no signal is present.
@@
 It is not recovered, as it is almost too bright for TESS, and the flux bleeds into too many pixels for a rotation period to be accurately obtained from the \qty{1.5}{\px} aperture used.
-An exception to the methods is not made here, no reanalysis is attempted.
+An exception to the methods is not made here, and no reanalysis is attempted.
+However, the periodogram and lightcurve are given in the appendix, \autoref{ApFig:Ceres}.
 This object is the main reason for the arrow on the colour bar indicating higher mean flux values without a change in colour.
-(1) Ceres has a mean flux is over 100000 counts which, if included, completely washes out the gradient.
+(1) Ceres has a mean flux is over 150000 counts which, if included, completely washes out the gradient.
 200 counts was chosen as a replacement value because of the range of detail that can be identified.
@@
 If a circle is dark blue, it may only just be above the minimum requirements for average flux.
-The bright circles in the lower right of \autoref{Fig:LCDBcomp} are cause for concern, as these are bright asteroids that are passing the checks with an incorrect period.
 A range of brightness pass the checks, indicating sensitivity to a wide range of distances and sizes.
@@
 
+The yellow circles in the lower right of \autoref{Fig:LCDBcomp} are cause for concern, as these are bright asteroids that are passing the checks with an incorrect period.
+There are a dozen asteroids that pass the checks with large detected periods when compared to their known period.
+These points imply the checks are not perfect.
+Some experimentation was done on the limiting values, and the ones presented here achieved the highest recovery of accurate asteroid periods, while minimising the number that were incorrectly recovered.
+
 The $y=2x$ line of \autoref{Fig:LCDBcomp} is of interest here, the known periods being twice the detected period.
 This line corresponds to the calculated frequency being an alias of double the known frequency, as frequency and period are inverses of each other.
-This is expected \citep{McNeill2023}, this line is what the periodogram should return due to double-peaked lightcurves.
+This is expected, the periodogram should return values on this line due to double-peaked lightcurves.
 As seen in \autoref{Fig:PeriodEx}, (5254) Ulysses falls on this line, highlighted by text in \autoref{Fig:LCDBcomp}.
@@
 
-There are about a dozen asteroids that pass the checks with large detected periods when compared to their known period.
-These points in the lower right of \autoref{Fig:LCDBcomp} imply the checks are not perfect.
-Some experimentation was done on the limiting values, and the ones presented here achieved the highest recovery of accurate asteroid periods, while minimising the number that were incorrectly recovered.
-
-The idea of doubling the detected period of all the asteroids and calling that number the rotation period has had some doubt cast on it, as there are a couple of points recovered nicely on the $y=x$ line, indicating that the true period is found.
-Either the LCDB has not accounted for the alias, which is unlikely as I only took the periods with high quality codes as they suggest \citep{Warner2009}, or the rotation signal sometimes returns the true period.
+The idea of doubling the detected period was called into question due to the two of points recovered nicely on the $y=x$ line, indicating that the true period is found.
+Either the LCDB has not accounted for the alias,  or the rotation signal sometimes returns the true period.
+The former is unlikely as I only took the periods with high quality codes as recommended by \citet{Warner2009}.
 The latter implies the lightcurve is not double-peaked, as is normally expected.
 On the whole, there are enough periods that pass the checks with incorrect periods that these extra erroneous periods will not significantly change the fraction of reliable detections.
+
+In total, 18 asteroids have periods that are not compatible with the LCDB.
+With the 59 objects that pass the quality checks, this means \qty{70}{\percent} of the rotation periods calculated are accurate.
+Including those periods that find a different alias of the rotation rate, falling near the $y=x$ or $y=0.5x$ lines, this value only increase to \qty{75}{\percent}. 
+Considering that the detected periods are doubled to obtain the rotation period, the \qty{70}{\percent} accuracy is a better reflection of the data.
 
@@
 The exception being centaurs, which have an $a$ too larger for the detail in the distribution of the majority of the objects to be seen, and comets, which have an undefined $a$.
-The 374 objects that passed all the quality checks have a similar distribution of orbital parameters to the total sample of 5664 small bodies in the sector.
-This shows just how beneficial TESS is, and why its full sky coverage is so important to solarsystem science.
+Shown in green dots, the 374 objects that passed all the quality checks have a similar distribution of orbital parameters to the total sample of 5664 small bodies in the sector.
+This shows just how beneficial TESS is, and why its full sky coverage is so important to solar system science.
 
-The distinction between the resonances in the main belt can be seen.
+The distinction between the resonances in the main belt can be seen in \autoref{Fig:AEI}.
 There are vertical strips separating the middle belt from the inner and outer belt.
@@
 A some Mars crossers and near earth asteroids (NEAs) at progressively smaller $a$ and a few more objects that are scattered with high $e$ and/or $i$ complete the distribution.
-Many objects have failed the quality checks on their lightcurve and periodogram, but the ones that pass have a similar distribution in orbital parameters to the sample as a whole.
 
@@
   \includegraphics[width=\textwidth]{./Figures/AEIplotqualCut.pdf}
-  \caption[aei distribution]{
+  \caption[Proper Element Distribution]{
     The $a$ against both $e$ and $i$ of the objects that \texttt{SkyBoT} found in Sector 29 with $V <\qty{20}{\mag}$.
@@
 As expected from the orbital elements in \autoref{Fig:AEI}, the majority of the minor planets fall in the three regions of the main belt.
-There are some objects that are not asteroids, 20 comets were in the field and 1 centaur.
-While they have a V mag less than 20, these objects are often not bright enough to have lightcurves with high enough mean fluxes to be detected well by TESS.
+There are some objects that are not asteroids, 20 comets were in the field, along with one centaur.
+While they have a V \unit{\mag} less than 20, these objects are not bright enough to have lightcurves with high enough mean fluxes to be detected well by TESS. 
 As seen by the green bars in \autoref{Fig:NumPerClass}, around \qty{5}{\percent} of any class have lightcurves and periodograms that pass the quality checks, and this is relatively class independent.
-The highest fraction is the detection of 1 object in the Amor family of NEAs.
-However, with only 9 total objects in the dataset, the sample is too small to make claims on the detectability of this class as a whole.
+The highest fraction is the detection of one object in the Amor family of NEAs.
+However, with only nine total objects in the dataset, the sample is too small to make claims on the detectability of this class as a whole.
 
@@
   \includegraphics[width=\textwidth]{./Figures/classesBarqualCut.pdf}
-  \caption[Class distribution]{A bar chart distribution of the total number of objects per class that \texttt{SkyBoT} found in the Sector (orange) and the same distribution of those that passed all the quality checks (green).
+  \caption[Asteroid Class Distribution]{A bar chart distribution of the total number of objects per class that \texttt{SkyBoT} found in the Sector (orange) and the same distribution of those that passed all the quality checks (green).
     The numbers above each class (coloured the same) correspond to the height of the bar, as some classes have too few objects to be seen.
@@
   \includegraphics[width=\textwidth]{./Figures/pointsMatchesNumberHistdoubleChangedbound.pdf}
-  \caption[Number of observations or matches]{\textit{Left}: A histogram of the number of observations in the lightcurve of each asteroid (orange solid line) in 100 observation bins.
+  \caption[Observation and Match Number Distribution]{\textit{Left}: A histogram of the number of observations in the lightcurve of each asteroid (orange solid line) in 100 observation bins.
 \textit{Right}: The number of matches each asteroid gets in the \texttt{TESSELLATE} detections (black dashed).
@@
 The quality cut of 200 observations in a lightcurve only cuts out $\sim 200$ asteroids.
-Most asteroids are well sampled, with the mode of the distribution in the 2900 to 3000 observation bin, which means the asteroids are in a large fraction of the about 3500 frames in a sector.
+Most asteroids are well sampled, with the mode of the distribution in the 2900 to 3000 observation bin, which means most of the asteroids are in a large fraction of the  $\sim3500$ frames in a sector of this cadence.
 
@@
 Yet the y-limit is too large for detail to be seen in any higher bin of this plot.
-There are only a total of 101 asteroids, out of the 5664 in the sector, that are matched more than 100 times to \texttt{TESSELLATE} detections.
+There are only a total of 99 asteroids, out of the 5664 in the sector, that are matched more than 100 times to \texttt{TESSELLATE} detections.
 So using the matches as lightcurves for periodogram analysis would mean most fail the minimum observations check, which limits the sample before any periods are calculated.
@@
 The other checks cut down in the middle of the brightness range of objects rather evenly.
-The brightest bins, which only have a handful of asteroids in them are not reliable data points, as the values they can take on are too discrete.
-(1) Ceres has $V<\qty{8}{\mag}$ and is not recovered, as discussed in \autoref{SubSec:QualCheckVer}, and is the only object in this magnitude bin, so a 0 recovery fraction is returned.
+The brightest bins are not reliable data points because they only have a few asteroids in them. The values they can take on are too discrete.
+(1) Ceres has $V<\qty{8}{\mag}$ and is not recovered, as discussed in \autoref{SubSec:QualCheckVer}, and is the only object in this magnitude bin, so a recovery fraction of 0 is returned.
 
@@
 where $y$ is the recovered fraction, $\alpha$ is how the steep the transition from 1 down to 0 is,  $m$ the magnitude and $m_{1/2}$ is the shift along the magnitude axis.
-The constant of 0.5 in \autoref{Eq:Atan} raises the point of inflection to be at $\frac 1/2$, and the $\frac{-1}{\pi}$ makes the function take the expected maximum at bright magnitudes at \qty{100}{\percent} recovered, down to no dim asteroids being detected.
+The constant of 0.5 in \autoref{Eq:Atan} raises the point of inflection to be at $\frac12$, and the $\frac{-1}{\pi}$ forces the range of the function to between the expected maximum at bright magnitudes at \qty{100}{\percent} recovered, down to \qty{0}{\percent} when dim asteroids are not detected.
 The $m_{1/2}$ parameter dictates the limiting magnitude, as dimmer than this value less than half of the asteroids are recovered.
@@
 There is a significant change in $m_{1/2}$ between the recovery curves in \autoref{Fig:RecovPassorFail}.
-When only the minimum average flux of 10 counts requirement is applied, the limiting magnitude is \qty{17.78(0.01)}{\mag}.
-When all the quality checks are imposed, this gets over a \unit{\mag} brighter, to \qty{16.46(0.37)}{\mag}.
+When only the minimum average flux of 10 counts requirement is applied, the limiting magnitude is $m_{1/2}= \qty{17.83(0.03)}{\mag}$.
+When all the quality checks are imposed, this gets over a \unit{\mag} brighter, to \qty{16.36(0.73)}{\mag}.
 It is expected that a higher threshold for quality will make the limiting magnitude brighter.
-The $m_{1/2}$ for the full suite of quality checks is rather  uncertain because of the scatter in the recovery fraction at brighter magnitudes, and the shallower decrease in efficiency.
+The $m_{1/2}$ for those that pass the full suite of quality checks is rather uncertain because of the scatter in the recovery fraction at brighter magnitudes.
+The shallower decrease in efficiency also affects the uncertainty.
+As it is no longer strictly a brightness cut-off that is being imposed on the data, the recovery fraction does not quite follow the $\arctan$ shape.
 
@@
 
-\begin{figure}[h]
+\begin{figure}[!b]
   \centering
@@
 
-For the 374 asteroids that pass the quality checks, relable rotation periods are obtained.
+For the 374 asteroids that pass the quality checks, reliable rotation periods are obtained.
 \autoref{Fig:PerDistr} shows the distribution of these rotation rates.
-The period plotted is double the detected period, to account for the double peaked nature of most lightcurves.
-The mode of the distribution is a period less that \qty{1}{\day}, by a factor of 10 over periods in the \qtyrange{1}{2}{\day} bin, the \nth{2} most populated.
+The period plotted is double the detected period, to account for the double-peaked nature of most lightcurves.
+The mode of the distribution is a period less than \qty{1}{\day}, by a factor of 10 over periods in the \qtyrange{1}{2}{\day} bin, the \nth{2} most populated range.
 
-\begin{figure}[!hb]
+\begin{figure}[!t]
   \centering
@@
 The high $\Delta m$ values are from lightcurves of dubious quality.
-Manually checking the analysis, it is clear that these points above $\Delta m =2$ have all been affected by background contamination, and have untrustworth photometry.
-Some examples are given in the appendix for asteroids with known periods.
+Manually checking the analysis, it is clear that these asteroids above $\Delta m =\qty{1.5}{\mag}$ have all been affected contamination in their lightcurves, and have untrustworth photometry.
+Some examples are given in the appendix, Figures \ref{ApFig:Takashimizuno} \ref{ApFig:Wuyeesun} and \ref{ApFig:Locarno}
 
@@
 For some objects, analysis of the residuals could show a pattern that leads to a period and possible fit of a model, but this is not the case for all of them.
-For only one sector worth of data, it was managable for me to preform a manual check on these few outliers
-But this will not be the case for all the whole dataset \texttt{TESSELLATE} will analyse.
+For only one sector worth of data, it was manageable to preform a manual check on these few outliers.
+This will not be the case for all the whole dataset \texttt{TESSELLATE} will analyse.
 
-\begin{figure}[h]
+\begin{figure}[!h]
   \centering
@@
 All 374 of the objects that pass the quality checks are plotted here, and no trend is observed.
-The variation is quite constant accros the range of rotation periods, although there are less objects at longer periods.
+The variation is quite constant across the range of rotation periods, although there are fewer objects at longer periods.
 
 The modes of each of the distributions in \autoref{Fig:PerDistr} and \autoref{Fig:ampDistr} do not fully overlap, as indicated by \autoref{Fig:perVSamp}.
-Those asteroids with rotation periods $<\qty{24}{\hour}$ have a range of rotation periods.
-Converly, the objects with a $\Delta m \approx 0.3$ span the entire period range.
+Those asteroids with rotation periods $<\qty{24}{\hour}$ have a wide range of amplitudes.
+Conversely, the objects with a $\Delta m \approx 0.3$ span the entire period range.
 
-The position in these axes of \omuamua is shown in \autoref{Fig:perVSamp} by the black caltrop and is highlighted with text.
+The position in the space of these axes for \omuamua is shown in \autoref{Fig:perVSamp} by the black caltrop and is highlighted with text.
 1I's rotation period of \qty{8.67(0.34)}{\hour} from \citet{Belton2018} and a $\Delta m$ of \qty{2.5}{\mag} from \citet{Meech2017} are used.
@@
   \includegraphics[width=\textwidth]{./Figures/Diam-FreqPlotThisWork.pdf}
-  \caption{This work's addition to \autoref{Fig:FreqVsDiam} and the LCDB of \citet{Warner2009}. The diameters are from NEOWISE where available.}
+  \caption[\autoref{Fig:FreqVsDiam} updated by this work and NEOWISE]{This work's addition to \autoref{Fig:FreqVsDiam} and the LCDB of \citet{Warner2009}.
+The diameters are from NEOWISE where available.}
   \label{Fig:FreqDiamUpdate}
@@
 
-\autoref{Fig:FreqDiamUpdate} displays my calculated rotation periods with those in the LCDB of \citet{Warner2009} (\autoref{Fig:FreqVsDiam}).
-269 of the asteroids that pass the quality checks have diameters in the near-Earth object wide-field infrared survey explorer (NEOWISE) V2.0 database \citep{Masiero2011, NEOWISE2019}.
-It is only this sub-sample of the 374 asteroids that are plotted in \autoref{Fig:FreqDiamUpdate}.
-The diameters in the LCBD are not calculated using NEOWISE data, so there could be discrepancies in those asteroids that I have calculated periods for that are also in the LCDB.
-The NEOWISE data is self-consistent and encompasses more asteroids that the diameters from just the LCDB, so it was preferred for the over plotting of the rotation data from this work.
+\autoref{Fig:FreqDiamUpdate} overlays the rotation periods calculated here with those presented in \autoref{Fig:FreqVsDiam}, the LCDB of \citet{Warner2009}.
+269 of the asteroids that pass the quality checks have diameters available. 
+The sizes are from the near-Earth object wide-field infrared survey explorer (NEOWISE) V2.0 database \citep{Masiero2011, NEOWISE2019}.
+It is only this sub-sample of the 374 asteroids with derived periods that are plotted in \autoref{Fig:FreqDiamUpdate}.
+The diameters in the LCBD are not calculated using NEOWISE data, so there could be discrepancies in those asteroids that have calculated periods by this analysis and are also in the LCDB.
+The NEOWISE data is self-consistent and encompasses more asteroids that the diameters from just the LCDB, so it was preferred for this figure.
 
@@
 This is most likely due to these bodies being too dim for the telescope.
-Those that are seen must not pass the quality checks I have imposed.
+Those that are seen have not passed the quality checks imposed upon them.
 
 The spin barrier of \citet{Pravec2000} is not violated by any of the periods calculated.
-The quality checks to do exclued this, as the spin barrier is at $\sim \qty{2.4}{\hour}$, and I only impose the period must be $>\qty{1}{\hour}$.
-This cadence of FFI has the ability to probe the forbidden region, to find more objects that violate the spin barrier.
-Unfortunalty, none were found in sector 29.
+The quality checks to do not exclude a violation, as the spin barrier is at $\sim \qty{2.4}{\hour}$, and the checks only remove periods $<\qty{1}{\hour}$.
+This cadence of FFIs have the ability to probe the forbidden region, to find more objects that violate the spin barrier.
+However, none were found in sector 29.
 
@@
 
-As mentioned in \autoref{SubSec:Asteroid}, this is not the first work done analysing asteroids in TESS data.
-This work has 18 objects in common with \citet{McNeill2023} and an overlap of 13 asteroids with the data of \citet{Pal2020}, shown in \autoref{Fig:TESSLit}.
-There is agreement in the derivered periods for some of the objects in both datasets, as seen by the points on or near the $y=x$ line in each subplot.
-These datasets are in broad, but not complete, agreement for the asteroids they both measure \citep[see the comparison in ][]{McNeill2023}, so it is not surpising that this work is not in complete agreement.
+As mentioned in \autoref{SubSec:TESS}, this is not the first work analysing asteroids in TESS data.
+This work has 18 objects in common with \citet{McNeill2023} and 13 asteroids with the data of \citet{Pal2020}.
+A comparison of the calculated periods is displayed in \autoref{Fig:TESSLit}.
+There is agreement in the derived periods for some objects in both datasets, as seen by the points on or near the $y=x$ line in each sub-plot.
+Both literature datasets are in broad, but not complete, agreement for the asteroids they both measure \citep[see the comparison in][]{McNeill2023}, so it is not surprising that this work is not in complete agreement with either of them.
 
@@
 Both in of these comparison datasets, and in the check against the LCDB in \autoref{Fig:LCDBcomp}, there are periods on either extreme.
-Sometimes the periods found here are much shorter than those found in the external data, and othertimes much longer, but this is in about equal proportions.
-The sample size is really too small to make any statistical statments, once again more sectors need to be analysed.
+Sometimes the periods found here are much shorter than those found in the external data, and others are much longer.
+However, this is in about equal proportions.
+The sample size is too small to make any statistical statements, once again more sectors need to be analysed for a larger overlap to be found.
 
-There are 8 objects that appear in all three datasets.
-Two of them lie almost directly on top of each other in the \citeauthor{Pal2020} data on the right panels\autoref{Fig:TESSLit}, but there is a difference on the left hand side in the \citeauthor{McNeill2023} data.
+There are eight objects that appear in all three datasets, and are highlighted by black circles around the data points.
+Two of them lie almost directly on top of each other in the \citeauthor{Pal2020} data on the right panel of \autoref{Fig:TESSLit}, but there is a difference between them on the left-hand side, the \citeauthor{McNeill2023} data.
 This intersection is shown again in \autoref{Tab:AllPers}, where it is more clear how well some of these data line up.
-Both (13364) 1998 UK20 and (16151) 1999 XF230 have the same rotation period measured in all three analysies.
-My period for (36240) 1999 VN44 is two orders of magnitude larger than either of the two previos studies, so there must have been an issue with the data for this object in sector 29.
+Both (13364) 1998 UK20 and (16151) 1999 XF230 have the same rotation period measured in all three analyses.
+
+The period calculated for (36240) 1999 VN44 is two orders of magnitude larger than either of the two previous studies.
+A spot check (\autoref{ApFig:1999VN44}) shows that the average flux of the object is low, barely passing the quality check.
+This is yet another indication that these checks are not perfect.
+A comparison to these datasets of \citeauthor{Pal2020} and \citeauthor{McNeill2023} was not made until after the checks were finalised, in order to limit the biases from small sample sizes.
 
 I seem to calculate similar periods to \citeauthor{Pal2020} more frequently than with \citeauthor{McNeill2023}, but again this is too small of a sample size.
-I have only analysesed a sector of data, they both measure periods for the same 13 sectors. %!Check
-When more data has been run through both \texttt{TESSELLATE} and my pipelines, then the number of intersecting objects will increase.
+Only one sector of data is analysed here, they both measure periods for the same 13 sectors. 
+When more data has been run through both \texttt{TESSELLATE} and my pipelines, then the number of intersecting objects will increase, and a more thorough comparison can be done.
 
@@
   \centering
-  \caption[TESS Literature Comparison Table]{The 8 asteroids have rotation periods in this work rounded to two decimal places, as well as \citet{Pal2020} and \citet{McNeill2023}.
+  \caption[TESS Literature Comparison Table]{The eight asteroids have rotation periods in this work rounded to two decimal places.
+Their periods as calculated by \citet{Pal2020} and \citet{McNeill2023} are also given.
 All periods are measured in hours (\unit{\hour})} \label{Tab:AllPers}
@@
 
+The distributions of periods and amplitudes also agree with these studies. 
+My \autoref{Fig:PerDistr}  show similar distributions to Figures 11 and 12 in \citet{McNeill2023} and their Figure 13 peaks a little earlier than my \autoref{Fig:ampDistr}. 
+\citet{Pal2020} does not give a variation amplitude distribution, but my period distribution agrees broadly with their Figure 7.
+The reliability percentage of \qty{70}{\percent} derived by comparison to the LCDB (\autoref{Fig:LCDBcomp}) is almost as high as the $\sim \qty{80}{\percent}$ of \citeauthor{Pal2020} and the $\qty{85}{\percent}$ of \citeauthor{McNeill2023}.
+
 The limiting magnitude of the recovery fraction demonstrated in \autoref{Fig:RecovPassorFail} agrees well with other tests of \texttt{TESSELLATE}.
 Concurrent work by another Honours' student \citep{MontillaHons} achieves a similar recovery fraction detection efficiency of $\sim$\nth{17} \unit{\mag}.
-They have been preforming injection/recovery tests across many TESS sectors, and they find that the limiting magnitude is consistent across sectors
+They have been preforming injection/recovery tests across many TESS sectors, and they find that the limiting magnitude is consistent across sectors.
 The cadence of the FFIs also seems to have no effect on the depth of the images.
@@
 \subsection{Future work}
-
 %quality checks aren't great
-It is clear from the outliers in \autoref{Fig:perVSamp} and the clearly incorrect periods of know objects in \autoref{Fig:LCDBcomp} that the quality checks have room for improvement.
-Some of these checks were  taken to agree with the how others analyse similar data, like the maximum period from \citet{McNeill2023}.
-Others were more empherically derived, such as the minimum average flux value.
-There is a delecate balance between detecting as many known periods as possible and maintating strict enough cuts to keep out the junk.
+It is clear from the outliers in \autoref{Fig:perVSamp} and the clearly incorrect periods of known objects in \autoref{Fig:LCDBcomp} that the quality checks have room for improvement.
+Some of these checks were  taken to agree with how others analyse similar data, like the maximum period from \citet{McNeill2023}.
+Others were more empirically derived, such as the minimum average flux value.
+There is a delicate balance between detecting as many known periods as possible and maintaining strict enough cuts to keep unreliable periods out of the data.
 
 The periodogram noise at the low frequency end is quite strong throughout the range of rotation periods.
-Currently, it is responsible for a lot of undetected rotation rates, due to the \qty{90}{\percent} of the \qty{17}{\day} window check, and this noise region having a larger power than any signal.
+Currently, it is responsible for a lot of undetected rotation rates, due to the maximum period check, and this noise region having a larger power than any signal.
 Modelling this excess to try and remove the noise and boost the signal from the rotation could be done to improve the maximum power of the periodogram.
@@
 
-Lomb-Scargle periodograms can deal with large gaps in the time data, but would require the flux values to have the same normilisation.
-\texttt{TESSELLATE} is self consistent in the data reduction for a sector, but comparing fluxes between sectors may have some difficulties.
-Assuming the mean flux of the asteroid should be constant, subtracting the mean and
+Lomb-Scargle periodograms can deal with large gaps in the time data, but would require the flux values to have the same normalisation.
+\texttt{TESSELLATE} is self-consistent in the data reduction for a sector, but comparing fluxes between sectors may have some difficulties.
+By assuming the mean flux of the asteroid should be physically constant, subtracting this off the lightcurves before the periodogram is run, as is already done, will start to mitigate this difference.
+There are also phase and viewing angle effects on the lightcurve that are important to correct for with large time gaps. %????
+These were not considered for the single sector analysis because the change in these values is not significant over a month.
 
-There are also phase and viewing angle effects on the lightcurve that are important to correct for with large time gaps.
-These were not considered for the single sector analysis because
+TESS is sensitive to many classes of solar system objects, as seen in \autoref{Fig:AEI} and \autoref{Fig:NumPerClass}. 
+This class agnostic approach is beneficial for gathering rotation statistics across a wide range of objects. 
+Asteroids physically further away, such as the Jupiter Trojans, will have their detections biased towards the larger, brighter objects, but this is the case of any survey.
+While it is not a TESS specific problem, the high limiting magnitude found here does impose a cut-off in the detection of objects to an even larger size.
 
-%run on rest of TESSELLATE data
+High-cadence imaging of all classes of asteroids will be helpful, as missions are being sent to further away populations.
+For example, the Lucy mission \citep{Olkin2021} is en route to the Jupiter Trojan asteroids.
+None of the targets are in sector 29, but they will be in a TESS sector. 
+Extra knowledge of their properties from the TESS data would enhance the analysis of the flybys.
 
@@
 
-It works ok
+This work has analysed the asteroid population in sector 29 of TESS.
+Asteroid positions and properties have been gathered from public sources, \texttt{SkyBoT} and \texttt{JPL Horizons}.
+These positions have been interpolated to match the cadence of TESS, \qty{10}{\min} for the sector analysed here. 
+This interpolation has been shown to be accurate to within a TESS pixel.
+The COM of a small region around the interpolated position, aperture photometry was performed on this position, using a \texttt{TESSELLATE} standard \qty{1.5}{\px} aperture. 
+These lightcurves then underwent minor cleaning to make sure there was no double counting of any position, and no stellar contamination.
 
-but could be better
+A Lomb-Scargle periodogram was used to detect the periods of each object. 
+From the model that \texttt{Astropy} determined was the best fit to the data, a $\Delta m$ was calculated to quantify the amplitude of the lightcurve. 
+The distributions of these broadly agree with the literature.
 
-summerise findings
+Of the 5664 objects with $V<\qty{20}{\mag}$ in found in sector 29, the rotation properties of 374 asteroids are calculated. 
+These are the objects that have met all the requirements placed on their lightcurve and periodogram, and the checks were found to be \qty{70}{\percent} accurate.
+59 of these asteroids have rotation periods documented in the LCDB already \citep{Warner2009}.
+315 new asteroid rotation periods have been determined, as well as the amplitude of the variation in the lightcurve, \qty{80}{\percent} of the total asteroids that pass the checks.
+
+Some bodies appear to have more extreme variation of their lightcurves that the unusual \omuamuans.
+These were checked and have contamination in the lightcurves. 
+This contamination has managed to pass undetected through sigma-clipping and other quality checks on the periodogram. 
+There is evidently a slight issue with my checks
+
+I have shown that \texttt{TESSELLATE} detects asteroids far less often than they appear in the FFIs.
+Most objects are matches $<100$ times, while the forced photometry lightcurves average more than 2000 observations. 
+Those that are matched to TESSELLATE detections can be filtered out of further transient searches. 
+The forced photometry detects more than \qty{50}{\percent} of asteroids brighter than
+
+The main extension of this work is to analyse more TESS sectors. 
+Gathering a larger sample size is always beneficial. 
+Access to more data will also help improve the quality checks on the lightcurve. 
+It does take a long time to run an analysis, both the reductions and detections from \texttt{TESSELLATE} as well as the asteroid specific side of the pipeline I have developed.   
+My methods are easily transferable to more sectors and upon completion of this analysis, more sectors can be analysed.
 
@@
 \addcontentsline{toc}{section}{Acknowledgements}
-%OzSTAR asks for:
-\small
+
+I would like to thank my supervisors for the time and effort they spent making sure things were getting done, for help when I was stuck, and forever striving to improve the quality of my work. 
+I would also like to thank the rest of the ASTR480 cohort for their support throughout the year.
+
+There are a few entities used in this work that request acknowledgement and/or citation for use of their services:
+
+%OzSTAR wants
 This work was performed on the OzSTAR national facility at Swinburne University of Technology.
@@
 
+%MPC wants
 This research has made use of data and/or services provided by the International Astronomical Union's Minor Planet Center.
 
+%Horizon wants
 This work has used data and/or services provided by NASA's Solar System Dynamics website, \url{https://ssd.jpl.nasa.gov/}. Specifically \texttt{JPL Horizons}: \url{https://ssd.jpl.nasa.gov/horizons/}.
 
+%No where else worked
 The figures seen in this work have been plotted by \texttt{Matplotlib} \citep{Hunter2007}, specifically Version 3.9.2 \citep{PLT3.9.2}
 
+%Photutils wants this extra
 This research made use of Photutils, an Astropy package for detection and photometry of astronomical sources \citep{Bradley2024}.
 
+\section*{Appendix}
+\addcontentsline{toc}{section}{Appendix}
+\renewcommand{\thefigure}{A.\arabic{figure}}
+\setcounter{figure}{0}
+
+Here I present some more periodogram examples, Figures \ref{ApFig:Ceres} \ref{ApFig:Takashimizuno}, \ref{ApFig:Wuyeesun} and \ref{ApFig:Locarno}.
+This provides context to what I claim in text especially about the contamination of the lightcurves.
+As they are all so similar to \autoref{Fig:PeriodEx} and quite large, so I do not place them where they are referred to.
+
+\begin{figure}
+  \centering
+  \includegraphics[width = \textwidth]{Figures/PeriodogramCeresResid.pdf}
+  \caption[(1) Ceres Periodogram]{Same as \autoref{Fig:PeriodEx}, but for (1) Ceres.}
+  \label{ApFig:Ceres}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width = \textwidth]{Figures/PeriodogramTakashimizunoResid.pdf}
+  \caption[(6392) Takashimizuno Periodogram]{Same as \autoref{Fig:PeriodEx}, but for (6392) Takashimizuno.
+  This is the highest detected variation amplitude, at $\Delta m = 3.7$.
+  (6392) Takashimizuno is not in the LCDB, so a known period is not plotted.
+  }
+  \label{ApFig:Takashimizuno}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width = \textwidth]{Figures/PeriodogramWuyeesunResid.pdf}
+  \caption[(3570) Wuyeesun Periodogram]{Same as \autoref{Fig:PeriodEx}, but for (3570) Wuyeesun.
+  This is another high variation object, and it is clear a residuals' analysis would be beneficial here, but is not attempted.}
+  \label{ApFig:Wuyeesun}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width = \textwidth]{Figures/PeriodogramLocarnoResid.pdf}
+  \caption[(1937) Locarno Periodogram]{Same as \autoref{Fig:PeriodEx}, but for (1937) Locarno.
+  This is another high variation object, but a residuals' analysis would not help as much.}
+  \label{ApFig:Locarno}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width = \textwidth]{Figures/Periodogram1999 VN44Resid.pdf}
+  \caption[(36240) 1999 VN44 Periodogram]{Same as \autoref{Fig:PeriodEx}, but for (36240) 1999 VN44.
+  This object was in the data of both \citet{Pal2020} and \citet{McNeill2023}, but the calculated period here was not the same.}
+  \label{ApFig:1999VN44}
+\end{figure}
+
+\newpage
 %*Bib

### 2024-10-20 FinalReport/FinalReportDraft.tex
Final Report Done

@@
 \makeatother
-
-\usepackage{chngcntr}
 
@@
 Asteroids move across the TESS field, and their intrinsic rotation varies their brightness periodically. 
-Here, I present an analysis of the asteroids in TESS sector 29, as an addition to the \texttt{TESSELLATE} pipeline's analysis.
+Here, an analysis of the asteroids in TESS sector 29 is presented, as an addition to the \texttt{TESSELLATE} pipeline's analysis.
 
@@
 
-\subsection{Asteroids}\label{SubSec:Asteroid}
+\subsection{Asteroids and their Rotation Properties}\label{SubSec:Asteroid}
 
@@
 
-% The semi-major axis, $a$, is the largest distance from the sun that the asteroid achieves on its elliptical orbit.
-% Main belt asteroids have a typical $a$ of \qtyrange{2}{3.3}{\au} \citep{DeMeo2015}.
-% The bounds between the inner, middle and outer main belt at \qty{2.5}{\au} and \qty{2.82}{\au}, these separations are due to resonances with Jupiter (the 3:1 and 5:2 mean motion resonances mark the edges of middle main belt) and Saturn (the $\nu_6$ secular resonance cuts off the lowest $a$ edge of the inner belt) carving swathes where asteroids cannot be in stable orbits.
-
-% The eccentricity, $e$, of this orbit is the next proper element.
-$e=0$ implies a perfectly circular orbit, $0<e<1$ are ellipses, $e=1$ give a parabolic orbit, and $e>1$ are unbound, hyperbolic orbits.
-% Main belt asteroids have small $e$, in the range of  \citep{DeMeo2015}
-% Comets have larger $e$, with some long period comets having $e\approx 1$, but the typical range is \qtyrange{0.2}{0.7}{} \citep{Lewis2012}.
-% Interstellar objects (ISOs) have $e>1$ as they are on unbound hyperbolic trajectories.
-
-% The inclination, $i$, of asteroids is also a key orbital element, it is the angle relative to the ecliptic plane.
-% In principle asteroids can span the whole range of inclinations, though they rarely do.
-% Main belt asteroids have typical inclinations of \qtyrange{0}{30}{\degree} \citep{DeMeo2015}, so the TESS cameras closer to the ecliptic plane are expected to have the most asteroids in their fields.
-
 Asteroids can be grouped into families based on  the  clustering of the proper orbital elements and similarities in their colour or spectra \citep{Nesvorny2015}.
@@
 
-%TODO Rotation of asteroids
 As an asteroid rotates, the amount of sunlight reflected changes with time.
@@
 
-\subsection{TESS}\label{SubSec:TESS}
+\subsection{TESS and its History of Asteroid Analysis}\label{SubSec:TESS}
 
-%Why TESS
 TESS is a large area, high-cadence imaging, space telescope  \citep{Ricker2014}.
@@
 
-\subsection{Asteroid Properties}\label{SubSec:AstPropRes} %TODO fix name
+\subsection{The Distribution of Asteroid Properties}\label{SubSec:AstPropRes}
 
@@
 
-\subsection{Rotation Period Analysis Results}\label{SubSec:PerRes}
+\subsection{Rotation Period and Amplitude Variation Results}\label{SubSec:PerRes}
 
@@
 The high $\Delta m$ values are from lightcurves of dubious quality.
-Manually checking the analysis, it is clear that these asteroids above $\Delta m =\qty{1.5}{\mag}$ have all been affected contamination in their lightcurves, and have untrustworth photometry.
+Manually checking the analysis, it is clear that these asteroids above $\Delta m =\qty{1.5}{\mag}$ have all been affected contamination in their lightcurves, and have untrustworthy photometry.
 Some examples are given in the appendix, Figures \ref{ApFig:Takashimizuno} \ref{ApFig:Wuyeesun} and \ref{ApFig:Locarno}
@@
 \texttt{TESSELLATE} is self-consistent in the data reduction for a sector, but comparing fluxes between sectors may have some difficulties.
-By assuming the mean flux of the asteroid should be physically constant, subtracting this off the lightcurves before the periodogram is run, as is already done, will start to mitigate this difference.
 There are also phase and viewing angle effects on the lightcurve that are important to correct for with large time gaps. %????
@@
 From the model that \texttt{Astropy} determined was the best fit to the data, a $\Delta m$ was calculated to quantify the amplitude of the lightcurve. 
-The distributions of these broadly agree with the literature.
+The distributions of these properties broadly agree with the literature.
 
 Of the 5664 objects with $V<\qty{20}{\mag}$ in found in sector 29, the rotation properties of 374 asteroids are calculated. 
-These are the objects that have met all the requirements placed on their lightcurve and periodogram, and the checks were found to be \qty{70}{\percent} accurate.
+These are the objects that have met all the requirements placed on their lightcurve and periodogram, and the checks were found to be \qty{70}{\percent} reliable.
 59 of these asteroids have rotation periods documented in the LCDB already \citep{Warner2009}.
-315 new asteroid rotation periods have been determined, as well as the amplitude of the variation in the lightcurve, \qty{80}{\percent} of the total asteroids that pass the checks.
+So 315 new asteroid rotation periods have been determined, as well as the amplitude of the variation in the lightcurve, \qty{80}{\percent} of the total asteroids that pass the checks.
 
-Some bodies appear to have more extreme variation of their lightcurves that the unusual \omuamuans.
+Some bodies appear to have more extreme variation of their lightcurves than the unusual \omuamuans.
 These were checked and have contamination in the lightcurves. 
 This contamination has managed to pass undetected through sigma-clipping and other quality checks on the periodogram. 
-There is evidently a slight issue with my checks
+There is evidently a issue with my checks, and more data will help to refine them.
 
@@
 Those that are matched to TESSELLATE detections can be filtered out of further transient searches. 
-The forced photometry detects more than \qty{50}{\percent} of asteroids brighter than
+The forced photometry detects more than \qty{50}{\percent} of asteroids brighter than $V = \qty{16.36(0.73)}{\mag}$ from analysis of the recovery fraction.
+This agrees with other work on the detection efficiency of \texttt{TESSELLATE}
 
@@
 
-Here I present some more periodogram examples, Figures \ref{ApFig:Ceres} \ref{ApFig:Takashimizuno}, \ref{ApFig:Wuyeesun} and \ref{ApFig:Locarno}.
-This provides context to what I claim in text especially about the contamination of the lightcurves.
+Here I present some more periodogram examples, Figures \ref{ApFig:Ceres} \ref{ApFig:Takashimizuno}, \ref{ApFig:Wuyeesun} \ref{ApFig:Locarno}, and \ref{ApFig:1999VN44}.
+This provides context to what I claim in the text, especially about the contamination of the lightcurves.
 As they are all so similar to \autoref{Fig:PeriodEx} and quite large, so I do not place them where they are referred to.
@@
   \caption[(36240) 1999 VN44 Periodogram]{Same as \autoref{Fig:PeriodEx}, but for (36240) 1999 VN44.
-  This object was in the data of both \citet{Pal2020} and \citet{McNeill2023}, but the calculated period here was not the same.}
+  This object was in the data of both \citet{Pal2020} and \citet{McNeill2023}, but period the calculated here was not the same, due to the low mean flux.}
   \label{ApFig:1999VN44}

### 2025-08-28 PaperDraft/paperDraft.tex
started a paper draft

\documentclass[12pt]{article}
\usepackage{}
\usepackage[margin=2cm]{geometry}
\usepackage[authoryear,comma,round]{natbib}
\usepackage[separate-uncertainty=true,multi-part-units=single]{siunitx}
\usepackage{datatool,booktabs,hyperref,graphicx,caption,amsmath,amssymb}
\usepackage[list=true,listformat=simple]{subcaption}
\usepackage[para]{footmisc}
\usepackage[super]{nth}
\usepackage[utf8]{inputenc}

\DeclareSIUnit{\arcsec}{''}
\DeclareSIUnit{\au}{\mathrm{AU}}
\DeclareSIUnit{\mag}{\mathrm{mag}}
\DeclareSIUnit{\px}{\mathrm{px}}

\newcommand{\ttt}{\texttt}
\newcommand{\tess}{\textit{TESS}}
\newcommand{\tessellate}{\texttt{TESSELLATE}}
\newcommand{\red}[1]{\textcolor{red}{#1}}

\title{TESSELLATE asteroids paper}
\author{Brayden Leicester, Ryan Ridden-Harper, Michele Bannister and Others}
\begin{document}
\maketitle

\section{Abstract}\label{sec:Abs}
%TODO Later

%\twocolumn %To check size

\section{Introduction}\label{sec:Intro}

%TODO general opening statment about asteroids as transiet events/ moving objects

Since 2017, the Transiting Exoplanet Survey Satellite \citep[\tess,][]{Ricker2014} has been obseving large swathes ($\qty{96}{\degree}\times\qty{24}{\degree}$) of sky at a very high cadence.
These month long blocks of full frame images (FFIs) are named sectors, with the cadence of the sectors dependant on the cycle.
The cadence started at \qty{30}{\minute}, then increased to \qty{10}{\minute}, and has now increased again to \qty{200}{\second} FFIs.

The large archival dataset of \tess\ has gone largely unanalysied.
This is until the \tessellate\ project \citep{TESSELLATE}, which seeks to find all the transient events that were serendipitously observed by \tess.
\tessellate\ searches every pixel of every FFI from \tess, using the \ttt{TESSreduce} \citep{Ridden-Harper2021a} to looking for changes in brightness that could be related to a physical event.

Asteroids are the most numerous source of non-transient detections in \tessellate, and need to be filtered out of the rest of the data.
This also allows for analysis of the properties of these planetesimals.
This letter presents our analysis of the asteroids found with the first set of \tessellate\ sectors, with a \qty{10}{\minute} cadence.
This cadence is fast enough that the Nyquist frequency of the observations is above the spin barrier for asteroids \citep{Pravec2000}, so \tess\ could detect new fast rotators.

There have been other studies done using \tess\ to characterise asteroids.
It was suggested by \citet{Pal2018} that it would be possible to get good photometry down to $V \lesssim \qty{19}{\mag}$.
They followed up with an analysis of the small bodies the first 12 sectors (cycle 1) \citep{Pal2020}, tripleing the number of asteroids with accurate rotation periods.
\citet{McNeill2023} re-analyses the same sectors of cycle 1, and is in broad agreement with \citeauthor{Pal2020}.

\tess\ is used for solar-system observation in many other ways aswell.
The Minor Planet Center (MPC) gets regular updates on asteroids in \tess\ from the \ttt{LINEAR-TESS} program \citep{Woods2021}.
\citet{Gowanlock2024} uses \tess\ photometry as well as observations by the Zwicky Transient Facility \citep[ZTF, ][]{Bellm2019} to get a longer baseline on mutually observed objects while combining ground and space-based observations.
Single object analysis is also possible, \citet{Humes2024} recover the period of an asteroid commensurate with Earth's rotation.
They also use the \tess\ data alongside other observations to determine the shape of the object.
\tess\ has been used to study comets as well \citep[e.g.][]{Ridden-Harper2021b}
Fainter and unknown solar system objects can be found by shift and stacking \citep{Holman2019, Payne2019, Rice2020} or taking a fast X-ray transform \citep{Nguyen2024} of the \tess\ FFI data.

For our analysis of the asteroids in \tess, we use sectors \red{TODO} %TODO
which have a \qty{10}{\minute} cadence.
Our methods are presented in \autoref{sec:Meth}, followed by the results in \autoref{sec:Res}.
A discussion of how our results compare to other \tess\ asteroid studies is given in \autoref{sec:Dis}, and we conclude in \autoref{sec:Conc}.

\section{Methods}\label{sec:Meth}

%-How they are found in TESSELLATE accidentally

\tessellate\ provides a new way to find and characterize asteroids: detecting them as transient events.
The lightcurves of these detections often have spike only a few frames long, as the asteroid moves over the pixel of interest.
Asteroids move at about \qty{1}{\px} per frame for the \qty{30}{\min} cadence \citep{Pal2018}, and thus are found as many different transient events as they cross a sector.
Both \ttt{SourceDetect} and \ttt{Starfind}, the detection methods employed by \tessellate, detect planetesimals.

Characterizing the spikes in the single pixel lightcurves as an ``asteroid'' has it's problems, as real transient events can look similar.
We have implemented another way of finding and removing asteroids from the set of detected events.
The reduction of full sectors during a \tessellate\ run also allows for forced photometry at known asteroid positions.
These can then be matched with the detections to classify the spikes, and properties of the asteroids can be measured at the same time.

%-Conesearch w/ SkyBoT and names w/ Horizons

The position of known asteroids is important for matching them to detections.
We use \ttt{SkyBoT} \citep{Berthier2006} conesearch to find all the asteroids with $V\leq \qty{20}{\mag}$, the limit suggested by \citep{Pal2018}.
We query every \qty{12}{\hour} over the month of the sector, and interpolate the positions to the times of the \tess\ observations.
With \tess's large pixel scale, these interpolations are accurate enough for our purposes, and save on time and resources, as the query can be expensive.
%? They were checked for accuracy against JPL Horizons

%-Matching to TESSELLATE
The positions of these asteroids can be matched with the detections, using a KDTree \citep{Maneewongvatana1999}.
Setting boundaries of \qty{0.1}{\day} and \qty{1}{\px} %TODO check it is 1 px.
on the nearest match keeps only objects that are coincident temporally and spatially.
If two such objects  match to the same detection, the shortest distance between the positions is taken.
The matched events are then classified as asteroids, and are not considered further by \tessellate.%Clunky end to sentance

%-Forced Photometry Lightcurve

Forced photometry along the tracks of the asteroids allows for the construction of lightcurves.
% The positions are readjusted to the center of mass (COM) of the object.
Aperture photometry, using \texttt{Photutils} \citep{Bradley2024}, is preformed at this COM position with a \qty{1.5}{\px} aperture.
The lightcurves are then sigma-clipped to $3\sigma$ from the mean remove background stellar contamination.

%-Lomb-Scargle Periodogram

From their lightcurves, the rotation period of the object can be determined by using a Lomb-Scargle periodogram \citep[\citet{Lomb1976, Scargle1982}, but see][ for a review]{VanderPlas2018}.
The new \ttt{NIFTY-LS} package \citep{Garrison2024},  as implemented in \ttt{astropy} \citep{Astropy2013,Astropy2018,Astropy2022}, was used to calculate the periodogram.
The periodogram searched for rotation periods between the Nyquist limit of \qty{20}{\minute} and a maximum of \qty{17}{\day} \citep[the value used in][ due to the lenght of the lightcurve]{McNeill2023} .

-Example Figure

-Quality Checks on the lightcurves

Not every lightcurve and periodogram is good enough for analysis, so some quality cuts are applied.
A minimum brightness cut of 10 counts $\sim \qty{18}{mag}$ %?which way around is better? 
is applied, otherwise the lightcurve is indistinugshable from noise. %* This is emperical from match to LCDB, should that fig be here?
The periodograms could return the \qty{17}{\day} maximum value if no physical period was detected, so following \citet{McNeill2023} we discard any periods within \qty{10}{\percent} of this maximum.
Also following \citeauthor{McNeill2023}, we do not trust the periods of any lightcurve with $\leq 200$ datapoints.
%TODO are there more checks

%-Run all these are part of TESSELLATE
The above methods are integrated into the rest of the \tessellate\ pipeline. %*I should probably get around to doing that...
This allows for asteroids to be filtered out of the new detections as they happen.
The results given below will be improved upon as more sectors are run and a higher number of minor planets can have their rotation properties measured.

\section{Results}\label{sec:Res}

-Total Asteroids found

\red{A Number \#\#\#} of asteroids with $V\leq \qty{20}{\mag}$ were found in the sectors we analysed.
Most of these are too faint for the lightcurve to give a reliable and period.
Once the quality checks described above have been applied, we are left with \red{A smaller number \#\#\#} of planetesimals with accurate rotation rates.
\red{Some fraction} of these asteroids are not in the LCDB,%TODO check defind before
so we are the first to get the rotation period for these bodies.

-Properties of those that pass

%TODO table of asteroids by class and frac that pass.

\section{Discussion} \label{sec:Dis}

-Comp to Pal and McNeill

-Fast Rotator Plot

We do not find any fast rotators in our dataset.
While our Nyquist limit is well above the spin barrier of \citet{Pravec2000} thanks to \tess's fast cadence, most asteroids discoverd above the barrier are small%TODO CITE Warner2009 LCDB?
, and therefore dim.
The shallow magnitude of \tessellate\ \citep{TESSELLATE} makes it difficult for dim planetesimals to pass the quality checks.

-other discussion

The rotation periods of the brightest asteroids are not recovered.
This is due to the fixed aperture size of our forced photometry, \qty{1.5}{\px}.
The bright asteroids can start to saturate the \tess\ detector, and can bloom out to occupy more pixels.
We chose to keep a constant aperture to standardise the analysis, and to be agnostic to the properties of the asteroid being analysed.

Some asteroids slip through our detetcion methods, and end up unclassifed in the \tessellate\ data.
This is often because the detections of bright asteroids are offset from the predicted positions by further than the matching radius we set.
Asteroids are easy to spot for our Cosmics Cataclysms \ttt{Zooniverse}\footnote{\url{https://www.zooniverse.org/projects/cheerfuluser/cosmic-cataclysms}} volunteers, as they are at a different position in the ``1 hour later'' panel, so any that slip past can be caught here.

\section{Conclusion}\label{sec:Conc}
%TODO Later

\bibliographystyle{apalike}
\bibliography{paperBib.bib}

\end{document}

### 2025-08-28 PaperDraft/paperDraft.tex
A few small changes

@@
 
-%\twocolumn %To check size
+% \twocolumn %To check size
 
@@
 
-The large archival dataset of \tess\ has gone largely unanalysied.
-This is until the \tessellate\ project \citep{TESSELLATE}, which seeks to find all the transient events that were serendipitously observed by \tess.
+The large archival dataset of \tess\ has gone largely unanalysed.
+This is until the \tessellate\ project \defcitealias{TESSELLATE}{H.
+Roxburgh and R.
+Ridden-Harper et al.
+2025}\citepalias{TESSELLATE}, which seeks to find all the transient events that were serendipitously observed by \tess.
 \tessellate\ searches every pixel of every FFI from \tess, using the \ttt{TESSreduce} \citep{Ridden-Harper2021a} to looking for changes in brightness that could be related to a physical event.
@@
 
-For our analysis of the asteroids in \tess, we use sectors \red{TODO} %TODO
-which have a \qty{10}{\minute} cadence.
+For our analysis of the asteroids in \tess, we use the third year of observations from the spacecraft; sectors 27 to 39. 
+These were the first sectors to have a \qty{10}{\minute} cadence.
 Our methods are presented in \autoref{sec:Meth}, followed by the results in \autoref{sec:Res}.
@@
 Forced photometry along the tracks of the asteroids allows for the construction of lightcurves.
+The differenced images are used for the forced photometry, as the objects are moving they are not dimmed by the differencing. 
 % The positions are readjusted to the center of mass (COM) of the object.
-Aperture photometry, using \texttt{Photutils} \citep{Bradley2024}, is preformed at this COM position with a \qty{1.5}{\px} aperture.
+Aperture photometry, using \texttt{Photutils} \citep{Bradley2024}, is preformed at the centre of mass of the point spread function with a \qty{1.5}{\px} aperture, the standard size of an aperture in \tessellate.
 The lightcurves are then sigma-clipped to $3\sigma$ from the mean remove background stellar contamination.
@@
 The new \ttt{NIFTY-LS} package \citep{Garrison2024},  as implemented in \ttt{astropy} \citep{Astropy2013,Astropy2018,Astropy2022}, was used to calculate the periodogram.
-The periodogram searched for rotation periods between the Nyquist limit of \qty{20}{\minute} and a maximum of \qty{17}{\day} \citep[the value used in][ due to the lenght of the lightcurve]{McNeill2023} .
+The periodogram searched for rotation periods between the Nyquist limit of \qty{20}{\minute} and a maximum of \qty{17}{\day} \citep[the value used in][ due to the length of the lightcurve]{McNeill2023}.
 
 -Example Figure
+
+An example of the periodogram fit to the lightcurve is given by \autoref{fig:perEx}. 
+A clear maximum power is found in the top panel, and the associated frequency is fed to the model shown in the middle panel.
+The residuals in the lower panel are small for the most part, with large deviations from 0 only where the lightcurve has large uncertainty in its flux. 
+Thanks to the known frequency from LCDB plotted, we note that we recover the double frequency alias, this is explored further below.
 
@@
 A minimum brightness cut of 10 counts $\sim \qty{18}{mag}$ %?which way around is better? 
-is applied, otherwise the lightcurve is indistinugshable from noise. %* This is emperical from match to LCDB, should that fig be here?
+is applied, otherwise the lightcurve is indistinguishable from noise. %* This is emperical from match to LCDB, should that fig be here?
 The periodograms could return the \qty{17}{\day} maximum value if no physical period was detected, so following \citet{McNeill2023} we discard any periods within \qty{10}{\percent} of this maximum.
+%TODO min period of \qty{1}{\hour}
+%! This and the doubling of the periods means we can't get fast rotators at all. Need to think this through
 Also following \citeauthor{McNeill2023}, we do not trust the periods of any lightcurve with $\leq 200$ datapoints.
@@
 We do not find any fast rotators in our dataset.
-While our Nyquist limit is well above the spin barrier of \citet{Pravec2000} thanks to \tess's fast cadence, most asteroids discoverd above the barrier are small%TODO CITE Warner2009 LCDB?
+While our Nyquist limit is well above the spin barrier of \citet{Pravec2000} thanks to \tess's fast cadence, most asteroids discovered above the barrier are small%TODO CITE Warner2009 LCDB?
 , and therefore dim.
 The shallow magnitude of \tessellate\ \citep{TESSELLATE} makes it difficult for dim planetesimals to pass the quality checks.
+
+%! we also can't actually go below 2hrs even with a 10 min cadence because we double the period.
 
@@
 
-Some asteroids slip through our detetcion methods, and end up unclassifed in the \tessellate\ data.
+Some asteroids slip through our detection methods, and end up unclassified in the \tessellate\ data.
 This is often because the detections of bright asteroids are offset from the predicted positions by further than the matching radius we set.
-Asteroids are easy to spot for our Cosmics Cataclysms \ttt{Zooniverse}\footnote{\url{https://www.zooniverse.org/projects/cheerfuluser/cosmic-cataclysms}} volunteers, as they are at a different position in the ``1 hour later'' panel, so any that slip past can be caught here.
+Asteroids are easy to spot for our Cosmic Cataclysms \ttt{Zooniverse}\footnote{\url{https://www.zooniverse.org/projects/cheerfuluser/cosmic-cataclysms}} volunteers, as they are at a different position in the ``1 hour later'' panel, so any that slip past can be caught here.
 
@@
 
-\bibliographystyle{apalike}
+\bibliographystyle{aasjournal}
 \bibliography{paperBib.bib}

### 2025-09-04 PaperDraft/paperDraft.tex
some numbers added

@@
 
-\red{A Number \#\#\#} of asteroids with $V\leq \qty{20}{\mag}$ were found in the sectors we analysed.
-Most of these are too faint for the lightcurve to give a reliable and period.
-Once the quality checks described above have been applied, we are left with \red{A smaller number \#\#\#} of planetesimals with accurate rotation rates.
+\red{57426} of asteroids with $V\leq \qty{20}{\mag}$ were found in the year 3 sectors we analysed.
+Most of these are too faint for the lightcurve to give a reliable and period, while \citet{Pal2018} thinks good photometry down to $V\lesssim \qty{19}{\mag}$ should be achievable, we find the limit for \tessellate\ is closer to $m_{T} \approx \qty{17}{\mag}$ \citepalias{TESSELLATE}.
+Once the quality checks (described above) have been applied, we are left with \red{4328} of planetesimals with accurate rotation rates.
 \red{Some fraction} of these asteroids are not in the LCDB,%TODO check defind before

### 2025-09-11 PaperDraft/paperDraft.tex
added a few numbers

@@
 Once the quality checks (described above) have been applied, we are left with \red{4328} of planetesimals with accurate rotation rates.
-\red{Some fraction} of these asteroids are not in the LCDB,%TODO check defind before
+\red{\qty{80}{\percent}} of these asteroids do not have reliable periods reported in the LCDB, %TODO check defind before
 so we are the first to get the rotation period for these bodies.
@@
 
-%TODO table of asteroids by class and frac that pass.
+%? TODO table of asteroids by class and frac that pass.
+
+%TODO comp to LCDB
+\red{57426 total, 1530 in LCDB, 868 of those pass check, 581 within 5 percent of y=2x, 28 within 5 percent of y=x}
 

### 2025-09-16 PaperDraft/paperDraft.tex
figures added

@@
 The periodogram searched for rotation periods between the Nyquist limit of \qty{20}{\minute} and a maximum of \qty{17}{\day} \citep[the value used in][ due to the length of the lightcurve]{McNeill2023}.
+The uncertainty in the frequency is calculated as half full width half maximum of the peak found in the periodogram, crudely approximating the peaks as Gaussian.
 
@@
 
+\begin{figure}[!t]
+    \centering
+    \includegraphics[width=\textwidth]{PeriodogramUlyssesPaperFigDraft.pdf}
+    \caption{An example of the periodogram analysis we conduct for each object. 
+    This figure uses the lightcurve of the asteroid (5254) Ulysses from sector 29. 
+    Top panel: A periodogram of the lightcurve, with the known frequency and aliases indicated (from the LCDB).
+    Middle panel: The lightcurve (orange error bars) and model(dashed black). 
+    Lower panel: Residuals of the difference between the model and the lightcurve.}
+    \label{fig:perEx}
+\end{figure}
+
 An example of the periodogram fit to the lightcurve is given by \autoref{fig:perEx}. 
 A clear maximum power is found in the top panel, and the associated frequency is fed to the model shown in the middle panel.
-The residuals in the lower panel are small for the most part, with large deviations from 0 only where the lightcurve has large uncertainty in its flux. 
+The residuals in the lower panel are small for the most part, with large deviations from zero only where the lightcurve has large uncertainty in its flux. 
 Thanks to the known frequency from LCDB plotted, we note that we recover the double frequency alias, this is explored further below.
@@
 
-Not every lightcurve and periodogram is good enough for analysis, so some quality cuts are applied.
-A minimum brightness cut of 10 counts $\sim \qty{18}{mag}$ %?which way around is better? 
+Not every lightcurve is good enough for analysis, so some quality cuts are applied to the lightcurves.
+A minimum brightness cut of 10 counts ($m_T \sim \qty{18}{mag}$) %?which way around is better? 
 is applied, otherwise the lightcurve is indistinguishable from noise. %* This is emperical from match to LCDB, should that fig be here?
-The periodograms could return the \qty{17}{\day} maximum value if no physical period was detected, so following \citet{McNeill2023} we discard any periods within \qty{10}{\percent} of this maximum.
+Following \citeauthor{McNeill2023}, we do not trust the periods of any lightcurve with $\leq 200$ data points.
+
+There were also quality cuts applied to the periodograms. 
+The methods would return the \qty{17}{\day} maximum value if no physical period was detected, so also following \citet{McNeill2023} we discard any periods within \qty{10}{\percent} of this maximum.
 %TODO min period of \qty{1}{\hour}
+They find that periods of $\leq \qty{3}{\hour}$ is also unreliable.
+We are using data with a factor of three higher cadence, so we reduce this condition to $\leq \qty{1}{\hour}$ for the same number of observations sampled per rotation. 
 %! This and the doubling of the periods means we can't get fast rotators at all. Need to think this through
-Also following \citeauthor{McNeill2023}, we do not trust the periods of any lightcurve with $\leq 200$ datapoints.
 %TODO are there more checks
@@
 \red{57426 total, 1530 in LCDB, 868 of those pass check, 581 within 5 percent of y=2x, 28 within 5 percent of y=x}
+%TODO say we think only 70% are accurate.
+
+We present double the periods we detect from periodograms as the rotation rate of these objects.
+This is because $\approx \qty{70}{\percent}$ of the asteroid that pass the quality checks in \autoref{fig:LCDBComp} line within \qty{5}{\percent} of the $y=2x$ line; they have double peaked lightcurves.
+A few percent have the same period detected as in the LCDB, but a larger fraction have periods detected with no relation to their known period.
+Extrapolating this trend to the \red{4328} asteroids that pass the checks, we present an accurate rotation period for $\sim \red{3030}$ of them (unfortunately we cannot say which ones).
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{LCDBCompTotalPaperFigDraft.pdf}
+    \caption{Comparison of the detected period and known period for all the asteroids analysed that are in the LCDB.
+    Those that failed the quality check are shown with crosses, and those that passed are filled circles. They are coloured by the measured mean \tess\ magnitude ($m_T$).
+    The 1:1 correspondence line (blue), and the factor of two aliases (purple ($\times 2$) and red ($\times 0.5$)) are plotted as dotted lines.
+    }
+    \label{fig:LCDBComp}
+\end{figure}
 
@@
 
+\autoref{fig:otherWorksComp}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{M23andP20CompPaperFigDraft.pdf}
+    \caption{Comparing the rotation period of asteroids found in this work to asteroids in other studies that used TESS to investigate asteroids. 
+    Left: The comparison to \citet{McNeill2023} plotted as pink crosses, there are \red{139} objects.
+    Right: The comparison to \citet{Pal2020} plotted as purple pluses, there are \red{117} objects. 
+    Those planetesimals found in all three works are circled (\red{54} asteroids).
+    Lines of 1:1 agreement are plotted (dotted blue).}
+    \label{fig:otherWorksComp}
+\end{figure}
+
 -Fast Rotator Plot
 
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{Diam-FreqPlotThisWorkPaperFigDraft.pdf}
+    \caption{A rotation rate against size plot after the LCDB \citep{Warner2009} (grey
+circles) with this work's addition (blue squares). The
+diameters are from NEOWISE \citep{Masiero2011,NEOWISE2019} where available. The red dashed lines represent the spin barrier of \citep{Pravec2000}.}
+    \label{fig:FreqDiam}
+\end{figure}
+
 We do not find any fast rotators in our dataset.
+This is encapsulated by none of our objects being above the spin barrier shown in \autoref{fig:FreqDiam}.
+
 While our Nyquist limit is well above the spin barrier of \citet{Pravec2000} thanks to \tess's fast cadence, most asteroids discovered above the barrier are small%TODO CITE Warner2009 LCDB?

### 2025-09-18 PaperDraft/paperDraft.tex
"group work"
added far more words

@@
 This also allows for analysis of the properties of these planetesimals.
-This letter presents our analysis of the asteroids found with the first set of \tessellate\ sectors, with a \qty{10}{\minute} cadence.
+This letter presents our analysis of the asteroids found with the third year of \tess\ sectors as analysed by \tessellate, with a \qty{10}{\minute} cadence. %?? Cite in prep something 
 This cadence is fast enough that the Nyquist frequency of the observations is above the spin barrier for asteroids \citep{Pravec2000}, so \tess\ could detect new fast rotators.
@@
 It was suggested by \citet{Pal2018} that it would be possible to get good photometry down to $V \lesssim \qty{19}{\mag}$.
-They followed up with an analysis of the small bodies the first 12 sectors (cycle 1) \citep{Pal2020}, tripleing the number of asteroids with accurate rotation periods.
+They followed up with an analysis of the small bodies the first 12 sectors (cycle 1) \citep{Pal2020}, tripling the number of asteroids with accurate rotation periods.
 \citet{McNeill2023} re-analyses the same sectors of cycle 1, and is in broad agreement with \citeauthor{Pal2020}.
 
-\tess\ is used for solar-system observation in many other ways aswell.
+\tess\ is used for solar-system observation in many other ways as well.
 The Minor Planet Center (MPC) gets regular updates on asteroids in \tess\ from the \ttt{LINEAR-TESS} program \citep{Woods2021}.
@@
 
-%-How they are found in TESSELLATE accidentally
-
+%How they are found in TESSELLATE accidentally
 \tessellate\ provides a new way to find and characterize asteroids: detecting them as transient events.
@@
 
-%-Conesearch w/ SkyBoT and names w/ Horizons
-
+%Conesearch w/ SkyBoT and names w/ Horizons
 The position of known asteroids is important for matching them to detections.
@@
 
-%-Matching to TESSELLATE
+%Matching to TESSELLATE
 The positions of these asteroids can be matched with the detections, using a KDTree \citep{Maneewongvatana1999}.
@@
 %-Forced Photometry Lightcurve
-
 Forced photometry along the tracks of the asteroids allows for the construction of lightcurves.
@@
 
-%-Lomb-Scargle Periodogram
-
+%Lomb-Scargle Periodogram
 From their lightcurves, the rotation period of the object can be determined by using a Lomb-Scargle periodogram \citep[\citet{Lomb1976, Scargle1982}, but see][ for a review]{VanderPlas2018}.
@@
 
--Example Figure
-
+%Example figure
 \begin{figure}[!t]
@@
 
--Quality Checks on the lightcurves
-
+%Quality Checks on the lightcurves
 Not every lightcurve is good enough for analysis, so some quality cuts are applied to the lightcurves.
@@
 is applied, otherwise the lightcurve is indistinguishable from noise. %* This is emperical from match to LCDB, should that fig be here?
+This check is implemented to gain a better agreement when comparing to the known periods of these objects from the LCDB (elaborated on below).
 Following \citeauthor{McNeill2023}, we do not trust the periods of any lightcurve with $\leq 200$ data points.
@@
 
--Total Asteroids found
+%TODO Total Asteroids found
 
@@
 Most of these are too faint for the lightcurve to give a reliable and period, while \citet{Pal2018} thinks good photometry down to $V\lesssim \qty{19}{\mag}$ should be achievable, we find the limit for \tessellate\ is closer to $m_{T} \approx \qty{17}{\mag}$ \citepalias{TESSELLATE}.
-Once the quality checks (described above) have been applied, we are left with \red{4328} of planetesimals with accurate rotation rates.
+Once the quality checks (described above) have been applied, we are left with \red{4328} of planetesimals with accurate rotation rates. %!check double ups, I think this might drop to 4314
 \red{\qty{80}{\percent}} of these asteroids do not have reliable periods reported in the LCDB, %TODO check defind before
@@
 
--Properties of those that pass
-
+%TODO Properties of those that pass
 %? TODO table of asteroids by class and frac that pass.
@@
 %TODO comp to LCDB
-\red{57426 total, 1530 in LCDB, 868 of those pass check, 581 within 5 percent of y=2x, 28 within 5 percent of y=x}
-%TODO say we think only 70% are accurate.
-
-We present double the periods we detect from periodograms as the rotation rate of these objects.
-This is because $\approx \qty{70}{\percent}$ of the asteroid that pass the quality checks in \autoref{fig:LCDBComp} line within \qty{5}{\percent} of the $y=2x$ line; they have double peaked lightcurves.
-A few percent have the same period detected as in the LCDB, but a larger fraction have periods detected with no relation to their known period.
-Extrapolating this trend to the \red{4328} asteroids that pass the checks, we present an accurate rotation period for $\sim \red{3030}$ of them (unfortunately we cannot say which ones).
-
 \begin{figure}
@@
 
+%\red{57426 total, 1530 in LCDB, 868 of those pass check, 581 within 5 percent of y=2x, 28 within 5 percent of y=x}
+
+We find \red{1530} of our total population of asteroids are in the LCDB with reliable periods. 
+A comparison to our detected periods is given in \autoref{fig:LCDBComp}, and \red{868} asteroids pass the quality checks we describe above. 
+The majority of these (\red{581} planetesimals) lie within \qty{5}{\percent} of the $y=2x$ line, indicating a double peaked lightcurve.
+Only \red{28} objects have periods in direct agreement to the LCDB, far more of them are in disagreement. 
+Most of the disagreement comes from systematically higher periods detected, and we are not sure why the known period is not found.
+
+%// TODO say we think only 70% are accurate. 
+Following this comparison, we present double the periods we detect from periodograms as the rotation rate of these objects.
+This is because $\approx \qty{70}{\percent}$ of the asteroid that pass the quality checks in \autoref{fig:LCDBComp} line within \qty{5}{\percent} of the $y=2x$ line; they have double peaked lightcurves.
+A few percent have the same period detected as in the LCDB, but a larger fraction have periods detected with no relation to their known period.
+Extrapolating this trend to the \red{4328} asteroids that pass the checks, we present an accurate rotation period for $\sim \red{3030}$ of them (unfortunately we cannot say which ones).
+
+%// TODO justify 5% val.
+We use an absolute difference between periods of $\leq\qty{5}{\percent}$ of our period as the agreement factor for a few reasons.
+Direct agreement is tricky, floating point values and how precise they are presented varies between datasets, so exact equality is rare.
+We did not want to have a set length of time to be the cut-off, as this would bias agreement to short periods. 
+We are comparing to multiple datasets, the LCDB here and later \citet{McNeill2023} and \citet{Pal2020}, so we use \qty{5}{\percent} of our value for consistency.
+
+The brightness of the object during the observation seems to be the main factor for the passing of the quality checks.
+We do apply a minimum brightness cut $(m_T \lesssim \qty{18}{\mag})$, there are plenty of objects with average $m_T$ between \nth{17} and \nth{18}\unit{\mag} in \autoref{fig:LCDBComp} that don't pass the other checks applied.
+This is in agreement with the recovery rates of \tessellate\ at around \qty{17}{\mag} \citepalias{TESSELLATE}.
+Raising the brightness cut we apply only removes some of the agreement with the LCDB, it doesn't drop the number of objects that pass the checks with periods far too large.
+
 \section{Discussion} \label{sec:Dis}
 
--Comp to Pal and McNeill
+%TODO Comp to Pal and McNeill
 
-\autoref{fig:otherWorksComp}
+Both \citet{Pal2020} and \citet{McNeill2023} published tables of the rotation properties for their asteroids.
+For those objects we have in common, we compare the periods we find in \autoref{fig:otherWorksComp}.
+There is good agreement in both cases, but our longer period detections haunt us here too.
+
+For \citet{McNeill2023}, the left panel of \autoref{fig:otherWorksComp}, we agree (to within \qty{5}{\percent} of our derived period) on \red{45} of the \red{139} objects. 
+These are the asteroids that lie on the dotted $y=x$ line, and are close to one third of the asteroids.
+They find higher periods for about a third of the asteroids, and then our incorrect high periods make the last third.
+
+We find a higher agreement with \citet{Pal2020}, \red{63} of \red{117} objects within the \qty{5}{\percent} limits. 
+So over half of our measurements agree with theirs, which is intriguing as our methods followed \citet{McNeill2023} more closely.
+Most of the disagreement comes from out high period scatter and the systematic lower period found by \citet{Pal2020}.
+
+There are \red{54} asteroids in both all three datasets. 
+\red{37} of them lie close to the $y=x$ line in the right \citeauthor{Pal2020} panel of \autoref{fig:otherWorksComp}, while only \red{14} agree with the measurements of \citeauthor{McNeill2023}. 
+There are an extra \red{5} objects that \citet{Pal2020} and \citet{McNeill2023} find similar periods for, but we calculate a different period.
 
@@
 
--Fast Rotator Plot
+%TODO Fast Rotator Plot and discussion
 
@@
 This is encapsulated by none of our objects being above the spin barrier shown in \autoref{fig:FreqDiam}.
-
-While our Nyquist limit is well above the spin barrier of \citet{Pravec2000} thanks to \tess's fast cadence, most asteroids discovered above the barrier are small%TODO CITE Warner2009 LCDB?
+While our Nyquist limit is well above the spin barrier of \citet{Pravec2000} thanks to \tess's fast cadence, most asteroids discovered above the barrier are small%? TODO CITE Warner2009 LCDB?
 , and therefore dim.
-The shallow magnitude of \tessellate\ \citep{TESSELLATE} makes it difficult for dim planetesimals to pass the quality checks.
+The shallow magnitude of \tessellate\ \citepalias{TESSELLATE} makes it difficult for dim planetesimals to pass the quality checks.
 
 %! we also can't actually go below 2hrs even with a 10 min cadence because we double the period.
+Our presented periods could never be above the spin barrier, due to the minimum period quality check. 
+We individually examine the \red{4} objects that pass the other checks, but have a period detected of less than an hour, and it is clear that the check is required.
+\red{3} three of the objects have known periods in the LCDB, 2 of these are much longer than those we detect, and for the \red{\nth{3}} we find the $\times 4$ alias, instead of the usual $\times 2$.
+The quality of the lightcurves of these objects is bad, they have large gaps and the fluxes vary wildly, even with the sigma-clipping done to them.
 
--other discussion
+Through this extra analysis, we trust the quality cuts we have made even more. 
+Unfortunately, this does mean we do not find any fast rotators.
+The fast cadence of \tess\ and the multi-sector search with \tessellate\ were not enough to overcome the low data quality and the rarity of these objects.
+
+%TODO other discussion
 
@@
 This is often because the detections of bright asteroids are offset from the predicted positions by further than the matching radius we set.
-Asteroids are easy to spot for our Cosmic Cataclysms \ttt{Zooniverse}\footnote{\url{https://www.zooniverse.org/projects/cheerfuluser/cosmic-cataclysms}} volunteers, as they are at a different position in the ``1 hour later'' panel, so any that slip past can be caught here.
+Asteroids are easy to spot for our Cosmic Cataclysms \ttt{Zooniverse}\footnote{\url{https://www.zooniverse.org/projects/cheerfuluser/cosmic-cataclysms}} volunteers, as they are at a different position in the ``1 hour later'' panel, so any that slip past can be caught there.
+
+%? TODO Why we detect long periods
+%! I don't know. Ive tried supressing the exponetial that appears at low frequencies and adding the long cutoff. This kind of works, but only so much. 
+%! Can't just take out anything we get above P=X hrs, as there are some slow rotators that we get correct.
 
@@
 %TODO Later
+
+%We find 43XX asteroids with good rotation properties
+
+%Agree with McNeil and Pal in different amounts
+
+%Do not find fast rotators
 

### 2025-09-18 PaperDraft/paperDraft.tex
a bit of editing

@@
 
+As opposed to the population studies, single object analysis is also possible with \tess.
+\citet{Humes2024} recover the period of an asteroid commensurate with Earth's rotation.
+They also use the \tess\ data alongside other observations to determine the shape of the object.
+\tess\ has been used to study comets as well \citep[e.g.][]{Ridden-Harper2021b}
+
 \tess\ is used for solar-system observation in many other ways as well.
@@
 \citet{Gowanlock2024} uses \tess\ photometry as well as observations by the Zwicky Transient Facility \citep[ZTF, ][]{Bellm2019} to get a longer baseline on mutually observed objects while combining ground and space-based observations.
-Single object analysis is also possible, \citet{Humes2024} recover the period of an asteroid commensurate with Earth's rotation.
-They also use the \tess\ data alongside other observations to determine the shape of the object.
-\tess\ has been used to study comets as well \citep[e.g.][]{Ridden-Harper2021b}
 Fainter and unknown solar system objects can be found by shift and stacking \citep{Holman2019, Payne2019, Rice2020} or taking a fast X-ray transform \citep{Nguyen2024} of the \tess\ FFI data.
@@
 The position of known asteroids is important for matching them to detections.
-We use \ttt{SkyBoT} \citep{Berthier2006} conesearch to find all the asteroids with $V\leq \qty{20}{\mag}$, the limit suggested by \citep{Pal2018}.
+We preform a conesearch with \ttt{SkyBoT} \citep{Berthier2006} to find all the asteroids with $V\leq \qty{20}{\mag}$, the limit suggested by \citep{Pal2018}.
 We query every \qty{12}{\hour} over the month of the sector, and interpolate the positions to the times of the \tess\ observations.
@@
 If two such objects  match to the same detection, the shortest distance between the positions is taken.
-The matched events are then classified as asteroids, and are not considered further by \tessellate.%Clunky end to sentance
+The matched events are then classified as asteroids, so other analyses of the \tessellate data can filter them out.
 
@@
 %Quality Checks on the lightcurves
-Not every lightcurve is good enough for analysis, so some quality cuts are applied to the lightcurves.
+Not every lightcurve is good enough for analysis, so some quality cuts are applied to them.
 A minimum brightness cut of 10 counts ($m_T \sim \qty{18}{mag}$) %?which way around is better? 
 is applied, otherwise the lightcurve is indistinguishable from noise. %* This is emperical from match to LCDB, should that fig be here?
+%! No, fits the flow better in Results
 This check is implemented to gain a better agreement when comparing to the known periods of these objects from the LCDB (elaborated on below).
@@
 The methods would return the \qty{17}{\day} maximum value if no physical period was detected, so also following \citet{McNeill2023} we discard any periods within \qty{10}{\percent} of this maximum. 
-%TODO min period of \qty{1}{\hour}
-They find that periods of $\leq \qty{3}{\hour}$ is also unreliable.
+\citet{McNeill2023} find that periods of $\leq \qty{3}{\hour}$ is also unreliable.
 We are using data with a factor of three higher cadence, so we reduce this condition to $\leq \qty{1}{\hour}$ for the same number of observations sampled per rotation.
-%! This and the doubling of the periods means we can't get fast rotators at all. Need to think this through
-%TODO are there more checks
+%// ! This and the doubling of the periods means we can't get fast rotators at all. Need to think this through
+%* Addressed later
 
